### 2015-03-26 Notes/DES_rec_phi_paper.tex
26 March notes improved and DES analysis added.

\input{/Users/alessandromanzotti/Work/preamble_onecol_notitle.tex}
\usepackage{hyperref}
\usepackage{amsmath}
\def\ublu{\bl'}
\newcommand{\bx}{{\bf x}}
\newcommand{\bk}{{\bf k}}
\newcommand{\bq}{{\bf q}}
\newcommand{\bP}{{\bf\Psi}}
\newcommand{\bs}{{\bf s}}
\newcommand{\cs}{{\cal S}}
\newcommand{\by}{{\bf y}}
\newcommand{\deltar}{\delta_{\rm recon}}%\def\lcdm{$\Lambda$CDM}
\def\arcsec{$^{\prime\prime}$}
\def\nl{N_\ell}
\def\sl{S_\ell}
\newcommand{\conv}[2]{\left(\frac{#1}{#2}\right)}
\def\arctanh{\mathop{\rm arctanh}\nolimits}
%\renewcommand{\eqref}[1] {equation $($\ref{#1}$)$}
\newcommand{\comment}[1]{{}}
\def\beq{\begin{equation}}
\def\eeq{\end{equation}}
\def\beqn{\begin{eqnarray}}
\def\eeqn{\end{eqnarray}}
\def\a{\alpha}
\def\cl{C_{l}}
\def\h{\mathrm{h}}
\def\d{\rmn{d}}
\def\pa{\partial}
\def\deldel#1#2{\frac{\pa{#1}}{\pa{#2}}}
\def\ba{\bm{\alpha}}
\def\fracj#1#2{{\textstyle{#1\over#2}}}
\def\bxi{\bm{\xi}}
\def\half{\frac{1}{2}}
\def\ti{\widetilde}
\def\O{\Omega}
\def\OL{\Omega_\Lambda}
\def\Om{\ensuremath{\Omega_{\mathrm{m}}}}
\def\Ob{\ensuremath{\Omega_{\mathrm{b}}}}
\def\Oc{\ensuremath{\Omega_{\mathrm{CDM}}}}
\def\msol{\ensuremath{M_\odot}}
\def\l{\left}
\def\r{\right}
\def\o{\omega}
\def\gcm{\textrm{g cm$^{-3}$}}
\def\2gcm{\textrm{g cm$^{-2}$}}
\def\Scr{\Sigma_{\mathrm{crit}}}
\def\rcr{\rho_{\mathrm{crit}}}
\def\phidot{\ensuremath{\dot\phi_{\bl,\bl'}}}
\def\ddelta{\ensuremath{\dot\delta}}
\def\modu#1{\l |{#1}\r |}
\def\av#1{\l \langle{#1}\r \rangle}
\def\hmpc{\:{h}^{-1}\mathrm{Mpc}}
\def\th{\Theta}
\def\tth{\tilde\Theta}
\def\sg{\sigma}
\def\Sig{\Sigma}
\def\cf{{\cal F}}
\def\k{\kappa}
\def\P{{P}}
\def\pnl{{P}_{{\!\textrm{\tiny NL}}}}
\def\dnl{\Delta_{\text{\scriptsize NL}}}
\def\kmin{\k_{\mathrm{min}}}
\def\kmax{\k_{\mathrm{max}}}
\def\hires{the \emph{highRes} experiment}
\def\lores{the \emph{lowRes} experiment}
\def\ktot{\k_{\mathrm{tot}}}
\def\hunit{\ensuremath{\mathrm{km}{\mathrm{s}^{-1}} \mathrm{Mpc}^{-1}}}
\def\H0{\ensuremath{\mathrm{H}_0}}
\def\nn{\nonumber}
\def\lin{\mathrm{lin}}
\def\ISW{\mathrm{ISW}}
\def\bl{\bmm{l}}
\def\bL{\bmm{L}}
\def\fsky{f_{\mathrm{sky}}}
\newcommand{\E}[1]{\times 10^{#1}}
\newcommand{\bmm}[1]{{\mathbf{#1}}}
\newcommand{\new}[1]{{\color{blue} #1}}
\newcommand{\bsection}[1]{\section{\uppercase{#1}}}

\begin{document}

\title{D.E.S reconstructed lensing potential template and application to delensing and cross correlation }         % Enter your title between curly braces
\author{A. Manzotti}        % Enter your name between curly braces
\date{\today}

        % Enter your date or \today between curly braces
\begin{abstract}
Goal: test if, at least on a small patch of the sky $~500$ deg$^{2}$ LSS survey like D.E.S or DESI can improve the reconstruction of the lensing potential that lenses the CMB photons.
This would be crucial to build a template of the B-mode signal coming from the lensing of the primordial E-mode. This will be at multipoles higher than 200 the main contaminants of the primordial B-mode.
\end{abstract}

\vspace{1cm}

\maketitle

\section{Theory} \label{sec:th}

%=======================================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/spectra.pdf}
\caption{Correlation factor between galaxies survey and CMB lensing potential.}
\label{fig:corr}
\end{center}
\end{figure}
%=======================================

%=======================================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/clbb_res.pdf}
\caption{Residual lensing B modes power spectrum using different large scale structure.}
\label{fig:bres}
\end{center}
\end{figure}
%=======================================

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/compare_kernel.pdf}
\caption{Comparison of the different kernels used in this analysis. This allow to understand  
Redshift distribution of D.E.S galaxies (I suspect this is the benchmark, anyway taken from Giannantionio et al.). }
\label{fig:corr}
\end{center}
\end{figure}
As for the temperature, the intensity map of photons on the sky, also the Q and U mode decomponsition of their polarization is modified by lensing as: 
\be
Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\mathbf{d});~~
U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\mathbf{d})
\ee
where $\mathbf{d}$ is the deflection angle directly related to the lensing potential $\phi$.

As a first approximation, the B mode resulting from the lensing of primordial E mode by a convergence field $\kappa$ is:
\be \label{eq}
B^{\mathrm{lens}}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') E(\ublu) \kappa(\bl - \bl')
\ee
where
\be
W(\bl,\bl') = \frac{2 \bl' \cdot (\bl-\bl')}{|\bl-\bl'|^2} \sin(2\varphi_{\bl,\bl'}),
\ee

As usual we define the power spectrum as:
\be
\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l 
\ee

From this we get that the power spectrum:
\be
C_l^{BB,\mathrm{lens}}  = \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl')C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}.
\ee

Now the full B-mode power spectrum measured on the sky is
\be
C_l^{BB,\mathrm{full}} = C_l^{BB,r} + C_l^{BB,\mathrm{lens}} + N_l^{BB}.
\ee

Now, if we have an LSS measurements I that traces the lensing potential responsible for the lensing of the CMB we can build a template of the true lensing B mode on the sky with a weighted convolution:
\be
\hat{B}^{\mathrm{lens}}(\bl) = \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') f(\bl,\bl') E^N(\ublu) I(\bl - \bl')
\ee
where $f(\bl,\bl')$ is a weight that must be determined.

The residual lensing b mode will be 
\beqn
B^\mathrm{res}(\bl) &=&  B^\mathrm{lens}(\bl) - \hat{B}^\mathrm{lens}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') \times \nonumber \\ &&\left( E(\ublu) \kappa(\bl - \bl' ) -  f(\bl,\bl') E^N(\ublu) I(\bl - \bl' ) \right) 
\eeqn
and its power spectrum
\beqn
C_{{l}}^{BB,\mathrm{res}} &=&   \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl') 
 [ C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|} \nonumber \\ && -(f(\bl,\bl')+f^*(\bl,\bl')) C^{EE}_{l'} C^{\kappa I}_{|\bl-\bl'|} \\ \nonumber && + f^*(\bl,\bl') f(\bl,\bl') (C^{EE}_{{l'}}+ N^{EE}_{{l'}}) C^{II}_{|\bl-\bl'|} ] 
 \eeqn

We can now easily choose $f(\bl,\bl')$ so that the residual lensing B mode power is minimized. We find:
\be
f(\bl,\bl') = \left(\frac{C^{EE}_{{l'}}}{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right)  \frac{C^{\kappa I}_{|\bl-\bl'|}}{C^{II}_{|\bl-\bl'|} } 
\ee
Notice that the first term consists in the usual inverse variance filter applied to the measured E-mode and the second minimize the difference between the reconstructed $\phi$ and the CMB lensing potential.

We finally have that the residual power is:
\beqn
C_{{l}}^{BB,\mathrm{res}} &=&   \int \frac{d^2 \bl'}{(2 \pi)^2}  W^2 (\bl,\bl') 
 C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}  \\
&\times&  \left[1 - \left(\frac{C^{EE}_{{l'}} }{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right) \rho^2_{|\bl-\bl'|} \right] \nonumber
\eeqn
with
\be
\rho_l= \frac{\cl^{\kappa I}}{\sqrt{\cl^{\kappa \kappa} \cl^{I I}}}.
\ee

The bigger $\rho_l$ is for a LSS field the more it is correlated with the lensing potential acting on the CMB photons. An higher correlation allows for a better reconstruction of the $\phi^{CMB}$ and, as a consequence, of $B^{lens}$.

\section{Data} \label{sec:data}

D.E.S, Planck CIB, DESI?
\cite{hall:2010,sherwin:2015,ade:2014,hanson:2013,szapudi:2001,0004-637X-567-1-2,planck-collaboration:2014,planck-collaboration:2014a,planck-collaboration:2011,plank-collaboration:2014,boulanger:1996,lewis:2006}

\section{Forecast} \label{sec:for}

\bibliographystyle{abbrv}
\bibliography{DES_rec_phi_paper}

\end{document}

### 2015-03-26 plotting/plot_Blens.py

@@ -1,0 +1,148 @@
+
+from __future__ import division
+# from __future__ import print_function
+print "\n Python start up script from /home/manzotti/.py_startup.py.\n"
+import scipy as sp
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+
+import matplotlib.pyplot as plt
+
+try:
+    import pyfits
+    import pdb
+
+except:
+    pass
+
+try:
+    import palettable
+
+except Exception, e:
+    pass
+import os
+import os
+import sys
+import time
+import subprocess
+
+from pprint import pprint
+import __builtin__
+import pyximport
+pyximport.install(reload_support=True)
+
+try:
+    cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+    plt.rcParams['image.cmap'] = cmap
+except Exception, e:
+    print 'probably palettable not found'
+
+
+datadir = '../Data/limber_spectra/'
+
+image_dir = '../images/'
+
+
+# Import data
+clbb = np.loadtxt(datadir + 'cbb.txt')
+cbb_res_des = np.loadtxt(datadir + 'cbb_res_des.txt')
+cbb_res_cib = np.loadtxt(datadir + 'cbb_res_cib.txt')
+cbb_res_cib_525 = np.loadtxt(datadir + 'cbb_res_cib_525.txt')
+ell = np.loadtxt(datadir + 'cbb_res_cib.txt')
+
+
+
+
+
+# ============================================
+# SIZE OF THE PICTURE
+# ============================================
+WIDTH = 462.0  # the number latex spits out
+# the fraction of the width you'd like the figure to occupy, 0.5 if double
+# column cause
+FACTOR = 0.6
+fig_width_pt = WIDTH * FACTOR
+
+inches_per_pt = 1.0 / 72.27
+# golden_ratio  = (np.sqrt(5) - 1.0) / 2.0  # because it looks good
+ratio = 0.8
+fig_width_in = fig_width_pt * inches_per_pt  # figure width in inches
+fig_height_in = fig_width_in * ratio   # figure height in inches
+fig_dims = [fig_width_in, fig_height_in]  # fig dims as a list
+# ============================================
+# ============================================
+
+
+# ============================================
+# SET LATEX
+plt.rc('text', usetex=True)
[…]

### 2015-03-26 plotting/plot_corr_fac.py

@@ -1,0 +1,158 @@
+
+from __future__ import division
+# from __future__ import print_function
+print "\n Python start up script from /home/manzotti/.py_startup.py.\n"
+import scipy as sp
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+
+import matplotlib.pyplot as plt
+
+try:
+    import pyfits
+    import pdb
+
+except:
+    pass
+
+try:
+    import palettable
+
+except Exception, e:
+    pass
+import os
+import os
+import sys
+import time
+import subprocess
+
+from pprint import pprint
+import __builtin__
+import pyximport
+pyximport.install(reload_support=True)
+
+try:
+    cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+    plt.rcParams['image.cmap'] = cmap
+except Exception, e:
+    print 'probably palettable not found'
+
+
+datadir = '../Data/limber_spectra/'
+
+image_dir = '../images/'
+
+
+# Import data
+clkappades = np.loadtxt(datadir + 'cldesk.txt')
+clkappacib = np.loadtxt(datadir + 'clcibk.txt')
+cldes = np.loadtxt(datadir + 'cldes.txt')
+clcib = np.loadtxt(datadir + 'clcib.txt')
+clkappa = np.loadtxt(datadir + 'clk.txt')
+ell = np.loadtxt(datadir + 'ells_.txt')
+
+# ADD NOISE
+# using table 6 of 1309.0382v1
+# note: you need the power spectrum in the right units to add the noise this way
+# clcib+=225.
+
+
+# DES NOISE Evaluation
+
+degree_sq = 500
+rad_sq= 500 * (np.pi/180)**2
+fsky = rad_sq/4. /np.pi
+n_gal = 3207184
+nlgg = 1./(3207184./rad_sq) * np.ones_like(cldes) # they mention N=2.1 10^-8 in Tommaso's paper
+
+
+
+
+# ============================================
+# SIZE OF THE PICTURE
+# ============================================
+WIDTH = 462.0  # the number latex spits out
+# the fraction of the width you'd like the figure to occupy, 0.5 if double
+# column cause
+FACTOR = 0.6
+fig_width_pt = WIDTH * FACTOR
+
+inches_per_pt = 1.0 / 72.27
+# golden_ratio  = (np.sqrt(5) - 1.0) / 2.0  # because it looks good
+ratio = 0.8
[…]

### 2015-03-26 plotting/plot_dndz.py

@@ -1,0 +1,134 @@
+
+from __future__ import division
+# from __future__ import print_function
+print "\n Python start up script from /home/manzotti/.py_startup.py.\n"
+import scipy as sp
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+
+import matplotlib.pyplot as plt
+
+try:
+    import pyfits
+    import pdb
+
+except:
+    pass
+
+try:
+    import palettable
+
+except Exception, e:
+    pass
+import os
+import os
+import sys
+import time
+import subprocess
+
+from pprint import pprint
+import __builtin__
+import pyximport
+pyximport.install(reload_support=True)
+
+try:
+    cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+    plt.rcParams['image.cmap'] = cmap
+except Exception, e:
+    print 'probably palettable not found'
+
+
+datadir = '/Users/alessandromanzotti/Work/Software/cosmosis_new/cosmosis-standard-library/structure/PS_limber/output/limber_spectra/'
+
+image_dir = '../images/'
+
+
+# Import data
+des_dndz=np.loadtxt("/Users/alessandromanzotti/Work/Software/cosmosis_new/cosmosis-standard-library/structure/PS_limber/data_input/DES/N_z_wavg_spread_model_0.2_1.2_tpz.txt")
+
+
+
+# ============================================
+# SIZE OF THE PICTURE
+# ============================================
+WIDTH = 462.0  # the number latex spits out
+# the fraction of the width you'd like the figure to occupy, 0.5 if double
+# column cause
+FACTOR = 0.6
+fig_width_pt = WIDTH * FACTOR
+
+inches_per_pt = 1.0 / 72.27
+# golden_ratio  = (np.sqrt(5) - 1.0) / 2.0  # because it looks good
+ratio = 0.8
+fig_width_in = fig_width_pt * inches_per_pt  # figure width in inches
+fig_height_in = fig_width_in * ratio   # figure height in inches
+fig_dims = [fig_width_in, fig_height_in]  # fig dims as a list
+# ============================================
+# ============================================
+
+
+# ============================================
+# SET LATEX
+plt.rc('text', usetex=True)
+plt.rc('font', **{'family': 'serif', 'serif': ['Computer Modern Roman']})
+# ============================================
+
+
+font_size = 10.
[…]

### 2015-03-26 plotting/plot_kernel.py

@@ -1,0 +1,172 @@
+
+from __future__ import division
+# from __future__ import print_function
+print "\n Python start up script from /home/manzotti/.py_startup.py.\n"
+import scipy as sp
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+
+import matplotlib.pyplot as plt
+
+try:
+    import pyfits
+    import pdb
+
+except:
+    pass
+
+try:
+    import palettable
+
+except Exception, e:
+    pass
+import os
+import os
+import sys
+import time
+import subprocess
+
+from pprint import pprint
+import __builtin__
+import pyximport
+pyximport.install(reload_support=True)
+
+try:
+    cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+    plt.rcParams['image.cmap'] = cmap
+except Exception, e:
+    print 'probably palettable not found'
+
+
+datadir = '../Data/limber_spectra/'
+
+image_dir = '../images/'
+
+
+# Import data
+kappakernel_l30 = np.loadtxt(datadir + 'kappa_kernel_l30.txt')
+kappakernel_l100 = np.loadtxt(datadir + 'kappa_kernel_l100.txt')
+kappakernel_l500 = np.loadtxt(datadir + 'kappa_kernel_l500.txt')
+
+kappakernel_l30_h = np.loadtxt(datadir + 'kappa_kernel_l30_h.txt')
+kappakernel_l100_h = np.loadtxt(datadir + 'kappa_kernel_l100_h.txt')
+kappakernel_l500_h = np.loadtxt(datadir + 'kappa_kernel_l500_h.txt')
+
+cibkernel_l30 = np.loadtxt(datadir + 'cib_kernel_l30.txt')
+cibkernel_l100 = np.loadtxt(datadir + 'cib_kernel_l100.txt')
+cibkernel_l500 = np.loadtxt(datadir + 'cib_kernel_l500.txt')
+
+cibkernel_l30_h = np.loadtxt(datadir + 'cib_kernel_l30_h.txt')
+cibkernel_l100_h = np.loadtxt(datadir + 'cib_kernel_l100_h.txt')
+cibkernel_l500_h= np.loadtxt(datadir + 'cib_kernel_l500_h.txt')
+
+
+deskernel_l30 = np.loadtxt(datadir + 'des_kernel_l30.txt')
[…]

### 2015-03-27 Notes/DES_rec_phi_paper.tex
DESI input add

@@
 \usepackage{amsmath}
-\def\ublu{\bl'}
+\usepackage{natbib}
 \newcommand{\bx}{{\bf x}}
@@
 \def\sl{S_\ell}
+\def\ublu{\bl'}
 \newcommand{\conv}[2]{\left(\frac{#1}{#2}\right)}
@@
 This will be at multipoles higher than 200 the main contaminants of the primordial B-mode.
+Preliminary: CIB performs as an equivalent $\rho_{\text{eff}}$  (correlation coefficient constant over $\ell$) of 0.8.
+Both D.E.S and DESI are significantly worse, $\rho_{\text{eff}}<0.6$.
+DESI is slightly better than D.E.S.
+To delens we need the bulk of the redshift distribution to follow the CMB kernel, having a few outliers at redshift $z>1.5$ is not enough.
+We also have to keep in mind the results of \citet{smith:2012}, even a perfect LSS survey can not help too much the delensing technique, with the possible exception of futuristic 21 cm.
 \end{abstract}
@@
 
-\bibliographystyle{abbrv}
+\bibliographystyle{unsrtnat}
 \bibliography{DES_rec_phi_paper}

### 2015-03-27 plotting/plot_Blens.py

@@ -47,11 +47,12 @@ image_dir = '../images/'
 # Import data
 clbb = np.loadtxt(datadir + 'cbb.txt')
 cbb_res_des = np.loadtxt(datadir + 'cbb_res_des.txt')
+cbb_res_desi = np.loadtxt(datadir + 'cbb_res_desi.txt')
 cbb_res_cib = np.loadtxt(datadir + 'cbb_res_cib.txt')
 cbb_res_cib_525 = np.loadtxt(datadir + 'cbb_res_cib_525.txt')
 ell = np.loadtxt(datadir + 'cbb_res_cib.txt')
 
-
+print cbb_res_cib_525
 
 
 
@@ -125,11 +126,13 @@ fg = plt.figure(figsize=fig_dims)
 ax1 = fg.add_subplot(111)
 
 plt.loglog(ell,clbb,label='$C^{BB}$')
-plt.loglog(ell,clbb*(1-0.7**2),linestyle='-.',label=r'$\rho=0.7$')
-plt.loglog(ell,clbb*(1-0.8**2),linestyle='-.',label=r'$\rho=0.8$')
-plt.loglog(ell,clbb*(1-0.6**2),linestyle='-.',label=r'$\rho=0.6$')
+# plt.loglog(ell,clbb*(1-0.7**2),linestyle='-.',label=r'$\rho_{eff}=0.7$')
+plt.loglog(ell,clbb*(1-0.8**2),linestyle='-.',label=r'$\rho_{eff}=0.8$')
+plt.loglog(ell,clbb*(1-0.5**2),linestyle='-.',label=r'$\rho_{eff}=0.5$')
 plt.loglog(ell,cbb_res_des,label=r'D.E.S')
+plt.loglog(ell,cbb_res_desi,label=r'DESI')
 plt.loglog(ell, cbb_res_cib_525,label=r'CIB')
+
 
 
 

### 2015-03-27 plotting/plot_corr_fac.py

@@ -45,12 +45,18 @@ image_dir = '../images/'
 
 
 # Import data
-clkappades = np.loadtxt(datadir + 'cldesk.txt')
-clkappacib = np.loadtxt(datadir + 'clcibk.txt')
-cldes = np.loadtxt(datadir + 'cldes.txt')
-clcib = np.loadtxt(datadir + 'clcib.txt')
-clkappa = np.loadtxt(datadir + 'clk.txt')
-ell = np.loadtxt(datadir + 'ells_.txt')
+clkappades = np.loadtxt(datadir + 'cldeskdes.txt')
+clkappacib = np.loadtxt(datadir + 'clcibkdes.txt')
+cldes = np.loadtxt(datadir + 'cldesdes.txt')
+clcib = np.loadtxt(datadir + 'clcibdes.txt')
+clkappa = np.loadtxt(datadir + 'clkdes.txt')
+ell = np.loadtxt(datadir + 'ells_des.txt')
+
+desi_ell = np.loadtxt(datadir + 'ells.txt')
+cldesi = np.loadtxt(datadir + 'cldesi.txt')
+clkappadesi = np.loadtxt(datadir + 'cldesik.txt')
+desi_clkappa = np.loadtxt(datadir + 'clk.txt')
+
 
 # ADD NOISE
 # using table 6 of 1309.0382v1
@@ -61,12 +67,10 @@ ell = np.loadtxt(datadir + 'ells_.txt')
 # DES NOISE Evaluation
 
 degree_sq = 500
-rad_sq= 500 * (np.pi/180)**2
-fsky = rad_sq/4. /np.pi
+rad_sq = 500 * (np.pi / 180) ** 2
+fsky = rad_sq / 4. / np.pi
 n_gal = 3207184
-nlgg = 1./(3207184./rad_sq) * np.ones_like(cldes) # they mention N=2.1 10^-8 in Tommaso's paper
-
-
+nlgg = 1. / (3207184. / rad_sq) * np.ones_like(cldes)  # they mention N=2.1 10^-8 in Tommaso's paper
 
 
 # ============================================
@@ -100,7 +104,7 @@ plt.rcParams['font.size'] = font_size
 
 plt.rcParams['axes.labelsize'] = font_size * 1.3
 plt.rcParams['axes.linewidth'] = font_size / 22.
-plt.rcParams['axes.titlesize'] = font_size*1.3
[…]

### 2015-03-27 plotting/plot_kernel.py

@@ -70,7 +70,13 @@ deskernel_l30_h = np.loadtxt(datadir + 'des_kernel_l30_h.txt')
 deskernel_l100_h = np.loadtxt(datadir + 'des_kernel_l100_h.txt')
 deskernel_l500_h= np.loadtxt(datadir + 'des_kernel_l500_h.txt')
 
+desikernel_l30 = np.loadtxt(datadir + 'desi_kernel_l30.txt')
+desikernel_l100 = np.loadtxt(datadir + 'desi_kernel_l100.txt')
+desikernel_l500 = np.loadtxt(datadir + 'desi_kernel_l500.txt')
 
+desikernel_l30_h = np.loadtxt(datadir + 'desi_kernel_l30_h.txt')
+desikernel_l100_h = np.loadtxt(datadir + 'desi_kernel_l100_h.txt')
+desikernel_l500_h= np.loadtxt(datadir + 'desi_kernel_l500_h.txt')
 
 # ============================================
 # SIZE OF THE PICTURE
@@ -152,6 +158,11 @@ plt.plot(cibkernel_l30[:,0],cibkernel_l500[:,1]/np.max(cibkernel_l500[:,1]),colo
 plt.plot(deskernel_l30[:,0],deskernel_l30[:,1]/np.max(deskernel_l30[:,1]),color='#4daf4a',label=r'$DES$')
 plt.plot(deskernel_l30[:,0],deskernel_l100[:,1]/np.max(deskernel_l100[:,1]),color='#4daf4a',linestyle=':')
 plt.plot(deskernel_l30[:,0],deskernel_l500[:,1]/np.max(deskernel_l500[:,1]),color='#4daf4a',linestyle='--')
+
+plt.plot(desikernel_l30[:,0],desikernel_l30[:,1]/np.max(desikernel_l30[:,1]),color='#984ea3',label=r'$DESI$')
+# plt.plot(desikernel_l30[:,0],deskernel_l100[:,1]/np.max(desikernel_l100[:,1]),color='#984ea3',linestyle=':')
+# plt.plot(desikernel_l30[:,0],desikernel_l500[:,1]/np.max(desikernel_l500[:,1]),color='#984ea3',linestyle='--')
+
 
 
 

### 2015-04-01 Notes/DES_rec_phi_paper.tex
add combined D.E.S and CIB forecast

@@
 \begin{abstract}
-Goal: test if, at least on a small patch of the sky $~500$ deg$^{2}$ LSS survey like D.E.S or DESI can improve the reconstruction of the lensing potential that lenses the CMB photons.
+Goal: test if LSS survey like D.E.S or DESI can improve the reconstruction of the lensing potential that lenses the CMB photons.
 This would be crucial to build a template of the B-mode signal coming from the lensing of the primordial E-mode.
-This will be at multipoles higher than 200 the main contaminants of the primordial B-mode.
-Preliminary: CIB performs as an equivalent $\rho_{\text{eff}}$  (correlation coefficient constant over $\ell$) of 0.8.
+The lensed B-mode component will be, at multipoles higher than 200, the main contaminants of the primordial B-mode.
+Preliminary: CIB performs as an equivalent $\rho_{\text{eff}}$  (correlation coefficient with cmb lensing potential constant over $\ell$) of 0.8.
 Both D.E.S and DESI are significantly worse, $\rho_{\text{eff}}<0.6$.
@@
 To delens we need the bulk of the redshift distribution to follow the CMB kernel, having a few outliers at redshift $z>1.5$ is not enough.
-We also have to keep in mind the results of \citet{smith:2012}, even a perfect LSS survey can not help too much the delensing technique, with the possible exception of futuristic 21 cm.
+We also have to keep in mind the results of \citet{smith:2012}: even a perfect LSS survey can not help too much the delensing process, with the possible exception of futuristic 21 cm.
 \end{abstract}
@@
 \includegraphics[scale=1.]{../images/spectra.pdf}
-\caption{Correlation factor between galaxies survey and CMB lensing potential.}
+\caption{Correlation factor between galaxies survey and CMB lensing potential as a function of the multipole $\ell$.}
 \label{fig:corr}
@@
 \includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{Comparison of the different kernels used in this analysis. This allow to understand  
-Redshift distribution of D.E.S galaxies (I suspect this is the benchmark, anyway taken from Giannantionio et al.). }
+\caption{Comparison of the different kernels used in this analysis. This allow to understand how well and where in redshift space different LSS surveys trace the CMB lensing potential. 
+Redshift distribution of D.E.S galaxies (I suspect this is the benchmark, anyway taken from Giannantionio et al.). DESI taken from their white paper. }
 \label{fig:corr}
@@
 \end{figure}
-As for the temperature, the intensity map of photons on the sky, also the Q and U mode decomponsition of their polarization is modified by lensing as: 
+As for the temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is modified by lensing as: 
 \be
@@
 
-Now, if we have an LSS measurements I that traces the lensing potential responsible for the lensing of the CMB we can build a template of the true lensing B mode on the sky with a weighted convolution:
+If we have an LSS measurements I($\hat n$) that traces the lensing potential responsible for the lensing of the CMB we can build a template of the lensing B mode on the sky with a weighted convolution:
 \be
@@
 
-The residual lensing b mode will be 
+The residual lensing B mode will be 
 \beqn
@@
 
+\subsection{Galaxies contribution at different redshift}\label{sec:galcontrib}
+
+Let's now assume that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
+It can be shown that the optimal way to combine them to obtain the optimal estimator of $\phi$ or, in other word maximizing the correlation factor $\rho$ is:
+\beqn
+I 		&=&	\sum_{i}c^{i}I^{i} \nonumber \\
+c_{i} 	&=& (C^{-1})_{ij}C^{\kappa I^{j}}
+\eeqn
+where C is the covariance matrix of the LSS tracers.
+
+\be
+\rho^{2} = \sum_{i,j}\frac{C^{\kappa i}~(C^{-1})_{ij}~C^{\kappa j}}{C^{\kappa\kappa}}
+\ee
+
+The gain we  have in adding a new tracer is not only proportional to its correlation with the CMB lensing but it also depends on how much it is correlated with the already used set of tracers.
+For example there s a very small gain in adding different CIB frequencies because they are highly correlated with each other.
+On the contrary a CMB reconstructed lensing potential or some low redshift survey like D.E.S can help.
+
 \section{Data} \label{sec:data}

### 2015-04-01 plotting/plot_Blens.py

@@ -49,10 +49,10 @@ clbb = np.loadtxt(datadir + 'cbb.txt')
 cbb_res_des = np.loadtxt(datadir + 'cbb_res_des.txt')
 cbb_res_desi = np.loadtxt(datadir + 'cbb_res_desi.txt')
 cbb_res_cib = np.loadtxt(datadir + 'cbb_res_cib.txt')
+cbb_res_comb = np.loadtxt(datadir + 'cbb_res_comb.txt')
 cbb_res_cib_525 = np.loadtxt(datadir + 'cbb_res_cib_525.txt')
 ell = np.loadtxt(datadir + 'cbb_res_cib.txt')
 
-print cbb_res_cib_525
 
 
 
@@ -127,12 +127,14 @@ ax1 = fg.add_subplot(111)
 
 plt.loglog(ell,clbb,label='$C^{BB}$')
 # plt.loglog(ell,clbb*(1-0.7**2),linestyle='-.',label=r'$\rho_{eff}=0.7$')
-plt.loglog(ell,clbb*(1-0.8**2),linestyle='-.',label=r'$\rho_{eff}=0.8$')
-plt.loglog(ell,clbb*(1-0.5**2),linestyle='-.',label=r'$\rho_{eff}=0.5$')
+plt.loglog(ell,clbb*(1-0.8**2),linestyle='-.')#,label=r'$\rho_{eff}=0.8$')
+plt.loglog(ell,clbb*(1-0.5**2),linestyle='-.')#,label=r'$\rho_{eff}=0.5$')
 plt.loglog(ell,cbb_res_des,label=r'D.E.S')
 plt.loglog(ell,cbb_res_desi,label=r'DESI')
 plt.loglog(ell, cbb_res_cib_525,label=r'CIB')
+plt.loglog(ell, cbb_res_comb,label=r'CIB+D.E.S')
 
+# plt.loglog(ell, cbb_res_cib,label=r'CIB')
 
 
 

### 2015-04-01 plotting/plot_corr_fac.py

@@ -56,6 +56,8 @@ desi_ell = np.loadtxt(datadir + 'ells.txt')
 cldesi = np.loadtxt(datadir + 'cldesi.txt')
 clkappadesi = np.loadtxt(datadir + 'cldesik.txt')
 desi_clkappa = np.loadtxt(datadir + 'clk.txt')
+# combined rho coming from multiple_survey_delens.py on midway
+rho_comb = np.loadtxt(datadir + 'rho_comb_des_cib.txt')
 
 
 # ADD NOISE
@@ -147,6 +149,7 @@ plt.plot(ell, clkappacib / np.sqrt(clkappa * (clcib)), color='#e41a1c', label='C
 plt.plot(ell, clkappades / np.sqrt(clkappa * cldes), color='#377eb8', label='D.E.S')
 plt.plot(ell, clkappades / np.sqrt(clkappa * (cldes + nlgg)), color='#377eb8', linestyle='--')
 plt.plot( desi_ell, clkappadesi / np.sqrt(desi_clkappa * cldesi), color='#4daf4a', label='DESI')
+plt.plot( rho_comb[:,0], rho_comb[:,1], color='#984ea3', linestyle='-.', label='CIB+D.E.S no noise')
 
 
 plt.xlabel(r'\ell')

### 2017-01-23 Notes/DES_rec_phi_paper.tex
first

@@
 To delens we need the bulk of the redshift distribution to follow the CMB kernel, having a few outliers at redshift $z>1.5$ is not enough.
-We also have to keep in mind the results of \citet{smith:2012}: even a perfect LSS survey can not help too much the delensing process, with the possible exception of futuristic 21 cm.
+We also have to keep in mind the results of \cite{smith:2012}: even a perfect LSS survey can not help too much the delensing process, with the possible exception of futuristic 21 cm.
 \end{abstract}
@@
 
-\section{Theory} \label{sec:th}
+\section{Theory} 
+\label{sec:th}
 
@@
 \includegraphics[scale=1.]{../images/spectra.pdf}
-\caption{Correlation factor between galaxies survey and CMB lensing potential as a function of the multipole $\ell$.}
+\caption{Correlation factor between galaxies survey and CMB lensing potential as a function of the multipole $\ell$. The ``CIB+D.E.S'' factor has been computed using \refeq{rho-combined}}
+\end{center}
 \label{fig:corr}
-\end{center}
+
 \end{figure}
+
 %=======================================
@@
 \includegraphics[scale=1.]{../images/clbb_res.pdf}
-\caption{Residual lensing B modes power spectrum using different large scale structure.}
+\caption{Residual lensing B modes power spectrum using different large scale structure. \refeq{combined}}
+\end{center}
 \label{fig:bres}
-\end{center}
 \end{figure}
+
 %=======================================
@@
 Redshift distribution of D.E.S galaxies (I suspect this is the benchmark, anyway taken from Giannantionio et al.). DESI taken from their white paper. }
-\label{fig:corr}
 \end{center}
+\label{fig:kernel}
 \end{figure}
@@
 As a first approximation, the B mode resulting from the lensing of primordial E mode by a convergence field $\kappa$ is:
-\be \label{eq}
+\be 
 B^{\mathrm{lens}}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') E(\ublu) \kappa(\bl - \bl')
+\label{eqn:}
 \ee
@@
 
-\subsection{Galaxies contribution at different redshift}\label{sec:galcontrib}
+\subsection{Galaxies contribution at different redshifts (CIB+D.E.S)}\label{sec:galcontrib}
 
 Let's now assume that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
-It can be shown that the optimal way to combine them to obtain the optimal estimator of $\phi$ or, in other word maximizing the correlation factor $\rho$ is:
+It can be shown that the optimal way to combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ is:
 \beqn
@@
 c_{i} 	&=& (C^{-1})_{ij}C^{\kappa I^{j}}
+\label{eqn:combined}
 \eeqn 
 where C is the covariance matrix of the LSS tracers.
-
+The ``effective'' correlation of these combined tracers with gravitational lensing is: 
 \be
-\rho^{2} = \sum_{i,j}\frac{C^{\kappa i}~(C^{-1})_{ij}~C^{\kappa j}}{C^{\kappa\kappa}}
+\rho^{2} = \sum_{i,j}\frac{C^{\kappa i}~(C^{-1})_{ij}~C^{\kappa j}}{C^{\kappa\kappa}}.
+\label{eqn:rho-combined}
 \ee
@@
 
-\section{Data} \label{sec:data}
+\section{Data} 
+\label{sec:data}
 
@@
 
-\section{Forecast} \label{sec:for}
+\section{Forecast} 
+\label{sec:for}
 
-\bibliographystyle{unsrtnat}
+\bibliographystyle{plain}
 \bibliography{DES_rec_phi_paper}

### 2017-01-23 plotting/plot_Blens.py

@@ -6,6 +6,7 @@ import scipy as sp
 import numpy as np
 import matplotlib
 matplotlib.use('Agg')
+from itertools import cycle
 
 import matplotlib.pyplot as plt
 
@@ -32,11 +33,11 @@ import __builtin__
 import pyximport
 pyximport.install(reload_support=True)
 
-try:
-    cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
-    plt.rcParams['image.cmap'] = cmap
-except Exception, e:
-    print 'probably palettable not found'
+# try:
+#     cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+#     plt.rcParams['image.cmap'] = cmap
+# except Exception, e:
+#     print 'probably palettable not found'
 
 
 datadir = '../Data/limber_spectra/'
@@ -49,17 +50,23 @@ clbb = np.loadtxt(datadir + 'cbb.txt')
 cbb_res_des = np.loadtxt(datadir + 'cbb_res_des.txt')
 cbb_res_desi = np.loadtxt(datadir + 'cbb_res_desi.txt')
 cbb_res_cib = np.loadtxt(datadir + 'cbb_res_cib.txt')
-cbb_res_comb = np.loadtxt(datadir + 'cbb_res_comb.txt')
+cbb_res_des_cib = np.loadtxt(datadir + 'cbb_res_des_cib.txt')
+cbb_res_des_cib_cmb = np.loadtxt(datadir + 'cbb_res_des_cib_cmb.txt')
 cbb_res_cib_525 = np.loadtxt(datadir + 'cbb_res_cib_525.txt')
-ell = np.loadtxt(datadir + 'cbb_res_cib.txt')
+ell = np.loadtxt(datadir + 'cbb_res_ls.txt')
 
+# LOAD primordial from camb
+# TODO check they have the same normalization
 
+cbb_prim_01 = np.loadtxt('/Users/alessandromanzotti/Work/Software/camb/manzotti_primordial_01_tensCls.dat')
+cbb_prim_001 = np.loadtxt('/Users/alessandromanzotti/Work/Software/camb/manzotti_primordial_001_tensCls.dat')
[…]

### 2017-01-23 plotting/plot_corr_fac.py

@@ -6,6 +6,7 @@ import scipy as sp
 import numpy as np
 import matplotlib
 matplotlib.use('Agg')
+from itertools import cycle
 
 import matplotlib.pyplot as plt
 
@@ -35,6 +36,7 @@ pyximport.install(reload_support=True)
 try:
     cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
     plt.rcParams['image.cmap'] = cmap
+
 except Exception, e:
     print 'probably palettable not found'
 
@@ -57,7 +59,13 @@ cldesi = np.loadtxt(datadir + 'cldesi.txt')
 clkappadesi = np.loadtxt(datadir + 'cldesik.txt')
 desi_clkappa = np.loadtxt(datadir + 'clk.txt')
 # combined rho coming from multiple_survey_delens.py on midway
-rho_comb = np.loadtxt(datadir + 'rho_comb_des_cib.txt')
+rho_comb_des_cib = np.loadtxt(datadir + 'rho_comb_des_cib.txt')
+rho_comb_des_cib_cmb= np.loadtxt(datadir + 'rho_comb_des_cib_cmb.txt')
+rho_des = np.loadtxt(datadir + 'rho_des.txt')
+rho_cib = np.loadtxt(datadir + 'rho_cib.txt')
+rho_cmb = np.loadtxt(datadir + 'rho_cmb.txt')
+
+
 
 
 # ADD NOISE
@@ -78,7 +86,7 @@ nlgg = 1. / (3207184. / rad_sq) * np.ones_like(cldes)  # they mention N=2.1 10^-
 # ============================================
 # SIZE OF THE PICTURE
 # ============================================
-WIDTH = 462.0  # the number latex spits out
+WIDTH = 700.0  # the number latex spits out
 # the fraction of the width you'd like the figure to occupy, 0.5 if double
 # column cause
 FACTOR = 0.6
@@ -110,8 +118,7 @@ plt.rcParams['axes.titlesize'] = font_size * 1.3
 plt.rcParams['legend.fontsize'] = font_size
 plt.rcParams['xtick.labelsize'] = font_size / 1.2
 plt.rcParams['ytick.labelsize'] = font_size / 1.2
[…]

### 2017-01-23 Code/compare_kernel.py
minor

@@ -1,13 +1,24 @@
+'''
+
+
+
+'''
+
+
+
 import numpy as np
 import kappa_cmb_kernel as kappa_kernel
 import gals_kernel
 import hall_CIB_kernel as cib_hall
 import scipy.integrate
 from scipy.interpolate import RectBivariateSpline, interp1d, InterpolatedUnivariateSpline
-zpower = np.loadtxt('output/matter_power_nl/z.txt')
-kpower = np.loadtxt('output/matter_power_nl/k_h.txt')
-powerarray = np.loadtxt('output/matter_power_nl/p_k.txt')
-powerarray = np.loadtxt('output/matter_power_nl/p_k.txt').reshape([np.size(zpower), np.size(kpower)]).T
+
+spectra_dir = '/home/manzotti/cosmosis/modules/limber/'
+
+zpower = np.loadtxt(spectra_dir+'output/matter_power_nl/z.txt')
+kpower = np.loadtxt(spectra_dir+'output/matter_power_nl/k_h.txt')
+powerarray = np.loadtxt(spectra_dir+'output/matter_power_nl/p_k.txt')
+powerarray = np.loadtxt(spectra_dir+'output/matter_power_nl/p_k.txt').reshape([np.size(zpower), np.size(kpower)]).T
 rbs = RectBivariateSpline(kpower, zpower, powerarray)
 
 omega_m = 0.3
@@ -15,16 +26,16 @@ omega_m = 0.3
 h0 = 0.7
 
 
-h = np.loadtxt('output/distances/h.txt')
+h = np.loadtxt(spectra_dir+'output/distances/h.txt')
 tmp = h[::-1]
 h = tmp
 
 xlss = 13615.317054155654
-zdist = np.loadtxt('output/distances/z.txt')
+zdist = np.loadtxt(spectra_dir+'output/distances/z.txt')
 tmp = zdist[::-1]
 zdist = tmp
 
-d_m = np.loadtxt('output/distances/d_m.txt')
+d_m = np.loadtxt(spectra_dir+'output/distances/d_m.txt')
 tmp = d_m[::-1]
 d_m = tmp
 
@@ -36,14 +47,14 @@ chispline = interp1d(zdist, d_m)
 z_chi_spline = interp1d(d_m, zdist)
 hspline = interp1d(zdist, h)
 
-dndz_filename = 'data_input/DES/N_z_wavg_spread_model_0.2_1.2_tpz.txt'
[…]

### 2017-03-15 Code/compare_kernel.py
several improvements

@@ -1,24 +1,33 @@
-'''
-
-
 
 '''
 
+
+
+'''
 
 
 import numpy as np
-import kappa_cmb_kernel as kappa_kernel
-import gals_kernel
-import hall_CIB_kernel as cib_hall
+# import kappa_cmb_kernel as kappa_kernel
+# import gals_kernel
+# import hall_CIB_kernel as cib_hall
 import scipy.integrate
 from scipy.interpolate import RectBivariateSpline, interp1d, InterpolatedUnivariateSpline
+import imp
 
-spectra_dir = '/home/manzotti/cosmosis/modules/limber/'
+gals_kernel = imp.load_source(
+    'gals_kernel', '/home/manzotti/cosmosis/modules/limber/gals_kernel.py')
+kappa_kernel = imp.load_source(
+    'kappa_kernel', '/home/manzotti/cosmosis/modules/limber/kappa_cmb_kernel.py')
+cib_hall = imp.load_source('cib_hall', '/home/manzotti/cosmosis/modules/limber/hall_CIB_kernel.py')
+kappa_gals_kernel = imp.load_source(
+    'kappa_gals_kernel', '/home/manzotti/cosmosis/modules/limber/kappa_gals_kernel.py')
+spectra_dir = '/home/manzotti/galaxies_delensing/Data/'
 
-zpower = np.loadtxt(spectra_dir+'output/matter_power_nl/z.txt')
-kpower = np.loadtxt(spectra_dir+'output/matter_power_nl/k_h.txt')
-powerarray = np.loadtxt(spectra_dir+'output/matter_power_nl/p_k.txt')
-powerarray = np.loadtxt(spectra_dir+'output/matter_power_nl/p_k.txt').reshape([np.size(zpower), np.size(kpower)]).T
+zpower = np.loadtxt(spectra_dir + 'matter_power_nl/z.txt')
+kpower = np.loadtxt(spectra_dir + 'matter_power_nl/k_h.txt')
+powerarray = np.loadtxt(spectra_dir + 'matter_power_nl/p_k.txt')
+powerarray = np.loadtxt(
+    spectra_dir + 'matter_power_nl/p_k.txt').reshape([np.size(zpower), np.size(kpower)]).T
[…]

### 2017-03-15 Code/multiple_survey_delens.py

@@ -70,7 +70,7 @@ rho_cib = np.zeros((np.size(lbins)))
 
 
 labels = ['k', 'euclid', 'des_weak', 'lsst', 'ska10', 'ska01',
-          'ska5', 'ska1', 'lkern', 'cib', 'desi', 'des']
+          'ska5', 'ska1', 'cib', 'desi', 'des']
 cl_cross_k = {}
 cl_auto = {}
 rho = {}
@@ -91,28 +91,54 @@ for label in labels:
 # Multiple surveys
 
 labels = ['cib', 'des']
-
+cmb = 'S4'
 cgk = np.zeros((len(labels) + 1, np.size(lbins)))
 cgg = np.zeros((len(labels) + 1, len(labels) + 1, np.size(lbins)))
 
-for i in np.arange(0, n_slices):
+for i in np.arange(0, len(labels)):
     cgk[i, :] = np.loadtxt(output_dir +
                            '/limber_spectra/cl_k' + labels[i] + '_delens.txt')
 
-    for j in np.arange(i, n_slices):
+    for j in np.arange(i, len(labels)):
         cgg[i, j, :] = np.loadtxt(output_dir +
                                   '/limber_spectra/cl_' + labels[i] + labels[j] + '_delens.txt')
         cgg[j, i, :] = cgg[i, j, :]
 
 
-# add cmb lensing
+if cmb == 'S3' :
+    noise = 7.0
+    beam = 3
+    noise_phi = np.loadtxt('./quicklens/min_var_noise_{}muk_{}beam.txt'.format(noise, beam))
+    noise_phi *= np.arange(0, len(noise_phi))**4. / 4.
+    # noise_cmb = nl(noise, beam, lmax=4000)
 
+    noise_cl[:, 1] = noise_cl[:, 1] * noise_cl[:, 0] ** 4 / \
+        4.  # because the power C_kk is l^4/4 C_phi
+    noise_fun = interp1d(np.arange(0, len(noise_phi)), noise_phi)
+    ckk_noise = np.zeros_like(ckk)
+    ckk_noise = noise_fun(lbins)
+
+
+elif cmb=='S4':
+    noise = 1.
+    beam = 1
+    noise_phi = np.loadtxt('./quicklens/min_var_noise_{}muk_{}beam.txt'.format(noise, beam))
[…]

### 2017-03-15 Code/plot_dndz.py

@@ -6,6 +6,7 @@ import scipy as sp
 import numpy as np
 import matplotlib
 matplotlib.use('Agg')
+import scipy.integrate
 
 import matplotlib.pyplot as plt
 
@@ -26,12 +27,12 @@ import os
 import sys
 import time
 import subprocess
-
+import imp
 from pprint import pprint
 import __builtin__
 import pyximport
+from scipy.interpolate import RectBivariateSpline, interp1d
 pyximport.install(reload_support=True)
-
 try:
     cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
     plt.rcParams['image.cmap'] = cmap
@@ -45,8 +46,13 @@ image_dir = '../images/'
 
 
 # Import data
-des_dndz = np.loadtxt("/home/manzotti/cosmosis/modules/limber/data_input/DES/N_z_wavg_spread_model_0.2_1.2_tpz.txt")
+des_dndz = np.loadtxt(
+    "/home/manzotti/cosmosis/modules/limber/data_input/DES/N_z_wavg_spread_model_0.2_1.2_tpz.txt")
 
+gals_kernel = imp.load_source(
+    'gals_kernel', '/home/manzotti/cosmosis/modules/limber/gals_kernel.py')
+# kappa_kenel = imp.load_source('kappa_kenel', '/home/manzotti/cosmosis/modules/limber/kappa_kenel.py')
+# gals_kern = imp.load_source('gals_kern', '/home/manzotti/cosmosis/modules/limber/gals_kernel.py')
 
 # ============================================
 # SIZE OF THE PICTURE
@@ -111,18 +117,119 @@ plt.rcParams['legend.handletextpad'] = 0.3
 # ============================================
 
 
+# LOAD DNDZ
+# =======================
+
+
+# =======================
+# DEFINE KERNELs
+# CIB
+# j2k = 1.e-6 / np.sqrt(83135.)  # for 353
+# lkern = kappa_kernel.kern(zdist, hspline, omega_m, h0, xlss)
+# cib = cib_hall.ssed_kern(
[…]

### 2017-03-16 Notes/DES_rec_phi_paper.tex
several improvements
tex fix

@@
-\input{/Users/alessandromanzotti/Work/preamble_onecol_notitle.tex}
+\documentclass[onecolumn,amsmath,amssymb,floatfix,superscriptaddress,notitlepage,aps,prd]{revtex4-1}
+\usepackage{bm}
+\usepackage{epsfig}
+\usepackage{natbib}
+\usepackage{graphicx,epsfig}
+\usepackage{hyperref}
+\usepackage{ifthen}
+\usepackage{xstring}
+\usepackage{graphicx}\usepackage[top=2.5 cm, bottom=2.5 cm, left=1.05in, right=1.05in]{geometry}
+\usepackage[applemac]{inputenc}
+\usepackage{amsmath,amssymb} 
+\usepackage{color}
+\usepackage{epstopdf}
+\usepackage[caption=false]{subcaption}
+\epstopdfDeclareGraphicsRule{.gif}{png}{.png}{convert gif:#1 png:\OutputFile}
+\AppendGraphicsExtensions{.gif}
+% notes to make comment
+
+\newcommand{\note}[1]{\textsc{#1}}
+
+% For various journals
+%\newcommand{\apj}{ApJ}
+\newcommand{\physrep}{Physics Reports}
+\newcommand{\jcap}{JCAP}
+\newcommand{\apjl}{ApJL}
+\newcommand{\apjs}{ApJS}
+\newcommand{\aap}{A\&A}
+\newcommand{\mnras}{MNRAS}
+%\newcommand{\prd}{Phys. Rev. D}
+\newcommand{\physrev}{Phys. Rev.}
+\newcommand{\physrevlett}{Phys. Rev. Lett.}
+
+%--------- NEW COMMAND STUFF
+
+\def\simlt{\lesssim}
+\def\simgt{\gtrsim}
+\newcommand{\Omegamzero}{\Omega_{{\rm m,0}}}
+\newcommand{\alm}{a_{lm}}
+
+%\newcommand{\l}{\ell}}
+
+\def\be{\begin{equation}}
+\def\ee{\end{equation}}
+\def\ben{\begin{equation*}}
+\def\een{\end{equation*}}
+
+\def\ba{\begin{eqnarray}}
+\def\ea{\end{eqnarray}}
+\def\ban{\begin{eqnarray*}}
+\def\ean{\end{eqnarray*}}
+
+\newcommand{\refsec}[1]{section~\ref{sec:#1}}
+\newcommand{\reftab}[1]{Tab.~\ref{tab:#1}}
+\newcommand{\refeq}[1]{Eq.~(\ref{eqn:#1})}
+\newcommand{\refssec}[1]{section~\ref{subsec:#1}}
+\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
+\newcommand{\refFig}[1]{Fig.~\ref{fig:#1}}
+
+\definecolor{darkgreen}{cmyk}{0.85,0.2,1.00,0.2} 
+\definecolor{purple}{cmyk}{0.5,1.0,0,0}
+
 \usepackage{hyperref}
@@
 
-\title{D.E.S reconstructed lensing potential template and application to delensing and cross correlation }         % Enter your title between curly braces
+\title{Future CMB delensing with galaxies surveys.}         % Enter your title between curly braces
 \author{A. Manzotti}        % Enter your name between curly braces
@@
 \label{sec:for}
+\subsection{CMB-S3 Era}
+\subsection{CMB-S4 Era}
+\subsection{Bias uncertainties degradation}
 
+\subsection{Bias uncertainties degradation}
 \bibliographystyle{plain}

### 2017-03-17 Notes/DES_rec_phi_paper.tex
tex
tex fix

@@
 % notes to make comment
+\newcommand{\al}[1]{\begin{align} #1 \end{align}}
 
 \newcommand{\note}[1]{\textsc{#1}}
+	\newcommand{\bR}[1]{{\bm {\mR{#1}}} }
 
+\def\l{\ell}
 % For various journals
@@
 
+\newcommand{\nver}{\hat{\mathbf{n}}}
+\newcommand{\cov}{\text{Cov}}
+\newcommand{\Nsim}{N_{\text{sim}}}
+\newcommand{\new}[1]{{\textcolor{blue}{\bf #1}}}
 %--------- NEW COMMAND STUFF
+	\newcommand{\PD}[2]{\dfrac{\partial #1}{\partial #2}}
 
@@
 \newcommand{\alm}{a_{lm}}
-
+	% * subscript
+	\def\rom#1{%
+		_{\mathrm{#1}}%
+	}%
 %\newcommand{\l}{\ell}}
@@
 \newcommand{\refFig}[1]{Fig.~\ref{fig:#1}}
+	\newcommand{\mR}[1]{\mathrm{#1}}   % mathrm
 
@@
 \newcommand{\bmm}[1]{{\mathbf{#1}}}
-\newcommand{\new}[1]{{\color{blue} #1}}
 \newcommand{\bsection}[1]{\section{\uppercase{#1}}}
@@
 
+\section{Introduction}
+
 \section{Theory}
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/spectra.pdf}
+\includegraphics[scale=0.6]{../images/spectra.pdf}
 \caption{Correlation factor between galaxies survey and CMB lensing potential as a function of the multipole $\ell$. The ``CIB+D.E.S'' factor has been computed using \refeq{rho-combined}}
@@
 
-\section{Data}
+\section{Tracers Model}
 \label{sec:data}
 
-D.E.S, Planck CIB, DESI?
+In this equation, $\chi(z)$ is the comoving distance to redshift $z$, $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, and $\Psi(\chi(z)\nver,z)$ is the three-dimensional gravitational potential at a point on the photon path given by $\chi(z)\nver$.
+Note that the deflection angle is given by $\mathbf{d}(\nver) = \nabla\phi(\nver)$, where $\nabla$ is the the two-dimensional gradient on the sphere.
+Because the lensing potential is an integrated measure of the projected gravitational potential, taking the two-dimensional Laplacian of the lensing potential we can define the lensing convergence $\kappa(\nver) = -\case{1}{2}\nabla^2\phi(\nver)$, which depends on the projected matter overdensity $\delta$ \citep{bart01}:
+\begin{equation}
+\kappa(\nver) = \int_0^{z_*} dz\, W^{\kappa}(z)\delta(\chi(z)\nver,z).
+\label{eqn:wkappa}
+\end{equation}
+The lensing kernel $W^{\kappa}$ is
+%
+\begin{equation}
+W^{\kappa}(z) = \frac{3\Omega_{\rm m}}{2c}\frac{H_0^2}{H(z)}(1+z)\chi(z)\frac{\chi_*-\chi(z)}{\chi_*},
+\end{equation}
+%
+where $\Omega_{\rm m}$ and $H_0$ are the present-day values of the Hubble and matter density parameters, respectively.
+
+%----------------------------------------------------------------------------------------------------%
+%\begin{eqnarray} \label{PS}
+%	C_\l^{XY} = \frac{2}{\pi}\Int{}{k}{k^2} P_{init}(k) \Delta_\l^X(k)\Delta_\l^Y(k) 
+%	\,, 
+%\end{eqnarray}
+%----------------------------------------------------------------------------------------------------%
+where $P_{init}(k)$ is the scalar power spectrum at an early time with $k$ being the Fourier wave 
+number. The functions $\Delta^X\l(k)$ and $\Delta^Y\l(k)$ are one of the following (see e.g., 
+\cite{Hu:2000ee,LoVerde:2006cj}):
+%----------------------------------------------------------------------------------------------------%
+%\begin{eqnarray}
+%
+%	\Delta^I_\l(k) &= k^2 \INT{}{z}{}{0}{z_{\rm max}} b(z) \D{N}{z}(z) D(z) j_\l(k\chi(z)) 
+%	\,, \label{eq:Delta^I} \\
+%	\Delta^\grad_\l(k) &= \frac{3\Omega\rom{m}H_0^2}{2}
+%	\notag \\
+%		&\times \INT{}{\chi}{}{0}{\chi_*} 
+%			\frac{\chi_*-\chi}{\chi_*\chi}\,\frac{D(z(\chi))}{a(\chi)}\,j_\l(k\chi)
+%	\,. 
+%	
+%\end{eqnarray}
+
+%----------------------------------------------------------------------------------------------------%
+
+\subsection{Galaxies}
+
+The galaxy overdensity $g(\nver)$ in a given direction on the sky is also expressed as a LOS integral of the matter overdensity:
+\begin{equation}
+g(\nver) = \int_0^{z_*} dz\, W^{g}(z)\delta(\chi(z)\nver,z),
+\end{equation}
+where the kernel is
+\begin{equation}
+\label{eqn:wg}
+\begin{split}
+W^{g}(z) &= \frac{b(z)\frac{dN}{dz}}{\Bigl(\int dz'\,\frac{dN}{dz'}\Bigr)} + \frac{3\Omega_{\rm m}}{2c}\frac{H_0^2}{H(z)}(1+z)\chi(z) \\
+&\times \int_z^{z_*}dz'\,\Bigl(1-\frac{\chi(z)}{\chi(z')}\Bigr)(\alpha(z')-1)\frac{dN}{dz'}.
+\end{split}
+\end{equation}
+
+\subsection{Cosmic infrared Background}
+\label{sec:cib}
+In this work, the CMB lensing potential is estimated from maps of the cosmic infrared background (CIB).
+
+Following \cite{Addison:2011se}, we model the CIB power as $C_{\ell}^{{\rm CIB}\mbox{-}{\rm CIB}} = 3500 (l/3000)^{-1.25} {\rm Jy^2 / sr}$.
+We test that this model provides an accurate fit for the power of the \textit{Herschel} $500\,\mu$ map used in this work.
+For the cross-spectrum $C_{\ell}^{{\rm CIB}\mbox{-}\phi}$, we use the single-SED model of~\cite{hall10}.
+This places the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$.
+This model is rescaled to agree with the results of
+~\citet{Holder:2013hqu} and \citet{planck2013XVIII} by choosing the corresponding linear bias parameter.
+Other multi-frequency CIB models are available \citep[e.g.,][]{Bethermin:2013nza}; however, given the level of noise, we are relatively insensitive to this choice.
+With these assumptions, depending on angular scale, $45-65\%$ of the CIB is correlated with the CMB lensing potential, as shown in \reffig{CIB}.
+
+\subsection{Weak Lensing}
 \cite{hall:2010,sherwin:2015,ade:2014,hanson:2013,szapudi:2001,0004-637X-567-1-2,planck-collaboration:2014,planck-collaboration:2014a,planck-collaboration:2011,plank-collaboration:2014,boulanger:1996,lewis:2006}
+\subsection{Internal CMB lensing}
 
@@
 
-\subsection{Bias uncertainties degradation}
+%----------------------------------------------------------------------------------------------------%
+%\al{
+%	F_{ij} &\equiv \sum_{\l=\l\rom{min}}^{100} \frac{2\l+1}{2[C_\l^{\rm BB,res}+N_\l^{\rm BB}]^2}
+%		\PD{C^{\rm BB,res}_\l}{\theta_i}\PD{C^{\rm BB,res}_\l}{\theta_j}
+%	\notag \\
+%	&+ \sum_{L=2}^{1000} \frac{2L+1}{2} 
+%		\Tr \left[\bR{C}^{-1}_L\PD{\bR{C}_L}{\theta_i}\bR{C}^{-1}_L\PD{\bR{C}_L}{\theta_j}\right]
+%	\,, 
+%}
+%----------------------------------------------------------------------------------------------------%
+where $\theta_i=r$, $b_i$ or $p_i$ \footnote{In our analysis, since the fiducial value of $r$ is zero, 
+the derivative of $C_\l^{\rm BB,res}$ is non-zero if $\theta_i=r$.}. 
+The covariance matrix of the mass tracers $\bR{C}_L$ is given in Eq.~\eqref{Eq:cov}. 
+We compute the derivatives of the power spectra as described in appendix \ref{sec:deriv}. 
+Note that we omit the correction of the partial sky coverage since it does not affect our results, i.e.,
+
+\section{Conclusions}
+
 \bibliographystyle{plain}

### 2017-03-22 Code/multiple_survey_delens.py
add

@@ -57,7 +57,8 @@ noise_phi = np.loadtxt('./quicklens/min_var_noise_{}muk_{}beam.txt'.format(noise
 noise_phi *= np.arange(0, len(noise_phi))**4. / 4.
 # noise_cmb = nl(noise, beam, lmax=4000)
 
-noise_cl[:, 1] = noise_cl[:, 1] * noise_cl[:, 0] ** 4 / 4.  # because the power C_kk is l^4/4 C_phi
+# noise_cl[:, 1] = noise_cl[:, 1] * noise_cl[:, 0] ** 4 / 4.  # because
+# the power C_kk is l^4/4 C_phi
 noise_fun = interp1d(np.arange(0, len(noise_phi)), noise_phi)
 ckk_noise = np.zeros_like(ckk)
 ckk_noise = noise_fun(lbins)
@@ -69,31 +70,34 @@ rho_des = np.zeros((np.size(lbins)))
 rho_cib = np.zeros((np.size(lbins)))
 
 
-labels = ['k', 'euclid', 'des_weak', 'lsst', 'ska10', 'ska01',
+surveys = ['k', 'euclid', 'des_weak', 'lsst', 'ska10', 'ska01',
           'ska5', 'ska1', 'cib', 'desi', 'des']
 cl_cross_k = {}
 cl_auto = {}
 rho = {}
-for label in labels:
-    cl_cross_k[label] = np.loadtxt(output_dir +'/limber_spectra/cl_k' + label + '_delens.txt')
+for label in surveys:
+    cl_cross_k[label] = np.loadtxt(output_dir + '/limber_spectra/cl_k' + label + '_delens.txt')
     cl_auto[label] = np.loadtxt(output_dir +
                                 '/limber_spectra/cl_' + label + label + '_delens.txt')
+    if label == 'cib':
+        cl_auto[label] = np.array([3500. * (1. * l / 3000.)**(-1.25) for l in lbins])
+
     rho[label] = cl_cross_k[label] / np.sqrt(ckk[:] * cl_auto[label])
 
 
 # single survey save
-for label in labels:
+for label in surveys:
     np.savetxt(output_dir + '/limber_spectra/rho_{}.txt'.format(label),
                np.vstack((lbins, rho[label])).T)
 
 
 # Multiple surveys
 
-labels = ['lsst', 'cib', 'desi', 'des']
-cmb = 'S4'
+# labels = ['lsst', 'cib', 'desi', 'des']
[…]

### 2017-03-22 Code/plots/plot_Blens.py

@@ -1,0 +1,258 @@
+
+from __future__ import division
+# from __future__ import print_function
+print "\n Python start up script from /home/manzotti/.py_startup.py.\n"
+import scipy as sp
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+from itertools import cycle
+
+import matplotlib.pyplot as plt
+
+try:
+    import pyfits
+    import pdb
+
+except:
+    pass
+
+try:
+    import palettable
+
+except Exception, e:
+    pass
+import os
+import os
+import sys
+import time
+import subprocess
+
+from pprint import pprint
+import __builtin__
+import pyximport
+pyximport.install(reload_support=True)
+
+# try:
+#     cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+#     plt.rcParams['image.cmap'] = cmap
+# except Exception, e:
+#     print 'probably palettable not found'
+
+
+datadir = '/home/manzotti/cosmosis/modules/limber/cib_des_delens/limber_spectra/'
+
+image_dir = '../images/'
+
+
+# Import data
+clbb = np.loadtxt(datadir + 'cbb.txt')
+cbb_res_des = np.loadtxt(datadir + 'cbb_res_des.txt')
+# cbb_res_desi = np.loadtxt(datadir + 'cbb_res_desi.txt')
+cbb_res_cib = np.loadtxt(datadir + 'cbb_res_cib.txt')
+cbb_res_des_cib = np.loadtxt(datadir + 'cbb_res_des_cib.txt')
+cbb_res_des_cib_cmb = np.loadtxt(datadir + 'cbb_res_des_cib_cmb.txt')
+# cbb_res_cib_525 = np.loadtxt(datadir + 'cbb_res_cib_525.txt')
+ell = np.loadtxt(datadir + 'cbb_res_ls.txt')
+
+# LOAD primordial from camb
+# TODO check they have the same normalization
+
+# cbb_prim_01 = np.loadtxt('/Users/alessandromanzotti/Work/Software/camb/manzotti_primordial_01_tensCls.dat')
+# cbb_prim_001 = np.loadtxt('/Users/alessandromanzotti/Work/Software/camb/manzotti_primordial_001_tensCls.dat')
[…]

### 2017-03-22 Code/plots/plot_alpha.py

@@ -1,0 +1,153 @@
+
+from __future__ import division
+# from __future__ import print_function
+import scipy as sp
+import numpy as np
+# import matplotlib
+# matplotlib.use('Agg')
+from itertools import cycle
+import matplotlib.pyplot as plt
+
+# try:
+#     import palettable
+
+# except ValueError:
+#     pass
+import os
+import os
+import sys
+import time
+import subprocess
+import imp
+from pprint import pprint
+import pyximport
+pyximport.install(reload_support=True)
+
+# try:
+#     cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+#     plt.rcParams['image.cmap'] = cmap
+
+# except ValueError:
+#     print('probably palettable not found')
+
+image_dir = '../../images/'
+
+
+# ============================================
+# SIZE OF THE PICTURE
+# ============================================
+WIDTH = 700.0  # the number latex spits out
+# the fraction of the width you'd like the figure to occupy, 0.5 if double
+# column cause
+FACTOR = 0.6
+fig_width_pt = WIDTH * FACTOR
+
+inches_per_pt = 1.0 / 72.27
+# golden_ratio  = (np.sqrt(5) - 1.0) / 2.0  # because it looks good
+ratio = 0.8
+fig_width_in = fig_width_pt * inches_per_pt  # figure width in inches
+fig_height_in = fig_width_in * ratio   # figure height in inches
+fig_dims = [fig_width_in, fig_height_in]  # fig dims as a list
+# ============================================
+# ============================================
+
+
+# ============================================
+# SET LATEX
+plt.rc('text', usetex=False)
+plt.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+plt.rc('font', **{'family': 'serif', 'serif': ['Computer Modern Roman']})
+# ============================================
+
+
+font_size = 10.
+plt.rcParams['font.size'] = font_size
+
+plt.rcParams['axes.labelsize'] = font_size * 1.3
+plt.rcParams['axes.linewidth'] = font_size / 22.
+plt.rcParams['axes.titlesize'] = font_size * 1.3
+plt.rcParams['legend.fontsize'] = font_size
[…]

### 2017-03-22 Code/plots/plot_corr_fac_now.py

@@ -1,0 +1,141 @@
+
+from __future__ import division
+# from __future__ import print_function
+import scipy as sp
+import numpy as np
+# import matplotlib
+# matplotlib.use('Agg')
+from itertools import cycle
+import matplotlib.pyplot as plt
+
+# try:
+#     import palettable
+
+# except ValueError:
+#     pass
+import os
+import os
+import sys
+import time
+import subprocess
+import imp
+from pprint import pprint
+import pyximport
+pyximport.install(reload_support=True)
+
+# try:
+#     cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+#     plt.rcParams['image.cmap'] = cmap
+
+# except ValueError:
+#     print('probably palettable not found')
+
+
+datadir = '/home/manzotti/cosmosis/modules/limber/cib_des_delens/limber_spectra/'
+
+image_dir = '../images/'
+
+labels = ['cib', 'des']
+cmb = 'now'
+
+
+# ============================================
+# SIZE OF THE PICTURE
+# ============================================
+WIDTH = 700.0  # the number latex spits out
+# the fraction of the width you'd like the figure to occupy, 0.5 if double
+# column cause
+FACTOR = 0.6
+fig_width_pt = WIDTH * FACTOR
+
+inches_per_pt = 1.0 / 72.27
+# golden_ratio  = (np.sqrt(5) - 1.0) / 2.0  # because it looks good
+ratio = 0.8
+fig_width_in = fig_width_pt * inches_per_pt  # figure width in inches
+fig_height_in = fig_width_in * ratio   # figure height in inches
+fig_dims = [fig_width_in, fig_height_in]  # fig dims as a list
+# ============================================
+# ============================================
+
+
+# ============================================
+# SET LATEX
+plt.rc('text', usetex=False)
+plt.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+plt.rc('font', **{'family': 'serif', 'serif': ['Computer Modern Roman']})
+# ============================================
+
+
+font_size = 10.
+plt.rcParams['font.size'] = font_size
+
+plt.rcParams['axes.labelsize'] = font_size * 1.3
+plt.rcParams['axes.linewidth'] = font_size / 22.
[…]

### 2017-03-22 Code/plots/plot_kernel.py

@@ -1,0 +1,179 @@
+
+from __future__ import division
+# from __future__ import print_function
+print "\n Python start up script from /home/manzotti/.py_startup.py.\n"
+import scipy as sp
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+
+import matplotlib.pyplot as plt
+
+try:
+    import pyfits
+    import pdb
+
+except:
+    pass
+
+try:
+    import palettable
+
+except Exception, e:
+    pass
+import os
+import os
+import sys
+import time
+import subprocess
+
+from pprint import pprint
+import __builtin__
+import pyximport
+pyximport.install(reload_support=True)
+
+try:
+    cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+    plt.rcParams['image.cmap'] = cmap
+except Exception, e:
+    print 'probably palettable not found'
+
+
+datadir = '../Data/limber_spectra/'
+
+image_dir = '../images/'
+
+
+# Import data
+kappakernel_l30 = np.loadtxt(datadir + 'kappa_kernel_l30.txt')
+kappakernel_l100 = np.loadtxt(datadir + 'kappa_kernel_l100.txt')
+kappakernel_l500 = np.loadtxt(datadir + 'kappa_kernel_l500.txt')
+
+kappakernel_l30_h = np.loadtxt(datadir + 'kappa_kernel_l30_h.txt')
+kappakernel_l100_h = np.loadtxt(datadir + 'kappa_kernel_l100_h.txt')
+kappakernel_l500_h = np.loadtxt(datadir + 'kappa_kernel_l500_h.txt')
+
+cibkernel_l30 = np.loadtxt(datadir + 'cib_kernel_l30.txt')
+cibkernel_l100 = np.loadtxt(datadir + 'cib_kernel_l100.txt')
+cibkernel_l500 = np.loadtxt(datadir + 'cib_kernel_l500.txt')
+
+cibkernel_l30_h = np.loadtxt(datadir + 'cib_kernel_l30_h.txt')
+cibkernel_l100_h = np.loadtxt(datadir + 'cib_kernel_l100_h.txt')
+cibkernel_l500_h = np.loadtxt(datadir + 'cib_kernel_l500_h.txt')
+
+
+deskernel_l30 = np.loadtxt(datadir + 'des_kernel_l30.txt')
+deskernel_l100 = np.loadtxt(datadir + 'des_kernel_l100.txt')
[…]

### 2017-04-09 Code/multiple_survey_delens.py
new img

@@ -71,7 +71,7 @@ rho_cib = np.zeros((np.size(lbins)))
 
 
 surveys = ['k', 'euclid', 'des_weak', 'lsst', 'ska10', 'ska01',
-          'ska5', 'ska1', 'cib', 'desi', 'des']
+           'ska5', 'ska1', 'cib', 'desi', 'des']
 cl_cross_k = {}
 cl_auto = {}
 rho = {}
@@ -115,6 +115,16 @@ for i in np.arange(0, len(labels)):
             cgg[i, j, :] = np.array([3500. * (1. * l / 3000.)**(-1.25) for l in lbins])
 
         cgg[j, i, :] = cgg[i, j, :]
+
+
+if cmb == 'Planck':
+
+    noise_phi = np.loadtxt('./quicklens/nlkk.dat')
+    # noise_cmb = nl(noise, beam, lmax=4000)
+    noise_fun = interp1d(noise_phi[:, 0], noise_phi[:, 1], bounds_error=False, fill_value=1e10)
+    ckk_noise = np.zeros_like(ckk)
+    ckk_noise = noise_fun(lbins)
+
 
 if cmb == 'S3':
     noise = 5.0

### 2017-04-09 Code/rho_to_Bres.py

@@ -5,7 +5,6 @@ import configparser as ConfigParser
 import numpy as np
 from joblib import Parallel, delayed
 import scipy.integrate as integrate
-from numba import jit
 
 
 def bl(fwhm_arcmin, lmax):
@@ -26,21 +25,20 @@ def nl(noise_uK_arcmin, fwhm_arcmin, lmax):
     return (noise_uK_arcmin * np.pi / 180. / 60.)**2 / bl(fwhm_arcmin, lmax)**2
 
 
-def compute_res_parallel(label_survey):
+def compute_res_parallel(rho_filename):
 
     print('start integration')
 
-    if label_survey == 'test':
-        @jit
+    if rho_filename == 'test':
         def integrand(theta, ell, L):
             clee = clee_fun(ell)
             return (ell / (2. * np.pi)**2 * (L * ell * np.cos(theta) - ell**2)**2 * clpp_fun(np.sqrt(L**2 + ell**2 - 2. * ell * L * np.cos(theta))) * clee * (np.sin(2. * theta))**2)
     else:
-        rho = np.loadtxt(
-            '/home/manzotti/galaxies_delensing/Data/limber_spectra/rho_' + label_survey + '.txt')
+        rho = np.loadtxt(rho_filename)
+        lbins = np.loadtxt('lbins.txt')
         rho_fun = InterpolatedUnivariateSpline(
-            rho[:, 0], np.nan_to_num(rho[:, 1]), ext='raise')
-        @jit
+            lbins, np.nan_to_num(rho), ext='raise')
+
         def integrand(theta, ell, L):
             clee = clee_fun(ell)
             return (ell / (2. * np.pi)**2 * (L * ell * np.cos(theta) - ell**2)**2 * clpp_fun(np.sqrt(L**2 + ell**2 - 2. * ell * L * np.cos(theta))) * clee * (np.sin(2. * theta))**2) * (1. - (clee / (clee + nle_fun(ell))) * rho_fun(ell) ** 2)
@@ -48,7 +46,7 @@ def compute_res_parallel(label_survey):
     clbb_res_ell = [integrate.dblquad(
         integrand, 4, 1500, lambda x: 0, lambda x: 2. * np.pi, args=(L,), epsabs=1.49e-08, epsrel=1.49e-07)[0] for L in np.arange(10, 700, 10)]
 
[…]

### 2017-05-05 Code/compare_kernel.py
new img

@@ -13,14 +13,17 @@ import numpy as np
 import scipy.integrate
 from scipy.interpolate import RectBivariateSpline, interp1d, InterpolatedUnivariateSpline
 import imp
+import pyximport
+pyximport.install(reload_support=True)
+import sys
+sys.path.append('/home/manzotti/cosmosis/modules/limber/')
+import gals_kernel
+import kappa_gals_kernel
+import kappa_cmb_kernel as kappa_kernel
 
-gals_kernel = imp.load_source(
-    'gals_kernel', '/home/manzotti/cosmosis/modules/limber/gals_kernel.py')
-kappa_kernel = imp.load_source(
-    'kappa_kernel', '/home/manzotti/cosmosis/modules/limber/kappa_cmb_kernel.py')
+
 cib_hall = imp.load_source('cib_hall', '/home/manzotti/cosmosis/modules/limber/hall_CIB_kernel.py')
-kappa_gals_kernel = imp.load_source(
-    'kappa_gals_kernel', '/home/manzotti/cosmosis/modules/limber/kappa_gals_kernel.py')
+
 spectra_dir = '/home/manzotti/galaxies_delensing/Data/'
 
 zpower = np.loadtxt(spectra_dir + 'matter_power_nl/z.txt')
@@ -52,9 +55,9 @@ d_m *= h0
 h /= h0
 xlss *= h0
 
-chispline = InterpolatedUnivariateSpline(zdist, d_m)
-z_chi_spline = InterpolatedUnivariateSpline(d_m, zdist)
-hspline = InterpolatedUnivariateSpline(zdist, h)
+chispline = InterpolatedUnivariateSpline(zdist[::-1], d_m[::-1])
+z_chi_spline = InterpolatedUnivariateSpline(d_m[::-1], zdist[::-1])
+hspline = InterpolatedUnivariateSpline(zdist[::-1], h[::-1])
 
 dndz_filename = '/home/manzotti/cosmosis/modules/limber/' + \
     'data_input/DES/N_z_wavg_spread_model_0.2_1.2_tpz.txt'
@@ -77,22 +80,24 @@ nu = 353e9
 zs = 2.
 b = 1.
 j2k = 1.e-6 / np.sqrt(83135.)  # for 353
[…]

### 2017-05-05 Code/multiple_survey_delens.py

@@ -125,6 +125,15 @@ if cmb == 'Planck':
     ckk_noise = np.zeros_like(ckk)
     ckk_noise = noise_fun(lbins)
 
+if cmb == 'LiteBird':
+    noise = 2.0
+    beam = 30
+    noise_phi = np.loadtxt('./quicklens/min_var_noise_{}muk_{}beam.txt'.format(noise, beam))
+    noise_phi *= np.arange(0, len(noise_phi))**4. / 4.
+    # noise_cmb = nl(noise, beam, lmax=4000)
+    noise_fun = interp1d(np.arange(0, len(noise_phi)), noise_phi)
+    ckk_noise = np.zeros_like(ckk)
+    ckk_noise = noise_fun(lbins)
 
 if cmb == 'S3':
     noise = 5.0

### 2017-05-05 Code/plots/plot_kernel.py

@@ -1,7 +1,6 @@
 
-from __future__ import division
+
 # from __future__ import print_function
-print "\n Python start up script from /home/manzotti/.py_startup.py.\n"
 import scipy as sp
 import numpy as np
 import matplotlib
@@ -19,7 +18,7 @@ except:
 try:
     import palettable
 
-except Exception, e:
+except Exception as e:
     pass
 import os
 import os
@@ -28,20 +27,20 @@ import time
 import subprocess
 
 from pprint import pprint
-import __builtin__
+import builtins
 import pyximport
 pyximport.install(reload_support=True)
 
 try:
     cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
     plt.rcParams['image.cmap'] = cmap
-except Exception, e:
-    print 'probably palettable not found'
+except Exception as e:
+    print('probably palettable not found')
 
 
-datadir = '../Data/limber_spectra/'
+datadir = '../../Data/limber_spectra/'
 
-image_dir = '../images/'
+image_dir = '../../images/'
 
 
 # Import data
@@ -49,49 +48,36 @@ kappakernel_l30 = np.loadtxt(datadir + 'kappa_kernel_l30.txt')
 kappakernel_l100 = np.loadtxt(datadir + 'kappa_kernel_l100.txt')
 kappakernel_l500 = np.loadtxt(datadir + 'kappa_kernel_l500.txt')
 
-kappakernel_l30_h = np.loadtxt(datadir + 'kappa_kernel_l30_h.txt')
-kappakernel_l100_h = np.loadtxt(datadir + 'kappa_kernel_l100_h.txt')
-kappakernel_l500_h = np.loadtxt(datadir + 'kappa_kernel_l500_h.txt')
 
-cibkernel_l30 = np.loadtxt(datadir + 'cib_kernel_l30.txt')
-cibkernel_l100 = np.loadtxt(datadir + 'cib_kernel_l100.txt')
-cibkernel_l500 = np.loadtxt(datadir + 'cib_kernel_l500.txt')
-
-cibkernel_l30_h = np.loadtxt(datadir + 'cib_kernel_l30_h.txt')
-cibkernel_l100_h = np.loadtxt(datadir + 'cib_kernel_l100_h.txt')
[…]

### 2017-05-05 Code/rho_to_Bres.py

@@ -44,10 +44,68 @@ def compute_res_parallel(rho_filename):
             return (ell / (2. * np.pi)**2 * (L * ell * np.cos(theta) - ell**2)**2 * clpp_fun(np.sqrt(L**2 + ell**2 - 2. * ell * L * np.cos(theta))) * clee * (np.sin(2. * theta))**2) * (1. - (clee / (clee + nle_fun(ell))) * rho_fun(ell) ** 2)
 
     clbb_res_ell = [integrate.dblquad(
-        integrand, 4, 1500, lambda x: 0, lambda x: 2. * np.pi, args=(L,), epsabs=1.49e-08, epsrel=1.49e-07)[0] for L in np.arange(10, 700, 10)]
+        integrand, 4, 2500, lambda x: 0, lambda x: 2. * np.pi, args=(L,), epsabs=1.49e-08, epsrel=1.49e-07)[0] for L in np.arange(4, 1500, 10)]
 
     np.savetxt(rho_filename.split('.txt')[0] + 'Cbb_res.txt', clbb_res_ell)
-    np.savetxt(datadir + 'limber_spectra/cbb_res_ls.txt', np.arange(10, 700, 10))
+    np.savetxt(datadir + 'limber_spectra/cbb_res_ls.txt', np.arange(4, 1500, 10))
+
+    return clbb_res_ell
+
+
+def compute_res(rho_filename, noise_pol=2., fwhm_beam=30.):
+
+    cosmosis_dir = '/home/manzotti/cosmosis/'
+    inifile = '/home/manzotti/cosmosis/modules/limber/galaxies_delens.ini'
+
+    Config_ini = ConfigParser.ConfigParser()
+    # values = ConfigParser.ConfigParser()
+    Config_ini.read(inifile)
+    # values_file = Config_ini.get('pipeline', 'values')
+    output_dir = Config_ini.get('test', 'save_dir')
+
+    datadir = output_dir
+
+    clpp = np.loadtxt(datadir + 'cmb_cl/pp.txt')
+    clee = np.loadtxt(datadir + 'cmb_cl/ee.txt')
+    ells_cmb = np.loadtxt(datadir + 'cmb_cl/ell.txt')
+
+    clee *= 2. * np.pi / (ells_cmb.astype(float) * (ells_cmb.astype(float) + 1.))
+    clpp = clpp * 2. * np.pi / (ells_cmb.astype(float) * (ells_cmb.astype(float) + 1.))
+
+    lbins = np.logspace(1, 3.5, 190)
+
[…]

### 2017-05-05 Code/rho_to_Bres_iter.py

@@ -6,6 +6,22 @@ import numpy as np
 from joblib import Parallel, delayed
 import scipy.integrate as integrate
 
+# import pyximport
+# pyximport.install()
+# import rho_to_Bres_iter_c
+import quicklens as ql
+
+
+def calc_nlqq(qest, clXX, clXY, clYY, flX, flY):
+
+    print("calculating full noise level for estimator of type")
+    clqq_fullsky = qest.fill_clqq(np.zeros(lmax + 1, dtype=np.complex),
+                                  clXX * flX * flX, clXY * flX * flY, clYY * flY * flY)
+    resp_fullsky = qest.fill_resp(qest, np.zeros(lmax + 1, dtype=np.complex), flX, flY)
+    nlqq_fullsky = clqq_fullsky / resp_fullsky**2
+
+    return nlqq_fullsky
+
 
 def bl(fwhm_arcmin, lmax):
     """ returns the map-level transfer function for a symmetric Gaussian beam.
@@ -97,20 +113,51 @@ def F_alpha(l1, l2, theta, x1='E', x2='E'):
 
 
 def reconstruction_noise_integrand(theta, ell, L, x1, x2):
+    # print(theta, ell, L, x1,x2)
     l2 = np.sqrt(L**2 + ell**2 - 2. * ell * L * np.cos(theta))
+    # print(np.log(ell / (2. * np.pi)**2 * fl1l2(ell, l2, theta, x1, x2) * F_alpha(ell, l2, theta, x1, x2)))
     if l2 < 4.:
-        return 1e30
+        return 1e20
     else:
-        return (ell / (2. * np.pi)**2 * fl1l2(ell, l2, theta, x1, x2) * F_alpha(ell, l2, theta, x1, x2))
+        return 1e-15 * (ell / (2. * np.pi)**2 * fl1l2(ell, l2, theta, x1, x2) * F_alpha(ell, l2, theta, x1, x2))
+
+# (ell / (2. * np.pi)**2 * fl1l2(ell, np.sqrt(L**2 + ell**2 - 2. * ell * L * np.cos(theta)), theta, x1, x2) * F_alpha(ell, np.sqrt(L**2 + ell**2 - 2. * ell * L * np.cos(theta)), theta, x1, x2))
 
 
-def reconstruction_noise(x1='E', x2='E'):
-    reconstruction_noise_ell = [integrate.dblquad(
[…]

### 2017-05-05 Notes/DES_rec_phi_paper.tex
add

@@
 \usepackage{epstopdf}
-\usepackage[caption=false]{subcaption}
-\epstopdfDeclareGraphicsRule{.gif}{png}{.png}{convert gif:#1 png:\OutputFile}
+%\usepackage[caption=false]{subcaption}
+%\epstopdfDeclareGraphicsRule{.gif}{png}{.png}{convert gif:#1 png:\OutputFile}
 \AppendGraphicsExtensions{.gif}
@@
 \begin{abstract}
-Goal: test if LSS survey like D.E.S or DESI can improve the reconstruction of the lensing potential that lenses the CMB photons.
+\begin{itemize}
+\item Worry : they might ask for iterative delensing
+\item Worry : they might ask for internal bias
+\item Possible TODO : $n_{T}$
+\item Possible TODO : temperature $N_{eff}$
+
+\end{itemize}
+
+Removing the lensing component from the measurement of CMB B-modes will be important to constraint the amplitude of the primordial signal.
+Here we discuss the role of large scale structure surveys survey like D.E.S or DESI in improving the reconstruction of the lensing potential that lenses the CMB photons.
 This would be crucial to build a template of the B-mode signal coming from the lensing of the primordial E-mode.
-The lensed B-mode component will be, at multipoles higher than 200, the main contaminants of the primordial B-mode.
-Preliminary: CIB performs as an equivalent $\rho_{\text{eff}}$  (correlation coefficient with cmb lensing potential constant over $\ell$) of 0.8.
+CIB performs as an equivalent $\rho_{\text{eff}}$  (correlation coefficient with cmb lensing potential constant over $\ell$) of 0.8.
 Both D.E.S and DESI are significantly worse, $\rho_{\text{eff}}<0.6$.
@@
 \begin{center}
-\includegraphics[scale=0.6]{../images/spectra.pdf}
-\caption{Correlation factor between galaxies survey and CMB lensing potential as a function of the multipole $\ell$. The ``CIB+D.E.S'' factor has been computed using \refeq{rho-combined}}
+\includegraphics[scale=0.5]{../images/actual_scenario.pdf}
+\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
 \end{center}
-\label{fig:corr}
+\label{fig:corrnow}
 
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=0.5]{../images/S3_scenario.pdf}
+\caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
+\end{center}
+\label{fig:corrS3}
+
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=0.5]{../images/S4_scenario.pdf}
+\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
+\end{center}
+\label{fig:corrs4}
+
+\end{figure}
 %=======================================
@@
 where $P_{init}(k)$ is the scalar power spectrum at an early time with $k$ being the Fourier wave 
-number. The functions $\Delta^X\l(k)$ and $\Delta^Y\l(k)$ are one of the following (see e.g., 
+number. The functions $\Delta^X_{\ell}(k)$ and $\Delta^Y_{\ell}(k)$ are one of the following (see e.g., 
 \cite{Hu:2000ee,LoVerde:2006cj}):
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=0.5]{../images/cmb_internal.pdf}
+\caption{Correlation factor between CMB reconstructed potential and the real lensing potential for different CMB experiments}
+\end{center}
+\label{fig:corr-internal}
+
+\end{figure}
+
 \subsection{Galaxies}
@@
 \begin{split}
-W^{g}(z) &= \frac{b(z)\frac{dN}{dz}}{\Bigl(\int dz'\,\frac{dN}{dz'}\Bigr)} + \frac{3\Omega_{\rm m}}{2c}\frac{H_0^2}{H(z)}(1+z)\chi(z) \\
-&\times \int_z^{z_*}dz'\,\Bigl(1-\frac{\chi(z)}{\chi(z')}\Bigr)(\alpha(z')-1)\frac{dN}{dz'}.
+W^{g}(z) &= \frac{b(z)\frac{dN}{dz}}{\Bigl(\int dz'\,\frac{dN}{dz'}\Bigr)}
 \end{split}
@@
 \cite{hall:2010,sherwin:2015,ade:2014,hanson:2013,szapudi:2001,0004-637X-567-1-2,planck-collaboration:2014,planck-collaboration:2014a,planck-collaboration:2011,plank-collaboration:2014,boulanger:1996,lewis:2006}
+
 \subsection{Internal CMB lensing}
+
+The ability of CMB experiments themselves to reconstruct the lensing potential with rapidly improve with the increasing sensitivity.
+
+Internal delensing is already competitive with galaxies delensing at very large scale.
+Potentially
+
+To derive the delensing performance let us assume that our lensing reconstruction, given certain instrumental noise level and beam, results in a lensing map with reconstruction noise $N_l^{\kappa \kappa}$, so that its power spectrum is
+\beq
+\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_l 
+\eeq
 
@@
 \label{sec:for}
+
+We defined the B-modes noise spectrum:
+\beq
+N_l^{BB} = \left({\Delta_P}/{T_{\mathrm{CMB}}}\right)^2  e^{ {l^2 \theta_\mathrm{FWHM}^2}/({8 \ln 2})}
+\eeq
+where $\theta_\mathrm{FWHM}$ is the full half width of the telescope  beam and $\Delta_P$ is the instrumental noise of the experiment.
+The gaussian covariance is then:
+\beq
+\sigma( C_{l}^{BB,\mathrm{full}} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
+\eeq
+With this we can simply quantify the constraints on the tensor to scalar ration r:
+\beqn
+&\sigma(r)&= \left[\sum_l \frac{\left(\frac{\partial {C}_l^{BB,r}}{\partial r}\right)^2}{\sigma^2( C_{l}^{BB,\mathrm{full}} )} \right]^{-\frac{1}{2}}  \\ &\approx&  \left[ \frac{\sum_l (2 {l}+1) \fsky \left(\frac{\partial {C}_l^{BB,r}}{\partial r}\right)^2}{2} \right]^{-\frac{1}{2}}\langle C_l^{BB,\mathrm{lens}} +N^{BB}_l \rangle_{l<100}\nonumber
+\eeqn
+where to go from the first to the second line we use the fact that most of the constraint comes from large scale modes at $\ell <100$, thus larger mode can be ignore here.
+We quantify the improvement due to delensing as the factor $\alpha$ defined as the ration of the error before and after delensing:
+\beq
+\alpha = \frac{ \langle C_l^{BB,\mathrm{lens}} +N^{BB}_l[\Delta_P] \rangle_{l<100}}{\langle C_l^{BB,\mathrm{res}}[\Delta_P]+N^{BB}_l[\Delta_P] \rangle_{l<100}}
+\eeq
+
+\subsection{Current generation}
 \subsection{CMB-S3 Era}
+The accuracy of CMB is rapidly improving.
+For example the next generation of the SPT telescope, SPT3G has been deployed and is currently taking data.
+Furthermore other experiments have considerably increased the number of detectors (AdvACT etc.)
+
 \subsection{CMB-S4 Era}
+
+An ambitious program for a generation 4 ground CMB experiment is currently under planning. Moreover satellite experiments have been proposed.
+
+Following we assume an instrumental level of noise of $1 \mu K$-arcmin and a telescope beam of 1 arcmin.
+
 \subsection{Bias uncertainties degradation}
+The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of inflationary constraint of delensed spectra.
+In this section we quantify this effect.
+We will now marginalize over unknown galaxies parameters but we will use a full dataset of CMB and galaxies data.
+The idea is that as shown in the low level of noise in galaxies surveys might allow us to internally calibrate them.
+We will use a Fisher approach the Fisher matrix is:
+\beqn
+F_{pq} &=& \sum_{l_{a} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}  \sum_{l_{b} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}   \frac{\partial {C}_{l_a}^{BB,\mathrm{del}}}{\partial \theta_p } \left[ \mathrm{Cov}^{BB,BB}\right]_{l_a, l_b} ^{-1} \frac{\partial {C}_{l_b}^{BB,\mathrm{del}}}{\partial \theta_q }\nonumber \\&+&  \sum_j \frac{\frac{\partial {C}_j^{\kappa I}}{\partial \theta_p } \frac{\partial {C}_j^{\kappa I}}{\partial \theta_q }}{(\Delta C_j^{\kappa I})^2}
++\sum_j \frac{\frac{\partial {C}_j^{I I}}{\partial \theta_p } \frac{\partial {C}_j^{I I}}{\partial \theta_q }}{(\Delta C_j^{I  I})^2}
+\eeqn
 
-%----------------------------------------------------------------------------------------------------%
-%\al{
-%	F_{ij} &\equiv \sum_{\l=\l\rom{min}}^{100} \frac{2\l+1}{2[C_\l^{\rm BB,res}+N_\l^{\rm BB}]^2}
-%		\PD{C^{\rm BB,res}_\l}{\theta_i}\PD{C^{\rm BB,res}_\l}{\theta_j}
-%	\notag \\
-%	&+ \sum_{L=2}^{1000} \frac{2L+1}{2} 
-%		\Tr \left[\bR{C}^{-1}_L\PD{\bR{C}_L}{\theta_i}\bR{C}^{-1}_L\PD{\bR{C}_L}{\theta_j}\right]
-%	\,, 
-%}
-%----------------------------------------------------------------------------------------------------%
-where $\theta_i=r$, $b_i$ or $p_i$ \footnote{In our analysis, since the fiducial value of $r$ is zero, 
+\beq
+\alpha_{\mathrm{marginalized}} = \sigma_0(r) / \sigma_{\mathrm{marginalized/delensed}}(r)
+\eeq
+
+where the parameter array contains both the tensor to scalar rate $\theta=r$, and the galaxies surveys parameters like the bias $b_i$ or $p_i$ \footnote{In our analysis, since the fiducial value of $r$ is zero,
 the derivative of $C_\l^{\rm BB,res}$ is non-zero if $\theta_i=r$.}.
-The covariance matrix of the mass tracers $\bR{C}_L$ is given in Eq.~\eqref{Eq:cov}. 
-We compute the derivatives of the power spectra as described in appendix \ref{sec:deriv}. 
-Note that we omit the correction of the partial sky coverage since it does not affect our results, i.e.,
+
+We compute the derivatives of the power spectra as described
 

### 2017-05-05 Code/compare_kernel.py
new img

@@ -88,7 +88,7 @@ desi = gals_kernel.kern(dndz[:, 0], dndzfun_desi, hspline, omega_m, h0)
 
 des_weak = kappa_gals_kernel.kern(dndz_des[:, 0], dndzfun_des, chispline, hspline, omega_m, h0)
 
-l = 500
+l = 30
 z_kappa = np.linspace(0, 13, 500)
 w_kappa = np.zeros_like(z_kappa)
 for i, z in enumerate(z_kappa):

### 2017-05-05 Code/plots/plot_kernel.py

@@ -172,7 +172,7 @@ fg.tight_layout()
 
 # ============================================
 # FINALLY SAVE
-plt.savefig(image_dir + 'compare_kernel.png', dpi=600, papertype='Letter',
-            format='png')
+plt.savefig(image_dir + 'compare_kernel.pdf', dpi=600, papertype='Letter',
+            format='pdf')
 
 plt.clf()

### 2017-05-05 Code/plots/plot_kernel.py
new img

@@ -67,7 +67,7 @@ desikernel_l500 = np.loadtxt(datadir + 'desi_kernel_l500.txt')
 # ============================================
 # SIZE OF THE PICTURE
 # ============================================
-WIDTH = 256.0  # the number latex spits out
+WIDTH = 500.0  # the number latex spits out
 # the fraction of the width you'd like the figure to occupy, 0.5 if double
 # column cause
 FACTOR = 0.6

### 2017-05-05 Notes/DES_rec_phi_paper.tex

@@
-\documentclass[onecolumn,amsmath,amssymb,floatfix,superscriptaddress,notitlepage,aps,prd]{revtex4-1}
+\documentclass[reprint,amsmath,amssymb,floatfix,aps,prd]{revtex4-1}
 \usepackage{bm}
@@
 \usepackage{xstring}
-\usepackage{graphicx}\usepackage[top=2.5 cm, bottom=2.5 cm, left=1.05in, right=1.05in]{geometry}
+% \usepackage{graphicx}\usepackage[top=2.5 cm, bottom=2.5 cm, left=1.05in, right=1.05in]{geometry}
 \usepackage[applemac]{inputenc}
@@
 %\epstopdfDeclareGraphicsRule{.gif}{png}{.png}{convert gif:#1 png:\OutputFile}
-\AppendGraphicsExtensions{.gif}
-% notes to make comment
+% \AppendGraphicsExtensions{.gif}
+% notes to make comment is non-zero if
 \newcommand{\al}[1]{\begin{align} #1 \end{align}}
@@
 \item Worry : they might ask for iterative delensing
-\item Worry : they might ask for internal bias
-\item Possible TODO : $n_{T}$
-\item Possible TODO : temperature $N_{eff}$
+\item Worry : they might ask for internal bias\
+item Possible TODO : temperature $N_{eff}$
 
@@
 
-Removing the lensing component from the measurement of CMB B-modes will be important to constraint the amplitude of the primordial signal.
-Here we discuss the role of large scale structure surveys survey like D.E.S or DESI in improving the reconstruction of the lensing potential that lenses the CMB photons.
-This would be crucial to build a template of the B-mode signal coming from the lensing of the primordial E-mode.
-CIB performs as an equivalent $\rho_{\text{eff}}$  (correlation coefficient with cmb lensing potential constant over $\ell$) of 0.8.
-Both D.E.S and DESI are significantly worse, $\rho_{\text{eff}}<0.6$.
-DESI is slightly better than D.E.S.
-To delens we need the bulk of the redshift distribution to follow the CMB kernel, having a few outliers at redshift $z>1.5$ is not enough.
-We also have to keep in mind the results of \cite{smith:2012}: even a perfect LSS survey can not help too much the delensing process, with the possible exception of futuristic 21 cm.
+The measurement of the CMB polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe.
+A particular component, the so called B-modes is indeed a direct signature of the presence of gravitational waves in the early universe.
+However improving the level of noise is not enough.
+This is even more true if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
+Removing the lensing component from the measurement of CMB B-modes will be important to constraint the amplitude of the primordial gravitational wave contribution to the signal.
+Here we discuss the role of current and future large scale structure surveys  in improving the reconstruction of the lensing potential that lenses the CMB photons.
+We find that..
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:intro}
+Inflation Physics with CMB
+
+Current constraints
+
+Now limiting by delensing
+
+Done what are the prospects/
+
+This paper is organized as follow:
 
@@
 %=======================================
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=0.5]{../images/actual_scenario.pdf}
-\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
-\end{center}
-\label{fig:corrnow}
 
-\end{figure}
+% %=======================================
+% \begin{figure}[htbp]
+% \begin{center}
+% \includegraphics[scale=1.]{../images/clbb_res.pdf}
+% \caption{Residual lensing B modes power spectrum using different large scale structure. \refeq{combined}}
+% \end{center}
+% \label{fig:bres}
+% \end{figure}
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=0.5]{../images/S3_scenario.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
-\end{center}
-\label{fig:corrS3}
+% %=======================================
 
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=0.5]{../images/S4_scenario.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
-\end{center}
-\label{fig:corrs4}
-
-\end{figure}
-%=======================================
-
-%=======================================
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/clbb_res.pdf}
-\caption{Residual lensing B modes power spectrum using different large scale structure. \refeq{combined}}
-\end{center}
-\label{fig:bres}
-\end{figure}
-
-%=======================================
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{Comparison of the different kernels used in this analysis. This allow to understand how well and where in redshift space different LSS surveys trace the CMB lensing potential.
-Redshift distribution of D.E.S galaxies (I suspect this is the benchmark, anyway taken from Giannantionio et al.). DESI taken from their white paper. }
-\end{center}
-\label{fig:kernel}
-\end{figure}
 As for the temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is modified by lensing as:
@@
 
-From this we get that the power spectrum:
+From this we get that the power spectrum of the lensing component of the B-modes:
 \be
@@
 
-\subsection{Galaxies contribution at different redshifts (CIB+D.E.S)}\label{sec:galcontrib}
+\subsection{Multiple tracers of the lensing potential}\label{sec:galcontrib}
 
@@
 
+% ===================
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/compare_kernel.pdf}
+\caption{Comparison of the different kernels used in this analysis. This allow to understand how well and where in redshift space different LSS surveys trace the CMB lensing potential.
+Redshift distribution of D.E.S galaxies (I suspect this is the benchmark, anyway taken from Giannantionio et al.). DESI taken from their white paper. }
+\end{center}
+\label{fig:kernel}
+\end{figure}
+% ===================
+
 \section{Tracers Model}
@@
 %
-where $\Omega_{\rm m}$ and $H_0$ are the present-day values of the Hubble and matter density parameters, respectively.
+where $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
 
@@
 \label{sec:for}
+In this section we forecast the expected delensing efficiency and the relative importance of galaxies tracers for current and future experiments.
+We focus on 3 different scenarios: the current stage, the 3G one and the futuristic CMB Stage 4.
+
+\subsection{Fisher method}
 
@@
 \subsection{Current generation}
+
+% \the\textwidth
+
+Very recently delensing has been proven possible on real data using both the internal CMB lensing potential reconstruction as well as external tracers.
+Here we discuss the improvement due to combining this two probes.
+This is already possible both on full sky using Planck data or in smaller sky patches combining Herschel or Planck CIB data with SPT and BK data.
+Furthermore we test how much current relative low redshift galaxies surveys like D.E.S can improve delensing efficiency on smaller patch of the sky like the Bicep-Keck SPT-Pol footprint.
+
+We model the current generation pf CMB experiments after 3 experiments: a full sky Planck experiment together with a deep low noise experiemtn and a high resolution one on a smaller patch of the sky.
+
+The correlation attainable using current experiments is shown in \reffig{corrnow}.
+
+At large angular scale the CMB internal reconstruction is clearly the best tracers for delensing.
+Because of the large level of noise in current experiment, its efficiency became comparable to the CIB one at few degrees ($\ell \simeq 400$).
+
+\begin{figure}[htbp]
+
+\begin{center}
+\includegraphics[scale=0.5]{../images/actual_scenario.pdf}
+\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
+\label{fig:corrnow}
+\end{center}
+\end{figure}
+
 \subsection{CMB-S3 Era}
@@
 For example the next generation of the SPT telescope, SPT3G has been deployed and is currently taking data.
-Furthermore other experiments have considerably increased the number of detectors (AdvACT etc.)
+Furthermore other experiments have considerably increased the number of detectors (AdvACT etc.).
+We will assume a level of noise of 3 $\mu$K-arcmin the level predicted for SPT3G.
+The correlation attainable using current experiments is shown in \reffig{corrS3},
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=0.5]{../images/S3_scenario.pdf}
+\caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
+\label{fig:corrS3}
+
+\end{center}
+
+\end{figure}
 
@@
 
-Following we assume an instrumental level of noise of $1 \mu K$-arcmin and a telescope beam of 1 arcmin.
+The correlation attainable using current experiments is shown in \reffig{corrs4},
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=0.5]{../images/S4_scenario.pdf}
+\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
+\label{fig:corrs4}
+
+\end{center}
+\end{figure}
+%=======================================
 
@@
 where the parameter array contains both the tensor to scalar rate $\theta=r$, and the galaxies surveys parameters like the bias $b_i$ or $p_i$ \footnote{In our analysis, since the fiducial value of $r$ is zero,
-the derivative of $C_\l^{\rm BB,res}$ is non-zero if $\theta_i=r$.}.
+the derivative of $C_\ell^{\rm BB,res}$ is non-zero if $\theta_i=r$.}.
 
@@
 
+\begin{table}
+\caption{$\alpha$: improvement on r constraint}
+  \begin{tabular}{ | c | c  | }
+\hline
+Surveys & $\alpha$\\ \hline \hline
+des & 1.34  \\ \hline
+cib & 1.71  \\ \hline
+cmb current & 1.79  \\ \hline
+gals current & 2.03  \\ \hline
+cmb S3 & 2.15  \\ \hline
+gals S3 & 2.16  \\ \hline
+gals S4 & 2.16  \\ \hline
+comb current & 2.66  \\ \hline
+comb S3 & 3.14  \\ \hline
+cmb S4 & 4.36  \\ \hline
+comb S4 & 5.27  \\ \hline
+  \end{tabular}
+\end{table}
+
 \section{Conclusions}
+Delensing or more in general the ability to separate the lensing component of the B-mode from a possible primordial inflationary signal is crucial to fully exploit the capabilities of future experiments.
+In this paper we studied the possible impact of large scale structure surveys in this important endeavor.
+For current experiment CIB data had already proven to be very efficient in delensing.
+If no High res experiment is avaiable to reconstruct the lensing potential phi even a low-z experiment like D.E.S will improve th delensign efficiency by $~10\%$ if combined with Planck lensing and a CIB tracers.
+As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential.
+Indeed using the CMB itself we reconstruct only the lenses that actually lens the CMB obtained an almost perfect cleaning in the absence of noise.
+This approach will need a careful study of possible biases coming from using the same source (the CMB) we are delensing to reconstruct the lensing effect itself.
+This has already been applied to data and studied.
+However the fairly good efficiency of galaxies tracers might come in end to cross-check these internal biases.
+This will probably be needed to confirm a possible detection of gravitation waves if this rely heavily on delensing.
 
 \bibliographystyle{plain}
-\bibliography{DES_rec_phi_paper}
+\bibliography{DES_rec_phi_paper,/home/manzotti/Astrophysics/sptpol_papers/BIBTEX/spt}
 

### 2017-05-05 Code/plots/plot_kernel.py
images

@@ -67,7 +67,7 @@ desikernel_l500 = np.loadtxt(datadir + 'desi_kernel_l500.txt')
 # ============================================
 # SIZE OF THE PICTURE
 # ============================================
-WIDTH = 500.0  # the number latex spits out
+WIDTH = 300.0  # the number latex spits out
 # the fraction of the width you'd like the figure to occupy, 0.5 if double
 # column cause
 FACTOR = 0.6

### 2017-05-05 Code/plots/plot_kernel.py
images

@@ -67,7 +67,7 @@ desikernel_l500 = np.loadtxt(datadir + 'desi_kernel_l500.txt')
 # ============================================
 # SIZE OF THE PICTURE
 # ============================================
-WIDTH = 300.0  # the number latex spits out
+WIDTH = 456.0  # the number latex spits out
 # the fraction of the width you'd like the figure to occupy, 0.5 if double
 # column cause
 FACTOR = 0.6

### 2017-05-05 Code/plots/plot_kernel.py
images

@@ -67,10 +67,10 @@ desikernel_l500 = np.loadtxt(datadir + 'desi_kernel_l500.txt')
 # ============================================
 # SIZE OF THE PICTURE
 # ============================================
-WIDTH = 456.0  # the number latex spits out
+WIDTH = 246.0  # the number latex spits out
 # the fraction of the width you'd like the figure to occupy, 0.5 if double
 # column cause
-FACTOR = 0.6
+FACTOR = 0.99
 fig_width_pt = WIDTH * FACTOR
 
 inches_per_pt = 1.0 / 72.27

### 2017-05-05 Code/plots/plot_kernel.py
images

@@ -90,7 +90,7 @@ plt.rc('font', **{'family': 'serif', 'serif': ['Computer Modern Roman']})
 # ============================================
 
 
-font_size = 5.
+font_size = 8.
 plt.rcParams['font.size'] = font_size
 plt.rcParams['axes.labelsize'] = font_size * 1.3
 plt.rcParams['axes.linewidth'] = font_size / 10.

### 2017-05-11 Notes/DES_rec_phi_paper.tex
-mtext

@@
-\documentclass[reprint,amsmath,amssymb,floatfix,aps,prd]{revtex4-1}
+\documentclass[reprint,superscriptaddress,amsmath,amssymb,floatfix,aps,prd]{revtex4-1}
 \usepackage{bm}
@@
 \author{A. Manzotti}        % Enter your name between curly braces
+\email{manzotti.alessandro@gmail.com}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, 5640 South Ellis Avenue, Chicago, IL 60637, USA}
+\affiliation{Department of Astronomy and Astrophysics, University of Chicago, 5640 South Ellis Avenue, Chicago, IL 60637, USA}
 \date{\today}
@@
 \begin{itemize}
-\item Worry : they might ask for iterative delensing
 \item Worry : they might ask for internal bias\
@@
 \eeqn
-and its power spectrum
-\beqn
+and its power spectru
+m\beqn
 C_{{l}}^{BB,\mathrm{res}} &=&   \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl')
@@
 
-The gain we  have in adding a new tracer is not only proportional to its correlation with the CMB lensing but it also depends on how much it is correlated with the already used set of tracers.
-For example there s a very small gain in adding different CIB frequencies because they are highly correlated with each other.
-On the contrary a CMB reconstructed lensing potential or some low redshift survey like D.E.S can help.
+Note that gain we have in adding a new tracer is not only proportional to its correlation with the CMB lensing but it also depends on how much it is correlated with the already used set of tracers.
+\reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in the following sections.
+Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST DES and DESI curves.
+However given the significant low noise of this measurement they can still help the delensing of the CMB.
+Furthermore the low-redshift tail of the lensing kernel is also the one affected by large scale structures non-linearities.
+They can indeed be extremely helpful if future CMB surveys will try to remove that contribution instead of attempting to model it.
+On the other hand CIB and 21 cm surveys are target to high redshift structures and independently from the model assumed they show a fairly good overlap with the CMB lensing kernel.
 
@@
 \includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{Comparison of the different kernels used in this analysis. This allow to understand how well and where in redshift space different LSS surveys trace the CMB lensing potential.
-Redshift distribution of D.E.S galaxies (I suspect this is the benchmark, anyway taken from Giannantionio et al.). DESI taken from their white paper. }
+\caption{Comparison of the different kernels used in this analysis.
+The bigger the overlap wit the CMB lensing kernel the better the reconstruction of the lensing potential will be.
+This will lead to a better delensing.
+}
+\label{fig:kernel}
+
 \end{center}
-\label{fig:kernel}
 \end{figure}
@@
 \label{sec:data}
+In this section we briefly describe how the power spectra of \refsec{th} are computed and what it is assumed for the different tracers considered in this work.
 
-In this equation, $\chi(z)$ is the comoving distance to redshift $z$, $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, and $\Psi(\chi(z)\nver,z)$ is the three-dimensional gravitational potential at a point on the photon path given by $\chi(z)\nver$.
-Note that the deflection angle is given by $\mathbf{d}(\nver) = \nabla\phi(\nver)$, where $\nabla$ is the the two-dimensional gradient on the sphere.
-Because the lensing potential is an integrated measure of the projected gravitational potential, taking the two-dimensional Laplacian of the lensing potential we can define the lensing convergence $\kappa(\nver) = -\case{1}{2}\nabla^2\phi(\nver)$, which depends on the projected matter overdensity $\delta$ \citep{bart01}:
+Given a 2D projected field $\delta^{i}$ this can be seen as a 3D field project along the line of site as:
+
 \begin{equation}
-\kappa(\nver) = \int_0^{z_*} dz\, W^{\kappa}(z)\delta(\chi(z)\nver,z).
+\delta^{i}(\theta) = \int_0^{\infty} dz\, W^{\kappa}(z)\delta^{i}(\chi(z)\nver,z).
 \label{eqn:wkappa}
 \end{equation}
+
+Using the Limber approximation \cite{limber53} we can compute the power spectra of 2 fields $i,j$ as:
+\begin{equation}
+C_{\ell}^{ij}= \int_0^{\infty} \frac{dz}{c} \frac{H(z)}{\chi(z)^2}\, W^{i}(z)W^{j}(z)P(k,z).
+\label{eqn:wkappa}{}
+\end{equation}
+
+In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, $\chi(z)$ the comoving distance $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ at redshift $z$.
+Furthermore $W^{i}(z)$ is the kernel function of the field i.
+These theoretical quantities have been computed using CAMB and HALOFIT.
+
+\subsection{CMB lensing potential}
+\label{sec:kappaCMB}
+CMB photons interact gravitationally with low-redshift large scale structure.
+The deflection angle is given by $\mathbf{d}(\nver) = \nabla\phi(\nver)$, where $\nabla$ is the the two-dimensional gradient on the sphere.
+Because the lensing potential is an integrated measure of the projected gravitational potential, taking the two-dimensional Laplacian of the lensing potential we can define the lensing convergence $\kappa(\nver) = -\case{1}{2}\nabla^2\phi(\nver)$, which depends on the projected matter overdensity $\delta$
+% \citep{bart01}:
 The lensing kernel $W^{\kappa}$ is
 %
+$\chi(z)$ is the comoving distance to redshift $z$, $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$
 \begin{equation}
@@
 where $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
-
-%----------------------------------------------------------------------------------------------------%
-%\begin{eqnarray} \label{PS}
-%	C_\l^{XY} = \frac{2}{\pi}\Int{}{k}{k^2} P_{init}(k) \Delta_\l^X(k)\Delta_\l^Y(k)
-%	\,,
-%\end{eqnarray}
-%----------------------------------------------------------------------------------------------------%
-where $P_{init}(k)$ is the scalar power spectrum at an early time with $k$ being the Fourier wave
-number. The functions $\Delta^X_{\ell}(k)$ and $\Delta^Y_{\ell}(k)$ are one of the following (see e.g.,
-\cite{Hu:2000ee,LoVerde:2006cj}):
-%----------------------------------------------------------------------------------------------------%
-%\begin{eqnarray}
-%
-%	\Delta^I_\l(k) &= k^2 \INT{}{z}{}{0}{z_{\rm max}} b(z) \D{N}{z}(z) D(z) j_\l(k\chi(z))
-%	\,, \label{eq:Delta^I} \\
-%	\Delta^\grad_\l(k) &= \frac{3\Omega\rom{m}H_0^2}{2}
-%	\notag \\
-%		&\times \INT{}{\chi}{}{0}{\chi_*}
-%			\frac{\chi_*-\chi}{\chi_*\chi}\,\frac{D(z(\chi))}{a(\chi)}\,j_\l(k\chi)
-%	\,.
-%
-%\end{eqnarray}
-
-%----------------------------------------------------------------------------------------------------%
+The CMB lensing potential is \textit{the} field that we need to reconstruct in order to reverse the effect of large scale structure and delens the CMB.
+This lensing potential can be reconstructed from the CMB.
+In that case it can be see as a noisy tracer of the true field.
+While the CMB internal reconstruction fiels as the same kernel as the true one $W^{\kappa}(z)$.
+Given certain instrumental noise level and beam, results in a lensing map with reconstruction noise $N_l^{\kappa \kappa}$, so that its power spectrum is
+\beq
+\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_l
+\eeq
 
@@
 \begin{center}
-\includegraphics[scale=0.5]{../images/cmb_internal.pdf}
+\includegraphics[scale=0.9]{../images/cmb_internal.pdf}
 \caption{Correlation factor between CMB reconstructed potential and the real lensing potential for different CMB experiments}
@@
 
-The galaxy overdensity $g(\nver)$ in a given direction on the sky is also expressed as a LOS integral of the matter overdensity:
+The galaxy overdensity $g(\nver)$ in a given direction on the sky is also expressed as a line of sight integral of the matter overdensity:
 \begin{equation}
@@
 \begin{split}
-W^{g}(z) &= \frac{b(z)\frac{dN}{dz}}{\Bigl(\int dz'\,\frac{dN}{dz'}\Bigr)}
+W^{g}(z) &= \frac{b(z)\frac{dN}{dz}}{\Bigl(\int dz'\,\frac{dN}{dz'}\Bigr)}.
 \end{split}
 \end{equation}
+Here $\frac{dN}{dz}$ is the number of galaxies as a function of redshift observed by the survey while $b(z)$ is the galaxy bias that connect at different redshift the amplitude of galaxies overdensity to the one of the dark matter.
+Different galaxy surveys in this work are fully characterized by their b(z) and $\frac{dN}{dz}$.
 
@@
 \label{sec:cib}
-In this work, the CMB lensing potential is estimated from maps of the cosmic infrared background (CIB).
-
-Following \cite{Addison:2011se}, we model the CIB power as $C_{\ell}^{{\rm CIB}\mbox{-}{\rm CIB}} = 3500 (l/3000)^{-1.25} {\rm Jy^2 / sr}$.
-We test that this model provides an accurate fit for the power of the \textit{Herschel} $500\,\mu$ map used in this work.
-For the cross-spectrum $C_{\ell}^{{\rm CIB}\mbox{-}\phi}$, we use the single-SED model of~\cite{hall10}.
+The definition of the kernel of the CIB is more complex and model dependent.
+Following \cite{Addison:2011se}, we model the CIB power directly as $C_{\ell}^{{\rm CIB}\mbox{-}{\rm CIB}} = 3500 (l/3000)^{-1.25} {\rm Jy^2 / sr}$.
+This model provides an accurate fit for the power of the \textit{Herschel} $500\,\mu$ map used in this work.
+For the cross-spectra with the CMB lensing or other galaxies tracers $C_{\ell}^{{\rm CIB}\mbox{-}j}$,we use the single-SED model of~\cite{hall10}.
 This places the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$.
@@
 Other multi-frequency CIB models are available \citep[e.g.,][]{Bethermin:2013nza}; however, given the level of noise, we are relatively insensitive to this choice.
-With these assumptions, depending on angular scale, $45-65\%$ of the CIB is correlated with the CMB lensing potential, as shown in \reffig{CIB}.
-
-\subsection{Weak Lensing}
+With these assumptions, depending on angular scale, $45-65\%$ of the CIB is correlated with the CMB lensing potential, as shown in \reffig{corrnow}.
+% \subsection{Weak Lensing}
 \cite{hall:2010,sherwin:2015,ade:2014,hanson:2013,szapudi:2001,0004-637X-567-1-2,planck-collaboration:2014,planck-collaboration:2014a,planck-collaboration:2011,plank-collaboration:2014,boulanger:1996,lewis:2006}
 
-\subsection{Internal CMB lensing}
+We test the delensing efficiency modelling  current survey after D.E.S and using as future probe galaxy clustering measured with D.E.S.I and LSST together with 21 cm measurement like SKA.
 
-The ability of CMB experiments themselves to reconstruct the lensing potential with rapidly improve with the increasing sensitivity.
+For DESI we used the ..
 
-Internal delensing is already competitive with galaxies delensing at very large scale.
-Potentially
+LSST is model following
 
-To derive the delensing performance let us assume that our lensing reconstruction, given certain instrumental noise level and beam, results in a lensing map with reconstruction noise $N_l^{\kappa \kappa}$, so that its power spectrum is
-\beq
-\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_l
-\eeq
+We model SKA following
 
@@
 \begin{center}
-\includegraphics[scale=0.5]{../images/actual_scenario.pdf}
+\includegraphics[scale=0.9]{../images/actual_scenario.pdf}
 \caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
@@
 \end{figure}
+
+\begin{figure}[htbp]
+
+\begin{center}
+\includegraphics[scale=0.9]{../images/actual_scenario_planck.pdf}
+\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
+\label{fig:corrnow}
+\end{center}
+\end{figure}
+
+\begin{table}
+\caption{$\alpha$: Current generation improvements}
+  \begin{tabular}{ | c | c  | }
+\hline
+Surveys & $\alpha$\\ \hline \hline
+des & 1.34  \\ \hline
+cib & 1.71  \\ \hline
+  \end{tabular}
+\end{table}
 
@@
 \begin{center}
-\includegraphics[scale=0.5]{../images/S3_scenario.pdf}
+\includegraphics[scale=0.9]{../images/S3_scenario.pdf}
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
@@
 
+\begin{table}
+\caption{$\alpha$: Stage-3 improvements}
+  \begin{tabular}{ | c | c  | }
+\hline
+Surveys & $\alpha$\\ \hline \hline
+des & 1.34  \\ \hline
+cib & 1.71  \\ \hline
+  \end{tabular}
+\end{table}
+
 \subsection{CMB-S4 Era}
@@
 
+\begin{table}
+\caption{$\alpha$: Stage-4 improvements}
+  \begin{tabular}{ | c | c  | }
+\hline
+Surveys & $\alpha$\\ \hline \hline
+des & 1.34  \\ \hline
+cib & 1.71  \\ \hline
+  \end{tabular}
+\end{table}
+
 \begin{figure}[htbp]
 \begin{center}
-\includegraphics[scale=0.5]{../images/S4_scenario.pdf}
+\includegraphics[scale=0.9]{../images/S4_scenario.pdf}
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
@@
 This will probably be needed to confirm a possible detection of gravitation waves if this rely heavily on delensing.
-
+\section*{Acknowledgment}
 \bibliographystyle{plain}

### 2017-05-11 Notes/DES_rec_phi_paper.tex
text

@@
 
-\title{Future CMB delensing with galaxies surveys.}         % Enter your title between curly braces
+\title{Future Cosmic Microwave Background delensing with galaxies surveys.}         % Enter your title between curly braces
 \author{A. Manzotti}        % Enter your name between curly braces
@@
 
-\section{Theory}
+\section{Gravitational lensing B-mode}
 \label{sec:th}
 
-%=======================================
+As for the CMB photons temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is remapped by lensing as:
+\be
+Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi);~~
+U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi)
+\ee
+where $\nabla\phi$ is the deflection angle directly related to the lensing potential potential integrated along the line of sight $\phi$.
+The CMB polarization is usually decompose into odd-parity Fourier modes E and B. As shown in \cite{} this for symmetry reason the B-modes configuration is mainly sourced by tensor perturbations.
+However this promising sign of primordial gravitational waves is partially obscured by a secondary mechanism that can produce B-modes in the late universe.
+Indeed gravitational interaction with large scale structure generates CMB B-modes by distorting primordial E-modes.
 
-% %=======================================
-% \begin{figure}[htbp]
-% \begin{center}
-% \includegraphics[scale=1.]{../images/clbb_res.pdf}
-% \caption{Residual lensing B modes power spectrum using different large scale structure. \refeq{combined}}
-% \end{center}
-% \label{fig:bres}
-% \end{figure}
-
-% %=======================================
-
-As for the temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is modified by lensing as:
-\be
-Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\mathbf{d});~~
-U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\mathbf{d})
-\ee
-where $\mathbf{d}$ is the deflection angle directly related to the lensing potential $\phi$.
-
-As a first approximation, the B mode resulting from the lensing of primordial E mode by a convergence field $\kappa$ is:
+As a first approximation, given a convergence field $\kappa= -\frac{1}{2}\nabla^2\phi $ the B mode resulting from the lensing of primordial E mode is:
 \be
 B^{\mathrm{lens}}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') E(\ublu) \kappa(\bl - \bl')
-\label{eqn:}
+\label{eqn:blens}
 \ee
-where
+where different modes contributes with a different weight given by:
 \be
@@
 
-Now the full B-mode power spectrum measured on the sky is
+The full B-mode power spectrum measured on the sky is given by a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and instrumental noise $N_l^{BB}$:
 \be
@@
 \ee
+Given the current constrains, the lensing component is a significant source of B-modes that, for GW searches purposes correspond to a white noise source of roughly $5 \mu K$-arcmin.
+This means that it is not only bigger than the biggest allowed GW contribution at scale bigger than several degrees but it is comparable with the decreasing level of instrumental noise.
+For this reason it is very important to characterize it and eventually remove it form the data.
+This can be done for example using \refeq{blens} (or its real-space analogue) given a measurement of the E-mode field and the lensing potential $\phi$.
+While E is measured directly, we ca estimate $\phi$ using "tracers" of the dark matter distribution that creates the potential.
 
-If we have an LSS measurements I($\hat n$) that traces the lensing potential responsible for the lensing of the CMB we can build a template of the lensing B mode on the sky with a weighted convolution:
+If we have a large scale structure field I($\nver$) that traces the lensing potential responsible for the lensing of the CMB the template of the lensing B mode on the sky is a weighted convolution:
 \be
@@
 \ee
-where $f(\bl,\bl')$ is a weight that must be determined.
+where $f(\bl,\bl')$ can be determined minimizing the difference with the true $B^{\mathrm{lens}}(\bl)$ defined in \refeq{blens}.
 
-The residual lensing B mode will be
+The residual lensing B mode due to an imperfect knowledge of the true E-mode and $\phi$ will be
 \beqn
@@
 \eeqn
-and its power spectru
-m\beqn
-C_{{l}}^{BB,\mathrm{res}} &=&   \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl')
- [ C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|} \nonumber \\ && -(f(\bl,\bl')+f^*(\bl,\bl')) C^{EE}_{l'} C^{\kappa I}_{|\bl-\bl'|} \\ \nonumber && + f^*(\bl,\bl') f(\bl,\bl') (C^{EE}_{{l'}}+ N^{EE}_{{l'}}) C^{II}_{|\bl-\bl'|} ]
- \eeqn
 
-We can now easily choose $f(\bl,\bl')$ so that the residual lensing B mode power is minimized. We find:
+The weights $f(\bl,\bl')$ so that the residual lensing B mode power is minimized are:
 \be
@@
 
-We finally have that the residual power is:
+With this choice of $f(\bl,\bl')$ we finally have that the residual power is:
 \beqn
@@
 
-Given a 2D projected field $\delta^{i}$ this can be seen as a 3D field project along the line of site as:
+A 2D projected field $\delta^{i}$ that traces dark matter overdensities can be seen as a 3D field projected along the line of site as:
 
 \begin{equation}
-\delta^{i}(\theta) = \int_0^{\infty} dz\, W^{\kappa}(z)\delta^{i}(\chi(z)\nver,z).
+\delta^{i}(\nver) = \int_0^{\infty} dz\, W^{i}(z)\delta(\chi(z)\nver,z).
 \label{eqn:wkappa}
 \end{equation}
-
+where $\delta(\chi(z)\nver,z)$ correspond to the dark matter overdesity field at a comoving distance $\chi(z)$ at redshift $z$ in the angular direction $\nver$.
 Using the Limber approximation \cite{limber53} we can compute the power spectra of 2 fields $i,j$ as:
@@
 where $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
+
 The CMB lensing potential is \textit{the} field that we need to reconstruct in order to reverse the effect of large scale structure and delens the CMB.
-This lensing potential can be reconstructed from the CMB.
+However the lensing potential can also be reconstructed from the CMB itself.
 In that case it can be see as a noisy tracer of the true field.
 While the CMB internal reconstruction fiels as the same kernel as the true one $W^{\kappa}(z)$.
-Given certain instrumental noise level and beam, results in a lensing map with reconstruction noise $N_l^{\kappa \kappa}$, so that its power spectrum is
+Given the instrumental noise level and the beam, we can compute the reconstruction noise $N_l^{\kappa \kappa}$, so that its power spectrum is
 \beq
@@
 \eeq
+In this work the level of noise is computed assuming an iterative quadratic estimator is used in the reconstruction.
 
@@
 \includegraphics[scale=0.9]{../images/cmb_internal.pdf}
-\caption{Correlation factor between CMB reconstructed potential and the real lensing potential for different CMB experiments}
+\caption{\textbf{CMB alone:} Correlation factor between CMB internal reconstructed potential and the true lensing potential for different CMB experiments and corresponding noise reconstructions. }
 \end{center}
@@
 Here $\frac{dN}{dz}$ is the number of galaxies as a function of redshift observed by the survey while $b(z)$ is the galaxy bias that connect at different redshift the amplitude of galaxies overdensity to the one of the dark matter.
-Different galaxy surveys in this work are fully characterized by their b(z) and $\frac{dN}{dz}$.
+When computing the auto-spectrum of a galaxies density the shot noise term needs to be taken into account.
+This is done by adding a constant term to the power spectrum equal to the inverse of the number of galaxies per steradians.
+
+Different galaxy surveys in this work are fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxies density.
+We test the delensing efficiency modelling current survey like WISE or D.E.S as well as future galaxy probes like D.E.S.I  LSST together with 21 cm measurement like SKA.
+
+WISE redsfhit distribution is taken from 
+DES is modeled using
+For DESI we used the ..
+
+LSST is model following
+
+We model SKA following
 
@@
 
-We test the delensing efficiency modelling  current survey after D.E.S and using as future probe galaxy clustering measured with D.E.S.I and LSST together with 21 cm measurement like SKA.
-
-For DESI we used the ..
-
-LSST is model following
-
-We model SKA following
+%=================
+% FORECAST
+%=================
 
@@
 This will probably be needed to confirm a possible detection of gravitation waves if this rely heavily on delensing.
-\section*{Acknowledgment}
+
+%=================================================================
+\section{Acknowledgments}
+We thank 
+This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.
+%%%=================================================================
+
 \bibliographystyle{plain}

### 2017-05-11 Notes/DES_rec_phi_paper.tex
text

@@
 The measurement of the CMB polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe.
-A particular component, the so called B-modes is indeed a direct signature of the presence of gravitational waves in the early universe.
+A particular component, the so-called B-modes is indeed a direct signature of the presence of gravitational waves in the early universe.
 However improving the level of noise is not enough.
-This is even more true if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
-Removing the lensing component from the measurement of CMB B-modes will be important to constraint the amplitude of the primordial gravitational wave contribution to the signal.
+This is even truer if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
+Removing the lensing component from the measurement of CMB B-modes will be important to constrain the amplitude of the primordial gravitational wave contribution to the signal.
 Here we discuss the role of current and future large scale structure surveys in improving the reconstruction of the lensing potential that lenses the CMB photons.
@@
 \ee
-Given the current constrains, the lensing component is a significant source of B-modes that, for GW searches purposes correspond to a white noise source of roughly $5 \mu K$-arcmin.
-This means that it is not only bigger than the biggest allowed GW contribution at scale bigger than several degrees but it is comparable with the decreasing level of instrumental noise.
-For this reason it is very important to characterize it and eventually remove it form the data.
-This can be done for example using \refeq{blens} (or its real-space analogue) given a measurement of the E-mode field and the lensing potential $\phi$.
+Given the current constraints, the lensing component is a significant source of B-modes that, for GW searches purposes correspond to a white noise source of roughly $5 \mu K$-arcmin.
+This means that it is not only bigger than the biggest allowed GW contribution at scales bigger than several degrees but it is comparable with the decreasing level of instrumental noise.
+For this reason, it is very important to characterize it and eventually remove it from the data.
+This can be done for example using \refeq{blens} (or its real-space analog) given a measurement of the E-mode field and the lensing potential $\phi$.
 While E is measured directly, we ca estimate $\phi$ using "tracers" of the dark matter distribution that creates the potential.
@@
 However given the significant low noise of this measurement they can still help the delensing of the CMB.
-Furthermore the low-redshift tail of the lensing kernel is also the one affected by large scale structures non-linearities.
+Furthermore, the low-redshift tail of the lensing kernel is also the one affected by large scale structures non-linearities.
 They can indeed be extremely helpful if future CMB surveys will try to remove that contribution instead of attempting to model it.
-On the other hand CIB and 21 cm surveys are target to high redshift structures and independently from the model assumed they show a fairly good overlap with the CMB lensing kernel.
+On the other end, CIB and 21 cm surveys are probing to high redshift structures and independently from the model assumed they show a fairly good overlap with the CMB lensing kernel.
 
@@
 This is already possible both on full sky using Planck data or in smaller sky patches combining Herschel or Planck CIB data with SPT and BK data.
-Furthermore we test how much current relative low redshift galaxies surveys like D.E.S can improve delensing efficiency on smaller patch of the sky like the Bicep-Keck SPT-Pol footprint.
+Furthermore, we test how much current relative low redshift galaxies surveys like D.E.S can improve delensing efficiency on a smaller patch of the sky like the Bicep-Keck SPT-Pol footprint.
 
-We model the current generation pf CMB experiments after 3 experiments: a full sky Planck experiment together with a deep low noise experiemtn and a high resolution one on a smaller patch of the sky.
+We model the current generation of CMB experiments after 3 experiments: a full sky Planck experiment together with a deep low noise experiment and a high resolution one on a smaller patch of the sky.
 
@@
 Delensing or more in general the ability to separate the lensing component of the B-mode from a possible primordial inflationary signal is crucial to fully exploit the capabilities of future experiments.
-In this paper we studied the possible impact of large scale structure surveys in this important endeavor.
-For current experiment CIB data had already proven to be very efficient in delensing.
-If no High res experiment is avaiable to reconstruct the lensing potential phi even a low-z experiment like D.E.S will improve th delensign efficiency by $~10\%$ if combined with Planck lensing and a CIB tracers.
+In this paper, we studied the possible impact of large scale structure surveys in this important endeavor.
+For current experiment, CIB data had already proven to be very efficient in delensing.
+If no High res experiment is available to reconstruct the lensing potential phi even a low-z experiment like D.E.S will improve the delensign efficiency by $~10\%$ if combined with Planck lensing and a CIB tracers.
 As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential.
@@
 This has already been applied to data and studied.
-However the fairly good efficiency of galaxies tracers might come in end to cross-check these internal biases.
-This will probably be needed to confirm a possible detection of gravitation waves if this rely heavily on delensing.
-
+However, the fairly good efficiency of galaxies tracers might come in end to cross-check these internal biases.
+This will probably be needed to confirm a possible detection of gravitation waves if this relies heavily on delensing.
 %=================================================================

### 2017-05-11 Notes/DES_rec_phi_paper.tex
text

@@
 \newcommand{\mnras}{MNRAS}
-%\newcommand{\prd}{Phys. Rev. D}
 \newcommand{\physrev}{Phys. Rev.}
 \newcommand{\physrevlett}{Phys. Rev. Lett.}
+\newcommand*\aj{AJ}
+%\newcommand*\prd{Phys.~Rev.~D}
 
@@
 Different galaxy surveys in this work are fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxies density.
-We test the delensing efficiency modelling current survey like WISE or D.E.S as well as future galaxy probes like D.E.S.I  LSST together with 21 cm measurement like SKA.
+We test the delensing efficiency modeling current survey like WISE or DES as well as future galaxy probes like DESI and LSST together with 21 cm measurement like SKA.
 
-WISE redsfhit distribution is taken from 
-DES is modeled using
-For DESI we used the ..
+WISE redshift distribution is taken from \cite{yan:2013} (see Fig. 4 therein). 
+To compute the noise term we assume that the available scale after masking WISE should be around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015}.
 
-LSST is model following
+DES is modeled after the DES Science Verification data public release.
 
-We model SKA following
+For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report.
+
+LSST is modeled following \cite{schaan:2016} as $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
+Furthermore we assume a density of 26 galaxies per arcmin squared.
+
+We model both the SKA redshift distribution and bias survey following \cite{namikawa:2016a}.
 
@@
 This model is rescaled to agree with the results of
-~\citet{Holder:2013hqu} and \citet{planck2013XVIII} by choosing the corresponding linear bias parameter.
+~\cite{Holder:2013hqu} and \cite{planck2013XVIII} by choosing the corresponding linear bias parameter.
 Other multi-frequency CIB models are available \citep[e.g.,][]{Bethermin:2013nza}; however, given the level of noise, we are relatively insensitive to this choice.
@@
 \bibliographystyle{plain}
-\bibliography{DES_rec_phi_paper,/home/manzotti/Astrophysics/sptpol_papers/BIBTEX/spt}
+%\bibliography{DES_rec_phi_paper,/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
 
+\bibliography{DES_rec_phi_paper}
 \end{document}

### 2017-05-22 Code/compare_kernel.py
-mmove code to CAMB

@@ -76,6 +76,16 @@ norm = dndzfun.integral(dndz[0, 0], dndz[-2, 0])
 dndzfun_desi = InterpolatedUnivariateSpline(dndz[:, 0], dndz[:, 1] / norm)
 
 
+wise_dn_dz = np.loadtxt('/home/manzotti/galaxies_delensing/wise_dn_dz.txt')
+dndzwise = InterpolatedUnivariateSpline(wise_dn_dz[:, 0], wise_dn_dz[:, 1], k=3, ext='zeros')
+norm = dndzwise.integral(0, 2)
+dndzwise = InterpolatedUnivariateSpline(
+    wise_dn_dz[:, 0], wise_dn_dz[:, 1] / norm, ext='zeros')
+# Biased was measured equal to 1 in Feerraro et al. WISE ISW measureament
+# by cross correlating with planck lensing
+wise = gals_kernel.kern(wise_dn_dz[:, 0], dndzwise, hspline, omega_m, h0, b=1.)
+
+
 nu = 353e9
 zs = 2.
 b = 1.
@@ -88,7 +98,7 @@ desi = gals_kernel.kern(dndz[:, 0], dndzfun_desi, hspline, omega_m, h0)
 
 des_weak = kappa_gals_kernel.kern(dndz_des[:, 0], dndzfun_des, chispline, hspline, omega_m, h0)
 
-l = 100
+l = 50
 z_kappa = np.linspace(0, 13, 500)
 w_kappa = np.zeros_like(z_kappa)
 for i, z in enumerate(z_kappa):
@@ -138,11 +148,13 @@ for i, z in enumerate(z_ska):
 
 z_kappa_gal = np.linspace(0, des_weak.zmax, 500)
 w_kappa_gal = np.zeros_like(z_kappa_gal)
+w_structure = np.zeros_like(z_kappa_gal)
+
 for i, z in enumerate(z_kappa_gal):
 
     x = chispline(z)
-    # w_kappa_gal[i] = 1. / x * des_weak.w_lxz(l, x, z) * np.sqrt(rbs.ev((l + 0.5) / x, z))
-    w_kappa_gal[i] = des_weak.w_lxz(l, x, z)
+    w_kappa_gal[i] = 1. / x * des_weak.w_lxz(l, x, z) * np.sqrt(rbs.ev((l + 0.5) / x, z))
[…]

### 2017-05-22 Code/multiple_survey_delens.py

@@ -26,176 +26,158 @@ def nl(noise_uK_arcmin, fwhm_arcmin, lmax):
     return (noise_uK_arcmin * np.pi / 180. / 60.)**2 / bl(fwhm_arcmin, lmax)**2
 
 
-cosmosis_dir = '/home/manzotti/cosmosis/'
-inifile = '/home/manzotti/cosmosis/modules/limber/galaxies_delens.ini'
-n_slices = 2
+def main(labels,cmb):
+
+    cosmosis_dir = '/home/manzotti/cosmosis/'
+    inifile = '/home/manzotti/cosmosis/modules/limber/galaxies_delens.ini'
+    Config_ini = ConfigParser.ConfigParser()
+    values = ConfigParser.ConfigParser()
+    Config_ini.read(inifile)
+    values_file = Config_ini.get('pipeline', 'values')
+    output_dir = Config_ini.get('test', 'save_dir')
+    values.read(cosmosis_dir + values_file)
+    lbins = np.loadtxt(output_dir + '/limber_spectra/ells__delens.txt')
+    ckk = np.loadtxt(output_dir + '/limber_spectra/cl_kk_delens.txt')
+    # initialize
+    rho_comb = np.zeros((np.size(lbins)))
+
+    surveys = ['wise', 'k', 'euclid', 'des_weak', 'lsst', 'ska10', 'ska01',
+               'ska5', 'ska1', 'cib', 'desi', 'des']
+    cl_cross_k = {}
+    cl_auto = {}
+    rho = {}
+    for label in surveys:
+        cl_cross_k[label] = np.loadtxt(output_dir + '/limber_spectra/cl_k' + label + '_delens.txt')
+        cl_auto[label] = np.loadtxt(output_dir +
+                                    '/limber_spectra/cl_' + label + label + '_delens.txt')
+        if label == 'wise':
+            cl_cross_k[label][np.where(lbins < 100)] = 0.
+            #  from Simone Our conservative masking leaves f sky = 0.47 and about
+            # 50 million galaxies.
+            steradians_on_sphere = 4 * np.pi
+            fsky = 0.447
+            n_gal = 50e6
+            gal_per_ster = n_gal / (steradians_on_sphere * fsky)
+            cl_auto[label] += 1 / gal_per_ster * np.ones_like(cl_auto[label])
+
[…]

### 2017-05-22 Code/rho_to_Bres.py

@@ -25,7 +25,7 @@ def nl(noise_uK_arcmin, fwhm_arcmin, lmax):
     return (noise_uK_arcmin * np.pi / 180. / 60.)**2 / bl(fwhm_arcmin, lmax)**2
 
 
-def compute_res_parallel(rho_filename):
+def compute_res_parallel(rho_filename, output_dir, clee_fun, clpp_fun, nle_fun):
 
     print('start integration')
 
@@ -46,23 +46,12 @@ def compute_res_parallel(rho_filename):
             return (ell / (2. * np.pi)**2 * (L * ell * np.cos(theta) - ell**2)**2 * clpp_fun(np.sqrt(L**2 + ell**2 - 2. * ell * L * np.cos(theta))) * clee * (np.sin(2. * theta))**2) * (1. - (clee / (clee + nle_fun(ell))) * rho_fun(ell) ** 2)
 
     clbb_res_ell = [integrate.dblquad(
-        integrand, 8, 3000, lambda x: 0, lambda x: 2. * np.pi, args=(L,), epsabs=1.49e-08, epsrel=1.49e-07)[0] for L in np.arange(4, 2500, 10)]
+        integrand, 8, 1500, lambda x: 0, lambda x: 2. * np.pi, args=(L,), epsabs=1.49e-08, epsrel=1.49e-07)[0] for L in np.arange(4, 2500, 10)]
 
     np.savetxt(rho_filename.split('.txt')[0] + 'Cbb_res.txt', clbb_res_ell)
-    np.savetxt(datadir + 'limber_spectra/cbb_res_ls.txt', np.arange(4, 2500, 10))
+    np.savetxt(output_dir + 'limber_spectra/cbb_res_ls.txt', np.arange(4, 2500, 10))
 
     return clbb_res_ell
-
-
-
-
-
-
-
-
-
-
-
 
 
 def compute_res(rho_filename, noise_pol=2., fwhm_beam=30.):
@@ -131,8 +120,8 @@ def load_res(labels):
         res_list.append(np.loadtxt(datadir + 'limber_spectra/cbb_res_' + label + 'test3.txt'))
     return res_list
 
-if __name__ == "__main__":
 
+def main(rho_names, nle):
     cosmosis_dir = '/home/manzotti/cosmosis/'
     inifile = '/home/manzotti/cosmosis/modules/limber/galaxies_delens.ini'
 
@@ -174,69 +163,10 @@ if __name__ == "__main__":
[…]

### 2017-05-22 Notes/DES_rec_phi_paper.tex
-mtext

@@
-\documentclass[reprint,superscriptaddress,amsmath,amssymb,floatfix,aps,prd]{revtex4-1}
+\documentclass[reprint,superscriptaddress,amsmath,amssymb,floatfix,aps,prd,nofootinbib]{revtex4-1}
 \usepackage{bm}
@@
 \newcommand{\apjl}{ApJL}
+\def\araa{\ref@jnl{ARA\&A}}
 \newcommand{\apjs}{ApJS}
@@
 \email{manzotti.alessandro@gmail.com}
-\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, 5640 South Ellis Avenue, Chicago, IL 60637, USA}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago}
 \affiliation{Department of Astronomy and Astrophysics, University of Chicago, 5640 South Ellis Avenue, Chicago, IL 60637, USA}
@@
 \begin{abstract}
-\begin{itemize}
-\item Worry : they might ask for internal bias\
-item Possible TODO : temperature $N_{eff}$
+%\begin{itemize}
+%\item Worry : they might ask for internal bias\
+%item Possible TODO : temperature $N_{eff}$
+%
+%\end{itemize}
+%
 
-\end{itemize}
-
-The measurement of the CMB polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe.
-A particular component, the so-called B-modes is indeed a direct signature of the presence of gravitational waves in the early universe.
-However improving the level of noise is not enough.
+The measurement of the cosmic microwave background (CMB) polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe. 
+A particular component, the so-called B-modes is indeed a direct signature of the presence of gravitational waves in the early universe. However improving the level of noise is not enough. 
 This is even truer if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
 Removing the lensing component from the measurement of CMB B-modes will be important to constrain the amplitude of the primordial gravitational wave contribution to the signal.
-Here we discuss the role of current and future large scale structure surveys in improving the reconstruction of the lensing potential that lenses the CMB photons.
+Here we discuss the role of current and future large scale structure surveys in improving the reconstruction of the lensing potential that lenses the CMB photons and how this reflects into an improved delensing efficiency.
+We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$)
 We find that..
@@
 \label{sec:intro}
-Inflation Physics with CMB
+In the standard cosmological paradigm, the universe underwent a period of near-exponential expansion in its early phase;
+this period is called ``cosmic inflation.''
 
-Current constraints
+Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski15}.
+These primordial gravitational waves in the early universe imprint a unique signature on the polarized anisotropies of the CMB.
+CMB polarization is usually decomposed fields into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
+In the standard scenario, the PGW background is the only source of B-mode polarization at the epoch of recombination.
 
-Now limiting by delensing
+The amplitude of this primordial B-mode component is parametrized by the ratio of the amplitudes of the primordial tensor and scalar perturbations, $r$.
+The shape of this component, i.e.
+the amplitude as a function of the angular scales is encoded in the tensor tilt $n_{t}$.
+Most of the inflationary models directly relate the value of r to the energy scale of inflation.
+CMB B modes provide the cleanest known observational window into the primordial gravitational waves background.
+Measuring or constraint the value of $r$ and $n_t$ will put tight constraint on the allowed models of inflation.
 
-Done what are the prospects/
+Indeed improving the CMB B-modes measurement is a major objective of current and future CMB experiments.
+Even if inflationary B-modes have not been detected yet, a possibly natural value $r \gtrsim 10^{-3}$, which should be reachable given the near future level of noise expected in CMB experiments \citep{kamionkowski15}.
 
-This paper is organized as follow:
+However reducing the level of noise will not be enough to reach this goal.
+Unfortunately the observed B modes are not solely due primordial gravitational waves but also by late time effects. In this work we will focus on gravitational interactions with LSS.
+Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
+This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,polarbear2014c,vanengelen14b,planck2015XV} and directly from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
 
-\section{Gravitational lensing B-mode}
+This component acts as a source of confusion  for searches of the PGW background.
+Indeed, current experiment are already limited by the contamination from lensing B modes since the instrument noise is below the lensing B mode $rms$ fluctuations \citep{bk14}.
+The optimal way to approach this source of noise is to build the specific realization of lensing B modes on the sky can be and used it to clean the data in a process called ``delensing''.
+
+Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
+Furthermore recently it has been performed on CMB temperature data using the cosmic infrared background in \cite{larsen:2016}, on CMB temperature and polarization data using internal CMB reconstruction \cite{ carron17}.
+Finally the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in  where $~28\%$ of the lensing power has been removed.
+\vspace{0.5cm}
+
+In this paper we
+
+\vspace{0.5cm}
+
+This paper is organized as follow: in \refsec{th} we describe the lensing B-mode components and the residual power after delensing with tracers of the lensing potential.
+We describe the LSS tracers used in this analysis in \refsec{model}.
+\refsec{for} is the main results of this work: the improvement to inflationary parameters due to delensing with CMB and LSS.
+We conclude in \refsec{concl}.
+
+\section{Gravitational lensing B-mode and delensing}
 \label{sec:th}
@@
 \ee
-where $\nabla\phi$ is the deflection angle directly related to the lensing potential potential integrated along the line of sight $\phi$.
-The CMB polarization is usually decompose into odd-parity Fourier modes E and B. As shown in \cite{} this for symmetry reason the B-modes configuration is mainly sourced by tensor perturbations.
-However this promising sign of primordial gravitational waves is partially obscured by a secondary mechanism that can produce B-modes in the late universe.
-Indeed gravitational interaction with large scale structure generates CMB B-modes by distorting primordial E-modes.
+where the deflection angle is the gradient of lensing potential potential integrated along the line of sight $\nabla\phi$.
+The CMB polarization is usually decomposed into odd-parity Fourier modes E and B. 
+As shown in \cite{} this for symmetry reason tensor perturbations are the main source of the B-modes configuration.
+For this reason, B-modes are a promising signature of early universe tensor perturbations.
+However, this promising sign of primordial gravitational waves is partially obscured by a secondary mechanism that can produce B-modes also in the late universe.
+Indeed the gravitational interaction with large scale structure generates CMB B-modes by distorting primordial E-modes.
 
@@
 \be
-W(\bl,\bl') = \frac{2 \bl' \cdot (\bl-\bl')}{|\bl-\bl'|^2} \sin(2\varphi_{\bl,\bl'}),
+W(\bl,\bl') = \frac{2 \bl' \cdot (\bl-\bl')}{|\bl-\bl'|^2} \sin(2\varphi_{\bl,\bl'}).
 \ee
@@
 \be
-\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l
+\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l .
 \ee
@@
 
-The full B-mode power spectrum measured on the sky is given by a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and instrumental noise $N_l^{BB}$:
+The full B-mode power spectrum measured on the sky is given by a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_l^{BB}$:
 \be
@@
 This means that it is not only bigger than the biggest allowed GW contribution at scales bigger than several degrees but it is comparable with the decreasing level of instrumental noise.
-For this reason, it is very important to characterize it and eventually remove it from the data.
-This can be done for example using \refeq{blens} (or its real-space analog) given a measurement of the E-mode field and the lensing potential $\phi$.
+For this reason, it is critical to characterize and eventually remove it from the data.
+To do so, we need to build a template \refeq{blens} (or its real-space analog) of the actual lensing modes given a measurement of the E-mode field and the lensing potential $\phi$.
 While E is measured directly, we ca estimate $\phi$ using "tracers" of the dark matter distribution that creates the potential.
 
-If we have a large scale structure field I($\nver$) that traces the lensing potential responsible for the lensing of the CMB the template of the lensing B mode on the sky is a weighted convolution:
+We will now show how much the delensing efficiency is related to the fidelity of the lensing tracers.
+If we have a large scale structure field I($\nver$) that traces the lensing potential responsible for the lensing of the CMB we can build a template of the lensing B mode on the sky by a weighted convolution:
 \be
@@
 \ee
-where $f(\bl,\bl')$ can be determined minimizing the difference with the true $B^{\mathrm{lens}}(\bl)$ defined in \refeq{blens}.
+where $f(\bl,\bl')$ can be determined byminimizing the difference with the true $B^{\mathrm{lens}}(\bl)$ defined in \refeq{blens}.
 
@@
 \ee
-Notice that the first term consists in the usual inverse variance filter applied to the measured E-mode and the second minimize the difference between the reconstructed $\phi$ and the CMB lensing potential.
+Notice that the first term consists in an inverse variance filter applied to the measured E-mode and the second minimize the difference between the reconstructed $\phi$ and the CMB lensing potential.
 
@@
 \beqn
+\label{eqn:Bres}
 C_{{l}}^{BB,\mathrm{res}} &=&   \int \frac{d^2 \bl'}{(2 \pi)^2}  W^2 (\bl,\bl')
@@
 Let's now assume that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
-It can be shown that the optimal way to combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ is:
+We can optimally combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ with:
 \beqn
@@
 where C is the covariance matrix of the LSS tracers.
-The ``effective'' correlation of these combined tracers with gravitational lensing is:
+The residual B-mode power can we computed using \refeq{Bres} using an ``effective'' correlation of these combined tracers with gravitational lensing is:
 \be
@@
 
-Note that gain we have in adding a new tracer is not only proportional to its correlation with the CMB lensing but it also depends on how much it is correlated with the already used set of tracers.
-\reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in the following sections.
-Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST DES and DESI curves.
+Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing but it also depends on how much it is correlated with the already used set of tracers.
+\reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
+Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST,DES and DESI curves.
 However given the significant low noise of this measurement they can still help the delensing of the CMB.
@@
 \includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{Comparison of the different kernels used in this analysis.
+\caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
 The bigger the overlap wit the CMB lensing kernel the better the reconstruction of the lensing potential will be.
@@
 
-\section{Tracers Model}
-\label{sec:data}
-In this section we briefly describe how the power spectra of \refsec{th} are computed and what it is assumed for the different tracers considered in this work.
+\section{Lensing potential tracers}
+\label{sec:model}
+In this section, we describe how the power spectra defined in \refsec{th} are computed and how we characterize different tracers considered in this work.
 
-A 2D projected field $\delta^{i}$ that traces dark matter overdensities can be seen as a 3D field projected along the line of site as:
+Large scale structure surveys usually probe the 3D dark matter overdensities as a 2D field projected along the line of sight:
 
@@
 \delta^{i}(\nver) = \int_0^{\infty} dz\, W^{i}(z)\delta(\chi(z)\nver,z).
-\label{eqn:wkappa}
+\label{eqn:wkernel}
 \end{equation}
-where $\delta(\chi(z)\nver,z)$ correspond to the dark matter overdesity field at a comoving distance $\chi(z)$ at redshift $z$ in the angular direction $\nver$.
-Using the Limber approximation \cite{limber53} we can compute the power spectra of 2 fields $i,j$ as:
+
+where $\delta(\chi(z)\nver,z)$ corresponds to the dark matter overdensity field at a comoving distance $\chi(z)$ and at a redshift $z$ in the angular direction $\nver$.
+Using the Limber approximation \cite{limber53} we can compute the power spectra of two large scale structure fields $i,j$ as:
 \begin{equation}
@@
 
-In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, $\chi(z)$ the comoving distance $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ at redshift $z$.
-Furthermore $W^{i}(z)$ is the kernel function of the field i.
-These theoretical quantities have been computed using CAMB and HALOFIT.
+In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, $\chi(z)$ is the comoving distance and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$.
+Furthermore $W^{i}(z)$ is the kernel function of the field $i$ defined in \refeq{wkernel}.
+We compute these quantities using CAMB and HALOFIT.
 
@@
 \label{sec:kappaCMB}
-CMB photons interact gravitationally with low-redshift large scale structure.
-The deflection angle is given by $\mathbf{d}(\nver) = \nabla\phi(\nver)$, where $\nabla$ is the the two-dimensional gradient on the sphere.
-Because the lensing potential is an integrated measure of the projected gravitational potential, taking the two-dimensional Laplacian of the lensing potential we can define the lensing convergence $\kappa(\nver) = -\case{1}{2}\nabla^2\phi(\nver)$, which depends on the projected matter overdensity $\delta$
-% \citep{bart01}:
-The lensing kernel $W^{\kappa}$ is
-%
-$\chi(z)$ is the comoving distance to redshift $z$, $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$
+The first field we compute is the CMB lensing potential.
+The lensing kernel $W^{\kappa}$ is:
 \begin{equation}
@@
 %
-where $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
+where $\chi(z)$ is the comoving distance to redshift $z$, $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$, $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
 
-The CMB lensing potential is \textit{the} field that we need to reconstruct in order to reverse the effect of large scale structure and delens the CMB.
-However the lensing potential can also be reconstructed from the CMB itself.
-In that case it can be see as a noisy tracer of the true field.
-While the CMB internal reconstruction fiels as the same kernel as the true one $W^{\kappa}(z)$.
-Given the instrumental noise level and the beam, we can compute the reconstruction noise $N_l^{\kappa \kappa}$, so that its power spectrum is
+The CMB lensing potential is \textit{the} field that we need to reconstruct to reverse the effect of large scale structure and delens the CMB.
+However, the lensing potential can also be reconstructed from the CMB itself.
+In that case, we can treat it as a noisy tracer of the true field.
+Both the CMB lensing fields and its noisy reconstructed counterparts have the same kernel $W^{\kappa}(z)$.
+However, when computing the power spectrum of the latter, we need to add a noise component.
+Given the instrumental noise level and the beam, we can compute the reconstruction noise $N_l^{\kappa \kappa}$, and so
 \beq
@@
 \eeq
-In this work the level of noise is computed assuming an iterative quadratic estimator is used in the reconstruction.
+In this work, the level of noise is computed assuming an iterative quadratic estimator is used in the reconstruction
 
@@
 \begin{center}
-\includegraphics[scale=0.9]{../images/cmb_internal.pdf}
-\caption{\textbf{CMB alone:} Correlation factor between CMB internal reconstructed potential and the true lensing potential for different CMB experiments and corresponding noise reconstructions. }
+\includegraphics[scale=1.]{../images/cmb_internal.pdf}
+\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the true lensing potential for different CMB experiments and corresponding noise reconstructions.
+}
 \end{center}
@@
 
-The galaxy overdensity $g(\nver)$ in a given direction on the sky is also expressed as a line of sight integral of the matter overdensity:
+We can also write the galaxy overdensity field $g(\nver)$ in a given direction on the sky as a line of sight integral of the matter overdensity:
 \begin{equation}
@@
 \end{equation}
-Here $\frac{dN}{dz}$ is the number of galaxies as a function of redshift observed by the survey while $b(z)$ is the galaxy bias that connect at different redshift the amplitude of galaxies overdensity to the one of the dark matter.
-When computing the auto-spectrum of a galaxies density the shot noise term needs to be taken into account.
-This is done by adding a constant term to the power spectrum equal to the inverse of the number of galaxies per steradians.
+Here $\frac{dN}{dz}$ is the number of galaxies as a function of redshift observed by the survey while $b(z)$ is the galaxy bias that connects the amplitude of galaxies overdensity to the one of the dark matter at different redshift.
+When computing the auto-spectrum of a galaxies density, a shot noise term needs to be taken into account. 
+To do so, we add a constant term to the power spectrum equal to the inverse of the number of galaxies per steradians.
 
 Different galaxy surveys in this work are fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxies density.
-We test the delensing efficiency modeling current survey like WISE or DES as well as future galaxy probes like DESI and LSST together with 21 cm measurement like SKA.
+We test the delensing efficiency modeling both current survey like WISE or DES as well as future galaxy probes like DESI and LSST together with 21 cm measurement like SKA.
 
-WISE redshift distribution is taken from \cite{yan:2013} (see Fig. 4 therein). 
-To compute the noise term we assume that the available scale after masking WISE should be around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015}.
-
+We defined the WISE redshift distribution following \cite{yan:2013} (see Fig. 4 therein). 
+To compute the noise term, we assume that the available sky after masking WISE is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxies density is approximately uniform.
 DES is modeled after the DES Science Verification data public release.
-
 For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report.
-
 LSST is modeled following \cite{schaan:2016} as $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
 Furthermore we assume a density of 26 galaxies per arcmin squared.
+Finally, we model both the SKA redshift distribution and bias survey following \cite{namikawa:2016a}.
 
-We model both the SKA redshift distribution and bias survey following \cite{namikawa:2016a}.
+\subsection{Cosmic infrared Background (CIB)}
+\label{sec:cib}
+Following \cite{Addison:2011se}, we model the CIB power directly as $C_{\ell}^{{\rm CIB}\mbox{-}{\rm CIB}} = 3500 (l/3000)^{-1.25} {\rm Jy^2 / sr}$.
+This model provides an accurate fit to several experimental results.
+For the cross-spectra with the CMB lensing or other galaxies tracers $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall10}.
+It corresponds to the kernel:
+\begin{align}\label{eq:12}
+W^{\text{CIB}}(z) = b_c\ \frac{\chi^2(z)}{H(z)(1+z)^2}\ e^{-\frac{(z-z_c)^2}{2\sigma^2_z} } f_{\nu(1+z)},
+\end {align}
+for
+\begin{equation}
+f_{\nu} = 
+\begin{cases}
+\Big( e^{\frac{h\nu}{kT}} - 1 \Big)^{-1} \nu^{\beta+3} & (\nu \leq v^{\prime}) \\ \Big( e^{\frac{h\nu^{\prime}}{kT}} - 1 \Big)^{-1} \nu^{\prime \beta+3} \Big( \frac{\nu}{\nu^{\prime}} \Big)^{-\alpha} & (\nu > v^{\prime})
+\end{cases}
+\end{equation}
 
-\subsection{Cosmic infrared Background}
-\label{sec:cib}
-The definition of the kernel of the CIB is more complex and model dependent.
-Following \cite{Addison:2011se}, we model the CIB power directly as $C_{\ell}^{{\rm CIB}\mbox{-}{\rm CIB}} = 3500 (l/3000)^{-1.25} {\rm Jy^2 / sr}$.
-This model provides an accurate fit for the power of the \textit{Herschel} $500\,\mu$ map used in this work.
-For the cross-spectra with the CMB lensing or other galaxies tracers $C_{\ell}^{{\rm CIB}\mbox{-}j}$,we use the single-SED model of~\cite{hall10}.
-This places the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$.
-This model is rescaled to agree with the results of
-~\cite{Holder:2013hqu} and \cite{planck2013XVIII} by choosing the corresponding linear bias parameter.
-Other multi-frequency CIB models are available \citep[e.g.,][]{Bethermin:2013nza}; however, given the level of noise, we are relatively insensitive to this choice.
-With these assumptions, depending on angular scale, $45-65\%$ of the CIB is correlated with the CMB lensing potential, as shown in \reffig{corrnow}.
-% \subsection{Weak Lensing}
+We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$. Also $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
+
+%With these assumptions, depending on angular scale, $45-65\%$ of the CIB is correlated with the CMB lensing potential, as shown in \reffig{corrnow}.
+
 \cite{hall:2010,sherwin:2015,ade:2014,hanson:2013,szapudi:2001,0004-637X-567-1-2,planck-collaboration:2014,planck-collaboration:2014a,planck-collaboration:2011,plank-collaboration:2014,boulanger:1996,lewis:2006}
@@
 
-\section{Forecast}
+\section{Parameter constraints improvement after delensing}
 \label{sec:for}
-In this section we forecast the expected delensing efficiency and the relative importance of galaxies tracers for current and future experiments.
-We focus on 3 different scenarios: the current stage, the 3G one and the futuristic CMB Stage 4.
+In this section, we forecast the expected delensing efficiency and the relative importance of galaxies tracers for delensing current and future experiments.
+We will use a Fisher method and measured the delensign efficiency by the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
+We assume a CMB experimental scenario composed of a high-resolution CMB experiments which define the internal reconstruction efficiency, together with a low-noise, low-resolution experiment which B-modes will be delensed and used to constrain the inflationary B-modes.
+We focus on three different scenarios: the current stage, the 3G one and the futuristic CMB Stage 4.
 
 \subsection{Fisher method}
-
-We defined the B-modes noise spectrum:
+In this framework, the error on the cosmological parameters constrained by a B-mode spectrum measurement is given as usual by:
+\begin{equation}
+\label{eqn:fisher}
+\sigma(p)= \left[\sum_l^{\ell_{\rm{max}}} \left(\frac{\partial {C}_{\ell}^{BB,r,n_{t}}}{\partial p}\right)^2/\mathrm{Cov}^2( C_{\ell}^{BB} ) \right]^{-\frac{1}{2}}
+\end{equation}
+where we consider the parameters $p =\{r,n_{t}\}$. 
+We will assume a gaussian covariance:
+\beq
+\sigma( C_{l}^{BB} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_l^{BB,r} + C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
+\eeq
+Here the B-modes noise spectrum is given by:
 \beq
@@
 where $\theta_\mathrm{FWHM}$ is the full half width of the telescope  beam and $\Delta_P$ is the instrumental noise of the experiment.
-The gaussian covariance is then:
-\beq
-\sigma( C_{l}^{BB,\mathrm{full}} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
-\eeq
-With this we can simply quantify the constraints on the tensor to scalar ration r:
-\beqn
-&\sigma(r)&= \left[\sum_l \frac{\left(\frac{\partial {C}_l^{BB,r}}{\partial r}\right)^2}{\sigma^2( C_{l}^{BB,\mathrm{full}} )} \right]^{-\frac{1}{2}}  \\ &\approx&  \left[ \frac{\sum_l (2 {l}+1) \fsky \left(\frac{\partial {C}_l^{BB,r}}{\partial r}\right)^2}{2} \right]^{-\frac{1}{2}}\langle C_l^{BB,\mathrm{lens}} +N^{BB}_l \rangle_{l<100}\nonumber
-\eeqn
-where to go from the first to the second line we use the fact that most of the constraint comes from large scale modes at $\ell <100$, thus larger mode can be ignore here.
-We quantify the improvement due to delensing as the factor $\alpha$ defined as the ration of the error before and after delensing:
-\beq
-\alpha = \frac{ \langle C_l^{BB,\mathrm{lens}} +N^{BB}_l[\Delta_P] \rangle_{l<100}}{\langle C_l^{BB,\mathrm{res}}[\Delta_P]+N^{BB}_l[\Delta_P] \rangle_{l<100}}
-\eeq
+Even if the covariance structure has known non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012} this approximation is good enough to show the improvement due to delensing.
 
-\subsection{Current generation}
+It can be seen from \refeq{fisher}, that the parameter constraints are inversely proportional to the covariance of the measurement.
+Since lensing B-mode $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them trough delensing will improve parameter constraints.
+
+Note that we are fixing the other cosmological parameters.
+Uncertainties in those will however propagate to uncertainties in $\{r,n_{t}\}$.
+While the constraints will be slightly optimistic, this has no real impact on the estimate of the improvement due to delensing which is the main focus of this work.
+We define the improvement due to delensing as the ration of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$
+%===================
+\subsection{Delensing with current CMB and LSS}
 
@@
 
-Very recently delensing has been proven possible on real data using both the internal CMB lensing potential reconstruction as well as external tracers.
-Here we discuss the improvement due to combining this two probes.
-This is already possible both on full sky using Planck data or in smaller sky patches combining Herschel or Planck CIB data with SPT and BK data.
-Furthermore, we test how much current relative low redshift galaxies surveys like D.E.S can improve delensing efficiency on a smaller patch of the sky like the Bicep-Keck SPT-Pol footprint.
+Very recently delensing has been proven possible on real data using as tracers both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as the CIB \cite{larsen:2016a,manzotti:2017}.
+Here we discuss the improvement due to combining these and other available probes (see \cite{yu:2017} for publicly averrable multi-tracers data-products).
 
-We model the current generation of CMB experiments after 3 experiments: a full sky Planck experiment together with a deep low noise experiment and a high resolution one on a smaller patch of the sky.
+On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with a noise equal to \textcolor{red}{$3\mu$K-arcmin} and a beam of \textcolor{red}{30 arcmin}.
+The CMB lensing reconstruction will be performed by a higher resolution CMB experiment. We explore two possibilities.
+First, we test the efficiency with internal reconstruction performed by Planck \cite{2016A&A...594A..15P} using the CMB lensing noise curves publicly available
+\footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
+The second one uses a ground-based experiment with the final noise level of SPT-Pol which correspond to \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
 
-The correlation attainable using current experiments is shown in \reffig{corrnow}.
+Furthermore, we test how much not only CIB maps but also current low redshift galaxies surveys like DES and WISE can improve the delensing efficiency.
 
-At large angular scale the CMB internal reconstruction is clearly the best tracers for delensing.
-Because of the large level of noise in current experiment, its efficiency became comparable to the CIB one at few degrees ($\ell \simeq 400$).
+The correlation attainable using current experiments is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol and.
+In general, we observe that CMB lensing reconstruction correlate very well at low $\ell$  and it then falls rapidly because of the the raise of the reconstruction noise.
+For the noise levels of Planck, the internal reconstruction is at most $70\%$ correlated at very large scales end falls rapidly to $40\%$ at $\ell=200$.
+Its correlation is at the level of the one coming from LSS at almost all scales.
+On the other end, the CMB lensing reconstruction \cite{} from SPT-Pol will be more than 80\% correlated with the true field at large scales $\ell<300$ where the correlation become comparable with the current LSS surveys.
+
+Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by very large Galactic dust residuals.
+Optical surveys galaxies like DES are less affected by that systematics and can be used at larger scales.
+Furthermore, in \reffig{corrnow} DES contribution to delensing is higher than WISE because the redshift distributions peaks at slightly higher redshift thus overlapping better with lensing.
 
 \begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/actual_scenario_planck.pdf}
+\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
+\label{fig:corrnow-planck}
+\end{center}
+\end{figure}
 
+%At large angular scale, the CMB internal reconstruction is clearly the best tracers for delensing.
+Because of the large level of noise in current experiments, its efficiency became comparable to the CIB one at few degrees ($\ell \simeq 400$).
+
+\begin{figure}[htbp]
 \begin{center}
-\includegraphics[scale=0.9]{../images/actual_scenario.pdf}
+\includegraphics[scale=1.]{../images/actual_scenario.pdf}
 \caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
@@
 
+Using this correlation level we can, using \refeq{Bres} compute the residual B-mode power after delensing and test the consequent improvement on parameter constraints with \refeq{fisher}. 
+For current surveys we only consider $r$ since the broad beam of the deep BK-like survey does not allow enough leverage to constrain the shape of the B-mode spectra.
+We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
+The results are summarized in \reftab{current}
+
+\begin{table}
+\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to ration of the error before and after delensing.}
+  \vspace{0.2cm}
+  \begin{tabular}{ | c | c | c|}
+\hline
+Surveys & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+WISE & 1.10 & 1.04 \\ \hline
+DES & 1.19  & 1.07 \\ \hline
+CIB & 1.46  & 1.18\\ \hline
+LSS & 1.64  & 1.24\\ \hline
+CMB & 1.14  & 1.04 \\ \hline
+LSS+CMB & 1.71& 1.26  \\ \hline
+
+  \end{tabular}
+  \label{tab:current}
+\end{table}
+
+The results are in agreement with \cite{yu:2017,larsen:2016a}, with the CIB being the predominant source of delensing with an improvement of almost 50\%.
+Planck and DES are essential to reconstruct the largest scales of the lensing potential and as such the play an important role in delensing.
+The improvement due to WISE is quite marginal, but given the big sky coverage, it might be useful to delens large CMB polarization experiments.
+For reference we also test a scenario where GW are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.
+As expected the importance of delensing itself is now reduced given that the lensing component constitutes a smaller portion of the total B-modes variance.
+
+As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS. 
+This can be seen in \reftab{current-spt}. 
+In particular optical surveys will rapidly lose the role of filling in large-scale modes and they start, as the CIB, to supplement information about lensing at higher multipoles.
+
+ \begin{table}
+\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to ration of the error before and after delensing.}
+  \vspace{0.2cm}
+  \begin{tabular}{ | c | c | c|}
+\hline
+Surveys & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+WISE & 1.10 &  \\ \hline
+DES & 1.21  & \\ \hline
+CIB & 1.52  & \\ \hline
+LSS & 1.78  & \\ \hline
+CMB & 1.62  &  \\ \hline
+LSS+CMB & 2.35 &   \\ \hline
+
+  \end{tabular}
+  \label{tab:current-spt}
+\end{table}
+
+\subsection{CMB-S3 Era}
+The accuracy of CMB is rapidly improving and the next generation ground based telescopes have been already deployed and are currently taking data.
+For the CMB at the beginning of the next decade, we will assume a deep experiment with level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
+This is a reasonable level for a combination of the next generation of BK and SPT3G.
+This level of noise will also be assumed for the internal noise reconstruction.
+
+The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
+Not only will CMB experiments improved in the near future: possibly D.E.S.I will start taking data. For this reason we add DESI to the previously mentioned LSS tracers. 
+DESI will be slightly more efficient than a DES-like surveys because of the slightly higher redshift that it is able to probe.
+However the CIB will still play the major role among LSS.
+
+This generation of experiments will start exploring the small scales B-modes thus constraining, even if loosely, the value of $n_{t}$. 
+The improvement in cosmological parameters are summarized in \reftab{S3}
+
 \begin{figure}[htbp]
-
 \begin{center}
-\includegraphics[scale=0.9]{../images/actual_scenario_planck.pdf}
-\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
-\label{fig:corrnow}
+\includegraphics[scale=1.]{../images/S3_scenario.pdf}
+\caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
+\label{fig:corrS3}
 \end{center}
@@
 \begin{table}
-\caption{$\alpha$: Current generation improvements}
-  \begin{tabular}{ | c | c  | }
+\caption{$\alpha$: S3 improvements. Note that the absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is  $\sigma(r) =$  }
+  \begin{tabular}{ | c | c  |  c |}
 \hline
-Surveys & $\alpha$\\ \hline \hline
-des & 1.34  \\ \hline
-cib & 1.71  \\ \hline
+Surveys & $\alpha(r)$ & $\alpha(n_{t})$\\ \hline \hline
+DESI & 1.34 & \\ \hline
+LSS & 1.71  &\\ \hline
+CMB & 1.71  &\\ \hline
 \end{tabular}
-\end{table}
-
-\subsection{CMB-S3 Era}
-The accuracy of CMB is rapidly improving.
-For example the next generation of the SPT telescope, SPT3G has been deployed and is currently taking data.
-Furthermore other experiments have considerably increased the number of detectors (AdvACT etc.).
-We will assume a level of noise of 3 $\mu$K-arcmin the level predicted for SPT3G.
-The correlation attainable using current experiments is shown in \reffig{corrS3},
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=0.9]{../images/S3_scenario.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
-\label{fig:corrS3}
-
-\end{center}
-
-\end{figure}
-
-\begin{table}
-\caption{$\alpha$: Stage-3 improvements}
-  \begin{tabular}{ | c | c  | }
-\hline
-Surveys & $\alpha$\\ \hline \hline
-des & 1.34  \\ \hline
-cib & 1.71  \\ \hline
-  \end{tabular}
+\label{tab:S3}
 \end{table}
@@
 An ambitious program for a generation 4 ground CMB experiment is currently under planning. Moreover satellite experiments have been proposed.
-
 The correlation attainable using current experiments is shown in \reffig{corrs4},
-
-\begin{table}
-\caption{$\alpha$: Stage-4 improvements}
-  \begin{tabular}{ | c | c  | }
-\hline
-Surveys & $\alpha$\\ \hline \hline
-des & 1.34  \\ \hline
-cib & 1.71  \\ \hline
-  \end{tabular}
-\end{table}
+In this very futuristic but reasonable scenario we add to the avaialable large scale structures both 21 cm surveys like SKA and large scale structures like LSST and Euclid.
 
@@
 \begin{center}
-\includegraphics[scale=0.9]{../images/S4_scenario.pdf}
+\includegraphics[scale=1.]{../images/S4_scenario.pdf}
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
@@
 %=======================================
+
+\begin{table}
+\caption{$\alpha$: Gen 4 experiments}
+  \begin{tabular}{ | c | c  |  c |}
+\hline
+Surveys & $\alpha(r)$ & $\alpha(n_{t})$\\ \hline \hline
+Euclid & 1.53 & \\ \hline
+LSST & 1.65  &\\ \hline
+SKA & 4.35  &\\ \hline
+CMB S4 & 3.7  &\\ \hline
+LSS+CMB & 6.3  &\\ \hline
+
+  \end{tabular}
+  \label{tab:S4}
+\end{table}
 
@@
 
-\begin{table}
-\caption{$\alpha$: improvement on r constraint}
-  \begin{tabular}{ | c | c  | }
-\hline
-Surveys & $\alpha$\\ \hline \hline
-des & 1.34  \\ \hline
-cib & 1.71  \\ \hline
-cmb current & 1.79  \\ \hline
-gals current & 2.03  \\ \hline
-cmb S3 & 2.15  \\ \hline
-gals S3 & 2.16  \\ \hline
-gals S4 & 2.16  \\ \hline
-comb current & 2.66  \\ \hline
-comb S3 & 3.14  \\ \hline
-cmb S4 & 4.36  \\ \hline
-comb S4 & 5.27  \\ \hline
-  \end{tabular}
-\end{table}
+%\begin{table}
+%\caption{$\alpha$: improvement on r constraint}
+%  \begin{tabular}{ | c | c  | }
+%\hline
+%Surveys & $\alpha$\\ \hline \hline
+%des & 1.34  \\ \hline
+%cib & 1.71  \\ \hline
+%cmb current & 1.79  \\ \hline
+%gals current & 2.03  \\ \hline
+%cmb S3 & 2.15  \\ \hline
+%gals S3 & 2.16  \\ \hline
+%gals S4 & 2.16  \\ \hline
+%comb current & 2.66  \\ \hline
+%comb S3 & 3.14  \\ \hline
+%cmb S4 & 4.36  \\ \hline
+%comb S4 & 5.27  \\ \hline
+%  \end{tabular}
+%\end{table}
 
 \section{Conclusions}
-Delensing or more in general the ability to separate the lensing component of the B-mode from a possible primordial inflationary signal is crucial to fully exploit the capabilities of future experiments.
-In this paper, we studied the possible impact of large scale structure surveys in this important endeavor.
+\label{sec:concl}
+Delensing or more, in general, the ability to separate the lensing component of the B-mode from a possible primordial inflationary signal is important to fully exploit the capabilities of future experiments.
+In this paper, we studied the possible impact of large-scale structure surveys in this important endeavor.
 For current experiment, CIB data had already proven to be very efficient in delensing.
-If no High res experiment is available to reconstruct the lensing potential phi even a low-z experiment like D.E.S will improve the delensign efficiency by $~10\%$ if combined with Planck lensing and a CIB tracers.
+If no high resolution CMB experiment is available to reconstruct the lensing potential $\phi$ a low-z optical experiment like D.E.S will improve the delensign efficiency by $~10\%$ if combined with Planck lensing and a CIB tracers.
 As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential.
@@
 This will probably be needed to confirm a possible detection of gravitation waves if this relies heavily on delensing.
+
+Another possible application of LSS delensing is the removal of the low-z lensing of the CMB that possibly adds the complexity of modeling non linearities in the study of CMB lensing
+
 %=================================================================
 \section{Acknowledgments}
-We thank 
+We thank S. Dodelson for useful discussion.
 This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.
@@
 \bibliographystyle{plain}
-%\bibliography{DES_rec_phi_paper,/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
+\bibliography{DES_rec_phi_paper,/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
 
-\bibliography{DES_rec_phi_paper}
+%\bibliography{DES_rec_phi_paper}
 \end{document}

### 2017-05-29 Notes/DES_rec_phi_paper.tex
order

@@
 We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$)
-We find that..
+We find that current generation experiments...
 \end{abstract}
@@
 
-The amplitude of this primordial B-mode component is parametrized by the ratio of the amplitudes of the primordial tensor and scalar perturbations, $r$.
+The amplitude of this primordial B-mode component is parameterized by the ratio of the amplitudes of the primordial tensor and scalar perturbations, $r$.
 The shape of this component, i.e.
@@
 Most of the inflationary models directly relate the value of r to the energy scale of inflation.
-CMB B modes provide the cleanest known observational window into the primordial gravitational waves background.
-Measuring or constraint the value of $r$ and $n_t$ will put tight constraint on the allowed models of inflation.
+CMB B-modes provide the cleanest known observational window into the primordial gravitational waves background.
+Measuring or constraint the value of $r$ and $n_t$ will put a tight constraint on the allowed models of inflation.
 
@@
 However reducing the level of noise will not be enough to reach this goal.
-Unfortunately the observed B modes are not solely due primordial gravitational waves but also by late time effects. In this work we will focus on gravitational interactions with LSS.
+Unfortunately, the observed B modes are not solely due to primordial gravitational waves but also by late time effects. In this work, we will focus on gravitational interactions with LSS.
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
@@
 This component acts as a source of confusion for searches of the PGW background.
-Indeed, current experiment are already limited by the contamination from lensing B modes since the instrument noise is below the lensing B mode $rms$ fluctuations \citep{bk14}.
+Indeed, current experiments are already limited by the contamination from lensing B modes since the instrument noise is below the lensing B-mode $rms$ fluctuations \citep{bk14}.
 The optimal way to approach this source of noise is to build the specific realization of lensing B modes on the sky can be and used it to clean the data in a process called ``delensing''.
@@
 Furthermore recently it has been performed on CMB temperature data using the cosmic infrared background in \cite{larsen:2016}, on CMB temperature and polarization data using internal CMB reconstruction \cite{ carron17}.
-Finally the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in  where $~28\%$ of the lensing power has been removed.
-\vspace{0.5cm}
+Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in  where $~28\%$ of the lensing power has been removed.
 
-In this paper we
+In this paper, we study the possible role of future galaxies surveys as tracers of the lensing potential.
+These will translate into a better reconstruction of the B-modes in the measured patch and as a consequence improve the constraints on inflation trough delensing.
+We model several actual and future surveys, and after computing the residual B-modes, we forecast the statistical uncertainness on $r$ and $n_T$ for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
 
-\vspace{0.5cm}
-
-This paper is organized as follow: in \refsec{th} we describe the lensing B-mode components and the residual power after delensing with tracers of the lensing potential.
+This article is organized as follow: in \refsec{th} we describe the lensing B-mode components and the residual power after delensing with tracers of the lensing potential.
 We describe the LSS tracers used in this analysis in \refsec{model}.
-\refsec{for} is the main results of this work: the improvement to inflationary parameters due to delensing with CMB and LSS.
+\refsec{for} is the main results of this work: the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
 We conclude in \refsec{concl}.
@@
 The CMB polarization is usually decomposed into odd-parity Fourier modes E and B. 
-As shown in \cite{} this for symmetry reason tensor perturbations are the main source of the B-modes configuration.
+As shown in \cite{} this for symmetry reason tensor perturbations are the principal source of the B-modes configuration.
 For this reason, B-modes are a promising signature of early universe tensor perturbations.
@@
 Given the current constraints, the lensing component is a significant source of B-modes that, for GW searches purposes correspond to a white noise source of roughly $5 \mu K$-arcmin.
-This means that it is not only bigger than the biggest allowed GW contribution at scales bigger than several degrees but it is comparable with the decreasing level of instrumental noise.
+This means that it is not only bigger than the biggest allowed GW contribution at scales bigger than several degrees, but it is comparable with the decreasing level of instrumental noise.
 For this reason, it is critical to characterize and eventually remove it from the data.
@@
 
-The bigger $\rho_l$ is for a LSS field the more it is correlated with the lensing potential acting on the CMB photons.
-An higher correlation allows for a better reconstruction of the $\phi^{CMB}$ and, as a consequence, of $B^{lens}$.
+The bigger $\rho_l$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
+A higher correlation allows for a better reconstruction of the $\phi^{CMB}$ and, as a consequence, of $B^{lens}$.
 
@@
 
-Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing but it also depends on how much it is correlated with the already used set of tracers.
+Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing, but it also depends on how much it is correlated with the already used set of tracers.
 \reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
-Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST,DES and DESI curves.
+Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
 However given the significant low noise of this measurement they can still help the delensing of the CMB.
-Furthermore, the low-redshift tail of the lensing kernel is also the one affected by large scale structures non-linearities.
+Furthermore, the low-redshift tail of the lensing kernel is also the one affected by large-scale structures non-linearities.
 They can indeed be extremely helpful if future CMB surveys will try to remove that contribution instead of attempting to model it.
-On the other end, CIB and 21 cm surveys are probing to high redshift structures and independently from the model assumed they show a fairly good overlap with the CMB lensing kernel.
+On the other end, CIB and 21 cm surveys are probing to high redshift structures and independently from the model assumed they show a relatively good overlap with the CMB lensing kernel.
 
@@
 The CMB lensing potential is \textit{the} field that we need to reconstruct to reverse the effect of large scale structure and delens the CMB.
-However, the lensing potential can also be reconstructed from the CMB itself.
+However, the lensing potential can also be reconstructed using the CMB itself.
 In that case, we can treat it as a noisy tracer of the true field.
@@
 However, when computing the power spectrum of the latter, we need to add a noise component.
-Given the instrumental noise level and the beam, we can compute the reconstruction noise $N_l^{\kappa \kappa}$, and so
+Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_l^{\kappa \kappa}$, and so
 \beq
@@
 This model provides an accurate fit to several experimental results.
-For the cross-spectra with the CMB lensing or other galaxies tracers $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall10}.
+For the cross-spectra with the CMB lensing or other galaxies tracers, $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall10}.
 It corresponds to the kernel:
@@
 \label{sec:for}
-In this section, we forecast the expected delensing efficiency and the relative importance of galaxies tracers for delensing current and future experiments.
-We will use a Fisher method and measured the delensign efficiency by the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
-We assume a CMB experimental scenario composed of a high-resolution CMB experiments which define the internal reconstruction efficiency, together with a low-noise, low-resolution experiment which B-modes will be delensed and used to constrain the inflationary B-modes.
-We focus on three different scenarios: the current stage, the 3G one and the futuristic CMB Stage 4.
+In this section, we forecast the expected delensing efficiency and the relative importance of galaxies tracers for delensing in current and future experiments.
+We will use the Fisher information matrix to quantify the delensign efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
+We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction efficiency, together with a low-noise, low-resolution experiment which B-modes will be delensed and used to constrain the inflationary parameters.
+We focus on three different scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
 
 \subsection{Fisher method}
-In this framework, the error on the cosmological parameters constrained by a B-mode spectrum measurement is given as usual by:
+In this framework, the error on a cosmological parameter constrained by a B-mode spectrum measurement is given as usual by:
 \begin{equation}
@@
 where $\theta_\mathrm{FWHM}$ is the full half width of the telescope beam and $\Delta_P$ is the instrumental noise of the experiment.
-Even if the covariance structure has known non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012} this approximation is good enough to show the improvement due to delensing.
+Even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012} this approximation is good enough to show the improvement due to delensing.
 
-It can be seen from \refeq{fisher}, that the parameter constraints are inversely proportional to the covariance of the measurement.
-Since lensing B-mode $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them trough delensing will improve parameter constraints.
+The reason why removing the lensing contribution will improve parameters constraints can be seen from \refeq{fisher}.The parameter uncertainties are inversely proportional to the covariance of the measurement.
+Since lensing B-mode $C_{{l}}^{BB,\mathrm{lens}}$ is a substantial component of the covariance, removing part of them trough delensing will improve parameter constraints.
 
-Note that we are fixing the other cosmological parameters.
+Note that we are fixing the other cosmological parameters in \refeq{fisher}.
 Uncertainties in those will however propagate to uncertainties in $\{r,n_{t}\}$.
 While the constraints will be slightly optimistic, this has no real impact on the estimate of the improvement due to delensing which is the main focus of this work.
-We define the improvement due to delensing as the ration of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$
+We define that improvement as the ration of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$
 %===================
@@
 Very recently delensing has been proven possible on real data using as tracers both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as the CIB \cite{larsen:2016a,manzotti:2017}.
-Here we discuss the improvement due to combining these and other available probes (see \cite{yu:2017} for publicly averrable multi-tracers data-products).
+Here we discuss the improvement due to combining these and other available probes (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
 
@@
 The CMB lensing reconstruction will be performed by a higher resolution CMB experiment. We explore two possibilities.
-First, we test the efficiency with internal reconstruction performed by Planck \cite{2016A&A...594A..15P} using the CMB lensing noise curves publicly available
+First, we test the efficiency with an internal reconstruction performed by Planck \cite{2016A&A...594A..15P} using the CMB lensing noise curves publicly available
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
-The second one uses a ground-based experiment with the final noise level of SPT-Pol which correspond to \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
+Then we will assume a ground-based experiment with the final noise level of SPT-Pol which corresponds to \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
 
@@
 
-The correlation attainable using current experiments is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol and.
-In general, we observe that CMB lensing reconstruction correlate very well at low $\ell$  and it then falls rapidly because of the the raise of the reconstruction noise.
-For the noise levels of Planck, the internal reconstruction is at most $70\%$ correlated at very large scales end falls rapidly to $40\%$ at $\ell=200$.
-Its correlation is at the level of the one coming from LSS at almost all scales.
-On the other end, the CMB lensing reconstruction \cite{} from SPT-Pol will be more than 80\% correlated with the true field at large scales $\ell<300$ where the correlation become comparable with the current LSS surveys.
+The correlation attainable using current experiments is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
+In general, we observe that CMB lensing reconstruction correlate very well at low $\ell$  and it then falls rapidly because of the raise of the reconstruction noise.
+For the noise levels of Planck, the internal reconstruction is at most $70\%$ correlated at very large scales and falls rapidly to $40\%$ at $\ell=200$.
+Its correlation is comparable to the on of LSS at almost all scales.
+On the other end, the CMB lensing reconstruction \cite{} from SPT-Pol will be more than 80\% correlated with the true field at large scales $\ell<300$, where the correlation become comparable with the current LSS surveys.
 
-Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by very large Galactic dust residuals.
+Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals.
 Optical surveys galaxies like DES are less affected by that systematics and can be used at larger scales.
@@
 \includegraphics[scale=1.]{../images/actual_scenario_planck.pdf}
-\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
+\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the angular scale $\ell$. 
+The dashed curve corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$, showing which scales contribute to the $\ell<100$ B-mode power.}
 \label{fig:corrnow-planck}
@@
 Using this correlation level we can, using \refeq{Bres} compute the residual B-mode power after delensing and test the consequent improvement on parameter constraints with \refeq{fisher}. 
-For current surveys we only consider $r$ since the broad beam of the deep BK-like survey does not allow enough leverage to constrain the shape of the B-mode spectra.
+For current surveys we only consider $r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
 We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
@@
 
-The results are in agreement with \cite{yu:2017,larsen:2016a}, with the CIB being the predominant source of delensing with an improvement of almost 50\%.
-Planck and DES are essential to reconstruct the largest scales of the lensing potential and as such the play an important role in delensing.
+When the same tracers are used, the results are in agreement with \cite{yu:2017,larsen:2016a}, with the CIB being the predominant source of delensing with an improvement of almost 50\%.
+Planck and DES are essential to reconstruct the largest scales of the lensing potential and as such the play a major role in delensing.
 The improvement due to WISE is quite marginal, but given the big sky coverage, it might be useful to delens large CMB polarization experiments.
@@
  \begin{table}
-\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to ration of the error before and after delensing.}
+\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to the ratio of the error before and after delensing.}
   \vspace{0.2cm}
@@
 \subsection{CMB-S3 Era}
-The accuracy of CMB is rapidly improving and the next generation ground based telescopes have been already deployed and are currently taking data.
-For the CMB at the beginning of the next decade, we will assume a deep experiment with level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
+CMB measurements are rapidly improving, and the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
+For the CMB at the beginning of the next decade, we will assume a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
 This is a reasonable level for a combination of the next generation of BK and SPT3G.
@@
 The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
-Not only will CMB experiments improved in the near future: possibly D.E.S.I will start taking data. For this reason we add DESI to the previously mentioned LSS tracers. 
-DESI will be slightly more efficient than a DES-like surveys because of the slightly higher redshift that it is able to probe.
-However the CIB will still play the major role among LSS.
+Not only will CMB experiments improved shortly: possibly D.E.S.I will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
+DESI will be slightly more efficient than DES-like survey because of the slightly higher redshift that it can probe.
+However, the CIB will still play the major role among LSS.
 
@@
 
-An ambitious program for a generation 4 ground CMB experiment is currently under planning. Moreover satellite experiments have been proposed.
+An ambitious program for a generation 4 ground CMB experiment is currently under planning. Moreover, satellite experiments have been proposed.
 The correlation attainable using current experiments is shown in \reffig{corrs4},
-In this very futuristic but reasonable scenario we add to the avaialable large scale structures both 21 cm surveys like SKA and large scale structures like LSST and Euclid.
+In this very futuristic but reasonable scenario, we add to the available large scale structures both 21 cm surveys like SKA and large scale structures like LSST and Euclid.
 
@@
 \subsection{Bias uncertainties degradation}
-The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of inflationary constraint of delensed spectra.
-In this section we quantify this effect.
-We will now marginalize over unknown galaxies parameters but we will use a full dataset of CMB and galaxies data.
-The idea is that as shown in the low level of noise in galaxies surveys might allow us to internally calibrate them.
+The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of the inflationary constraint of delensed spectra.
+In this section, we quantify this effect.
+We will now marginalize over unknown galaxies parameters, but we will use a full dataset of CMB and galaxies data.
+The idea is that as shown in the low level of noise in galaxies surveys might allow us to calibrate them internally.
 We will use a Fisher approach the Fisher matrix is:
@@
 \label{sec:concl}
-Delensing or more, in general, the ability to separate the lensing component of the B-mode from a possible primordial inflationary signal is important to fully exploit the capabilities of future experiments.
-In this paper, we studied the possible impact of large-scale structure surveys in this important endeavor.
+Delensing or more, in general, the ability to separate the lensing component of the B-mode from a possible primordial inflationary signal is important to exploit the capabilities of future experiments fully.
+In this paper, we studied the potential impact of large-scale structure surveys in this important endeavor.
 For current experiment, CIB data had already proven to be very efficient in delensing.
-If no high resolution CMB experiment is available to reconstruct the lensing potential $\phi$ a low-z optical experiment like D.E.S will improve the delensign efficiency by $~10\%$ if combined with Planck lensing and a CIB tracers.
+If no high-resolution CMB experiment is available to reconstruct the lensing potential $\phi$ a low-z optical experiment like D.E.S will improve the delensign efficiency by $~10\%$ if combined with Planck lensing and a CIB tracers.
 As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential.
-Indeed using the CMB itself we reconstruct only the lenses that actually lens the CMB obtained an almost perfect cleaning in the absence of noise.
+Indeed using the CMB itself, we reconstruct only the lenses that actually lens the CMB obtained an almost perfect cleaning in the absence of noise.
 This approach will need a careful study of possible biases coming from using the same source (the CMB) we are delensing to reconstruct the lensing effect itself.
 This has already been applied to data and studied.
-However, the fairly good efficiency of galaxies tracers might come in end to cross-check these internal biases.
+However, the fairly good efficiency of galaxies tracers might be useful to cross-check these internal biases.
 This will probably be needed to confirm a possible detection of gravitation waves if this relies heavily on delensing.

### 2017-05-29 Code/Plot2.py

@@ -1,418 +1,0 @@
-
-
-import numpy as np
-import matplotlib.pyplot as plt
-import multiple_survey_delens
-
-
-# In[14]:
-
-# ============================================
-# SIZE OF THE PICTURE
-# ============================================
-WIDTH = 246.0  # the number latex spits out
-# the fraction of the width you'd like the figure to occupy, 0.5 if double
-# column cause
-FACTOR = 0.99
-fig_width_pt = WIDTH * FACTOR
-
-inches_per_pt = 1.0 / 72.27
-golden_ratio = (np.sqrt(5) - 1.0) / 2.0  # because it looks good
-ratio = 0.8
-fig_width_in = fig_width_pt * inches_per_pt  # figure width in inches
-fig_height_in = fig_width_in * ratio   # figure height in inches
-fig_dims = [fig_width_in, fig_height_in]  # fig dims as a list
-# ============================================
-# ============================================
-
-
-# ============================================
-# SET LATEX
-plt.rc('text', usetex=True)
-plt.rc('font', **{'family': 'serif', 'serif': ['Computer Modern Roman']})
-# ============================================
-
-
-font_size = 8.
-plt.rcParams['font.size'] = font_size
-plt.rcParams['axes.labelsize'] = font_size * 1.3
-plt.rcParams['axes.linewidth'] = font_size / 10.
-plt.rcParams['axes.titlesize'] = font_size * 1.3
-plt.rcParams['legend.fontsize'] = font_size
-plt.rcParams['xtick.labelsize'] = font_size / 1.2
-plt.rcParams['ytick.labelsize'] = font_size / 1.2
-plt.rcParams['xtick.major.width'] = font_size / 10.
-plt.rcParams['ytick.major.width'] = font_size / 10.
-# plt.rcParams['xtick.labelsize'] = font_size / 1.2
-# plt.rcParams['ytick.labelsize'] = font_size / 1.2
-
[…]

### 2017-05-29 Code/Plot_figures.py
order

@@ -1,0 +1,418 @@
+
+
+import numpy as np
+import matplotlib.pyplot as plt
+import multiple_survey_delens
+
+
+# In[14]:
+
+# ============================================
+# SIZE OF THE PICTURE
+# ============================================
+WIDTH = 246.0  # the number latex spits out
+# the fraction of the width you'd like the figure to occupy, 0.5 if double
+# column cause
+FACTOR = 0.99
+fig_width_pt = WIDTH * FACTOR
+
+inches_per_pt = 1.0 / 72.27
+golden_ratio = (np.sqrt(5) - 1.0) / 2.0  # because it looks good
+ratio = 0.8
+fig_width_in = fig_width_pt * inches_per_pt  # figure width in inches
+fig_height_in = fig_width_in * ratio   # figure height in inches
+fig_dims = [fig_width_in, fig_height_in]  # fig dims as a list
+# ============================================
+# ============================================
+
+
+# ============================================
+# SET LATEX
+plt.rc('text', usetex=True)
+plt.rc('font', **{'family': 'serif', 'serif': ['Computer Modern Roman']})
+# ============================================
+
+
+font_size = 8.
+plt.rcParams['font.size'] = font_size
+plt.rcParams['axes.labelsize'] = font_size * 1.3
+plt.rcParams['axes.linewidth'] = font_size / 10.
+plt.rcParams['axes.titlesize'] = font_size * 1.3
+plt.rcParams['legend.fontsize'] = font_size
+plt.rcParams['xtick.labelsize'] = font_size / 1.2
+plt.rcParams['ytick.labelsize'] = font_size / 1.2
+plt.rcParams['xtick.major.width'] = font_size / 10.
+plt.rcParams['ytick.major.width'] = font_size / 10.
+# plt.rcParams['xtick.labelsize'] = font_size / 1.2
+# plt.rcParams['ytick.labelsize'] = font_size / 1.2
+
[…]

### 2017-05-29 Code/limber_integrals.pyx
add limner int

@@ -1,0 +1,164 @@
+'''
+
+Module containing the limber integrals definitions used in the PS limber folder
+
+'''
+
+import numpy as np
+import scipy.integrate
+import sys
+from scipy.interpolate import  InterpolatedUnivariateSpline
+from joblib import Parallel, delayed
+
+def cl_limber_x(z_chi, p_kz, l, k1, k2=None, xmin=0.0, xmax=13000.):
+    """ calculate the cross-spectrum at multipole l between kernels k1 and k2 in the limber approximation. Comoving distance version. See  cl_limber_z for the redshift version.
+
+
+
+        Notes: Here everything is assumed in h units. Maybe not the best choice but that is it.
+
+        Args:
+          z_chi: z(chi) redshift as a function of comoving distance.
+          hspline: H(z). not used here kept to uniform to cl_limber_z
+          rbs: Power spectrum spline P(k,z) k and P in h units
+          l: angular multipole
+          k1: First kernel
+          k2: Optional Second kernel otherwise k2=k1
+          xmin: Min range of integration, comoving distance
+          xmax: Max range of integration, comoving distance
+
+
+        Returns:
+
+          cl_limber : C_l = \int_chi_min^chi_max d\chi {1/\chi^2} K_A(\chi) K_B(\chi)\times P_\delta(k=l/\chi;z)
+
+    """
+
+    if k2 == None:
+        k2 = k1
+
+        def integrand(double x):
+            cdef double z
+            z = z_chi(x)
+            return 1. / x /x * k1.w_lxz(l, x, z)**2 * p_kz(l / x, z)
+
+    else:
+
+        def integrand(double x):
+            cdef double z
+            z = z_chi(x)
+            return 1. / x /x * k1.w_lxz(l, x, z) * k2.w_lxz(l, x, z) * p_kz(l / x, z)
+
+
+    return scipy.integrate.quad(integrand, xmin, xmax, limit=300, epsrel=1.49e-06)[0]
+
+
+def  cl_limber_z(chi_z, hspline, rbs, l, kernel_1, kernel_2=None,  zmin=0.0,  zmax=1100.):
[…]

### 2017-05-29 Code/limber_integrals.pyx
order

@@ -139,26 +139,16 @@ def integratio_cross(ell,rbs,chi_z,hspline,kernel_1,kernel_2,integrand, zmin, zm
 
 def  cl_limber_z_ell_parallel(chi_z, hspline, rbs, lbins, kernel_1, kernel_2=None,  zmin=0.0,  zmax=1100.):
 
-    # print('lbins', lbins)
-    #  TODO check the H factor.
+
+  '''
+  As the one before but now different ells are computed in parallel
+  '''
+
+
     if kernel_2 == None:
         kernel_2 = kernel_1
-        return Parallel(n_jobs=6, verbose=0)(delayed(integratio_auto)(i,rbs,chi_z,hspline,kernel_1,integrand_auto, zmin, zmax)  for i in lbins)
+        return Parallel(n_jobs=-2, verbose=0)(delayed(integratio_auto)(i,rbs,chi_z,hspline,kernel_1,integrand_auto, zmin, zmax)  for i in lbins)
 
     else:
-        return Parallel(n_jobs=6, verbose=0)(delayed(integratio_cross)(i,rbs,chi_z,hspline,kernel_1,kernel_2,integrand_cross, zmin, zmax)  for i in lbins)
-
-
-
-    # print 'here' ,integrand(0.5)
-
-    # sys.exit()
-    # func = InterpolatedUnivariateSpline(np.linspace(zmin,zmax,100), np.vectorize(integrand)(np.linspace(zmin,zmax,100)), ext='zeros')
-    # print('')
-    # print(func(zmax-(zmax-zmin)/2.))
-    # print(scipy.integrate.quad(integrand, zmin, zmax, limit=300, epsrel=1.49e-06)[0])
-
-
-    # return scipy.integrate.quad(integrand, zmin, zmax, args=(ell,) limit=300, epsrel=1.49e-06)[0]
-
+        return Parallel(n_jobs=-2, verbose=0)(delayed(integratio_cross)(i,rbs,chi_z,hspline,kernel_1,kernel_2,integrand_cross, zmin, zmax)  for i in lbins)
 

### 2017-05-29 Code/run_galaxies_delens.py

@@ -1,4 +1,5 @@
 '''
+This script compute all the spectra needed for the delensing part.
 '''
 import pyximport
 import numpy as np
@@ -9,14 +10,14 @@ import hall_CIB_kernel as cib_hall
 import scipy.integrate
 from scipy.interpolate import RectBivariateSpline, interp1d, InterpolatedUnivariateSpline
 import limber_integrals
-import pickle as pk
+# import pickle as pk
 import camb
-from camb import model, initialpower
+from camb import model
 import configparser
 pyximport.install(reload_support=True)
-from joblib import Parallel, delayed
-from profiling.sampling import SamplingProfiler
-profiler = SamplingProfiler()
+# from joblib import Parallel, delayed
+# from profiling.sampling import SamplingProfiler
+# profiler = SamplingProfiler()
 
 
 def cl_limber_z_local(chi_z, hspline, rbs, l, kernel_1, kernel_2=None, zmin=0.0, zmax=1100.):
@@ -41,7 +42,7 @@ def cl_limber_z_local(chi_z, hspline, rbs, l, kernel_1, kernel_2=None, zmin=0.0,
     """
 
     #  TODO check the H factor.
-    if kernel_2 == None:
+    if kernel_2 is None:
         kernel_2 = kernel_1
 
         def integrand(z):
@@ -85,18 +86,7 @@ def setup(ini_file='./gal_delens_values.ini'):
     llmin = config.getfloat('spectra', 'llmin', fallback=0.)
     llmax = config.getfloat('spectra', 'llmax', fallback=3.)
     dlnl = config.getfloat('spectra', "dlnl", fallback=.1)
-    # # What matter power spectrum to use, linear Halofit etc
-    # dndz_filename = options.get_string(
-    # option_section, "dndz_filename",
-    # default="/home/manzotti/cosmosis/modules/limber/data_input/DES/N_z_wavg_spread_model_0.2_1.2_tpz.txt")
-
-    # nu = options[option_section, "nu"]
-    # # nu = np.fromstring(nu, dtype=float, sep=',')
-    # zc = options.get_double(option_section, "zc", default=2.)
-    # zs = options.get_double(option_section, "zs", default=2.)
-    # b = options.get_double(option_section, "b", default=1.)
-
[…]

### 2017-05-30 Notes/DES_rec_phi_paper.tex
work on text

@@
 this period is called ``cosmic inflation.''
-
 Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski15}.
 These primordial gravitational waves in the early universe imprint a unique signature on the polarized anisotropies of the CMB.
-CMB polarization is usually decomposed fields into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
-In the standard scenario, the PGW background is the only source of B-mode polarization at the epoch of recombination.
+For this reason, CMB polarization is a promising dataset to understand the physics of the early universe.
+Furthermore, CMB polarization is usually studied in Fourier space by decomposing the fields into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
+The B-mode polarization, in particular, is a great source of information because, in the standard scenario, the PGW background is the only source of B-modes at the epoch of recombination.
 
-The amplitude of this primordial B-mode component is parameterized by the ratio of the amplitudes of the primordial tensor and scalar perturbations, $r$.
+To connect data and theoretical models, the amplitude of this primordial B-mode component is parameterized by the ratio of the amplitudes of the primordial tensor and scalar perturbations, $r$.
 The shape of this component, i.e.
@@
 Most of the inflationary models directly relate the value of r to the energy scale of inflation.
-CMB B-modes provide the cleanest known observational window into the primordial gravitational waves background.
-Measuring or constraint the value of $r$ and $n_t$ will put a tight constraint on the allowed models of inflation.
+Measuring or constraining the value of $r$ and $n_t$ will put a tight constraint on the allowed models of inflation.
 
-Indeed improving the CMB B-modes measurement is a major objective of current and future CMB experiments.
-Even if inflationary B-modes have not been detected yet, a possibly natural value $r \gtrsim 10^{-3}$, which should be reachable given the near future level of noise expected in CMB experiments \citep{kamionkowski15}.
+Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational waves background, improving their measurement is a major objective of current and future CMB experiments.
+Even if inflationary B-modes have not been detected yet, a possibly natural value $r \gtrsim 10^{-3}$, should be reachable given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
 
 However reducing the level of noise will not be enough to reach this goal.
-Unfortunately, the observed B modes are not solely due to primordial gravitational waves but also by late time effects. In this work, we will focus on gravitational interactions with LSS.
+Indeed the observed B modes are not solely due to primordial gravitational waves but also by secondary effects taking place in the late-time low-redshift universe.
+In this work, we will focus on the effects of gravitational interactions with large scale structures (LSS).
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
-This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,polarbear2014c,vanengelen14b,planck2015XV} and directly from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
+This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,polarbear2014c,vanengelen14b,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
 
 This component acts as a source of confusion for searches of the PGW background.
-Indeed, current experiments are already limited by the contamination from lensing B modes since the instrument noise is below the lensing B-mode $rms$ fluctuations \citep{bk14}.
-The optimal way to approach this source of noise is to build the specific realization of lensing B modes on the sky can be and used it to clean the data in a process called ``delensing''.
+Indeed, the contamination from lensing B modes is already at the level of the instrumental noise of current experiments \citep{bk14}.
+Thus together with the experimental effort to reduce the amount of noise is important to study the effect of the spurious lensing component.
+The optimal way to reduce the lensing contributions is to reconstruct the expected realization of lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
+We can delens the observed B-modes combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
 
 Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
-Furthermore recently it has been performed on CMB temperature data using the cosmic infrared background in \cite{larsen:2016}, on CMB temperature and polarization data using internal CMB reconstruction \cite{ carron17}.
+Furthermore, recently it has been performed on CMB temperature data using as LSS tracers the cosmic infrared background in \cite{larsen:2016} and on CMB temperature and polarization data using internal CMB lensing reconstruction \cite{ carron17}.
 Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in where $~28\%$ of the lensing power has been removed.
@@
 In this paper, we study the possible role of future galaxies surveys as tracers of the lensing potential.
-These will translate into a better reconstruction of the B-modes in the measured patch and as a consequence improve the constraints on inflation trough delensing.
+These will translate into a better reconstruction of the B-modes in the measured patch, and as a consequence, it will improve the constraints on inflation trough delensing.
 We model several actual and future surveys, and after computing the residual B-modes, we forecast the statistical uncertainness on $r$ and $n_T$ for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
 
-This article is organized as follow: in \refsec{th} we describe the lensing B-mode components and the residual power after delensing with tracers of the lensing potential.
+We organize this article as follow: in \refsec{th} we describe the lensing B-mode components and the residual power after delensing with tracers of the lensing potential.
 We describe the LSS tracers used in this analysis in \refsec{model}.
@@
 \includegraphics[scale=1.]{../images/cmb_internal.pdf}
-\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the true lensing potential for different CMB experiments and corresponding noise reconstructions.
-}
+\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
+CMB measurements will rapidly improve providing the most efficient even if biased way to reconstruct the structures lensing the CMB.}
 \end{center}
@@
 
-Very recently delensing has been proven possible on real data using as tracers both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as the CIB \cite{larsen:2016a,manzotti:2017}.
-Here we discuss the improvement due to combining these and other available probes (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
+Very recently delensing has been proven possible on real data using as tracers both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as CIB data \cite{larsen:2016a,manzotti:2017}.
+Here we discuss the improvement due to combining these and other available surveys (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
 
-On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with a noise equal to \textcolor{red}{$3\mu$K-arcmin} and a beam of \textcolor{red}{30 arcmin}.
+On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin} and a beam of \textcolor{red}{30 arcmin}.
 The CMB lensing reconstruction will be performed by a higher resolution CMB experiment. We explore two possibilities.
@@
 \subsection{CMB-S3 Era}
-CMB measurements are rapidly improving, and the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
-For the CMB at the beginning of the next decade, we will assume a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
+CMB measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
+To model the CMB at the beginning of the next decade, we will assume a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
 This is a reasonable level for a combination of the next generation of BK and SPT3G.
-This level of noise will also be assumed for the internal noise reconstruction.
+We will assume this level of noise for the internal noise reconstruction.
 
 The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
-Not only will CMB experiments improved shortly: possibly D.E.S.I will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
+Not only will CMB experiments improved shortly: eventually  D.E.S.I will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
 DESI will be slightly more efficient than DES-like survey because of the slightly higher redshift that it can probe.
@@
 
-An ambitious program for a generation 4 ground CMB experiment is currently under planning. Moreover, satellite experiments have been proposed.
+An ambitious program for a generation 4 ground CMB experiment is currently under planning \cite{abazajian:2016}. Moreover, satellite experiments have been proposed.
 The correlation attainable using current experiments is shown in \reffig{corrs4},
-In this very futuristic but reasonable scenario, we add to the available large scale structures both 21 cm surveys like SKA and large scale structures like LSST and Euclid.
+In this futuristic but reasonable scenario, we add to the available large scale structures tracers  both 21 cm surveys like SKA and large scale structures like LSST and Euclid.
 
@@
 \label{fig:corrs4}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/S4_scenario_no_ska.pdf}
+\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments without SKA. }
+\label{fig:corrs4}
 
@@
 \end{figure}
+
 %=======================================
@@
 \label{sec:concl}
+
 Delensing or more, in general, the ability to separate the lensing component of the B-mode from a possible primordial inflationary signal is important to exploit the capabilities of future experiments fully.
@@
 As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential.
-Indeed using the CMB itself, we reconstruct only the lenses that actually lens the CMB obtained an almost perfect cleaning in the absence of noise.
+Indeed using the CMB itself, we reconstruct only the lenses that lens the CMB obtained an almost perfect cleaning in the absence of noise.
 This approach will need a careful study of possible biases coming from using the same source (the CMB) we are delensing to reconstruct the lensing effect itself.
 This has already been applied to data and studied.
-However, the fairly good efficiency of galaxies tracers might be useful to cross-check these internal biases.
+However, the relatively good efficiency of galaxies tracers might be useful to cross-check these internal biases.
 This will probably be needed to confirm a possible detection of gravitation waves if this relies heavily on delensing.
 
-Another possible application of LSS delensing is the removal of the low-z lensing of the CMB that possibly adds the complexity of modeling non linearities in the study of CMB lensing
+Another possible application of LSS delensing is the removal of the low-z lensing of the CMB that possibly adds the complexity of modeling nonlinearities in the study of CMB lensing
 
@@
 \section{Acknowledgments}
-We thank S. Dodelson for useful discussion.
+We thank S. Dodelson K. Wu and K. Story for useful discussion.
 This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.

### 2017-06-03 Notes/DES_rec_phi_paper.tex
work on writing

@@
 
-Delensing or more, in general, the ability to separate the lensing component of the B-mode from a possible primordial inflationary signal is important to exploit the capabilities of future experiments fully.
+Delensing or the ability to separate the lensing component of the CMB B-mode from a possible primordial inflationary signal is important to exploit the capabilities of future experiments thoroughly.
 In this paper, we studied the potential impact of large-scale structure surveys in this important endeavor.
-For current experiment, CIB data had already proven to be very efficient in delensing.
-If no high-resolution CMB experiment is available to reconstruct the lensing potential $\phi$ a low-z optical experiment like D.E.S will improve the delensign efficiency by $~10\%$ if combined with Planck lensing and a CIB tracers.
+For current experiment, CIB data had already proven to be very efficient in delensing \cite{larsen:2016,manzotti:2017}. 
+If no CMB experiment is available with higher resolution then Planck is used to reconstruct the lensing potential $\phi$, a low-z optical survey like D.E.S will improve the delensing efficiency by $~10\%$ when combined with Planck lensing and a CIB tracers.
 As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential.
-Indeed using the CMB itself, we reconstruct only the lenses that lens the CMB obtained an almost perfect cleaning in the absence of noise.
-This approach will need a careful study of possible biases coming from using the same source (the CMB) we are delensing to reconstruct the lensing effect itself.
-This has already been applied to data and studied.
-However, the relatively good efficiency of galaxies tracers might be useful to cross-check these internal biases.
+Indeed using the CMB itself, we reconstruct only the lenses that lens the CMB obtaining, as a result, an almost perfect cleaning in the absence of noise.
+This approach has already been applied to data \cite{carron:2017}.
+However it requires a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}.
+For this reason, the relatively good efficiency of galaxies tracers might be useful to cross-check these internal biases.
 This will probably be needed to confirm a possible detection of gravitation waves if this relies heavily on delensing.
 
-Another possible application of LSS delensing is the removal of the low-z lensing of the CMB that possibly adds the complexity of modeling nonlinearities in the study of CMB lensing
+Another possible application of LSS delensing, which we will explore in future works, is the possibility of removing just the low-z lensing of the CMB which will otherwise possibly adds the complexity of modeling nonlinearities in the study of CMB lensing.
 

### 2017-06-03 Notes/galaxies_delensing_uchicago_thesis.tex
add thesis format

--- 2017-06-03 Notes/DES_rec_phi_paper.tex
@@
-\documentclass[reprint,superscriptaddress,amsmath,amssymb,floatfix,aps,prd,nofootinbib]{revtex4-1}
+% ---- ETD Document Class and Useful Packages ---- %
+\documentclass{ucetd}
+\usepackage{subfigure,epsfig,amsfonts}
+\usepackage{natbib}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{natbib}
+\usepackage{enumitem}
+%\usepackage{hyperref}
+\input{./aas_abbr.tex}
 \usepackage{bm}
@@
 \usepackage{graphicx,epsfig}
-\usepackage{hyperref}
-\usepackage{ifthen}
 \usepackage{xstring}
 % \usepackage{graphicx}\usepackage[top=2.5 cm, bottom=2.5 cm, left=1.05in, right=1.05in]{geometry}
-\usepackage[applemac]{inputenc}
-\usepackage{amsmath,amssymb}
+%\usepackage[applemac]{inputenc}
+%\usepackage{amsmath,amssymb}
 \usepackage{color}
-\usepackage{epstopdf}
-%\usepackage[caption=false]{subcaption}
-%\epstopdfDeclareGraphicsRule{.gif}{png}{.png}{convert gif:#1 png:\OutputFile}
-% \AppendGraphicsExtensions{.gif}
-% notes to make comment is non-zero if
+%\usepackage{epstopdf}
+
 \newcommand{\al}[1]{\begin{align} #1 \end{align}}
@@
 \def\l{\ell}
-% For various journals
-%\newcommand{\apj}{ApJ}
-\newcommand{\physrep}{Physics Reports}
-\newcommand{\jcap}{JCAP}
-\newcommand{\apjl}{ApJL}
-\def\araa{\ref@jnl{ARA\&A}}
-\newcommand{\apjs}{ApJS}
-\newcommand{\aap}{A\&A}
-\newcommand{\mnras}{MNRAS}
-\newcommand{\physrev}{Phys. Rev.}
-\newcommand{\physrevlett}{Phys. Rev. Lett.}
-\newcommand*\aj{AJ}
-%\newcommand*\prd{Phys.~Rev.~D}
 
@@
 
-\usepackage{hyperref}
-\usepackage{amsmath}
-\usepackage{natbib}
 \newcommand{\bx}{{\bf x}}
@@
 \newcommand{\bmm}[1]{{\mathbf{#1}}}
-\newcommand{\bsection}[1]{\section{\uppercase{#1}}}
+\newcommand{\bsection}[1]{\chapter{\uppercase{#1}}}
+
+%% Use these commands to set biographic information for the title page:
+\title{Future Cosmic Microwave Background delensing with galaxies surveys}
+\author{Alessandro Manzotti}
+\department{Astronomy and Astrophysics}
+\division{Physical Sciences}
+\degree{Doctor of Phylosophy}
+\date{August 2017}
+\dedication{Dedication Text}
+\epigraph{No epigraph}
 
@@
 
-\title{Future Cosmic Microwave Background delensing with galaxies surveys.}         % Enter your title between curly braces
-\author{A. Manzotti}        % Enter your name between curly braces
-\email{manzotti.alessandro@gmail.com}
-\affiliation{Kavli Institute for Cosmological Physics, University of Chicago}
-\affiliation{Department of Astronomy and Astrophysics, University of Chicago, 5640 South Ellis Avenue, Chicago, IL 60637, USA}
-\date{\today}
+\maketitle
+\makecopyright
+%\makededication
+%\makeepigraph
+\tableofcontents
+\listoffigures
+\listoftables
 
-        % Enter your date or \today between curly braces
-\begin{abstract}
-%\begin{itemize}
-%\item Worry : they might ask for internal bias\
-%item Possible TODO : temperature $N_{eff}$
-%
-%\end{itemize}
-%
+\abstract
 
@@
 We find that current generation experiments...
-\end{abstract}
 
-\vspace{1cm}
+\mainmatter
 
-\maketitle
-
-\section{Introduction}
+\chapter{Introduction}
 \label{sec:intro}
@@
 
-\section{Gravitational lensing B-mode and delensing}
+\chapter{Gravitational lensing B-mode and delensing}
 \label{sec:th}
@@
 
-\subsection{Multiple tracers of the lensing potential}\label{sec:galcontrib}
+\section{Multiple tracers of the lensing potential}\label{sec:galcontrib}
 
@@
 
-\section{Lensing potential tracers}
+\chapter{Lensing potential tracers}
 \label{sec:model}
@@
 
-\subsection{CMB lensing potential}
+\section{CMB lensing potential}
 \label{sec:kappaCMB}
@@
 
-\subsection{Galaxies}
+\section{Galaxies}
 
@@
 
-\subsection{Cosmic infrared Background (CIB)}
+\section{Cosmic infrared Background (CIB)}
 \label{sec:cib}
@@
 
-\section{Parameter constraints improvement after delensing}
+\chapter{Parameter constraints improvement after delensing}
 \label{sec:for}
@@
 
-\subsection{Fisher method}
+\section{Fisher method}
 In this framework, the error on a cosmological parameter constrained by a B-mode spectrum measurement is given as usual by:
@@
 %===================
-\subsection{Delensing with current CMB and LSS}
+\section{Delensing with current CMB and LSS}
 
@@
 First, we test the efficiency with an internal reconstruction performed by Planck \cite{2016A&A...594A..15P} using the CMB lensing noise curves publicly available
-\footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
+\footnote{\textit{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
 Then we will assume a ground-based experiment with the final noise level of SPT-Pol which corresponds to \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
@@
 \begin{table}
+\centering
 \caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to ration of the error before and after delensing.}
@@
 \begin{table}
+\centering
 \caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to the ratio of the error before and after delensing.}
@@
 
-\subsection{CMB-S3 Era}
+\section{CMB-S3 Era}
 CMB measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
@@
 \begin{table}
+\centering
 \caption{$\alpha$: S3 improvements. Note that the absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is  $\sigma(r) =$  }
@@
 
-\subsection{CMB-S4 Era}
+\section{CMB-S4 Era}
 
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/S4_scenario_no_ska.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments without SKA. }
-\label{fig:corrs4}
-
-\end{center}
-\end{figure}
+%\begin{figure}[htbp]
+%\begin{center}
+%\includegraphics[scale=1.]{../images/S4_scenario_no_ska.pdf}
+%\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments without SKA. }
+%\label{fig:corrs4}
+%
+%\end{center}
+%\end{figure}
 
@@
 \begin{table}
+\centering
 \caption{$\alpha$: Gen 4 experiments}
@@
 
-\subsection{Bias uncertainties degradation}
+\section{Bias uncertainties degradation}
 The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of the inflationary constraint of delensed spectra.
@@
 
-\section{Conclusions}
+\chapter{Conclusions}
 \label{sec:concl}
@@
 
-%=================================================================
-\section{Acknowledgments}
-We thank S. Dodelson K. Wu and K. Story for useful discussion.
-This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.
-%%%=================================================================
+\makebibliography
 
-\bibliographystyle{plain}
-\bibliography{DES_rec_phi_paper,/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
-
-%\bibliography{DES_rec_phi_paper}
 \end{document}

### 2017-06-07 Code/gals_kernel.pyx
add slides

@@ -55,6 +55,20 @@ def dNdZ_parametric_LSST(z, z0=0.5, alpha=1.27, beta=1.02):
     temp = (z / z0) ** alpha * np.exp(-(z / z0) ** beta)
     return temp
 
+def dNdZ_parametric_LSST_bins(z, z0=0.5, alpha=1.27, beta=1.02 , nbins = 10):
+    '''
+    usual paramteric form of galaxies distribution
+    '''
+    chunks = [z[i:i+nbins] for i in range(0, len(z), nbins)]
+    z_bins = []
+    for i in np.arange(0,nbins):
+        temp = np.zeros(len(z))
+        for j , z_val in enumerate(z):
+            temp[j] = (z_val / z0) ** alpha * np.exp(-(z_val / z0) ** beta) if z_val in chunks[i] else 0.
+        z_bins.append(temp)
+    return z_bins
+
+
 
 def dNdZ_parametric_LSST_2d_alpha(z, alpha, z0=0.5, beta=1.02):
     '''

### 2017-06-07 Code/rho_to_Bres.py

@@ -8,6 +8,7 @@ import scipy.integrate as integrate
 import profiling
 from profiling.sampling import SamplingProfiler
 
+
 def bl(fwhm_arcmin, lmax):
     """ returns the map-level transfer function for a symmetric Gaussian beam.
          * fwhm_arcmin      - beam full-width-at-half-maximum (fwhm) in arcmin.
@@ -32,14 +33,16 @@ def compute_BB(clee_fun, clpp_fun, ell_b_bin, ell_phi_bin):
         clee = clee_fun(ell)
         return (ell / (2. * np.pi)**2 * (L * ell * np.cos(theta) - ell**2)**2 * clpp_fun(np.sqrt(L**2 + ell**2 - 2. * ell * L * np.cos(theta))) * clee * (np.sin(2. * theta))**2)
 
+    options1 = {'limit': 4900, 'epsabs': 0., 'epsrel': 1.e-7}
+    options2 = {'limit': 4900, 'epsabs': 0., 'epsrel': 1.e-3}
 
-    options1 = {'limit': 900, 'epsabs': 0., 'epsrel': 1.49e-3}
-    options2 = {'limit': 900, 'epsabs': 0., 'epsrel': 1.49e-3}
-
-    lbins_int = np.arange(ell_b_bin[0], ell_b_bin[1], 5)
+    lbins_int = np.arange(ell_b_bin[0], ell_b_bin[1], 15)
 
     clbb_ell = [integrate.nquad(
         integrand, [[0., 2. * np.pi], [4, 3000]], args=(L,), opts=[options1, options2])[0] for L in lbins_int]
+    # print(clbb_ell, 'errors on b', [integrate.nquad(
+    # integrand, [[0., 2. * np.pi], [4, 3000]], args=(L,), opts=[options1,
+    # options2])[1] for L in lbins_int])
 
     # reconstruction_noise_ell = [integrate.nquad(reconstruction_noise_integrand, [[0., 2. * np.pi], [6, 2500]], args=(
     #     L, x1, x2), opts=[options1, options2])[0] for L in np.arange(10, 1000, 50)]
@@ -80,7 +83,7 @@ def compute_res_parallel(rho_filename, output_dir, clee_fun, clpp_fun, nle_fun):
     return clbb_res_ell
 
 
-def compute_deriv(ells, clee_fun, clpp, l_phi, l_bb):
[…]

### 2017-06-07 Code/run_galaxies_delens.py

@@ -188,6 +188,21 @@ def main(ini_par):
         z_lsst, dndzlsst / norm * 1. * np.sqrt(1. + z_lsst), ext='zeros')
     lsst = gals_kernel.kern(z_lsst, dndzlsst, hspline, pars.omegac, h, b=1.)
 
+    nbins = 5
+    lsst_bins = []
+    z_bins = [z_lsst[i:i + int(200 / nbins)] for i in range(0, len(z_lsst), int(200 / nbins))]
+    for z in z_bins:
+        dndzlsst = gals_kernel.dNdZ_parametric_LSST(z)
+        dndzfun = interp1d(z, dndzlsst)
+        # print(z,dndzlsst / norm * 1. * np.sqrt(1. + z))
+        dndzlsst = InterpolatedUnivariateSpline(
+            z, dndzlsst / norm * 1. * np.sqrt(1. + z), ext='zeros')
+        lsst_bins.append(gals_kernel.kern(z, dndzlsst, hspline, pars.omegac, h, b=1.))
+
+    # ======
+    # WEAK
+    # =======
+
     des_weak = kappa_gals_kernel.kern(z_lsst, dndzlsst, chispline, hspline, pars.omegac, h)
 
     # ======
@@ -218,9 +233,13 @@ def main(ini_par):
     # Compute Cl implicit loops on ell
     # =======================
 
-    kernels = [lkern, wise, euclid, des_weak, lsst, ska10, ska01, ska5, ska1, cib, desi, des]
-    names = ['k', 'wise', 'euclid', 'des_weak', 'lsst', 'ska10',
-             'ska01', 'ska5', 'ska1', 'cib', 'desi', 'des']
+    # kernels = [lkern, wise, euclid, des_weak, lsst, ska10, ska01, ska5, ska1, cib, desi, des]
+    # names = ['k', 'wise', 'euclid', 'des_weak', 'lsst', 'ska10',
+    #          'ska01', 'ska5', 'ska1', 'cib', 'desi', 'des']
+
+    kernels = [lkern, lsst_bins[0], lsst_bins[1], lsst_bins[2], lsst_bins[3], lsst_bins[4]]
+    names = ['k', 'lsst0', 'lsst1', 'lsst2', 'lsst3', 'lsst4']
+
 
     labels = []
     kernel_list = []
@@ -238,7 +257,7 @@ def main(ini_par):
 
     # profiler.stop()
     # profiler.run_viewer()
-
[…]

### 2017-06-07 Notes/DES_rec_phi_paper.tex
text

@@
 
-We can also write the galaxy overdensity field $g(\nver)$ in a given direction on the sky as a line of sight integral of the matter overdensity:
-\begin{equation}
-g(\nver) = \int_0^{z_*} dz\, W^{g}(z)\delta(\chi(z)\nver,z),
-\end{equation}
-where the kernel is
+We can also write the galaxy overdensity field in a given direction on the sky as a line of sight integral of the matter overdensity.
+The galaxies kernel is:
 \begin{equation}
@@
 The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
-Not only will CMB experiments improved shortly: eventually  D.E.S.I will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
+Not only will CMB experiments improved shortly: eventually  DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
 DESI will be slightly more efficient than DES-like survey because of the slightly higher redshift that it can probe.

### 2017-06-11 Notes/DES_rec_phi_paper.tex
work on writing

@@
 Here we discuss the role of current and future large scale structure surveys in improving the reconstruction of the lensing potential that lenses the CMB photons and how this reflects into an improved delensing efficiency.
-We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$)
-We find that current generation experiments...
+We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
+We find that current generation experiments... we also point out that a tomographic approach..
 \end{abstract}
@@
 
-We organize this article as follow: in \refsec{th} we describe the lensing B-mode components and the residual power after delensing with tracers of the lensing potential.
-We describe the LSS tracers used in this analysis in \refsec{model}.
+We organize this article as follow: we describe the LSS tracers used in this analysis in \refsec{model}.
+In \refsec{th} we define the lensing B-mode component and the residual power after delensing with tracers of the lensing potential.
 \refsec{for} is the main results of this work: the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
@@
 
-\section{Gravitational lensing B-mode and delensing}
-\label{sec:th}
-
-As for the CMB photons temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is remapped by lensing as:
-\be
-Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi);~~
-U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi)
-\ee
-where the deflection angle is the gradient of lensing potential potential integrated along the line of sight $\nabla\phi$.
-The CMB polarization is usually decomposed into odd-parity Fourier modes E and B. 
-As shown in \cite{} this for symmetry reason tensor perturbations are the principal source of the B-modes configuration.
-For this reason, B-modes are a promising signature of early universe tensor perturbations.
-However, this promising sign of primordial gravitational waves is partially obscured by a secondary mechanism that can produce B-modes also in the late universe.
-Indeed the gravitational interaction with large scale structure generates CMB B-modes by distorting primordial E-modes.
-
-As a first approximation, given a convergence field $\kappa= -\frac{1}{2}\nabla^2\phi $ the B mode resulting from the lensing of primordial E mode is:
-\be
-B^{\mathrm{lens}}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') E(\ublu) \kappa(\bl - \bl')
-\label{eqn:blens}
-\ee
-where different modes contributes with a different weight given by:
-\be
-W(\bl,\bl') = \frac{2 \bl' \cdot (\bl-\bl')}{|\bl-\bl'|^2} \sin(2\varphi_{\bl,\bl'}).
-\ee
-
-As usual we define the power spectrum as:
-\be
-\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l .
-\ee
-
-From this we get that the power spectrum of the lensing component of the B-modes:
-\be
-C_l^{BB,\mathrm{lens}}  = \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl')C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}.
-\ee
-
-The full B-mode power spectrum measured on the sky is given by a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_l^{BB}$:
-\be
-C_l^{BB,\mathrm{full}} = C_l^{BB,r} + C_l^{BB,\mathrm{lens}} + N_l^{BB}.
-\ee
-Given the current constraints, the lensing component is a significant source of B-modes that, for GW searches purposes correspond to a white noise source of roughly $5 \mu K$-arcmin.
-This means that it is not only bigger than the biggest allowed GW contribution at scales bigger than several degrees, but it is comparable with the decreasing level of instrumental noise.
-For this reason, it is critical to characterize and eventually remove it from the data.
-To do so, we need to build a template \refeq{blens} (or its real-space analog) of the actual lensing modes given a measurement of the E-mode field and the lensing potential $\phi$.
-While E is measured directly, we ca estimate $\phi$ using "tracers" of the dark matter distribution that creates the potential.
-
-We will now show how much the delensing efficiency is related to the fidelity of the lensing tracers.
-If we have a large scale structure field I($\nver$) that traces the lensing potential responsible for the lensing of the CMB we can build a template of the lensing B mode on the sky by a weighted convolution:
-\be
-\hat{B}^{\mathrm{lens}}(\bl) = \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') f(\bl,\bl') E^N(\ublu) I(\bl - \bl')
-\ee
-where $f(\bl,\bl')$ can be determined byminimizing the difference with the true $B^{\mathrm{lens}}(\bl)$ defined in \refeq{blens}.
-
-The residual lensing B mode due to an imperfect knowledge of the true E-mode and $\phi$ will be
-\beqn
-B^\mathrm{res}(\bl) &=&  B^\mathrm{lens}(\bl) - \hat{B}^\mathrm{lens}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') \times \nonumber \\ &&\left( E(\ublu) \kappa(\bl - \bl' ) -  f(\bl,\bl') E^N(\ublu) I(\bl - \bl' ) \right)
-\eeqn
-
-The weights $f(\bl,\bl')$ so that the residual lensing B mode power is minimized are:
-\be
-f(\bl,\bl') = \left(\frac{C^{EE}_{{l'}}}{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right)  \frac{C^{\kappa I}_{|\bl-\bl'|}}{C^{II}_{|\bl-\bl'|} }
-\ee
-Notice that the first term consists in an inverse variance filter applied to the measured E-mode and the second minimize the difference between the reconstructed $\phi$ and the CMB lensing potential.
-
-With this choice of $f(\bl,\bl')$ we finally have that the residual power is:
-\beqn
-\label{eqn:Bres}
-C_{{l}}^{BB,\mathrm{res}} &=&   \int \frac{d^2 \bl'}{(2 \pi)^2}  W^2 (\bl,\bl')
- C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}  \\
-&\times&  \left[1 - \left(\frac{C^{EE}_{{l'}} }{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right) \rho^2_{|\bl-\bl'|} \right] \nonumber
-\eeqn
-with
-\be
-\rho_l= \frac{\cl^{\kappa I}}{\sqrt{\cl^{\kappa \kappa} \cl^{I I}}}.
-\ee
-
-The bigger $\rho_l$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
-A higher correlation allows for a better reconstruction of the $\phi^{CMB}$ and, as a consequence, of $B^{lens}$.
-
-\subsection{Multiple tracers of the lensing potential}\label{sec:galcontrib}
-
-Let's now assume that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
-We can optimally combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ with:
-\beqn
-I         &=&    \sum_{i}c^{i}I^{i} \nonumber \\
-c_{i}     &=& (C^{-1})_{ij}C^{\kappa I^{j}}
-\label{eqn:combined}
-\eeqn
-where C is the covariance matrix of the LSS tracers.
-The residual B-mode power can we computed using \refeq{Bres} using an ``effective'' correlation of these combined tracers with gravitational lensing is:
-\be
-\rho^{2} = \sum_{i,j}\frac{C^{\kappa i}~(C^{-1})_{ij}~C^{\kappa j}}{C^{\kappa\kappa}}.
-\label{eqn:rho-combined}
-\ee
-
-Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing, but it also depends on how much it is correlated with the already used set of tracers.
-\reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
-Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
-However given the significant low noise of this measurement they can still help the delensing of the CMB.
-Furthermore, the low-redshift tail of the lensing kernel is also the one affected by large-scale structures non-linearities.
-They can indeed be extremely helpful if future CMB surveys will try to remove that contribution instead of attempting to model it.
-On the other end, CIB and 21 cm surveys are probing to high redshift structures and independently from the model assumed they show a relatively good overlap with the CMB lensing kernel.
-
-% ===================
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
-The bigger the overlap wit the CMB lensing kernel the better the reconstruction of the lensing potential will be.
-This will lead to a better delensing.
-}
-\label{fig:kernel}
-
-\end{center}
-\end{figure}
-% ===================
-
 \section{Lensing potential tracers}
 \label{sec:model}
-In this section, we describe how the power spectra defined in \refsec{th} are computed and how we characterize different tracers considered in this work.
+In this section, we introduce the different large scale structure tracers considered in this work to reconstruct the lensing potential that distorts primordial E-modes generating a non-primordial B-mode component.
+Also, we define the power spectra that we will use later in \refsec{th}.
 
@@
 where $\delta(\chi(z)\nver,z)$ corresponds to the dark matter overdensity field at a comoving distance $\chi(z)$ and at a redshift $z$ in the angular direction $\nver$.
-Using the Limber approximation \cite{limber53} we can compute the power spectra of two large scale structure fields $i,j$ as:
+Using the Limber approximation \cite{limber53} we can compute the power spectra of two large-scale structure fields $i,j$ as:
 \begin{equation}
@@
 
-In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, $\chi(z)$ is the comoving distance and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$.
-Furthermore $W^{i}(z)$ is the kernel function of the field $i$ defined in \refeq{wkernel}.
-We compute these quantities using CAMB and HALOFIT.
+In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$. 
+Furthermore $W^{i}(z)$ is the kernel function of the field $i$ introduced in \refeq{wkernel}.
+We will now introduce the kernels for each of the tracers used in this work.
 
@@
 \label{sec:kappaCMB}
-The first field we compute is the CMB lensing potential.
+We start from the CMB lensing potential.
 The lensing kernel $W^{\kappa}$ is:
@@
 %
-where $\chi(z)$ is the comoving distance to redshift $z$, $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$, $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
+where $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$, $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
 
@@
 In that case, we can treat it as a noisy tracer of the true field.
-Both the CMB lensing fields and its noisy reconstructed counterparts have the same kernel $W^{\kappa}(z)$.
+Both the CMB lensing field and its noisy reconstructed counterpart have the same kernel $W^{\kappa}(z)$.
 However, when computing the power spectrum of the latter, we need to add a noise component.
@@
 \eeq
-In this work, the level of noise is computed assuming an iterative quadratic estimator is used in the reconstruction
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/cmb_internal.pdf}
-\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
-CMB measurements will rapidly improve providing the most efficient even if biased way to reconstruct the structures lensing the CMB.}
-\end{center}
-\label{fig:corr-internal}
-
-\end{figure}
+In this work, the level of noise is computed assuming an iterative quadratic estimator is used in the reconstruction \cite{}.
 
@@
 
-We can also write the galaxy overdensity field in a given direction on the sky as a line of sight integral of the matter overdensity.
-The galaxies kernel is:
+The galaxies clustering kernel is:
 \begin{equation}
@@
 \end{equation}
-Here $\frac{dN}{dz}$ is the number of galaxies as a function of redshift observed by the survey while $b(z)$ is the galaxy bias that connects the amplitude of galaxies overdensity to the one of the dark matter at different redshift.
-When computing the auto-spectrum of a galaxies density, a shot noise term needs to be taken into account. 
+Here $\frac{dN}{dz}$ is the number of galaxies observed by the survey as a function of redshift while $b(z)$ is the galaxy bias that connects the amplitude of galaxies overdensity to the one of the dark matter underneath.
+When computing the auto-spectrum of the galaxies density, a shot noise term needs to be taken into account. 
 To do so, we add a constant term to the power spectrum equal to the inverse of the number of galaxies per steradians.
-
  Different galaxy surveys in this work are fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxies density.
-We test the delensing efficiency modeling both current survey like WISE or DES as well as future galaxy probes like DESI and LSST together with 21 cm measurement like SKA.
+We test the delensing efficiency taking into account both current survey like WISE or DES as well as future galaxy probes like DESI and LSST together with 21 cm measurement like SKA.
 
 We defined the WISE redshift distribution following \cite{yan:2013} (see Fig. 4 therein). 
-To compute the noise term, we assume that the available sky after masking WISE is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxies density is approximately uniform.
+To compute the noise term, we assume that the available sky after masking is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxies density is approximately uniform.
 DES is modeled after the DES Science Verification data public release.
 For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report.
-LSST is modeled following \cite{schaan:2016} as $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
+For LSST we follow \cite{schaan:2016}: $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
 Furthermore we assume a density of 26 galaxies per arcmin squared.
@@
 
-%With these assumptions, depending on angular scale, $45-65\%$ of the CIB is correlated with the CMB lensing potential, as shown in \reffig{corrnow}.
+\section{Gravitational lensing B-mode and delensing}
+\label{sec:th}
+The large scale structures described in \refsec{model} have an important impact on the search of primordial CMB B-modes: they lens the primordial E-modes generating non inflationary B-modes that constitues an important source of noise.
 
-\cite{hall:2010,sherwin:2015,ade:2014,hanson:2013,szapudi:2001,0004-637X-567-1-2,planck-collaboration:2014,planck-collaboration:2014a,planck-collaboration:2011,plank-collaboration:2014,boulanger:1996,lewis:2006}
+Indeed the Q and U mode decompositions of the CMB photons polarization are remapped by lensing as:
+\be
+Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi);~~
+U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi)
+\ee
+where the deflection angle is the gradient of the lensing potential integrated along the line of sight $\nabla\phi$.
+The CMB polarization is usually decomposed into odd-parity Fourier modes E and B. 
+As shown in \cite{}, because of the symmetry of the problem, tensor perturbations are the principal source of the B-modes configuration.
+For this reason, B-modes are a promising signature of early universe tensor perturbations.
+
+However, this promising sign of primordial gravitational waves is partially obscured by gravitational interaction with large scale structure that generates CMB B-modes by distorting primordial E-modes.
+At first order, given the convergence field $\kappa= -\frac{1}{2}\nabla^2\phi$ introduced in \refsec{kappaCMB} the B-modes resulting from the lensing of primordial E mode are:
+\be
+B^{\mathrm{lens}}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') E(\ublu) \kappa(\bl - \bl')
+\label{eqn:blens}
+\ee
+where different modes contributes with a different weight:
+\be
+W(\bl,\bl') = \frac{2 \bl' \cdot (\bl-\bl')}{|\bl-\bl'|^2} \sin(2\varphi_{\bl,\bl'}).
+\ee
+As usual we define the power spectrum as:
+\be
+\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l .
+\ee
+
+From this we get that the power spectrum of the lensing component of the B-modes:
+\be
+C_l^{BB,\mathrm{lens}}  = \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl')C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}.
+\ee
+
+The full B-mode power spectrum measured on the sky is given by a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_l^{BB}$:
+\be
+C_l^{BB,\mathrm{full}} = C_l^{BB,r} + C_l^{BB,\mathrm{lens}} + N_l^{BB}.
+\ee
+Given the current constraints, the lensing component is a significant source of B-modes that, at large scales, correspond to a white noise source of roughly $5 \mu K$-arcmin.
+This means that it is not only bigger than the allowed inflationary component at scales smaller than several degrees, but it is also comparable with current levels of instrumental noise.
+For this reason, it is critical to characterize and eventually remove it from the data.
+To do so, we can build a template \refeq{blens} of the actual lensing B-modes given a measurement of the E-mode field and the lensing potential $\phi$ in the observed patch.
+While E is measured directly, we ca estimate $\phi$ using "tracers" of the dark matter distribution that creates the potential.
+
+We will now show how the delensing efficiency is related to the fidelity of the lensing tracers and instrumental noise in the CMB E-modes.
+If we have a large scale structure field I($\nver$) that traces the lensing potential responsible for the lensing of the CMB we can build a template of the lensing B-modes on the sky by a weighted convolution:
+\be
+\hat{B}^{\mathrm{lens}}(\bl) = \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') f(\bl,\bl') E^N(\ublu) I(\bl - \bl'),
+\ee
+where $f(\bl,\bl')$ can be determined by minimizing the difference with the true $B^{\mathrm{lens}}(\bl)$ defined in \refeq{blens}.
+We introduce the instrumental noise in the CMB E-modes ($E^N$) that will also limits the ability to fully reconstruct the lensing B-modes.
+
+The residual lensing B-modes due to an imperfect knowledge of the true E-mode and $\phi$ will be
+\beqn
+B^\mathrm{res}(\bl) &=&  B^\mathrm{lens}(\bl) - \hat{B}^\mathrm{lens}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') \times \nonumber \\ &&\left( E(\ublu) \kappa(\bl - \bl' ) -  f(\bl,\bl') E^N(\ublu) I(\bl - \bl' ) \right).
+\eeqn
+
+The weights $f(\bl,\bl')$ so that the residual lensing B mode power is minimized are:
+\be
+\label{eqn:fweight}
+f(\bl,\bl') = \left(\frac{C^{EE}_{{l'}}}{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right)  \frac{C^{\kappa I}_{|\bl-\bl'|}}{C^{II}_{|\bl-\bl'|} }.
+\ee
+Here $C^{\kappa I}$ and $C^{II}$ are the cross correlation spectrum of the tracer $I$ with the lensing convergence $\kappa$ and its autospectrum and are described for each LSS field in \refsec{model}.
+Notice that the first term consists of an inverse variance filter applied to the measured E-mode and the second minimize the difference between the reconstructed $\phi$ and the CMB lensing potential.
+
+With this choice of $f(\bl,\bl')$ we finally have that the residual power is:
+\beqn
+\label{eqn:Bres}
+C_{{l}}^{BB,\mathrm{res}} &=&   \int \frac{d^2 \bl'}{(2 \pi)^2}  W^2 (\bl,\bl')
+ C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}  \\
+&\times&  \left[1 - \left(\frac{C^{EE}_{{l'}} }{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right) \rho^2_{|\bl-\bl'|} \right] \nonumber
+\eeqn
+with
+\be
+\rho_l= \frac{\cl^{\kappa I}}{\sqrt{\cl^{\kappa \kappa} \cl^{I I}}}.
+\ee
+The bigger the $\rho_l$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
+A higher correlation allows for a better reconstruction of the $\phi^{CMB}$ and, as a consequence, of $B^{lens}$.
+
+\subsection{Multiple tracers of the lensing potential}\label{sec:galcontrib}
+
+In this section we extend the formalism to the case where multiple tracers are used to reconstruct the lensing potential.
+Furthermore we show that splitting galaxy surveys into multiple tomographic bins can significantly improve their efficiency.
+
+Let's start by assuming that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
+We can optimally combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ with:
+\beqn
+I         &=&    \sum_{i}c^{i}I^{i} \nonumber \\
+c_{i}     &=& (C_{II}^{-1})_{ij}C^{\kappa I^{j}}
+\label{eqn:combined}
+\eeqn
+where $C_{II}$ is the covariance matrix of the LSS tracers.
+The residual B-mode power can be derived from \refeq{Bres} using an ``effective'' correlation $\rho^{2}$ of these combined tracers with gravitational lensing:
+\be
+\rho^{2} = \sum_{i,j}\frac{C^{\kappa i}~(C^{-1})_{ij}~C^{\kappa j}}{C^{\kappa\kappa}}.
+\label{eqn:rho-combined}
+\ee
+
+Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing, but it also depends on how much it is correlated with the already used set of tracers.
+\reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
+Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
+However given the significant low noise of this measurement they can still help the delensing of the CMB.
+On the other end, CIB and 21 cm surveys are probing to high redshift structures and independently from the model assumed they show a relatively good overlap with the CMB lensing kernel.
+
+\subsection{Improving efficiency with tomographic binning}\label{sec:tomo-bin}
+When we weight a tracer with $\frac{C^{\kappa I}}{C^{II}}$ in \refeq{fweight} to maximize its ability to reconstruct the lensing potential we are only using redshift averaged information about the survey.
+However, as can be seen in \reffig{kernel}, the overlap varies as a function of redshifts.
+For this reason, the optimal approach is to weight each redshift bin with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
+We can see this with a simple example. Let's split a single survey I into two non-overlapping redshift bins $I_{1}$ and ${I_{2}}$ with $I=I_{1}+I_{2}$.
+For the full survey the effective cross correlation is equal to 
+\be
+\rho_{\rm{full}}^{2}= \frac{(\cl^{\kappa I})^2}{\cl^{\kappa \kappa} \cl^{I I}}=\frac{(\cl^{\kappa I_{1}}+\cl^{\kappa I_{2}})^2}{\cl^{\kappa \kappa} (\cl^{I_{1} I_{1}}+\cl^{I_{2} I_{2}})}
+\ee
+while for the split survey it will be 
+\be
+\rho_{\rm{split}}^{2}= \frac{(\cl^{\kappa I_{1}})^2}{\cl^{\kappa \kappa} \cl^{I_{1} I_{1}}}+\frac{(\cl^{\kappa I_{2}})^2}{\cl^{\kappa \kappa} \cl^{I_{2} I_{2}}}.
+\ee
+Now it can be show that $\rho_{\rm{split}}\ge \rho_{\rm{full}}$ being:
+\be
+\rho_{\rm{split}}^{2} - \rho_{\rm{full}}^{2}= 
+\ee
+The two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the some for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
+Of course, this is an idealized situation: increasing the number of bins will reduce the number of galaxies in the bin thus increasing the shot noise. 
+An optimal balance needs to be found given the survey specifics, but the general idea is that the optimal approach consists in binning the galaxies in the survey at different redshift. 
+Of course, this is possible for galaxies surveys, but it is harder for a tracer with no redshift information like the CIB.
+
+In this work, we then bin galaxy surveys by splitting the window function \refeq{wg} into different slices so that all the bins contain the same number of galaxies. 
+For photometric surveys like DES and LSST, we assume a photometric redshift estimation gaussianly distributed around the true value with an rms fluctuation $\sigma(z)$.
+
+In that case the i$^{th}$ slice has a galaxy distribution \cite{hu:2004}:
+\begin{eqnarray}
+\label{eqn:zbin}
+W_{i}(z) \propto b(z) {dN(z) \over dz} 
+\Bigl[{\,\rm erfc}\left({\Delta(i-1) -z \over \sigma(z)\sqrt{2}}\right) 
+\\ \nonumber - {\,\rm erfc}
+\left({\Delta i  - z \over \sigma(z)\sqrt{2}}\right)\Bigr].
+\end{eqnarray}
+
+% ===================
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/compare_kernel.pdf}
+\caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
+The bigger the overlap wit the CMB lensing kernel the better the reconstruction of the lensing potential will be.
+This will lead to a better delensing.
+}
+\label{fig:kernel}
+
+\end{center}
+\end{figure}
+% ===================
 
@@
 In this section, we forecast the expected delensing efficiency and the relative importance of galaxies tracers for delensing in current and future experiments.
-We will use the Fisher information matrix to quantify the delensign efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
+We will use the Fisher information matrix to quantify the delensing efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
 We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction efficiency, together with a low-noise, low-resolution experiment which B-modes will be delensed and used to constrain the inflationary parameters.
 We focus on three different scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/cmb_internal.pdf}
+\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
+CMB measurements will rapidly improve providing the most efficient even if biased way to reconstruct the structures lensing the CMB.}
+\end{center}
+\label{fig:corr-internal}
+\end{figure}
 
@@
 The correlation attainable using current experiments is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
+We include a dashed curve that corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$, showing which scales contribute to the $\ell<100$ B-mode power.
 In general, we observe that CMB lensing reconstruction correlate very well at low $\ell$  and it then falls rapidly because of the raise of the reconstruction noise.

### 2017-06-21 Notes/galaxy_delensing.tex

--- 2017-06-11 Notes/DES_rec_phi_paper.tex
@@
 The measurement of the cosmic microwave background (CMB) polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe. 
-A particular component, the so-called B-modes is indeed a direct signature of the presence of gravitational waves in the early universe. However improving the level of noise is not enough. 
-This is even truer if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
+A particular component, the so-called B-modes is indeed a direct signature of a prediction of inflation: the presence of gravitational waves in the early universe.
+However improving the level of noise in future experiments will not be enough.
+%This is even truer if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
+Secondary effects in the low redshift universe will also produce non-primordial B-modes.
+In particular the gravitational interactions of CMB photons with large scale structures will distort the primordial E-modes adding a lensing B-mode component to the primordial signal.
 Removing the lensing component from the measurement of CMB B-modes will be important to constrain the amplitude of the primordial gravitational wave contribution to the signal.
@@
 \label{sec:intro}
-In the standard cosmological paradigm, the universe underwent a period of near-exponential expansion in its early phase;
-this period is called ``cosmic inflation.''
+In the standard cosmological paradigm, in its early phase, the universe underwent a period of near-exponential expansion called "cosmic inflation." 
+So far all the cosmological observations agree with this picture, making it a compelling and elegant description of the Universe initial conditions.
+Despite the experimental effort, other possible explanations are still valid, and a conclusive evidence of inflation is still to be found.
 Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski15}.
-These primordial gravitational waves in the early universe imprint a unique signature on the polarized anisotropies of the CMB.
-For this reason, CMB polarization is a promising dataset to understand the physics of the early universe.
-Furthermore, CMB polarization is usually studied in Fourier space by decomposing the fields into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
-The B-mode polarization, in particular, is a great source of information because, in the standard scenario, the PGW background is the only source of B-modes at the epoch of recombination.
-
-To connect data and theoretical models, the amplitude of this primordial B-mode component is parameterized by the ratio of the amplitudes of the primordial tensor and scalar perturbations, $r$.
-The shape of this component, i.e.
-the amplitude as a function of the angular scales is encoded in the tensor tilt $n_{t}$.
-Most of the inflationary models directly relate the value of r to the energy scale of inflation.
-Measuring or constraining the value of $r$ and $n_t$ will put a tight constraint on the allowed models of inflation.
+This prediction set inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologist are looking for.
+These primordial gravitational waves in the early universe would imprint a unique signature on the polarized anisotropies of the CMB.
+For this reason, CMB polarization is a promising dataset to understand the physics of the early universe and ultimately test inflation.
+In particular, if we can decompose the CMB polarization fields in Fourier space into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
+ In the standard scenario, the B-mode polarization is a clean probe of primordial gravitational waves, because these are the only source of B-modes at the epoch of recombination.
 
 Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational waves background, improving their measurement is a major objective of current and future CMB experiments.
-Even if inflationary B-modes have not been detected yet, a possibly natural value $r \gtrsim 10^{-3}$, should be reachable given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
-
-However reducing the level of noise will not be enough to reach this goal.
-Indeed the observed B modes are not solely due to primordial gravitational waves but also by secondary effects taking place in the late-time low-redshift universe.
+Even if inflationary B-modes have not been detected yet, a natural value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
+However, reducing the level of noise will not be enough to attain this goal.
+Indeed the observed B modes are not solely sourced by early universe physics, but they are also produced by secondary effects taking place in the late-time low-redshift universe. 
 In this work, we will focus on the effects of gravitational interactions with large scale structures (LSS).
@@
 
-This component acts as a source of confusion for searches of the PGW background.
+This component acts as a source of confusion for searches of the primordial gravitational waves background.
 Indeed, the contamination from lensing B modes is already at the level of the instrumental noise of current experiments \citep{bk14}.
 Thus together with the experimental effort to reduce the amount of noise is important to study the effect of the spurious lensing component.
+
 The optimal way to reduce the lensing contributions is to reconstruct the expected realization of lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
 We can delens the observed B-modes combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
-
 Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
-Furthermore, recently it has been performed on CMB temperature data using as LSS tracers the cosmic infrared background in \cite{larsen:2016} and on CMB temperature and polarization data using internal CMB lensing reconstruction \cite{ carron17}.
+Furthermore, recently it has been performed on CMB temperature data using as LSS tracer the cosmic infrared background in \cite{larsen:2016} and on CMB temperature and polarization data using internal CMB lensing reconstruction \cite{ carron17}.
 Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in where $~28\%$ of the lensing power has been removed.
 
-In this paper, we study the possible role of future galaxies surveys as tracers of the lensing potential.
+For future experiments, we need to increase the delensing efficiency by almost a factor of 3 to fully exploit the instrumental capabilities \cite{abazajian:2016}.
+In this paper, we propose and study a possible way: using future galaxies surveys as tracers of the lensing potential in addition to other probes like 21 cm, cosmic infrared background and internal CMB reconstruction.
+Furthermore, we point out how using redshift information trough a clever tomographic binning can improve the delensing efficiency of galaxies survey.
 These will translate into a better reconstruction of the B-modes in the measured patch, and as a consequence, it will improve the constraints on inflation trough delensing.
-We model several actual and future surveys, and after computing the residual B-modes, we forecast the statistical uncertainness on $r$ and $n_T$ for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
+We model several actual and future surveys, and after computing the residual B-modes, we forecast the statistical uncertainties the amplitude and the shape of the inflationary tensor perturbations for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
 
@@
 Furthermore $W^{i}(z)$ is the kernel function of the field $i$ introduced in \refeq{wkernel}.
-We will now introduce the kernels for each of the tracers used in this work.
+We will now describe the kernels for each of the tracers used in this work.
 
@@
 \end{equation}
-Here $\frac{dN}{dz}$ is the number of galaxies observed by the survey as a function of redshift while $b(z)$ is the galaxy bias that connects the amplitude of galaxies overdensity to the one of the dark matter underneath.
+Here $\frac{dN}{dz}$ is the number of galaxies observed by the survey as a function of redshift while $b(z)$ is the galaxy bias that connects the amplitude of galaxy overdensities to the ones of the dark matter underneath.
 When computing the auto-spectrum of the galaxies density, a shot noise term needs to be taken into account. 
 To do so, we add a constant term to the power spectrum equal to the inverse of the number of galaxies per steradians.
- Different galaxy surveys in this work are fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxies density.
+ Different galaxy surveys in this work are then fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxies density.
 We test the delensing efficiency taking into account both current survey like WISE or DES as well as future galaxy probes like DESI and LSST together with 21 cm measurement like SKA.
@@
 
-However, this promising sign of primordial gravitational waves is partially obscured by gravitational interaction with large scale structure that generates CMB B-modes by distorting primordial E-modes.
-At first order, given the convergence field $\kappa= -\frac{1}{2}\nabla^2\phi$ introduced in \refsec{kappaCMB} the B-modes resulting from the lensing of primordial E mode are:
+However, this promising sign of primordial gravitational waves is partially obscured by gravitational interactions with large scale structures that generate CMB B-modes by distorting primordial E-modes.
+At first order, given the convergence field $\kappa= -\frac{1}{2}\nabla^2\phi$ introduced in \refsec{kappaCMB} the B-modes resulting from the lensing of primordial E-modes are:
 \be
@@
 \ee
-As usual we define the power spectrum as:
-\be
-\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l .
-\ee
+%As usual we define the power spectrum as:
+%\be
+%\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l .
+%\ee
 
-From this we get that the power spectrum of the lensing component of the B-modes:
+From this we get the power spectrum of the lensing component of the B-modes:
 \be
@@
 
-The full B-mode power spectrum measured on the sky is given by a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_l^{BB}$:
+The B-mode power spectrum measured on the sky is composed of a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_l^{BB}$ (defined in \refeq{noise}):
 \be
@@
 \ee
-Given the current constraints, the lensing component is a significant source of B-modes that, at large scales, correspond to a white noise source of roughly $5 \mu K$-arcmin.
+The lensing component is a significant source of B-modes that, at large scales, correspond to a white noise source of roughly $5 \mu K$-arcmin.
 This means that it is not only bigger than the allowed inflationary component at scales smaller than several degrees, but it is also comparable with current levels of instrumental noise.
 For this reason, it is critical to characterize and eventually remove it from the data.
-To do so, we can build a template \refeq{blens} of the actual lensing B-modes given a measurement of the E-mode field and the lensing potential $\phi$ in the observed patch.
-While E is measured directly, we ca estimate $\phi$ using "tracers" of the dark matter distribution that creates the potential.
+To do so, we build a template \refeq{blens} of the lensing B-modes in the observed patch given a measurement of the E-mode field and the lensing potential $\phi$.
+While E is measured directly, we can estimate $\phi$ using "tracers" of the dark matter distribution that sources the potential.
 
-We will now show how the delensing efficiency is related to the fidelity of the lensing tracers and instrumental noise in the CMB E-modes.
-If we have a large scale structure field I($\nver$) that traces the lensing potential responsible for the lensing of the CMB we can build a template of the lensing B-modes on the sky by a weighted convolution:
+We will now show how the delensing efficiency is related to the fidelity of the lensing tracers and the instrumental noise in the CMB E-modes.
+If we have a large scale structure field I($\nver$) that traces the lensing potential we can build a template of the lensing B-modes on the sky by a weighted convolution:
 \be
@@
 Here $C^{\kappa I}$ and $C^{II}$ are the cross correlation spectrum of the tracer $I$ with the lensing convergence $\kappa$ and its autospectrum and are described for each LSS field in \refsec{model}.
+The power spectrum of the E-modes noise $N^{EE}$ is the same as the B-modes one in \refeq{noise}.
+
 Notice that the first term consists of an inverse variance filter applied to the measured E-mode and the second minimize the difference between the reconstructed $\phi$ and the CMB lensing potential.
@@
 The bigger the $\rho_l$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
-A higher correlation allows for a better reconstruction of the $\phi^{CMB}$ and, as a consequence, of $B^{lens}$.
+An higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$.
 
@@
 In this section we extend the formalism to the case where multiple tracers are used to reconstruct the lensing potential. 
-Furthermore we show that splitting galaxy surveys into multiple tomographic bins can significantly improve their efficiency.
+%Furthermore we show that splitting galaxy surveys into multiple tomographic bins can significantly improve their efficiency.
 
@@
 \reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
-Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
-However given the significant low noise of this measurement they can still help the delensing of the CMB.
-On the other end, CIB and 21 cm surveys are probing to high redshift structures and independently from the model assumed they show a relatively good overlap with the CMB lensing kernel.
+CIB and 21 cm surveys probes the high redshift structures and, independently from the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
+On the other end, galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
+However, given the low noise of this measurement and their small overlap with other probes they still improve the delensing efficiency even if they overlap with the CMB lensing potential is not optimal.
 
 \subsection{Improving efficiency with tomographic binning}\label{sec:tomo-bin}
+Furthermore the delensing efficiency of galaxy surveys can be improved by taking into account redshift information.
 When we weight a tracer with $\frac{C^{\kappa I}}{C^{II}}$ in \refeq{fweight} to maximize its ability to reconstruct the lensing potential we are only using redshift averaged information about the survey.
 However, as can be seen in \reffig{kernel}, the overlap varies as a function of redshifts.
-For this reason, the optimal approach is to weight each redshift bin with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
+For this reason, the optimal approach is to weight galaxies at different redshifts with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
 We can see this with a simple example. Let's split a single survey I into two non-overlapping redshift bins $I_{1}$ and ${I_{2}}$ with $I=I_{1}+I_{2}$.
@@
 \be
-\rho_{\rm{split}}^{2} - \rho_{\rm{full}}^{2}= 
+\rho_{\rm{split}}^{2} - \rho_{\rm{full}}^{2}\propto  (\cl^{I_{1} I_{1}}\cl^{\kappa I_{2}}-\cl^{I_{2} I_{2}}\cl^{\kappa I_{1}})^2
 \ee
-The two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the some for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
+Then $\rho^2$ is always bigger in the tomographic case and the two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the some for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
 Of course, this is an idealized situation: increasing the number of bins will reduce the number of galaxies in the bin thus increasing the shot noise. 
 An optimal balance needs to be found given the survey specifics, but the general idea is that the optimal approach consists in binning the galaxies in the survey at different redshift. 
-Of course, this is possible for galaxies surveys, but it is harder for a tracer with no redshift information like the CIB.
+Of course, this is possible for galaxies surveys, but it is harder for a tracer with poor redshift information like the CIB.
 
@@
 \end{eqnarray}
+
+\textbf{ .. add binning used ...}
 
@@
 \label{eqn:fisher}
-\sigma(p)= \left[\sum_l^{\ell_{\rm{max}}} \left(\frac{\partial {C}_{\ell}^{BB,r,n_{t}}}{\partial p}\right)^2/\mathrm{Cov}^2( C_{\ell}^{BB} ) \right]^{-\frac{1}{2}}
+\sigma(p)= \left[\sum_l^{\ell_{\rm{max}}} \left(\frac{\partial {C}_{\ell}^{BB,r,n_{t}}}{\partial p}\right)^2/\sigma( C_{l}^{BB} )^2 \right]^{-\frac{1}{2}}
 \end{equation}
@@
 Here the B-modes noise spectrum is given by:
-\beq
+\beq \label{eqn:noise}
 N_l^{BB} = \left({\Delta_P}/{T_{\mathrm{CMB}}}\right)^2  e^{ {l^2 \theta_\mathrm{FWHM}^2}/({8 \ln 2})}
@@
 \bibliographystyle{plain}
-\bibliography{DES_rec_phi_paper,/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
+\bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
 

### 2017-06-23 Notes/galaxy_delensing.tex
text Jenny suggestions

@@
 Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski15}.
-This prediction set inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologist are looking for.
+This prediction set inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologists are looking for.
 These primordial gravitational waves in the early universe would imprint a unique signature on the polarized anisotropies of the CMB.
@@
 For future experiments, we need to increase the delensing efficiency by almost a factor of 3 to fully exploit the instrumental capabilities \cite{abazajian:2016}.
-In this paper, we propose and study a possible way: using future galaxies surveys as tracers of the lensing potential in addition to other probes like 21 cm, cosmic infrared background and internal CMB reconstruction.
-Furthermore, we point out how using redshift information trough a clever tomographic binning can improve the delensing efficiency of galaxies survey.
+In this paper, we propose and study a possible way: using future galaxies surveys as tracers of the lensing potential in addition to other probes like SKA radio continuum survey, cosmic infrared background and internal CMB reconstruction.
+Furthermore, we point out how using redshift information through a clever tomographic binning can improve the delensing efficiency of galaxies survey.
 These will translate into a better reconstruction of the B-modes in the measured patch, and as a consequence, it will improve the constraints on inflation trough delensing.
@@
 We organize this article as follow: we describe the LSS tracers used in this analysis in \refsec{model}.
-In \refsec{th} we define the lensing B-mode component and the residual power after delensing with tracers of the lensing potential.
-\refsec{for} is the main results of this work: the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
+In \refsec{th} we define the lensing B-mode component and the residual power after delensing with tracers of the lensing potential.The main results of this work is \refsec{for} : the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
 We conclude in \refsec{concl}.
@@
 \end{equation}
-Here $\frac{dN}{dz}$ is the number of galaxies observed by the survey as a function of redshift while $b(z)$ is the galaxy bias that connects the amplitude of galaxy overdensities to the ones of the dark matter underneath.
+Here $\frac{dN}{dz}$ is the number of galaxies observed by the survey as a function of redshift while $b(z)$ is the galaxy bias that connects the amplitude of galaxy overdensities to the one of the dark matter density underneath.
 When computing the auto-spectrum of the galaxies density, a shot noise term needs to be taken into account. 
@@
 
-We defined the WISE redshift distribution following \cite{yan:2013} (see Fig. 4 therein). 
+The WISE survey observed the entire sky in the infrared \cite{wright:2010}.
+We defined the redshift distribution of the WISE infrared galaxy samples following \cite{yan:2013} (see Fig. 4 therein). 
 To compute the noise term, we assume that the available sky after masking is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxies density is approximately uniform.
+Furthermore we adopt a liner bias $b_{\rm{WISE}}=1.41$ obtained by \cite{ferraro:2015} cross correlating WISE with Planck lensing potential.
+
 DES is modeled after the DES Science Verification data public release.
-For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report.
+For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy denisty of 0.63 galaxies per arcmin squared.
+
 For LSST we follow \cite{schaan:2016}: $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
 Furthermore we assume a density of 26 galaxies per arcmin squared.
-Finally, we model both the SKA redshift distribution and bias survey following \cite{namikawa:2016a}.
+Finally, we model both the redshift distribution and bias for the Radio Continuum (RC) survey conducted by the Square Kilometre Array (SKA) following \cite{namikawa:2016a}.
 
@@
 where $f(\bl,\bl')$ can be determined by minimizing the difference with the true $B^{\mathrm{lens}}(\bl)$ defined in \refeq{blens}.
-We introduce the instrumental noise in the CMB E-modes ($E^N$) that will also limits the ability to fully reconstruct the lensing B-modes.
+We introduce the instrumental noise in the CMB E-modes ($E^N$) that will also limit the ability to fully reconstruct the lensing B-modes.
 
@@
 \ee
-Here $C^{\kappa I}$ and $C^{II}$ are the cross correlation spectrum of the tracer $I$ with the lensing convergence $\kappa$ and its autospectrum and are described for each LSS field in \refsec{model}.
+Here $C^{\kappa I}$ and $C^{II}$ are the cross correlation spectrum of the tracer $I$ with the lensing convergence $\kappa$ and its autospectrum; they are described for each LSS field in \refsec{model}.
 The power spectrum of the E-modes noise $N^{EE}$ is the same as the B-modes one in \refeq{noise}.
@@
 \ee
-Then $\rho^2$ is always bigger in the tomographic case and the two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the some for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
+Then $\rho^2$ is always bigger in the tomographic case, and the two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the same for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
 Of course, this is an idealized situation: increasing the number of bins will reduce the number of galaxies in the bin thus increasing the shot noise. 
 An optimal balance needs to be found given the survey specifics, but the general idea is that the optimal approach consists in binning the galaxies in the survey at different redshift. 
-Of course, this is possible for galaxies surveys, but it is harder for a tracer with poor redshift information like the CIB.
+Of course, this is possible for galaxies surveys, but it is harder for a tracer with poor redshift information like the CIB or  radio continuum surveys.
 
@@
 We will use the Fisher information matrix to quantify the delensing efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
-We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction efficiency, together with a low-noise, low-resolution experiment which B-modes will be delensed and used to constrain the inflationary parameters.
+We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction efficiency, together with a low-noise, low-resolution experiment the B-modes of which will be delensed and used to constrain the inflationary parameters.
 We focus on three different scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
@@
 The improvement due to WISE is quite marginal, but given the big sky coverage, it might be useful to delens large CMB polarization experiments.
-For reference we also test a scenario where GW are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.
+For reference we also test a scenario where primordial gravitational waves are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.
 As expected the importance of delensing itself is now reduced given that the lensing component constitutes a smaller portion of the total B-modes variance.
@@
 The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
-Not only will CMB experiments improved shortly: eventually  DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
+Not only will CMB experiments improve shortly: eventually  DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
 DESI will be slightly more efficient than DES-like survey because of the slightly higher redshift that it can probe.
@@
 \section{Acknowledgments}
-We thank S. Dodelson K. Wu and K. Story for useful discussion.
+We thank S. Dodelson K. Wu and K. Story for useful discussion. Furthermore we thank D. Alonso for his feedback on SKA-like surveys.
 This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.

### 2017-06-26 Notes/galaxy_delensing.tex
fisher text and references

@@
 
-\title{Future Cosmic Microwave Background delensing with galaxies surveys.}         % Enter your title between curly braces
+\title{Future Cosmic Microwave Background delensing with galaxy surveys.}         % Enter your title between curly braces
 \author{A. Manzotti}        % Enter your name between curly braces
@@
 
-The measurement of the cosmic microwave background (CMB) polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe. 
-A particular component, the so-called B-modes is indeed a direct signature of a prediction of inflation: the presence of gravitational waves in the early universe.
-However improving the level of noise in future experiments will not be enough.
+The cosmic microwave background (CMB) polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe. 
+A particular component, the so-called B-modes, is indeed a direct signature of a prediction of inflation: the presence of gravitational waves in the early universe.
+However, improving the level of noise in future experiments will not be enough.
 %This is even truer if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
 Secondary effects in the low redshift universe will also produce non-primordial B-modes.
-In particular the gravitational interactions of CMB photons with large scale structures will distort the primordial E-modes adding a lensing B-mode component to the primordial signal.
+In particular, the gravitational interactions of CMB photons with large scale structures will distort the primordial E-modes adding a lensing B-mode component to the primordial signal.
 Removing the lensing component from the measurement of CMB B-modes will be important to constrain the amplitude of the primordial gravitational wave contribution to the signal.
@@
 \label{sec:intro}
-In the standard cosmological paradigm, in its early phase, the universe underwent a period of near-exponential expansion called "cosmic inflation." 
-So far all the cosmological observations agree with this picture, making it a compelling and elegant description of the Universe initial conditions.
+In the standard cosmological paradigm the early universe underwent a period of near-exponential expansion called "cosmic inflation." 
+All the cosmological observations agree with this picture, making it a compelling and elegant description of the Universe initial conditions.
 Despite the experimental effort, other possible explanations are still valid, and a conclusive evidence of inflation is still to be found.
@@
 For this reason, CMB polarization is a promising dataset to understand the physics of the early universe and ultimately test inflation.
-In particular, if we can decompose the CMB polarization fields in Fourier space into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
+In particular, we can decompose the CMB polarization fields in Fourier space into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
  In the standard scenario, the B-mode polarization is a clean probe of primordial gravitational waves, because these are the only source of B-modes at the epoch of recombination.
@@
 Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational waves background, improving their measurement is a major objective of current and future CMB experiments.
-Even if inflationary B-modes have not been detected yet, a natural value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
-However, reducing the level of noise will not be enough to attain this goal.
-Indeed the observed B modes are not solely sourced by early universe physics, but they are also produced by secondary effects taking place in the late-time low-redshift universe. 
-In this work, we will focus on the effects of gravitational interactions with large scale structures (LSS).
+Even if inflationary B-modes have not been detected yet, a natural value $r \gtrsim 10^{-3}$ should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
+However, just reducing the level of noise will not be enough to attain this goal.
+Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe. 
+In this work, we will focus on the effect of gravitational interactions with large scale structures (LSS).
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
@@
 
-This component acts as a source of confusion for searches of the primordial gravitational waves background.
+This component acts as a source of confusion for searches of the primordial gravitational wave background.
 Indeed, the contamination from lensing B modes is already at the level of the instrumental noise of current experiments \citep{bk14}.
-Thus together with the experimental effort to reduce the amount of noise is important to study the effect of the spurious lensing component.
+Thus, together with the experimental effort to reduce the amount of noise, the effect of the spurious lensing component must be understood.
 
-The optimal way to reduce the lensing contributions is to reconstruct the expected realization of lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
-We can delens the observed B-modes combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
+The optimal way to reduce the lensing contributions to the B-modes is to reconstruct the expected realization of lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
+We can delens the observed B-modes by combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
 Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
-Furthermore, recently it has been performed on CMB temperature data using as LSS tracer the cosmic infrared background in \cite{larsen:2016} and on CMB temperature and polarization data using internal CMB lensing reconstruction \cite{ carron17}.
-Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in where $~28\%$ of the lensing power has been removed.
+Furthermore, it has recently been performed on CMB temperature data using the cosmic infrared background as LSS tracer \cite{larsen:2016} and on CMB temperature and polarization data using CMB data to internally reconstruct the LSS lensing potential \cite{ carron17}.
+Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in which $~28\%$ of the lensing power was removed.
 
-For future experiments, we need to increase the delensing efficiency by almost a factor of 3 to fully exploit the instrumental capabilities \cite{abazajian:2016}.
-In this paper, we propose and study a possible way: using future galaxies surveys as tracers of the lensing potential in addition to other probes like SKA radio continuum survey, cosmic infrared background and internal CMB reconstruction.
-Furthermore, we point out how using redshift information through a clever tomographic binning can improve the delensing efficiency of galaxies survey.
-These will translate into a better reconstruction of the B-modes in the measured patch, and as a consequence, it will improve the constraints on inflation trough delensing.
-We model several actual and future surveys, and after computing the residual B-modes, we forecast the statistical uncertainties the amplitude and the shape of the inflationary tensor perturbations for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
+For future experiments, we need to increase the delensing efficiency by almost a factor of 3 to fully exploit the expected instrumental capabilities \cite{abazajian:2016}.
+In this paper, we propose and study a possible way: using future galaxy surveys as tracers of the lensing potential in addition to other probes such as the SKA radio continuum survey, cosmic infrared background and internal CMB reconstruction.
+Furthermore, we point out how using redshift information through tomographic binning can improve the delensing efficiency of galaxy survey.
+These will translate into a better reconstruction of the B-modes in the measured patch, and, as a consequence will improve the constraints on inflation trough delensing.
+We model several actual and future surveys, and after computing the residual B-modes, we forecast the resulting statistical uncertainties on the amplitude and the shape of the inflationary tensor perturbations for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
 
 We organize this article as follow: we describe the LSS tracers used in this analysis in \refsec{model}.
-In \refsec{th} we define the lensing B-mode component and the residual power after delensing with tracers of the lensing potential.The main results of this work is \refsec{for} : the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
+In \refsec{th} we define the lensing B-mode component and the residual power after delensing with tracers of the lensing potential.
+The main result of this work is \refsec{for} : the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
 We conclude in \refsec{concl}.
@@
 \label{sec:model}
-In this section, we introduce the different large scale structure tracers considered in this work to reconstruct the lensing potential that distorts primordial E-modes generating a non-primordial B-mode component.
+In this section, we introduce the different large scale structure tracers considered in this work to reconstruct the lensing potential.
+Gravitational lensing distorts primordial E-modes generating a non-primordial B-mode component.
 Also, we define the power spectra that we will use later in \refsec{th}.
@@
 \eeq
-In this work, the level of noise is computed assuming an iterative quadratic estimator is used in the reconstruction \cite{}.
+In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}.
 
@@
 
-The galaxies clustering kernel is:
+The galaxy clustering kernel is:
 \begin{equation}
@@
 \end{equation}
-Here $\frac{dN}{dz}$ is the number of galaxies observed by the survey as a function of redshift while $b(z)$ is the galaxy bias that connects the amplitude of galaxy overdensities to the one of the dark matter density underneath.
-When computing the auto-spectrum of the galaxies density, a shot noise term needs to be taken into account. 
+Here $\frac{dN}{dz}$ is the number of galaxies observed by the survey as a function of redshift while $b(z)$ is the galaxy bias that connects the amplitude of galaxy overdensities to the underlying dark matter density.
+When computing the auto-spectrum of the galaxy density, a shot noise term needs to be taken into account. 
 To do so, we add a constant term to the power spectrum equal to the inverse of the number of galaxies per steradians.
- Different galaxy surveys in this work are then fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxies density.
-We test the delensing efficiency taking into account both current survey like WISE or DES as well as future galaxy probes like DESI and LSST together with 21 cm measurement like SKA.
+ Different galaxy surveys in this work are then fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxy density.
+We test the delensing efficiency taking into account both current surveys like WISE or DES as well as future galaxy surveys like DESI and LSST together with 21 cm measurement like SKA.
 
@@
 We defined the redshift distribution of the WISE infrared galaxy samples following \cite{yan:2013} (see Fig. 4 therein). 
-To compute the noise term, we assume that the available sky after masking is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxies density is approximately uniform.
-Furthermore we adopt a liner bias $b_{\rm{WISE}}=1.41$ obtained by \cite{ferraro:2015} cross correlating WISE with Planck lensing potential.
+To compute the noise term, we assume that the available sky after masking is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxy density is approximately uniform.
+Furthermore we adopt a linear bias $b_{\rm{WISE}}=1.41$ obtained by \cite{ferraro:2015} cross correlating WISE with Planck lensing potential.
 
-DES is modeled after the DES Science Verification data public release.
+DES is modeled after the DES Science Verification public data release.
 For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy denisty of 0.63 galaxies per arcmin squared. 
+%As in \cite{passaglia:2017} we use a linear bias $b$
 
@@
 Furthermore we assume a density of 26 galaxies per arcmin squared.
-Finally, we model both the redshift distribution and bias for the Radio Continuum (RC) survey conducted by the Square Kilometre Array (SKA) following \cite{namikawa:2016a}.
+
+Finally, we consider the Square Kilometre Array (SKA).
+The SKA is a planned radio array that will survey large scale structure primarily by detecting the redshifted neutral hydrogen (HI) 21cm emission line from a large number of galaxies out to high redshift.
+We will take into account the intensity mapping survey mode where SKA will measure the large-scale fluctuations of the \textit{integrated} 21cm intensity from many unresolved galaxies.  
+We model both the redshift distribution and  bias of radio sources following \cite{namikawa:2016a}.
+
+%bias for the Radio Continuum (RC) survey conducted by SKA
 
@@
 \label{sec:cib}
+
+The CIB consists of diffuse extragalactic radiation generated by the unresolved emission from star-forming galaxies (see \cite{dole:2006} and references therein). 
+In these galaxies, the UV light from young stars heats the dust regions around them that then reradiates thermally in the infrared with a graybody spectrum of $T \simeq30$K.
+
 Following \cite{Addison:2011se}, we model the CIB power directly as $C_{\ell}^{{\rm CIB}\mbox{-}{\rm CIB}} = 3500 (l/3000)^{-1.25} {\rm Jy^2 / sr}$.
 This model provides an accurate fit to several experimental results.
-For the cross-spectra with the CMB lensing or other galaxies tracers, $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall10}.
+For the cross-spectra with the CMB lensing or other galaxy tracers, $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall10}.
 It corresponds to the kernel:
@@
 
-We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$. Also $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
+We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$ and we set $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
 
@@
 
-The weights $f(\bl,\bl')$ so that the residual lensing B mode power is minimized are:
+The optimal weights $f(\bl,\bl')$ such that the residual lensing B mode power is minimized are:
 \be
@@
 
-Notice that the first term consists of an inverse variance filter applied to the measured E-mode and the second minimize the difference between the reconstructed $\phi$ and the CMB lensing potential.
+Notice that the first term consists of an inverse variance filter applied to the measured E-mode and the second minimizes the difference between the reconstructed $\phi$ and the CMB lensing potential.
 
@@
 
-Let's start by assuming that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
+We start by assuming that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
 We can optimally combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ with:
@@
 CIB and 21 cm surveys probes the high redshift structures and, independently from the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
-On the other end, galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
-However, given the low noise of this measurement and their small overlap with other probes they still improve the delensing efficiency even if they overlap with the CMB lensing potential is not optimal.
+On the other end, galaxy clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
+However, given the low noise of this measurement and their small overlap with other probes they still improve the delensing efficiency even if their overlap with the CMB lensing potential is not optimal.
 
@@
 An optimal balance needs to be found given the survey specifics, but the general idea is that the optimal approach consists in binning the galaxies in the survey at different redshift. 
-Of course, this is possible for galaxies surveys, but it is harder for a tracer with poor redshift information like the CIB or  radio continuum surveys.
+Of course, this is possible for galaxy surveys, but it is harder for a tracer with poor redshift information like the CIB or  radio continuum surveys.
 
@@
 
-\textbf{ .. add binning used ...}
+We used 10 and 4 photometric bins for LSST and DES respectively and we split DESI into 4 spectroscopic bins with no overlap among each others.
 
@@
 
+% ===================
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/compare_kernel_tomo.pdf}
+\caption{\textbf{Kernels Comparison with tomographic bins}: Comparison of the 10 LSST different kernels for every tomographic bin with the CMB lensing kernel. 
+This will lead to a better delensing. }
+\label{fig:kernel}
+
+\end{center}
+\end{figure}
+% ===================
+
 %=================
@@
 \label{sec:for}
-In this section, we forecast the expected delensing efficiency and the relative importance of galaxies tracers for delensing in current and future experiments.
+
+In this section, we forecast the expected delensing efficiency and the relative importance of galaxy tracers for delensing in current and future experiments.
 We will use the Fisher information matrix to quantify the delensing efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
-We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction efficiency, together with a low-noise, low-resolution experiment the B-modes of which will be delensed and used to constrain the inflationary parameters.
-We focus on three different scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
+We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction performance, together with a low-noise, low-resolution experiment the B-modes of which will be delensed and used to constrain the inflationary parameters.
+We focus on three distinct scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
 
@@
 \caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
-CMB measurements will rapidly improve providing the most efficient even if biased way to reconstruct the structures lensing the CMB.}
+CMB measurements will rapidly improve providing the most efficient (even if biased) way to reconstruct the structures lensing the CMB.}
 \end{center}
@@
 \subsection{Fisher method}
-In this framework, the error on a cosmological parameter constrained by a B-mode spectrum measurement is given as usual by:
-\begin{equation}
-\label{eqn:fisher}
-\sigma(p)= \left[\sum_l^{\ell_{\rm{max}}} \left(\frac{\partial {C}_{\ell}^{BB,r,n_{t}}}{\partial p}\right)^2/\sigma( C_{l}^{BB} )^2 \right]^{-\frac{1}{2}}
-\end{equation}
-where we consider the parameters $p =\{r,n_{t}\}$. 
-We will assume a gaussian covariance:
-\beq
+In the Fisher Information Matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtain from the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
+We constrain the inflationary parameters $p =\{r,n_{t}\}$ with a CMB B-mode spectrum measurement so the Fisher matrix is:
+\beq\label{eqn:fisher}
+F_{ij} = \sum^{\ell_{max}}_{\ell=\ell_{min}}\frac{1}{\sigma( C_{\ell}^{BB} )^2}\frac{\partial {C}_{\ell}^{BB}}{\partial p_{i}}\frac{\partial {C}_{\ell}^{BB}}{\partial p_{j}}
+\eeq
+where we assume a Gaussian covariance:
+\beq\label{eqn:cova}
 \sigma( C_{l}^{BB} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_l^{BB,r} + C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
 \eeq
-Here the B-modes noise spectrum is given by:
+The B-modes noise spectrum is given by \cite{knox:1995}:
 \beq \label{eqn:noise}
@@
 \eeq
-where $\theta_\mathrm{FWHM}$ is the full half width of the telescope beam and $\Delta_P$ is the instrumental noise of the experiment.
-Even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012} this approximation is good enough to show the improvement due to delensing.
+where $\theta_\mathrm{FWHM}$ is the full half width of the telescope beam, and $\Delta_P$ is the instrumental noise of the experiment.
+Even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}, this approximation is good enough to show the improvement due to delensing.
 
-The reason why removing the lensing contribution will improve parameters constraints can be seen from \refeq{fisher}.The parameter uncertainties are inversely proportional to the covariance of the measurement.
-Since lensing B-mode $C_{{l}}^{BB,\mathrm{lens}}$ is a substantial component of the covariance, removing part of them trough delensing will improve parameter constraints.
+It can be seen in \refeq{fisher} that removing the lensing contribution will improve parameters constraints.
+The parameter uncertainties are inversely proportional to the covariance of the measurement.
+Since the lensing B-modes $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them trough delensing will reduce the parameter statistical error.
+After delensing the error can $\sigma^{\rm{del}}(p)$ can still be obtained from \refeq{fisher} with $C_{{l}}^{BB,\mathrm{res}}<C_{{l}}^{BB,\mathrm{lens}$ in the covariance defined in \refeq{cova}.
+We define the improvement as the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$.
 
-Note that we are fixing the other cosmological parameters in \refeq{fisher}.
+In \refeq{fisher} we are making two important assumptions.
+Firstly we are fixing the all the other cosmological parameters.
 Uncertainties in those will however propagate to uncertainties in $\{r,n_{t}\}$.
 While the constraints will be slightly optimistic, this has no real impact on the estimate of the improvement due to delensing which is the main focus of this work.
-We define that improvement as the ration of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$
+Furthermore we are not considering the uncertainties on galaxy surveys internal parameters such as biases, source distributions and photometric redshift uncertainties.
+These parameters do have a significant impact in degrading inflationary constraints \cite{sherwin:2015,namikawa:2016a}.
+However as shown in \cite{sherwin:2015,namikawa:2016a} these can be auto-calibrated i.e. thy can be tightly constrained using galaxy surveys auto spectra and cross correlation.
+
 %===================
@@
 
-Furthermore, we test how much not only CIB maps but also current low redshift galaxies surveys like DES and WISE can improve the delensing efficiency.
+Furthermore, we test how much not only CIB maps but also current low redshift galaxy surveys like DES and WISE can improve the delensing efficiency.
 
@@
 \includegraphics[scale=1.]{../images/actual_scenario_planck.pdf}
-\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the angular scale $\ell$. 
+\caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the angular scale $\ell$. 
 The dashed curve corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$, showing which scales contribute to the $\ell<100$ B-mode power.}
@@
 \includegraphics[scale=1.]{../images/actual_scenario.pdf}
-\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
+\caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
 \label{fig:corrnow}
@@
 
-\subsection{Bias uncertainties degradation}
-The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of the inflationary constraint of delensed spectra.
-In this section, we quantify this effect.
-We will now marginalize over unknown galaxies parameters, but we will use a full dataset of CMB and galaxies data.
-The idea is that as shown in the low level of noise in galaxies surveys might allow us to calibrate them internally.
-We will use a Fisher approach the Fisher matrix is:
-\beqn
-F_{pq} &=& \sum_{l_{a} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}  \sum_{l_{b} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}   \frac{\partial {C}_{l_a}^{BB,\mathrm{del}}}{\partial \theta_p } \left[ \mathrm{Cov}^{BB,BB}\right]_{l_a, l_b} ^{-1} \frac{\partial {C}_{l_b}^{BB,\mathrm{del}}}{\partial \theta_q }\nonumber \\&+&  \sum_j \frac{\frac{\partial {C}_j^{\kappa I}}{\partial \theta_p } \frac{\partial {C}_j^{\kappa I}}{\partial \theta_q }}{(\Delta C_j^{\kappa I})^2}
-+\sum_j \frac{\frac{\partial {C}_j^{I I}}{\partial \theta_p } \frac{\partial {C}_j^{I I}}{\partial \theta_q }}{(\Delta C_j^{I  I})^2}
-\eeqn
-
-\beq
-\alpha_{\mathrm{marginalized}} = \sigma_0(r) / \sigma_{\mathrm{marginalized/delensed}}(r)
-\eeq
-
-where the parameter array contains both the tensor to scalar rate $\theta=r$, and the galaxies surveys parameters like the bias $b_i$ or $p_i$ \footnote{In our analysis, since the fiducial value of $r$ is zero,
-the derivative of $C_\ell^{\rm BB,res}$ is non-zero if $\theta_i=r$.}.
-
-We compute the derivatives of the power spectra as described
+%\subsection{Bias uncertainties degradation}
+%The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of the inflationary constraint of delensed spectra.
+%In this section, we quantify this effect.
+%We will now marginalize over unknown galaxies parameters, but we will use a full dataset of CMB and galaxies data.
+The idea is that as shown in the low level of noise in galaxy surveys might allow us to calibrate them internally.
+%We will use a Fisher approach the Fisher matrix is:
+%\beqn
+%F_{pq} &=& \sum_{l_{a} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}  \sum_{l_{b} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}   \frac{\partial {C}_{l_a}^{BB,\mathrm{del}}}{\partial \theta_p } \left[ \mathrm{Cov}^{BB,BB}\right]_{l_a, l_b} ^{-1} \frac{\partial {C}_{l_b}^{BB,\mathrm{del}}}{\partial \theta_q }\nonumber \\&+&  \sum_j \frac{\frac{\partial {C}_j^{\kappa I}}{\partial \theta_p } \frac{\partial {C}_j^{\kappa I}}{\partial \theta_q }}{(\Delta C_j^{\kappa I})^2}
+%+\sum_j \frac{\frac{\partial {C}_j^{I I}}{\partial \theta_p } \frac{\partial {C}_j^{I I}}{\partial \theta_q }}{(\Delta C_j^{I  I})^2}
+%\eeqn
+%
+%\beq
+%\alpha_{\mathrm{marginalized}} = \sigma_0(r) / \sigma_{\mathrm{marginalized/delensed}}(r)
+%\eeq
+%
+%
+%where the parameter array contains both the tensor to scalar rate $\theta=r$, and the galaxy surveys parameters like the bias $b_i$ or $p_i$ \footnote{In our analysis, since the fiducial value of $r$ is zero,
+%the derivative of $C_\ell^{\rm BB,res}$ is non-zero if $\theta_i=r$.}.
+%
+%We compute the derivatives of the power spectra as described
 
@@
 \section{Acknowledgments}
-We thank S. Dodelson K. Wu and K. Story for useful discussion. Furthermore we thank D. Alonso for his feedback on SKA-like surveys.
+We thank S. Dodelson S. Passaglia K. Wu and K. Story for useful discussion and their review of the draft. Furthermore we thank D. Alonso for his feedback on radio continuum surveys.
 This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.

### 2017-06-29 Notes/galaxy_delensing.tex
fisher text and references

@@
 \def\a{\alpha}
-\def\cl{C_{l}}
+\def\cl{C_{\ell}}
 \def\h{\mathrm{h}}
@@
 Furthermore, we point out how using redshift information through tomographic binning can improve the delensing efficiency of galaxy survey.
-These will translate into a better reconstruction of the B-modes in the measured patch, and, as a consequence will improve the constraints on inflation trough delensing.
+These will translate into a better reconstruction of the B-modes in the measured patch, and, as a consequence will improve the constraints on inflation through delensing.
 We model several actual and future surveys, and after computing the residual B-modes, we forecast the resulting statistical uncertainties on the amplitude and the shape of the inflationary tensor perturbations for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
@@
 However, when computing the power spectrum of the latter, we need to add a noise component.
-Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_l^{\kappa \kappa}$, and so
+Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_{\ell}^{\kappa \kappa}$, and so
 \beq
-\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_l
+\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_{\ell}
 \eeq
@@
 \be
-C_l^{BB,\mathrm{lens}}  = \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl')C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}.
+C_{\ell}^{BB,\mathrm{lens}}  = \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl')C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}.
 \ee
 
-The B-mode power spectrum measured on the sky is composed of a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_l^{BB}$ (defined in \refeq{noise}):
+The B-mode power spectrum measured on the sky is composed of a possible primordial component $C_{\ell}^{BB,r}$ together with the lensing $C_{\ell}^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_{\ell}^{BB}$ (defined in \refeq{noise}):
 \be
-C_l^{BB,\mathrm{full}} = C_l^{BB,r} + C_l^{BB,\mathrm{lens}} + N_l^{BB}.
+C_{\ell}^{BB,\mathrm{full}} = C_{\ell}^{BB,r} + C_{\ell}^{BB,\mathrm{lens}} + N_{\ell}^{BB}.
 \ee
@@
 \ee
-Here $C^{\kappa I}$ and $C^{II}$ are the cross correlation spectrum of the tracer $I$ with the lensing convergence $\kappa$ and its autospectrum; they are described for each LSS field in \refsec{model}.
+Here $C^{\kappa I}$ and $C^{II}$ are the cross-correlation spectrum of the tracer $I$ with the lensing convergence $\kappa$ and its autospectrum; they are described for each LSS field in \refsec{model}.
 The power spectrum of the E-modes noise $N^{EE}$ is the same as the B-modes one in \refeq{noise}.
@@
 with
-\be
-\rho_l= \frac{\cl^{\kappa I}}{\sqrt{\cl^{\kappa \kappa} \cl^{I I}}}.
+\be\label{eqn:rho1}
+\rho_{\ell}= \frac{\cl^{\kappa I}}{\sqrt{\cl^{\kappa \kappa} \cl^{I I}}}.
 \ee
-The bigger the $\rho_l$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
+The bigger the $\rho_{\ell}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
 An higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/BB_res.pdf}
+\caption{}
+\label{fig:bb_res}
+\end{center}
+\end{figure}
 
@@
 \be
-\rho^{2} = \sum_{i,j}\frac{C^{\kappa i}~(C^{-1})_{ij}~C^{\kappa j}}{C^{\kappa\kappa}}.
+\rho^{2}_{\ell} = \sum_{i,j}\frac{C^{\kappa i}_{\ell}~(C^{-1}_{\ell})_{ij}~C^{\kappa j}_{\ell}}{C^{\kappa\kappa}_{\ell}}.
 \label{eqn:rho-combined}
@@
 We can see this with a simple example. Let's split a single survey I into two non-overlapping redshift bins $I_{1}$ and ${I_{2}}$ with $I=I_{1}+I_{2}$.
-For the full survey the effective cross correlation is equal to 
+For the full survey the effective cross-correlation is equal to 
 \be
@@
 \label{fig:kernel}
-
 \end{center}
@@
 % ===================
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/B_res_bin.pdf}
+\caption{Tomographic bins improve the cross-correlation with lensing.
+Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line).}
+\label{fig:bb_res}
+\end{center}
+\end{figure}
 
@@
 We will use the Fisher information matrix to quantify the delensing efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
-We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction performance, together with a low-noise, low-resolution experiment the B-modes of which will be delensed and used to constrain the inflationary parameters.
+We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction performance, together with a low-noise, low-resolution experiment the B-modes of which \textcolor{red}{(?)} will be delensed and used to constrain the inflationary parameters.
 We focus on three distinct scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/cmb_internal.pdf}
-\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
-CMB measurements will rapidly improve providing the most efficient (even if biased) way to reconstruct the structures lensing the CMB.}
-\end{center}
-\label{fig:corr-internal}
-\end{figure}
+%\begin{figure}[htbp]
+%\begin{center}
+%\includegraphics[scale=1.]{../images/cmb_internal.pdf}
+%\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
+%CMB measurements will rapidly improve providing the most efficient (even if biased) way to reconstruct the structures lensing the CMB.}
+%\end{center}
+%\label{fig:corr-internal}
+%\end{figure}
 
-\subsection{Fisher method}
-In the Fisher Information Matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtain from the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
+\subsection{Fisher Information Matrix}
+In the Fisher Information Matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtain the from inverse of the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
 We constrain the inflationary parameters $p =\{r,n_{t}\}$ with a CMB B-mode spectrum measurement so the Fisher matrix is:
@@
 \beq\label{eqn:cova}
-\sigma( C_{l}^{BB} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_l^{BB,r} + C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
+\sigma( C_{\ell}^{BB} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_{\ell}^{BB,r} + C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
 \eeq
@@
 \beq \label{eqn:noise}
-N_l^{BB} = \left({\Delta_P}/{T_{\mathrm{CMB}}}\right)^2  e^{ {l^2 \theta_\mathrm{FWHM}^2}/({8 \ln 2})}
+N_{\ell}^{BB} = \left({\Delta_P}/{T_{\mathrm{CMB}}}\right)^2  e^{ {l^2 \theta_\mathrm{FWHM}^2}/({8 \ln 2})}
 \eeq
 where $\theta_\mathrm{FWHM}$ is the full half width of the telescope beam, and $\Delta_P$ is the instrumental noise of the experiment.
-Even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}, this approximation is good enough to show the improvement due to delensing.
 
@@
 Since the lensing B-modes $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them trough delensing will reduce the parameter statistical error.
-After delensing the error can $\sigma^{\rm{del}}(p)$ can still be obtained from \refeq{fisher} with $C_{{l}}^{BB,\mathrm{res}}<C_{{l}}^{BB,\mathrm{lens}$ in the covariance defined in \refeq{cova}.
+After delensing the error $\sigma^{\rm{del}}(p)$ can still be obtained from \refeq{fisher} with $C_{{l}}^{BB,\mathrm{res}}<C_{{l}}^{BB,\mathrm{lens}}$ in the covariance defined in \refeq{cova}.
 We define the improvement as the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$.
 
-In \refeq{fisher} we are making two important assumptions.
-Firstly we are fixing the all the other cosmological parameters.
-Uncertainties in those will however propagate to uncertainties in $\{r,n_{t}\}$.
-While the constraints will be slightly optimistic, this has no real impact on the estimate of the improvement due to delensing which is the main focus of this work.
-Furthermore we are not considering the uncertainties on galaxy surveys internal parameters such as biases, source distributions and photometric redshift uncertainties.
+In \refeq{fisher} we are making a few important assumptions.
+Firstly, we are fixing all the cosmological parameters except $\{r,n_{t}\}$.
+Uncertainties in those will, however, propagate to uncertainties in $\{r,n_{t}\}$.
+While neglecting this will lead to slightly optimistic constraints, it has no significant impact on the estimate of the improvement due to delensing.
+
+We are also neglecting an important contribution to measured CMB B-modes: Galactic polarized foregrounds. 
+The amplitude of these has been constrained in \cite{bicep2/keck-collaboration:2015, planck-collaboration:2015} and it strongly varies in different parts of the sky.
+Future experiments will use multi-frequency data to exploit the frequency dependency of these contaminants to remove them from the data.
+The amount of residual foregrounds depends both on uncertain foreground properties and experimental choices (see a review in \cite{abazajian:2016}).
+For this reason, accurately treating foreground requires the use of simulations and knowledge about experimental details. 
+We decided to focus on an ideal situation of no foregrounds or perfect cleaning even if the importance of delensing will be overestimated.
+
+Furthermore, we are not considering the uncertainties on galaxy survey internal parameters such as biases, source distributions and photometric redshift uncertainties.
 These parameters do have a significant impact in degrading inflationary constraints \cite{sherwin:2015,namikawa:2016a}.
-However as shown in \cite{sherwin:2015,namikawa:2016a} these can be auto-calibrated i.e. thy can be tightly constrained using galaxy surveys auto spectra and cross correlation.
+However, as shown in \cite{sherwin:2015,namikawa:2016a}, these can be auto-calibrated i.e.
+they can be tightly constrained using galaxy surveys auto spectra and cross-correlation.
+We checked this for a few of the tracers combinations used here and find it particularly true once several tracers are jointly taken into account.
+Finally, we have assumed a Gaussian covariance even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
+This approximation is good enough to show the improvement due to delensing.
 
@@
 
-Very recently delensing has been proven possible on real data using as tracers both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as CIB data \cite{larsen:2016a,manzotti:2017}.
-Here we discuss the improvement due to combining these and other available surveys (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
+Only recently delensing has been performed on data using both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as CIB maps \cite{larsen:2016a,manzotti:2017} as large scale structure tracers.
+Here we discuss the improvement that can be obtained by combining these and other currently available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
 
-On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin} and a beam of \textcolor{red}{30 arcmin}.
-The CMB lensing reconstruction will be performed by a higher resolution CMB experiment. We explore two possibilities.
-First, we test the efficiency with an internal reconstruction performed by Planck \cite{2016A&A...594A..15P} using the CMB lensing noise curves publicly available
+On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} CMB experiment.
+For the deep experiment we use an instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin}, a beam of \textcolor{red}{30 arcmin} and an angular scale range of $50<\ell<500$.
+The CMB lensing reconstruction will be performed by the higher resolution CMB experiment. We explore two possibilities.
+First, we test the efficiency with an internal reconstruction performed by Planck \cite{planck-collaboration:2016a} using the CMB lensing noise curves publicly available
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
-Then we will assume a ground-based experiment with the final noise level of SPT-Pol which corresponds to \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
+Then we assume a ground-based experiment with noise levels consistent with SPT-Pol: \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
+In this case we set the largest achievable scale at $\ell_{min}=300$.
+However, the results are robust against this choice since the neglected scales contribute only moderately to the reconstruction of the CMB lensing potential.
 
-Furthermore, we test how much not only CIB maps but also current low redshift galaxy surveys like DES and WISE can improve the delensing efficiency.
+We combined these CMB experiments with the CIB and current low redshift galaxy surveys like DES and WISE and we compute the improvement in the delensing efficiency.
+Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals.
+Optical surveys galaxies like DES are less affected by dust and can be used at larger scales.
+The achievable correlation is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
+In these figures we include a dashed curve that corresponds to (with arbitrary scale) $<C_{\ell}^{\kappa \kappa} \times \frac{\partial C^{BB}_{\ell'}}{\partial{C^{\kappa \kappa}_{\ell}}}>_{\ell ' <100}$, showing which scales contribute to the $\ell<100$ B-mode power.
 
-The correlation attainable using current experiments is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
-We include a dashed curve that corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$, showing which scales contribute to the $\ell<100$ B-mode power.
-In general, we observe that CMB lensing reconstruction correlate very well at low $\ell$ and it then falls rapidly because of the raise of the reconstruction noise.
-For the noise levels of Planck, the internal reconstruction is at most $70\%$ correlated at very large scales and falls rapidly to $40\%$ at $\ell=200$.
-Its correlation is comparable to the on of LSS at almost all scales.
-On the other end, the CMB lensing reconstruction \cite{} from SPT-Pol will be more than 80\% correlated with the true field at large scales $\ell<300$, where the correlation become comparable with the current LSS surveys.
+In both cases the CMB lensing reconstruction correlates very well at low $\ell$ and it then falls rapidly at smaller scales because of the raise of the reconstruction noise.
+For Planck, the internal reconstruction is at most $70\%$ correlated at very large scales and then it falls rapidly to $40\%$ at $\ell=200$.
+Its correlation is comparable to the one of LSS at almost all scales.
+On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 80\% correlated with the true field at $\ell<300$, and only then the correlation become comparable with the current LSS surveys.
 
-Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals.
-Optical surveys galaxies like DES are less affected by that systematics and can be used at larger scales.
-Furthermore, in \reffig{corrnow} DES contribution to delensing is higher than WISE because the redshift distributions peaks at slightly higher redshift thus overlapping better with lensing.
+\reffig{corrnow} shows that the DES is an effective tracer of the LSS and can, at least in the near future, be used to improve delensing for CMB experiments that overlap with it.
+Indeed its delensing efficiency is higher than WISE because the redshift distributions peaks at slightly higher redshift thus it better overlaps with CMB lensing.
 
@@
 
-%At large angular scale, the CMB internal reconstruction is clearly the best tracers for delensing.
-Because of the large level of noise in current experiments, its efficiency became comparable to the CIB one at few degrees ($\ell \simeq 400$).
-
 \begin{figure}[htbp]
@@
 
-Using this correlation level we can, using \refeq{Bres} compute the residual B-mode power after delensing and test the consequent improvement on parameter constraints with \refeq{fisher}. 
-For current surveys we only consider $r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
+Using these correlation levels we can  compute the residual B-mode power after delensing using \refeq{Bres}, and test the consequent improvement on parameter constraints with \refeq{fisher}. 
+For current surveys we only consider $p=r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
 We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
-The results are summarized in \reftab{current}
+The results are summarized in \reftab{current} and \reftab{current-spt} for the Planck and SPTPol case respectively.
 
 \begin{table}
-\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to ration of the error before and after delensing.}
+\caption{$\alpha(r)$: Improvements on $\sigma(r)$ due to delensing for current generation and Planck lensing reconstruction. Values correspond to ratio of the error before and after delensing.}
   \vspace{0.2cm}
@@
 \end{table}
-
-When the same tracers are used, the results are in agreement with \cite{yu:2017,larsen:2016a}, with the CIB being the predominant source of delensing with an improvement of almost 50\%.
-Planck and DES are essential to reconstruct the largest scales of the lensing potential and as such the play a major role in delensing.
+In agreement with the literature \cite{yu:2017,larsen:2016a}, the CIB is the predominant source of delensing with an improvement of almost 50\%.
+However Planck and DES are essential to reconstruct the largest scales of the lensing potential and as such the play a major role in delensing.
 The improvement due to WISE is quite marginal, but given the big sky coverage, it might be useful to delens large CMB polarization experiments.
-For reference we also test a scenario where primordial gravitational waves are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.
+We also test a scenario where primordial gravitational waves are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.
 As expected the importance of delensing itself is now reduced given that the lensing component constitutes a smaller portion of the total B-modes variance.
 
+\reftab{current-spt} that uses SPTPol instead of Planck for the internal reconstruction shows consistent results.
 As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS. 
 This can be seen in \reftab{current-spt}. 
-In particular optical surveys will rapidly lose the role of filling in large-scale modes and they start, as the CIB, to supplement information about lensing at higher multipoles.
+In particular optical surveys will rapidly lose the role of filling in large-scale modes and they will start to supplement information at higher multipoles.
 
@@
 
+%%%========================
+%%%========================
+
 \subsection{CMB-S3 Era}
@@
 Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
-DESI will be slightly more efficient than DES-like survey because of the slightly higher redshift that it can probe.
-However, the CIB will still play the major role among LSS.
+An interesting results is that DESI will be less efficient than a DES-like survey even if it can probe higher redshift.
+The reason is that, for the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES play a mojor role in the delensing efficiency.
+Even in the near future the CIB will still play the major role among LSS.
 
@@
 
+% ===================
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/errors_summary.pdf}
+\caption{}
+\label{fig:kernel}
+\end{center}
+\end{figure}
+% ===================
+
 \section{Conclusions}

### 2017-06-29 Notes/galaxy_delensing.tex
fisher text and references

@@
 However, improving the level of noise in future experiments will not be enough.
-%This is even truer if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
 Secondary effects in the low redshift universe will also produce non-primordial B-modes.
@@
 We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
-We find that current generation experiments... we also point out that a tomographic approach..
+We find that current generation experiments  we also point out that a tomographic approach..
 \end{abstract}
@@
 Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_{\ell}^{\kappa \kappa}$, and so
-\beq
+\beq\label{eqn:noisekappa}
 \cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_{\ell}
@@
 Then $\rho^2$ is always bigger in the tomographic case, and the two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the same for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
-Of course, this is an idealized situation: increasing the number of bins will reduce the number of galaxies in the bin thus increasing the shot noise. 
-An optimal balance needs to be found given the survey specifics, but the general idea is that the optimal approach consists in binning the galaxies in the survey at different redshift. 
-Of course, this is possible for galaxy surveys, but it is harder for a tracer with poor redshift information like the CIB or  radio continuum surveys.
+For photometric surveys, the maximum number of bins is dictated by the fact that the bin width can not be smaller than the photo-z accuracy. 
+Of course, binning is possible for galaxy surveys, but it is harder for a tracer with poor redshift information like the CIB or radio continuum surveys.
 
@@
 
-We are also neglecting an important contribution to measured CMB B-modes: Galactic polarized foregrounds. 
+We are also neglecting an important contribution to the measured CMB B-modes: galactic polarized foregrounds. 
 The amplitude of these has been constrained in \cite{bicep2/keck-collaboration:2015, planck-collaboration:2015} and it strongly varies in different parts of the sky.
-Future experiments will use multi-frequency data to exploit the frequency dependency of these contaminants to remove them from the data.
+Future experiments will use multi-band data to exploit the frequency dependence of these contaminants to remove them from the data.
 The amount of residual foregrounds depends both on uncertain foreground properties and experimental choices (see a review in \cite{abazajian:2016}).
-For this reason, accurately treating foreground requires the use of simulations and knowledge about experimental details. 
-We decided to focus on an ideal situation of no foregrounds or perfect cleaning even if the importance of delensing will be overestimated.
+For this reason, accurately treating foreground requires the use of simulations and the knowledge of several experimental details. 
+We decided to focus on an ideal situation assuming no foregrounds or perfect cleaning even if the importance of delensing will be slightly overestimated.
 
 Furthermore, we are not considering the uncertainties on galaxy survey internal parameters such as biases, source distributions and photometric redshift uncertainties.
-These parameters do have a significant impact in degrading inflationary constraints \cite{sherwin:2015,namikawa:2016a}.
+These uncertaintes can significantly degrade inflationary constraints \cite{sherwin:2015,namikawa:2016a}.
 However, as shown in \cite{sherwin:2015,namikawa:2016a}, these can be auto-calibrated i.e.
-they can be tightly constrained using galaxy surveys auto spectra and cross-correlation.
-We checked this for a few of the tracers combinations used here and find it particularly true once several tracers are jointly taken into account.
+they can be tightly constrained using galaxy survey auto- and cross-correlation spectra.
+We checked this for a few of the tracer combinations used here and find it particularly true once several tracers are jointly taken into account.
 Finally, we have assumed a Gaussian covariance even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
@@
 
-Only recently delensing has been performed on data using both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as CIB maps \cite{larsen:2016a,manzotti:2017} as large scale structure tracers.
+Recently delensing has been performed for the first time on data using both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as CIB maps \cite{larsen:2016a,manzotti:2017} as large scale structure tracers.
 Here we discuss the improvement that can be obtained by combining these and other currently available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
 
-On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} CMB experiment.
-For the deep experiment we use an instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin}, a beam of \textcolor{red}{30 arcmin} and an angular scale range of $50<\ell<500$.
-The CMB lensing reconstruction will be performed by the higher resolution CMB experiment. We explore two possibilities.
-First, we test the efficiency with an internal reconstruction performed by Planck \cite{planck-collaboration:2016a} using the CMB lensing noise curves publicly available
+On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment.
+For the deep experiment we assume an instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin}, a beam of \textcolor{red}{30 arcmin} and an angular scale range of $50<\ell<500$.
+We assume the CMB lensing reconstruction is  performed by the higher resolution CMB experiment. We explore two possibilities.
+First, we test the efficiency with an internal reconstruction of CMB lensing performed by Planck collaboration \cite{planck-collaboration:2016a}.
+We compute the noise in the CMB lensing map in \refeq{noisekappa} using the noise curves publicly available
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
@@
 The achievable correlation is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
-In these figures we include a dashed curve that corresponds to (with arbitrary scale) $<C_{\ell}^{\kappa \kappa} \times \frac{\partial C^{BB}_{\ell'}}{\partial{C^{\kappa \kappa}_{\ell}}}>_{\ell ' <100}$, showing which scales contribute to the $\ell<100$ B-mode power.
+In these figures we include a dashed curve that corresponds to (with arbitrary scale) $<C_{\ell}^{\kappa \kappa} \times \frac{\partial C^{BB}_{\ell'}}{\partial{C^{\kappa \kappa}_{\ell}}}>_{\ell ' <100}$, in order to show which scales in $\kappa$ contribute to the $\ell<100$ B-mode power.
 
@@
 Its correlation is comparable to the one of LSS at almost all scales.
-On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 80\% correlated with the true field at $\ell<300$, and only then the correlation become comparable with the current LSS surveys.
+On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 80\% correlated with the true field at $\ell<300$, and only then the correlation becomes comparable with the current LSS surveys.
 
-\reffig{corrnow} shows that the DES is an effective tracer of the LSS and can, at least in the near future, be used to improve delensing for CMB experiments that overlap with it.
-Indeed its delensing efficiency is higher than WISE because the redshift distributions peaks at slightly higher redshift thus it better overlaps with CMB lensing.
+\reffig{corrnow} and \reffig{corrnow-planck} show that the DES galaxies are effective tracers of the LSS and can, at least in the near future, be used to improve delensing for CMB experiments that overlap with it.
+For example, DES delensing efficiency is higher than WISE firstly because of the lower level of noise and furthermore because DES galaxies are located at slightly higher redshift thus they better overlap with CMB lensing.
 
@@
   \vspace{0.2cm}
-  \begin{tabular}{ | c | c | c|}
+  \begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
-WISE & 1.10 & 1.04 \\ \hline
-DES & 1.19  & 1.07 \\ \hline
-CIB & 1.46  & 1.18\\ \hline
-LSS & 1.64  & 1.24\\ \hline
-CMB & 1.14  & 1.04 \\ \hline
-LSS+CMB & 1.71& 1.26  \\ \hline
+Surveys& & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+WISE& & 1.10 & 1.04 \\ \hline
+DES && 1.19  & 1.07 \\ \hline
+CIB && 1.46  & 1.18\\ \hline
+LSS && 1.64  & 1.24\\ \hline
+CMB && 1.14  & 1.04 \\ \hline
+LSS+CMB& & 1.71& 1.26  \\ \hline
 
@@
 \end{table}
-In agreement with the literature \cite{yu:2017,larsen:2016a}, the CIB is the predominant source of delensing with an improvement of almost 50\%.
-However Planck and DES are essential to reconstruct the largest scales of the lensing potential and as such the play a major role in delensing.
-The improvement due to WISE is quite marginal, but given the big sky coverage, it might be useful to delens large CMB polarization experiments.
+
 We also test a scenario where primordial gravitational waves are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.

### 2017-07-03 Notes/galaxy_delensing.tex
text: section 4

@@
 \newcommand{\Nsim}{N_{\text{sim}}}
-\newcommand{\new}[1]{{\textcolor{blue}{\bf #1}}}
+\newcommand{\relook}[1]{{\textcolor{blue}{\bf #1}}}
 %--------- NEW COMMAND STUFF
@@
 However, improving the level of noise in future experiments will not be enough.
-Secondary effects in the low redshift universe will also produce non-primordial B-modes.
+Secondary effects in the low redshift universe will also produce non-primordial B-modes adding confusion to the inflationary signal.
 In particular, the gravitational interactions of CMB photons with large scale structures will distort the primordial E-modes adding a lensing B-mode component to the primordial signal.
-Removing the lensing component from the measurement of CMB B-modes will be important to constrain the amplitude of the primordial gravitational wave contribution to the signal.
-Here we discuss the role of current and future large scale structure surveys in improving the reconstruction of the lensing potential that lenses the CMB photons and how this reflects into an improved delensing efficiency.
+Removing the lensing component from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
+Here we discuss the role of current and future large scale structure surveys in multi-tracers approach that will improve the reconstruction of the lensing potential that lenses the CMB photons and and, as a consequence, the delensing efficiency.
 We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
-We find that current generation experiments  we also point out that a tomographic approach..
+We find that a galaxy survey like DES can remove about $14\%$ of the lensing signal. current generation CMB experiments will benefit from 
+With the improvement of the noise CMB internal reconstruction will become more important and the fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level.
+\relook{multitracers}
+Nevertheless optical galaxy surveys, in particular if split in different tomographic bin will still play an important role even for CMB S4.
+
 \end{abstract}
@@
 The SKA is a planned radio array that will survey large scale structure primarily by detecting the redshifted neutral hydrogen (HI) 21cm emission line from a large number of galaxies out to high redshift.
-We will take into account the intensity mapping survey mode where SKA will measure the large-scale fluctuations of the \textit{integrated} 21cm intensity from many unresolved galaxies.  
+We will assume an intensity mapping survey mode where SKA will measure the large-scale fluctuations of the \textit{integrated} 21cm intensity from many unresolved galaxies.  
 We model both the redshift distribution and  bias of radio sources following \cite{namikawa:2016a}.
@@
 \be\label{eqn:rho1}
-\rho_{\ell}= \frac{\cl^{\kappa I}}{\sqrt{\cl^{\kappa \kappa} \cl^{I I}}}.
+\rho_{\ell}^{2}= \frac{(\cl^{\kappa I})^{2}}{\cl^{\kappa \kappa}\cl^{I I}}.
 \ee
-The bigger the $\rho_{\ell}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
-An higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$.
+
+\refeq{Bres} directly relate the residual power after delensing with the cross correlation coefficients with CMB lensing of the tracers used.
+The bigger the $\rho_{\ell}^{2}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
+An higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$ leading to a smaller residual power $C_{{l}}^{BB,\mathrm{res}}$.
+\relook{We conclude this section showing in \reffig{bbplot} the expected residual lensing B-modes power spectrum for some of the tracers used in this work together with the primordial B-modes component and the instrumental noise for current and future experiments.
+}
 
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/BB_res.pdf}
-\caption{}
-\label{fig:bb_res}
+\includegraphics[scale=1.]{../images/BB_res_ell2.pdf}
+\caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
+The red solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed green line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
+The residual power left after delensing for some of the LSS tracers described in \refsec{model}. 
+The rapid improvement in the level of instrumental noise (dashed curves for SPTPol and CMB S4) will require a high level of delensing to exploit these experiments fully.}
+\label{fig:bbplot}
 \end{center}
@@
 Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing, but it also depends on how much it is correlated with the already used set of tracers.
+\relook{
 \reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
+Note that the cross correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
 CIB and 21 cm surveys probes the high redshift structures and, independently from the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
 On the other end, galaxy clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
-However, given the low noise of this measurement and their small overlap with other probes they still improve the delensing efficiency even if their overlap with the CMB lensing potential is not optimal.
+However, given the low noise of this measurement and their small overlap with other probes they can still play an important role in delensing even if their overlap with the CMB lensing potential is not optimal.
+}
+
+% ===================
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/compare_kernel.pdf}
+\caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
+The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential will be leading to a higher delensing efficiency.
+}
+\label{fig:kernel}
+\end{center}
+\end{figure}
+% ===================
 
 \subsection{Improving efficiency with tomographic binning}\label{sec:tomo-bin}
-Furthermore the delensing efficiency of galaxy surveys can be improved by taking into account redshift information.
-When we weight a tracer with $\frac{C^{\kappa I}}{C^{II}}$ in \refeq{fweight} to maximize its ability to reconstruct the lensing potential we are only using redshift averaged information about the survey.
-However, as can be seen in \reffig{kernel}, the overlap varies as a function of redshifts.
+The delensing efficiency of galaxy surveys can be improved by taking into account redshift information.
+When we weight a tracer with $\frac{C^{\kappa I}}{C^{II}}$ in \refeq{fweight} in order to maximize its ability to reconstruct the lensing potential we are only using redshift averaged information about the survey.
+However, as can be seen in \reffig{kernel}, the kernel overlap of a tracer with the CMB lensing varies as a function of redshifts.
 For this reason, the optimal approach is to weight galaxies at different redshifts with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
@@
 Then $\rho^2$ is always bigger in the tomographic case, and the two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the same for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
-For photometric surveys, the maximum number of bins is dictated by the fact that the bin width can not be smaller than the photo-z accuracy. 
-Of course, binning is possible for galaxy surveys, but it is harder for a tracer with poor redshift information like the CIB or radio continuum surveys.
+Of course, binning will improve the efficiency of galaxy surveys, but it is not applicable to tracers with poor redshift information like the CIB or radio continuum surveys.
 
-In this work, we then bin galaxy surveys by splitting the window function \refeq{wg} into different slices so that all the bins contain the same number of galaxies. 
+In this work, we bin both photometric and spectroscopic galaxy surveys by splitting the window function \refeq{wg} into different slices such as all the bins contain the same number of galaxies. 
 For photometric surveys like DES and LSST, we assume a photometric redshift estimation gaussianly distributed around the true value with an rms fluctuation $\sigma(z)$.
@@
 \label{eqn:zbin}
-W_{i}(z) \propto b(z) {dN(z) \over dz} 
-\Bigl[{\,\rm erfc}\left({\Delta(i-1) -z \over \sigma(z)\sqrt{2}}\right) 
+W_{i}(z) \propto b(z) {\frac{dN(z)}{dz}} 
+\Bigl[{\,\rm erfc}\left({\Delta(i-1) -z \frac{\sigma(z)}{\sqrt{2}}}\right) 
 \\ \nonumber - {\,\rm erfc}
-\left({\Delta i  - z \over \sigma(z)\sqrt{2}}\right)\Bigr].
+\left({\Delta i  - \frac{z}{\sigma(z)}\sqrt{2}}\right)\Bigr].
 \end{eqnarray}
 
-We used 10 and 4 photometric bins for LSST and DES respectively and we split DESI into 4 spectroscopic bins with no overlap among each others.
-
 % ===================
-
 \begin{figure}[htbp]
 \begin{center}
-\includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
-The bigger the overlap wit the CMB lensing kernel the better the reconstruction of the lensing potential will be.
-This will lead to a better delensing.
+\includegraphics[scale=1.]{../images/compare_kernel_tomo.pdf}
+\caption{\textbf{Kernels Comparison with tomographic bins}: Comparison of the 10 LSST  tomographic bins together with the full LSST survey and the CMB lensing kernel. 
+Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross correlation with the CMB lensing.
+This leads to a better delensing efficiency.
 }
-\label{fig:kernel}
-
+\label{fig:kerneltomo}
 \end{center}
@@
 
-% ===================
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/compare_kernel_tomo.pdf}
-\caption{\textbf{Kernels Comparison with tomographic bins}: Comparison of the 10 LSST different kernels for every tomographic bin with the CMB lensing kernel. 
-This will lead to a better delensing. }
-\label{fig:kernel}
-\end{center}
-\end{figure}
-% ===================
+For photometric surveys, the maximum number of bins is dictated by the fact that the bin width can not be smaller than the photo-z accuracy.
+We used 10 and 4 photometric bins for LSST and DES respectively with a common photo-z accuracy $\sigma(z) = 0.05(1+z)$. 
+\relook{In spectroscopic surveys, there are no limitations in increasing the number of bins. We split DESI into 4 spectroscopic bins with no overlap among each others.
+This number of bins is close to the saturation point where adding more bins does not improve delensing significantly while adding complexity to the analysis.} As an example, we show in \reffig{kerneltomo} the 10 bins and the full LSST redshift distributions together with the CMB lensing kernel.
+\relook{The improvement obtained by tomographic binning is illustrated in \reffig{rho-tomo}.
+In particular on large scales, binning can increase the value of $\rho$ by almost $30\%$ significantly improving the delensing efficiency of galaxy surveys.}
 
@@
 \includegraphics[scale=1.]{../images/B_res_bin.pdf}
-\caption{Tomographic bins improve the cross-correlation with lensing.
-Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line).}
-\label{fig:bb_res}
+\caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing.
+Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of angular scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line).}
+\label{fig:rho-tomo}
 \end{center}
@@
 \subsection{Fisher Information Matrix}
-In the Fisher Information Matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtain the from inverse of the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
+In the Fisher Information Matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtained from the inverse of the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
 We constrain the inflationary parameters $p =\{r,n_{t}\}$ with a CMB B-mode spectrum measurement so the Fisher matrix is:
@@
 
-Recently delensing has been performed for the first time on data using both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as CIB maps \cite{larsen:2016a,manzotti:2017} as large scale structure tracers.
+Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} as well as the internal CMB lensing potential reconstruction \cite{carron:2017} as large scale structure tracers.
 Here we discuss the improvement that can be obtained by combining these and other currently available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
@@
 For current surveys we only consider $p=r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
+Note that we still let $n_{t}$ to vary in our fisher calculation even if it is not very degenerate with $r$.
 We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
@@
 \begin{table}
-\caption{$\alpha(r)$: Improvements on $\sigma(r)$ due to delensing for current generation and Planck lensing reconstruction. Values correspond to ratio of the error before and after delensing.}
+\caption{$\alpha(r)$: Improvements on $\sigma(r)$ due to delensing for current generation and Planck lensing reconstruction.
+The values in parenthesis in the first column correspond to the fraction of lensing B-mode power removed using each LSS tracer. 
+The values in the other columns correspond to ratio of the error before and after delensing.}
   \vspace{0.2cm}
@@
 \hline
-Surveys& & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
-WISE& & 1.10 & 1.04 \\ \hline
-DES && 1.19  & 1.07 \\ \hline
-CIB && 1.46  & 1.18\\ \hline
-LSS && 1.64  & 1.24\\ \hline
-CMB && 1.14  & 1.04 \\ \hline
-LSS+CMB& & 1.71& 1.26  \\ \hline
-
+Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+WISE (6\%)& 1.09& 1.05 & 1.01 \\ \hline
+DES (14\%) &1.2 & 1.10  & 1.03 \\ \hline
+CIB (27\%) &1.45& 1.21  & 1.07\\ \hline
+LSS (36\%) &1.7& 1.29  & 1.09\\ \hline
+CMB Planck (8\%) &1.06& 1.05  & 1.01 \\ \hline
+LSS+CMB (41\%)& 1.77& 1.31& 1.10  \\ \hline
 \end{tabular}
@@
 Surveys & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
-WISE & 1.10 &  \\ \hline
-DES & 1.21  & \\ \hline
-CIB & 1.52  & \\ \hline
-LSS & 1.78  & \\ \hline
-CMB & 1.62  &  \\ \hline
-LSS+CMB & 2.35 &   \\ \hline
+CMB SPTPol (35\%) & 1.62  &  \\ \hline
+LSS+CMB (56\%) & 2.35 &   \\ \hline
 
@@
 \subsection{CMB-S3 Era}
-CMB measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
-To model the CMB at the beginning of the next decade, we will assume a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
-This is a reasonable level for a combination of the next generation of BK and SPT3G.
-We will assume this level of noise for the internal noise reconstruction.
+CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
+We model the CMB S3 stage with a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam. 
+We assume here that a deep CMB experiment and an overlapping high resolution one will be able to combine their measured modes.
+For this reason, we will also assume this level of noise for the internal noise reconstruction.
+With this experimental setup we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
 
+Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
 The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
-Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
-An interesting results is that DESI will be less efficient than a DES-like survey even if it can probe higher redshift.
-The reason is that, for the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES play a mojor role in the delensing efficiency.
-Even in the near future the CIB will still play the major role among LSS.
+An interesting finding is that DESI will be less efficient (removing 10\% of the power) than a DES-like survey despite the fact that it can probe slightly higher redshift.
+The reason is that, for the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES play a significant role in the delensing efficiency.
+Unfortunately, adding DESI will only bring the power removed using LSS fro 36 to 41\%.
+Indeed, even in the near future, CIB will still play the dominant role among LSS.
+Contrary to current experiments, th CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 550$.
 
-This generation of experiments will start exploring the small scales B-modes thus constraining, even if loosely, the value of $n_{t}$. 
-The improvement in cosmological parameters are summarized in \reftab{S3}
+%This generation of experiments will start exploring the small scales B-modes thus constraining, even if loosely, the value of $n_{t}$. 
+The improvement in cosmological parameters is summarized in \reftab{S3}. 
+The main difference from the current generation of experiments is that in the near future CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
+
+Despite this compared to CMB alone, galaxy surveys will still be able to remove an additional $20\%$ of power and to improve our constraint on the null hypothesis ($r=0$) by $15\%$.
+Note that apart from the addition of DESI, which plays a secondary role, the tracers used here are already available for the current generation.
+The ongoing effort can then allow this improvement even for the next generation of experiments.
+The combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor 1.8 or to improve the constraint of a possible detection (r=0.12) by 30\%.
+
+\relook{something about $n_t$}
 
@@
 \caption{$\alpha$: S3 improvements. Note that the absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is $\sigma(r) =$  }
-  \begin{tabular}{ | c | c  |  c |}
+  \begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r)$ & $\alpha(n_{t})$\\ \hline \hline
-DESI & 1.34 & \\ \hline
-LSS & 1.71  &\\ \hline
-CMB & 1.71  &\\ \hline
+Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+DESI (10\%) &1.12 & 1.08  & 1.03 \\ \hline
+LSS (40\%) &1.73& 1.42  & 1.2\\ \hline
+CMB 3G (56\%) &2& 1.57  & 1.21 \\ \hline
+LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
 \end{tabular}
@@
 An ambitious program for a generation 4 ground CMB experiment is currently under planning \cite{abazajian:2016}. Moreover, satellite experiments have been proposed.
+
 The correlation attainable using current experiments is shown in \reffig{corrs4},
@@
 
+\section{Conclusions}
+\label{sec:concl}
+
+The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal ("delensing") is necessary to test the inflationary paradigm using the next generation of CMB polarization experiments.
+To do so, we need to accurately reconstruct the large scale structures that lens the CMB in order to build a template of the lensing B-modes.
+In this paper, we studied the potential impact of large-scale structure galaxy surveys in this important endeavor.
+%For current experiment, the cosmic infrared background had already proven to be very efficient in delensing \cite{larsen:2016,manzotti:2017}. 
+For ongoing experiments, we find that an optical survey like DES alone is able to remove $14\%$ of the lensing component and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the contaminant signal using only LSS survey.
+Depending on the CMB noise and the amount of galactic foreground cleaning these can correspond to an improvement of 30\% in the constrain on the tensor to scalar ratio $r$.
+
+In the future, the improved level of noise of CMB experiments will allow to internally reconstruct the structures lensing the CMB. 
+The fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level. 
+Already in the 3G era, the CMB internal reconstruction will be the main source of delensing.
+However it will still be less efficient than galaxies at small scales$\ell>500$.
+For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $57\%$ to $68\%$.
+
+Furthermore internal delensing will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}.
+For this reason, efficient galaxies tracers are not only useful in the short term but will also play a role in cross-checking these internal biases. 
+For a future galaxy survey like LSST we find that using tomographic binning can improve its correlation with lensing by $30\%$.
+
+Giving the high level of correlation of future galaxy surveys and CMB lensing it is worth looking for possible application beside the detection of inflationary B-modes.
+For example LSS delensing allow to remove just the low-z lensing component from the CMB lensing.
+Removing this component can avoid the necessity of modelling nonlinearities in the study of CMB lensing power spectrum.
+
 % ===================
-
+%\onecolumngrid
 \begin{figure}[htbp]
@@
 \includegraphics[scale=1.]{../images/errors_summary.pdf}
-\caption{}
-\label{fig:kernel}
+\caption{\relook{to be filled with final results}}
+\label{fig}
 \end{center}
 \end{figure}
+\twocolumngrid
+
 % ===================
-
-\section{Conclusions}
-\label{sec:concl}
-
-Delensing or the ability to separate the lensing component of the CMB B-mode from a possible primordial inflationary signal is important to exploit the capabilities of future experiments thoroughly.
-In this paper, we studied the potential impact of large-scale structure surveys in this important endeavor.
-For current experiment, CIB data had already proven to be very efficient in delensing \cite{larsen:2016,manzotti:2017}. 
-If no CMB experiment is available with higher resolution then Planck is used to reconstruct the lensing potential $\phi$, a low-z optical survey like D.E.S will improve the delensing efficiency by $~10\%$ when combined with Planck lensing and a CIB tracers.
-As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential.
-Indeed using the CMB itself, we reconstruct only the lenses that lens the CMB obtaining, as a result, an almost perfect cleaning in the absence of noise.
-This approach has already been applied to data \cite{carron:2017}.
-However it requires a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}.
-For this reason, the relatively good efficiency of galaxies tracers might be useful to cross-check these internal biases.
-This will probably be needed to confirm a possible detection of gravitation waves if this relies heavily on delensing.
-
-Another possible application of LSS delensing, which we will explore in future works, is the possibility of removing just the low-z lensing of the CMB which will otherwise possibly adds the complexity of modeling nonlinearities in the study of CMB lensing.
 
@@
 \section{Acknowledgments}
-We thank S. Dodelson S. Passaglia K. Wu and K. Story for useful discussion and their review of the draft. Furthermore we thank D. Alonso for his feedback on radio continuum surveys.
+We thank S. Passaglia, S. Dodelson, K. Story and K. Wu for useful discussion and their review of the draft. 
+Furthermore we thank D. Alonso for his feedback on radio continuum surveys.
 This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.

### 2017-07-03 Notes/galaxy_delensing.tex
text: section 4

@@
 With this experimental setup we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
+Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers.
 
-Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
-The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
-An interesting finding is that DESI will be less efficient (removing 10\% of the power) than a DES-like survey despite the fact that it can probe slightly higher redshift.
+The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
+An interesting finding is that DESI will be less efficient (removing 10\% of the power) than a DES-like survey (14\%) despite the fact that it can probe slightly higher redshift.
 The reason is that, for the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES play a significant role in the delensing efficiency.
@@
 %This generation of experiments will start exploring the small scales B-modes thus constraining, even if loosely, the value of $n_{t}$.
-The improvement in cosmological parameters is summarized in \reftab{S3}. 
-The main difference from the current generation of experiments is that in the near future CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
-
-Despite this compared to CMB alone, galaxy surveys will still be able to remove an additional $20\%$ of power and to improve our constraint on the null hypothesis ($r=0$) by $15\%$.
-Note that apart from the addition of DESI, which plays a secondary role, the tracers used here are already available for the current generation.
-The ongoing effort can then allow this improvement even for the next generation of experiments.
-The combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor 1.8 or to improve the constraint of a possible detection (r=0.12) by 30\%.
-
-\relook{something about $n_t$}
 
@@
 \end{figure}
+
+The improvement in the constraint on $r$ is summarized in \reftab{S3}. 
+Given the improvement in the noise of the high resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
+Despite this, adding galaxy survey will still lead to a decent improvement with a fairly low effort.
+Indeed galaxy surveys will still be able to remove an additional $20\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by $15\%$.
+Note that apart from the addition of DESI, the tracers used here are already available for the current generation.
+The ongoing effort can then lead to a significant improvement even for the next generation of experiments.
+Finally, the combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor 1.8 or to improve the constraint of a possible detection (r=0.12) by 30\%.
+As expected the improvement on $n_{t}$ is similar to the one in $r$ since they are both proportional to the variance of the measured B-modes.
+We tested this in the scenario $r=0.12$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}/8}$.
+It is important to note that, independently from the level of delensing, for CMB 3G experiments this statistical error will still be several times bigger than the fiducial value.
 
@@
 
-An ambitious program for a generation 4 ground CMB experiment is currently under planning \cite{abazajian:2016}. Moreover, satellite experiments have been proposed.
+An ambitious program for a generation 4 ground CMB experiment is currently under planning \cite{abazajian:2016}. 
+Moreover, satellite and balloon experiments have been proposed and have the potential to extend the accessible B-mode measurement to the largest scales.
+Here we follow \cite{abazajian:2016} and assume a CMB-S4 ground experiment with  with a level of noise of \textcolor{red}{$1\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
+By the mid 2020s several next generation of LSS survey will be online.
+Here, as an example for future optical galaxy surveys we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers.
+We test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} will have similar delensing performance.
+Even if they will observe similar part of the sky, because of the different  strength and weakness of this experiments
+it will still important to combined them efficiently.
+We also consider a radio continuum survey modeled following the SKA planned experiment.
+Radio continuum observation of the 21 cm line is in its early stage and several experimental and analysis challanges
+need to be overcome. However this technique has the potential to map LSS with relatively low noise up to redshift $z=6$.
 
-The correlation attainable using current experiments is shown in \reffig{corrs4},
-In this futuristic but reasonable scenario, we add to the available large scale structures tracers both 21 cm surveys like SKA and large scale structures like LSST and Euclid.
+The correlation factor with CMB lensing of stage 4 experiments is shown in \reffig{corrs4}.
 
@@
 \label{fig:corrs4}
-
 \end{center}
 \end{figure}
+
+A tomographically binned LSST-like experiment will be a very efficient CMB lensing tracer.
+Indeed it will be more than 70\% correlated with CMB lensing for all the scales $\ell<600$ a performance similar to a Stage 3 CMB internal reconstruction.
+SKA very new but potentially amazing
+With SKA and LSST, for the first we will have LSS tracers way more efficient than the CIB.
+Even if galaxy survey will improve dramatically the CMB internal reconstruction will still be the main source of delensing.
+This, having a perfect kernel overlap with the true lensing potential we will benefit for the very low level of noise of CMB S4 experiments.
+CMB S4 delensing will probably be limited by small secondary effect like foregrounds contamination, missing modes due to filtering etc.
+
+\reftab{S4} shows the improvement on the inflationary constraint.
 
@@
 LSS+CMB & 6.3  &\\ \hline
+\end{tabular}
+\label{tab:S4}
+\end{table}
 
+%=======================================
+
+\begin{table}
+\caption{$\alpha$: Gen 4 experiments}
+  \begin{tabular}{ | c | c  |  c |}
+\hline
+Surveys & $\alpha(r)$ & $\alpha(n_{t})$\\ \hline \hline
+Euclid & 1.53 & \\ \hline
+LSST & 1.65  &\\ \hline
+SKA & 4.35  &\\ \hline
+CMB S4 & 3.7  &\\ \hline
+LSS+CMB & 6.3  &\\ \hline
 \end{tabular}

### 2017-07-05 Notes/galaxy_delensing.tex
text general

@@
 \begin{abstract}
-%\begin{itemize}
-%\item Worry : they might ask for internal bias\
-%item Possible TODO : temperature $N_{eff}$
-%
-%\end{itemize}
-%
 
@@
 A particular component, the so-called B-modes, is indeed a direct signature of a prediction of inflation: the presence of gravitational waves in the early universe.
-However, improving the level of noise in future experiments will not be enough.
+However, reducing the instrumental noise in future experiments will not be enough to detect this signal.
 Secondary effects in the low redshift universe will also produce non-primordial B-modes adding confusion to the inflationary signal.
 In particular, the gravitational interactions of CMB photons with large scale structures will distort the primordial E-modes adding a lensing B-mode component to the primordial signal.
-Removing the lensing component from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
-Here we discuss the role of current and future large scale structure surveys in multi-tracers approach that will improve the reconstruction of the lensing potential that lenses the CMB photons and and, as a consequence, the delensing efficiency.
+Removing the lensing part ('delensing') from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
+Here we discuss the role of current and future large scale structure surveys in a multi-tracers approach to delensing that will improve the reconstruction of the lensing potential that lenses the CMB photons and, as a consequence, the delensing efficiency.
 We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
-We find that a galaxy survey like DES can remove about $14\%$ of the lensing signal. current generation CMB experiments will benefit from 
-With the improvement of the noise CMB internal reconstruction will become more important and the fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level.
-\relook{multitracers}
+We find that in general splitting galaxy surveys into tomographic binning can improve the correlation factor with CMB lensing by 30\%. 
+Considering available survey, a galaxy survey like DES can remove about $14\%$ of the lensing signal. current generation CMB experiments will benefit from 
+With the improvement of the noise CMB internal reconstruction will become more important and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level.
 Nevertheless optical galaxy surveys, in particular if split in different tomographic bin will still play an important role even for CMB S4.
@@
 An higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$ leading to a smaller residual power $C_{{l}}^{BB,\mathrm{res}}$.
-\relook{We conclude this section showing in \reffig{bbplot} the expected residual lensing B-modes power spectrum for some of the tracers used in this work together with the primordial B-modes component and the instrumental noise for current and future experiments.
-}
+We conclude this section showing in \reffig{bbplot} the expected residual lensing B-modes power spectrum for some of the tracers used in this work together with the primordial B-modes component and the instrumental noise for current and future experiments.
 
@@
 Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing, but it also depends on how much it is correlated with the already used set of tracers.
-\relook{
 \reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
-Note that the cross correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
-CIB and 21 cm surveys probes the high redshift structures and, independently from the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
+Note that the cross-correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
+CIB and 21 cm surveys probe the high redshift structures and, independently from the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
 On the other end, galaxy clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
-However, given the low noise of this measurement and their small overlap with other probes they can still play an important role in delensing even if their overlap with the CMB lensing potential is not optimal.
-}
+However, given the low noise of this measurement and their small overlap with other probes they can still play a major role in delensing even if their overlap with the CMB lensing potential is not optimal.
 
@@
 We used 10 and 4 photometric bins for LSST and DES respectively with a common photo-z accuracy $\sigma(z) = 0.05(1+z)$. 
-\relook{In spectroscopic surveys, there are no limitations in increasing the number of bins. We split DESI into 4 spectroscopic bins with no overlap among each others.
-This number of bins is close to the saturation point where adding more bins does not improve delensing significantly while adding complexity to the analysis.} As an example, we show in \reffig{kerneltomo} the 10 bins and the full LSST redshift distributions together with the CMB lensing kernel.
-\relook{The improvement obtained by tomographic binning is illustrated in \reffig{rho-tomo}.
-In particular on large scales, binning can increase the value of $\rho$ by almost $30\%$ significantly improving the delensing efficiency of galaxy surveys.}
+In spectroscopic surveys, there are no limitations in increasing the number of bins. We split DESI into 4 spectroscopic bins with no overlap among each others.
+This number of bins is close to the saturation point where adding more bins does not improve delensing significantly while adding complexity to the analysis.
+As an example, we show in \reffig{kerneltomo} the 10 bins and the full LSST redshift distributions together with the CMB lensing kernel.
+\reffig{rho-tomo} illustrate the improvement obtained by tomographic binning.
+In particular on large scales, binning can increase the value of $\rho$ by almost $30\%$ significantly improving the delensing efficiency of galaxy surveys.
 
@@
 We define the improvement as the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$.
+The fiducial values used are described in \refsec{del_res}.
+Given a fiducial value for $r$, the value of $n_{t}$ is fixed imposing the consistency relation $n_t=-r/8$ \cite{LIDDLE1992391}.
 
@@
 \subsection{Delensing with current CMB and LSS}
+\label{sec:del_res}
 
@@
   \vspace{0.2cm}
-  \begin{tabular}{ | c | c | c|}
+  \begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
-CMB SPTPol (35\%) & 1.62  &  \\ \hline
-LSS+CMB (56\%) & 2.35 &   \\ \hline
+Surveys &$\alpha(r=0), N^{B}_{\ell}=0$& $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+CMB SPTPol (35\%) &1.46& 1.27  & 1.08 \\ \hline
+LSS+CMB (56\%) & 2.18 & 1.57 & 1.16  \\ \hline
 
@@
 \caption{$\alpha$: Gen 4 experiments}
-  \begin{tabular}{ | c | c  |  c |}
+  \begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r)$ & $\alpha(n_{t})$\\ \hline \hline
-Euclid & 1.53 & \\ \hline
-LSST & 1.65  &\\ \hline
-SKA & 4.35  &\\ \hline
-CMB S4 & 3.7  &\\ \hline
-LSS+CMB & 6.3  &\\ \hline
+Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+LSST (10\%) &1.12 & 1.08  & 1.03 \\ \hline
+SKA (40\%) &1.73& 1.42  & 1.2\\ \hline
+CMB S4 (56\%) &2& 1.57  & 1.21 \\ \hline
+LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
 \end{tabular}
@@
 Furthermore we thank D. Alonso for his feedback on radio continuum surveys.
+
+This work made use of computing resources and support provided by the Re- search Computing Center at the University of Chicago.
 This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.

### 2017-07-05 Notes/galaxy_delensing.tex
text Sect 4.

@@
 
-\subsection{Fisher Information Matrix}
+\subsection{Fisher Information Matrix}\label{sec:fisher}
 In the Fisher Information Matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtained from the inverse of the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
@@
 We assume the CMB lensing reconstruction is  performed by the higher resolution CMB experiment. We explore two possibilities.
-First, we test the efficiency with an internal reconstruction of CMB lensing performed by Planck collaboration \cite{planck-collaboration:2016a}.
+First, we test the efficiency with an internal reconstruction of CMB lensing performed by the Planck collaboration \cite{planck-collaboration:2016a}.
 We compute the noise in the CMB lensing map in \refeq{noisekappa} using the noise curves publicly available
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
-Then we assume a ground-based experiment with noise levels consistent with SPT-Pol: \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
+Furthermore, we test a second scenario with a ground-based experiment with noise levels consistent with SPT-Pol: \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
 In this case we set the largest achievable scale at $\ell_{min}=300$.
@@
 The results are summarized in \reftab{current} and \reftab{current-spt} for the Planck and SPTPol case respectively.
+
+The lensing reconstruction noise in Planck is such that internal delensing can only remove $8\%$ of the power, improving the constraints on $r$ by only a factor of 1.06 even in the ideal scenario of no instrumental noise in the B-modes.
+A similar performance is achieved with WISE.
+DES can do better, removing a level of 14\% of power.
+However, it covers a smaller fraction of the sky than the two previous probes.
+As shown in previous works \cite{sherwin:2015}, the best tracer is the CIB with a 27\% reduction in power.
+Nonetheless, it is still worth to consider all the tracers.
+A multi tracer-approach can bring the removed power on the overlapping area from 27\% of the CIB alone to 36\% using all the LSS and 41\% when Planck lensing reconstruction is added to the set.
+However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 30\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
+Furthermore, as explained in \refsec{fisher}, this is only an upper limit and will be even smaller when, for example, a non-perfect foreground cleaning is considered.
+Indeed current experiments are still significantly limited by foregrounds and instrumental noise together and not only by lensing.
 

### 2017-07-05 Notes/galaxy_delensing.tex
text Sect 4.

@@
 We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$ and we set $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
+Note that there are several available CIB observations that have already been used to delens the CMB.
+Given its full-sky coverage, a promising one is the CIB map derived in \cite{2016A&A...596A.109P} using multi-frequency Planck data and the Generalized Needlet Internal Linear Combination (GNILC) component separation algorithm.
 
@@
 
-The lensing reconstruction noise in Planck is such that internal delensing can only remove $8\%$ of the power, improving the constraints on $r$ by only a factor of 1.06 even in the ideal scenario of no instrumental noise in the B-modes.
+The lensing reconstruction noise in Planck is such that internal delensing can only remove $8\%$ of the power, improving the constraints on $r$ by a factor of 1.06 even in the ideal scenario of no instrumental noise in the B-modes.
 A similar performance is achieved with WISE.
@@
 Nonetheless, it is still worth to consider all the tracers.
-A multi tracer-approach can bring the removed power on the overlapping area from 27\% of the CIB alone to 36\% using all the LSS and 41\% when Planck lensing reconstruction is added to the set.
+Indeed a multi tracer-approach can bring the removed power on the overlapping area from 27\% of the CIB alone to 36\% using all the LSS and 41\% when Planck lensing reconstruction is added to the set.
 However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 30\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
@@
 Indeed current experiments are still significantly limited by foregrounds and instrumental noise together and not only by lensing.
+If on the other side we consider an high-resolution ground base experiment the internal lensing reconstruction improves significantly.
+With an instrumental noise at the expected level for SPTPol the CMB will soon be able to remove an amount of power (35\%) at the level of the CIB.
+Once combined with LSS tracer this will lead to removing 56\% of the power.
+The improvement of 1.57 in $\sigma(r)$ that will follow even when realistic noise is considered is such that \relook{here}
 

### 2017-07-05 Notes/galaxies_delensing_uchicago_thesis.tex
uchicago thesis

@@
 \usepackage{xstring}
+\newcommand{\relook}[1]{{\textcolor{black}{\bf #1}}}
+\usepackage[justification=centering]{caption}
+
 % \usepackage{graphicx}\usepackage[top=2.5 cm, bottom=2.5 cm, left=1.05in, right=1.05in]{geometry}
@@
 
-The measurement of the cosmic microwave background (CMB) polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe. 
-A particular component, the so-called B-modes is indeed a direct signature of the presence of gravitational waves in the early universe. However improving the level of noise is not enough. 
-This is even truer if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
-Removing the lensing component from the measurement of CMB B-modes will be important to constrain the amplitude of the primordial gravitational wave contribution to the signal.
-Here we discuss the role of current and future large scale structure surveys in improving the reconstruction of the lensing potential that lenses the CMB photons and how this reflects into an improved delensing efficiency.
-We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$)
-We find that current generation experiments...
-
+The cosmic microwave background (CMB) polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe. 
+A particular component, the so-called B-modes, is indeed a direct signature of a prediction of inflation: the presence of gravitational waves in the early universe.
+However, improving the level of noise in future experiments will not be enough.
+Secondary effects in the low redshift universe will also produce non-primordial B-modes adding confusion to the inflationary signal.
+In particular, the gravitational interactions of CMB photons with large scale structures will distort the primordial E-modes adding a lensing B-mode component to the primordial signal.
+Removing the lensing component from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
+Here we discuss the role of current and future large scale structure surveys in multi-tracers approach that will improve the reconstruction of the lensing potential that lenses the CMB photons and and, as a consequence, the delensing efficiency.
+We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
+We find that a galaxy survey like DES can remove about $14\%$ of the lensing signal. current generation CMB experiments will benefit from 
+With the improvement of the noise CMB internal reconstruction will become more important and the fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level.
+Nevertheless optical galaxy surveys, in particular if split in different tomographic bin will still play an important role even for CMB S4. 
 \mainmatter
@@
 \label{sec:intro}
-In the standard cosmological paradigm, the universe underwent a period of near-exponential expansion in its early phase;
-this period is called ``cosmic inflation.''
+In the standard cosmological paradigm the early universe underwent a period of near-exponential expansion called "cosmic inflation." 
+All the cosmological observations agree with this picture, making it a compelling and elegant description of the Universe initial conditions.
+Despite the experimental effort, other possible explanations are still valid, and a conclusive evidence of inflation is still to be found.
 Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski15}.
-These primordial gravitational waves in the early universe imprint a unique signature on the polarized anisotropies of the CMB.
-For this reason, CMB polarization is a promising dataset to understand the physics of the early universe.
-Furthermore, CMB polarization is usually studied in Fourier space by decomposing the fields into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
-The B-mode polarization, in particular, is a great source of information because, in the standard scenario, the PGW background is the only source of B-modes at the epoch of recombination.
-
-To connect data and theoretical models, the amplitude of this primordial B-mode component is parameterized by the ratio of the amplitudes of the primordial tensor and scalar perturbations, $r$.
-The shape of this component, i.e.
-the amplitude as a function of the angular scales is encoded in the tensor tilt $n_{t}$.
-Most of the inflationary models directly relate the value of r to the energy scale of inflation.
-Measuring or constraining the value of $r$ and $n_t$ will put a tight constraint on the allowed models of inflation.
+This prediction set inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologists are looking for.
+These primordial gravitational waves in the early universe would imprint a unique signature on the polarized anisotropies of the CMB.
+For this reason, CMB polarization is a promising dataset to understand the physics of the early universe and ultimately test inflation.
+In particular, we can decompose the CMB polarization fields in Fourier space into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
+ In the standard scenario, the B-mode polarization is a clean probe of primordial gravitational waves, because these are the only source of B-modes at the epoch of recombination.
 
 Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational waves background, improving their measurement is a major objective of current and future CMB experiments.
-Even if inflationary B-modes have not been detected yet, a possibly natural value $r \gtrsim 10^{-3}$, should be reachable given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
-
-However reducing the level of noise will not be enough to reach this goal.
-Indeed the observed B modes are not solely due to primordial gravitational waves but also by secondary effects taking place in the late-time low-redshift universe.
-In this work, we will focus on the effects of gravitational interactions with large scale structures (LSS).
+Even if inflationary B-modes have not been detected yet, a natural value $r \gtrsim 10^{-3}$ should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
+However, just reducing the level of noise will not be enough to attain this goal.
+Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe. 
+In this work, we will focus on the effect of gravitational interactions with large scale structures (LSS).
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
@@
 
-This component acts as a source of confusion for searches of the PGW background.
+This component acts as a source of confusion for searches of the primordial gravitational wave background.
 Indeed, the contamination from lensing B modes is already at the level of the instrumental noise of current experiments \citep{bk14}.
-Thus together with the experimental effort to reduce the amount of noise is important to study the effect of the spurious lensing component.
-The optimal way to reduce the lensing contributions is to reconstruct the expected realization of lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
-We can delens the observed B-modes combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
+Thus, together with the experimental effort to reduce the amount of noise, the effect of the spurious lensing component must be understood.
 
+The optimal way to reduce the lensing contributions to the B-modes is to reconstruct the expected realization of lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
+We can delens the observed B-modes by combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
 Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
-Furthermore, recently it has been performed on CMB temperature data using as LSS tracers the cosmic infrared background in \cite{larsen:2016} and on CMB temperature and polarization data using internal CMB lensing reconstruction \cite{ carron17}.
-Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in where $~28\%$ of the lensing power has been removed.
+Furthermore, it has recently been performed on CMB temperature data using the cosmic infrared background as LSS tracer \cite{larsen:2016} and on CMB temperature and polarization data using CMB data to internally reconstruct the LSS lensing potential \cite{ carron17}.
+Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in which $~28\%$ of the lensing power was removed.
 
-In this paper, we study the possible role of future galaxies surveys as tracers of the lensing potential.
-These will translate into a better reconstruction of the B-modes in the measured patch, and as a consequence, it will improve the constraints on inflation trough delensing.
-We model several actual and future surveys, and after computing the residual B-modes, we forecast the statistical uncertainness on $r$ and $n_T$ for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
+For future experiments, we need to increase the delensing efficiency by almost a factor of 3 to fully exploit the expected instrumental capabilities \cite{abazajian:2016}.
+In this paper, we propose and study a possible way: using future galaxy surveys as tracers of the lensing potential in addition to other probes such as the SKA radio continuum survey, cosmic infrared background and internal CMB reconstruction.
+Furthermore, we point out how using redshift information through tomographic binning can improve the delensing efficiency of galaxy survey.
+These will translate into a better reconstruction of the B-modes in the measured patch, and, as a consequence will improve the constraints on inflation through delensing.
+We model several actual and future surveys, and after computing the residual B-modes, we forecast the resulting statistical uncertainties on the amplitude and the shape of the inflationary tensor perturbations for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
 
-We organize this article as follow: in \refsec{th} we describe the lensing B-mode components and the residual power after delensing with tracers of the lensing potential.
-We describe the LSS tracers used in this analysis in \refsec{model}.
-\refsec{for} is the main results of this work: the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
+We organize this article as follow: we describe the LSS tracers used in this analysis in \refsec{model}.
+In \refsec{th} we define the lensing B-mode component and the residual power after delensing with tracers of the lensing potential.
+The main result of this work is \refsec{for} : the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
 We conclude in \refsec{concl}.
-
-\chapter{Gravitational lensing B-mode and delensing}
-\label{sec:th}
-
-As for the CMB photons temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is remapped by lensing as:
-\be
-Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi);~~
-U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi)
-\ee
-where the deflection angle is the gradient of lensing potential potential integrated along the line of sight $\nabla\phi$.
-The CMB polarization is usually decomposed into odd-parity Fourier modes E and B. 
-As shown in \cite{} this for symmetry reason tensor perturbations are the principal source of the B-modes configuration.
-For this reason, B-modes are a promising signature of early universe tensor perturbations.
-However, this promising sign of primordial gravitational waves is partially obscured by a secondary mechanism that can produce B-modes also in the late universe.
-Indeed the gravitational interaction with large scale structure generates CMB B-modes by distorting primordial E-modes.
-
-As a first approximation, given a convergence field $\kappa= -\frac{1}{2}\nabla^2\phi $ the B mode resulting from the lensing of primordial E mode is:
-\be
-B^{\mathrm{lens}}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') E(\ublu) \kappa(\bl - \bl')
-\label{eqn:blens}
-\ee
-where different modes contributes with a different weight given by:
-\be
-W(\bl,\bl') = \frac{2 \bl' \cdot (\bl-\bl')}{|\bl-\bl'|^2} \sin(2\varphi_{\bl,\bl'}).
-\ee
-
-As usual we define the power spectrum as:
-\be
-\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l .
-\ee
-
-From this we get that the power spectrum of the lensing component of the B-modes:
-\be
-C_l^{BB,\mathrm{lens}}  = \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl')C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}.
-\ee
-
-The full B-mode power spectrum measured on the sky is given by a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_l^{BB}$:
-\be
-C_l^{BB,\mathrm{full}} = C_l^{BB,r} + C_l^{BB,\mathrm{lens}} + N_l^{BB}.
-\ee
-Given the current constraints, the lensing component is a significant source of B-modes that, for GW searches purposes correspond to a white noise source of roughly $5 \mu K$-arcmin.
-This means that it is not only bigger than the biggest allowed GW contribution at scales bigger than several degrees, but it is comparable with the decreasing level of instrumental noise.
-For this reason, it is critical to characterize and eventually remove it from the data.
-To do so, we need to build a template \refeq{blens} (or its real-space analog) of the actual lensing modes given a measurement of the E-mode field and the lensing potential $\phi$.
-While E is measured directly, we ca estimate $\phi$ using "tracers" of the dark matter distribution that creates the potential.
-
-We will now show how much the delensing efficiency is related to the fidelity of the lensing tracers.
-If we have a large scale structure field I($\nver$) that traces the lensing potential responsible for the lensing of the CMB we can build a template of the lensing B mode on the sky by a weighted convolution:
-\be
-\hat{B}^{\mathrm{lens}}(\bl) = \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') f(\bl,\bl') E^N(\ublu) I(\bl - \bl')
-\ee
-where $f(\bl,\bl')$ can be determined byminimizing the difference with the true $B^{\mathrm{lens}}(\bl)$ defined in \refeq{blens}.
-
-The residual lensing B mode due to an imperfect knowledge of the true E-mode and $\phi$ will be
-\beqn
-B^\mathrm{res}(\bl) &=&  B^\mathrm{lens}(\bl) - \hat{B}^\mathrm{lens}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') \times \nonumber \\ &&\left( E(\ublu) \kappa(\bl - \bl' ) -  f(\bl,\bl') E^N(\ublu) I(\bl - \bl' ) \right)
-\eeqn
-
-The weights $f(\bl,\bl')$ so that the residual lensing B mode power is minimized are:
-\be
-f(\bl,\bl') = \left(\frac{C^{EE}_{{l'}}}{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right)  \frac{C^{\kappa I}_{|\bl-\bl'|}}{C^{II}_{|\bl-\bl'|} }
-\ee
-Notice that the first term consists in an inverse variance filter applied to the measured E-mode and the second minimize the difference between the reconstructed $\phi$ and the CMB lensing potential.
-
-With this choice of $f(\bl,\bl')$ we finally have that the residual power is:
-\beqn
-\label{eqn:Bres}
-C_{{l}}^{BB,\mathrm{res}} &=&   \int \frac{d^2 \bl'}{(2 \pi)^2}  W^2 (\bl,\bl')
- C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}  \\
-&\times&  \left[1 - \left(\frac{C^{EE}_{{l'}} }{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right) \rho^2_{|\bl-\bl'|} \right] \nonumber
-\eeqn
-with
-\be
-\rho_l= \frac{\cl^{\kappa I}}{\sqrt{\cl^{\kappa \kappa} \cl^{I I}}}.
-\ee
-
-The bigger $\rho_l$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
-A higher correlation allows for a better reconstruction of the $\phi^{CMB}$ and, as a consequence, of $B^{lens}$.
-
-\section{Multiple tracers of the lensing potential}\label{sec:galcontrib}
-
-Let's now assume that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
-We can optimally combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ with:
-\beqn
-I         &=&    \sum_{i}c^{i}I^{i} \nonumber \\
-c_{i}     &=& (C^{-1})_{ij}C^{\kappa I^{j}}
-\label{eqn:combined}
-\eeqn
-where C is the covariance matrix of the LSS tracers.
-The residual B-mode power can we computed using \refeq{Bres} using an ``effective'' correlation of these combined tracers with gravitational lensing is:
-\be
-\rho^{2} = \sum_{i,j}\frac{C^{\kappa i}~(C^{-1})_{ij}~C^{\kappa j}}{C^{\kappa\kappa}}.
-\label{eqn:rho-combined}
-\ee
-
-Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing, but it also depends on how much it is correlated with the already used set of tracers.
-\reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
-Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
-However given the significant low noise of this measurement they can still help the delensing of the CMB.
-Furthermore, the low-redshift tail of the lensing kernel is also the one affected by large-scale structures non-linearities.
-They can indeed be extremely helpful if future CMB surveys will try to remove that contribution instead of attempting to model it.
-On the other end, CIB and 21 cm surveys are probing to high redshift structures and independently from the model assumed they show a relatively good overlap with the CMB lensing kernel.
-
-% ===================
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
-The bigger the overlap wit the CMB lensing kernel the better the reconstruction of the lensing potential will be.
-This will lead to a better delensing.
-}
-\label{fig:kernel}
-
-\end{center}
-\end{figure}
-% ===================
 
@@
 \label{sec:model}
-In this section, we describe how the power spectra defined in \refsec{th} are computed and how we characterize different tracers considered in this work.
+In this section, we introduce the different large scale structure tracers considered in this work to reconstruct the lensing potential.
+Gravitational lensing distorts primordial E-modes generating a non-primordial B-mode component.
+Also, we define the power spectra that we will use later in \refsec{th}.
 
@@
 
-In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, $\chi(z)$ is the comoving distance and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$.
-Furthermore $W^{i}(z)$ is the kernel function of the field $i$ defined in \refeq{wkernel}.
-We compute these quantities using CAMB and HALOFIT.
+In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$. 
+Furthermore $W^{i}(z)$ is the kernel function of the field $i$ introduced in \refeq{wkernel}.
+We will now describe the kernels for each of the tracers used in this work.
 
@@
 In that case, we can treat it as a noisy tracer of the true field.
-Both the CMB lensing fields and its noisy reconstructed counterparts have the same kernel $W^{\kappa}(z)$.
+Both the CMB lensing field and its noisy reconstructed counterpart have the same kernel $W^{\kappa}(z)$.
 However, when computing the power spectrum of the latter, we need to add a noise component.
-Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_l^{\kappa \kappa}$, and so
-\beq
-\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_l
+Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_{\ell}^{\kappa \kappa}$, and so
+\beq\label{eqn:noisekappa}
+\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_{\ell}
 \eeq
-In this work, the level of noise is computed assuming an iterative quadratic estimator is used in the reconstruction
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/cmb_internal.pdf}
-\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
-CMB measurements will rapidly improve providing the most efficient even if biased way to reconstruct the structures lensing the CMB.}
-\end{center}
-\label{fig:corr-internal}
-
-\end{figure}
+In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}.
 
@@
 
-We can also write the galaxy overdensity field $g(\nver)$ in a given direction on the sky as a line of sight integral of the matter overdensity:
-\begin{equation}
-g(\nver) = \int_0^{z_*} dz\, W^{g}(z)\delta(\chi(z)\nver,z),
-\end{equation}
-where the kernel is
+The galaxy clustering kernel is:
 \begin{equation}
@@
 \end{equation}
-Here $\frac{dN}{dz}$ is the number of galaxies as a function of redshift observed by the survey while $b(z)$ is the galaxy bias that connects the amplitude of galaxies overdensity to the one of the dark matter at different redshift.
-When computing the auto-spectrum of a galaxies density, a shot noise term needs to be taken into account. 
+Here $\frac{dN}{dz}$ is the number of galaxies observed by the survey as a function of redshift while $b(z)$ is the galaxy bias that connects the amplitude of galaxy overdensities to the underlying dark matter density.
+When computing the auto-spectrum of the galaxy density, a shot noise term needs to be taken into account. 
 To do so, we add a constant term to the power spectrum equal to the inverse of the number of galaxies per steradians.
+ Different galaxy surveys in this work are then fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxy density.
+We test the delensing efficiency taking into account both current surveys like WISE or DES as well as future galaxy surveys like DESI and LSST together with 21 cm measurement like SKA.
 
-Different galaxy surveys in this work are fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxies density.
-We test the delensing efficiency modeling both current survey like WISE or DES as well as future galaxy probes like DESI and LSST together with 21 cm measurement like SKA.
+The WISE survey observed the entire sky in the infrared \cite{wright:2010}.
+We defined the redshift distribution of the WISE infrared galaxy samples following \cite{yan:2013} (see Fig. 4 therein). 
+To compute the noise term, we assume that the available sky after masking is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxy density is approximately uniform.
+Furthermore we adopt a linear bias $b_{\rm{WISE}}=1.41$ obtained by \cite{ferraro:2015} cross correlating WISE with Planck lensing potential.
 
-We defined the WISE redshift distribution following \cite{yan:2013} (see Fig. 4 therein). 
-To compute the noise term, we assume that the available sky after masking WISE is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxies density is approximately uniform.
-DES is modeled after the DES Science Verification data public release.
-For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report.
-LSST is modeled following \cite{schaan:2016} as $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
+DES is modeled after the DES Science Verification public data release.
+For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy denisty of 0.63 galaxies per arcmin squared. 
+%As in \cite{passaglia:2017} we use a linear bias $b$
+
+For LSST we follow \cite{schaan:2016}: $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
 Furthermore we assume a density of 26 galaxies per arcmin squared.
-Finally, we model both the SKA redshift distribution and bias survey following \cite{namikawa:2016a}.
+
+Finally, we consider the Square Kilometre Array (SKA).
+The SKA is a planned radio array that will survey large scale structure primarily by detecting the redshifted neutral hydrogen (HI) 21cm emission line from a large number of galaxies out to high redshift.
+We will assume an intensity mapping survey mode where SKA will measure the large-scale fluctuations of the \textit{integrated} 21cm intensity from many unresolved galaxies.  
+We model both the redshift distribution and  bias of radio sources following \cite{namikawa:2016a}.
+
+%bias for the Radio Continuum (RC) survey conducted by SKA
 
@@
 
-We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$. Also $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
+We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$ and we set $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
 
-%With these assumptions, depending on angular scale, $45-65\%$ of the CIB is correlated with the CMB lensing potential, as shown in \reffig{corrnow}.
+\chapter{Gravitational lensing B-mode and delensing}
+\label{sec:th}
+The large scale structures described in \refsec{model} have an important impact on the search of primordial CMB B-modes: they lens the primordial E-modes generating non inflationary B-modes that constitues an important source of noise.
 
-\cite{hall:2010,sherwin:2015,ade:2014,hanson:2013,szapudi:2001,0004-637X-567-1-2,planck-collaboration:2014,planck-collaboration:2014a,planck-collaboration:2011,plank-collaboration:2014,boulanger:1996,lewis:2006}
+Indeed the Q and U mode decompositions of the CMB photons polarization are remapped by lensing as:
+\be
+Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi);~~
+U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi)
+\ee
+where the deflection angle is the gradient of the lensing potential integrated along the line of sight $\nabla\phi$.
+The CMB polarization is usually decomposed into odd-parity Fourier modes E and B. 
+As shown in \cite{}, because of the symmetry of the problem, tensor perturbations are the principal source of the B-modes configuration.
+For this reason, B-modes are a promising signature of early universe tensor perturbations.
+
+However, this promising sign of primordial gravitational waves is partially obscured by gravitational interactions with large scale structures that generate CMB B-modes by distorting primordial E-modes.
+At first order, given the convergence field $\kappa= -\frac{1}{2}\nabla^2\phi$ introduced in \refsec{kappaCMB} the B-modes resulting from the lensing of primordial E-modes are:
+\be
+B^{\mathrm{lens}}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') E(\ublu) \kappa(\bl - \bl')
+\label{eqn:blens}
+\ee
+where different modes contributes with a different weight:
+\be
+W(\bl,\bl') = \frac{2 \bl' \cdot (\bl-\bl')}{|\bl-\bl'|^2} \sin(2\varphi_{\bl,\bl'}).
+\ee
+%As usual we define the power spectrum as:
+%\be
+%\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l .
+%\ee
+
+From this we get the power spectrum of the lensing component of the B-modes:
+\be
+C_{\ell}^{BB,\mathrm{lens}}  = \int \frac{d^2 \bl'}{(2 \pi)^2} W^2 (\bl,\bl')C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}.
+\ee
+
+The B-mode power spectrum measured on the sky is composed of a possible primordial component $C_{\ell}^{BB,r}$ together with the lensing $C_{\ell}^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_{\ell}^{BB}$ (defined in \refeq{noise}):
+\be
+C_{\ell}^{BB,\mathrm{full}} = C_{\ell}^{BB,r} + C_{\ell}^{BB,\mathrm{lens}} + N_{\ell}^{BB}.
+\ee
+The lensing component is a significant source of B-modes that, at large scales, correspond to a white noise source of roughly $5 \mu K$-arcmin.
+This means that it is not only bigger than the allowed inflationary component at scales smaller than several degrees, but it is also comparable with current levels of instrumental noise.
+For this reason, it is critical to characterize and eventually remove it from the data.
+To do so, we build a template \refeq{blens} of the lensing B-modes in the observed patch given a measurement of the E-mode field and the lensing potential $\phi$.
+While E is measured directly, we can estimate $\phi$ using "tracers" of the dark matter distribution that sources the potential.
+
+We will now show how the delensing efficiency is related to the fidelity of the lensing tracers and the instrumental noise in the CMB E-modes.
+If we have a large scale structure field I($\nver$) that traces the lensing potential we can build a template of the lensing B-modes on the sky by a weighted convolution:
+\be
+\hat{B}^{\mathrm{lens}}(\bl) = \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') f(\bl,\bl') E^N(\ublu) I(\bl - \bl'),
+\ee
+where $f(\bl,\bl')$ can be determined by minimizing the difference with the true $B^{\mathrm{lens}}(\bl)$ defined in \refeq{blens}.
+We introduce the instrumental noise in the CMB E-modes ($E^N$) that will also limit the ability to fully reconstruct the lensing B-modes.
+
+The residual lensing B-modes due to an imperfect knowledge of the true E-mode and $\phi$ will be
+\beqn
+B^\mathrm{res}(\bl) &=&  B^\mathrm{lens}(\bl) - \hat{B}^\mathrm{lens}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') \times \nonumber \\ &&\left( E(\ublu) \kappa(\bl - \bl' ) -  f(\bl,\bl') E^N(\ublu) I(\bl - \bl' ) \right).
+\eeqn
+
+The optimal weights $f(\bl,\bl')$ such that the residual lensing B mode power is minimized are:
+\be
+\label{eqn:fweight}
+f(\bl,\bl') = \left(\frac{C^{EE}_{{l'}}}{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right)  \frac{C^{\kappa I}_{|\bl-\bl'|}}{C^{II}_{|\bl-\bl'|} }.
+\ee
+Here $C^{\kappa I}$ and $C^{II}$ are the cross-correlation spectrum of the tracer $I$ with the lensing convergence $\kappa$ and its autospectrum; they are described for each LSS field in \refsec{model}.
+The power spectrum of the E-modes noise $N^{EE}$ is the same as the B-modes one in \refeq{noise}.
+
+Notice that the first term consists of an inverse variance filter applied to the measured E-mode and the second minimizes the difference between the reconstructed $\phi$ and the CMB lensing potential.
+
+With this choice of $f(\bl,\bl')$ we finally have that the residual power is:
+\beqn
+\label{eqn:Bres}
+C_{{l}}^{BB,\mathrm{res}} &=&   \int \frac{d^2 \bl'}{(2 \pi)^2}  W^2 (\bl,\bl')
+ C^{EE}_{l'} C^{\kappa \kappa}_{|\bl-\bl'|}  \\
+&\times&  \left[1 - \left(\frac{C^{EE}_{{l'}} }{C^{EE}_{{l'}}+N^{EE}_{{l'}}}\right) \rho^2_{|\bl-\bl'|} \right] \nonumber
+\eeqn
+with
+\be\label{eqn:rho1}
+\rho_{\ell}^{2}= \frac{(\cl^{\kappa I})^{2}}{\cl^{\kappa \kappa}\cl^{I I}}.
+\ee
+
+\refeq{Bres} directly relate the residual power after delensing with the cross correlation coefficients with CMB lensing of the tracers used.
+The bigger the $\rho_{\ell}^{2}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
+An higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$ leading to a smaller residual power $C_{{l}}^{BB,\mathrm{res}}$.
+\relook{We conclude this section showing in \reffig{bbplot} the expected residual lensing B-modes power spectrum for some of the tracers used in this work together with the primordial B-modes component and the instrumental noise for current and future experiments.
+}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.2,clip]{../images/BB_res_ell2.pdf}
+\caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
+The red solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed green line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
+The residual power left after delensing for some of the LSS tracers described in \refsec{model}. 
+The rapid improvement in the level of instrumental noise (dashed curves for SPTPol and CMB S4) will require a high level of delensing to exploit these experiments fully.}
+\label{fig:bbplot}
+\end{center}
+\end{figure}
+
+\section{Multiple tracers of the lensing potential}\label{sec:galcontrib}
+
+In this section we extend the formalism to the case where multiple tracers are used to reconstruct the lensing potential. 
+%Furthermore we show that splitting galaxy surveys into multiple tomographic bins can significantly improve their efficiency.
+
+We start by assuming that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
+We can optimally combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ with:
+\beqn
+I         &=&    \sum_{i}c^{i}I^{i} \nonumber \\
+c_{i}     &=& (C_{II}^{-1})_{ij}C^{\kappa I^{j}}
+\label{eqn:combined}
+\eeqn
+where $C_{II}$ is the covariance matrix of the LSS tracers.
+The residual B-mode power can be derived from \refeq{Bres} using an ``effective'' correlation $\rho^{2}$ of these combined tracers with gravitational lensing:
+\be
+\rho^{2}_{\ell} = \sum_{i,j}\frac{C^{\kappa i}_{\ell}~(C^{-1}_{\ell})_{ij}~C^{\kappa j}_{\ell}}{C^{\kappa\kappa}_{\ell}}.
+\label{eqn:rho-combined}
+\ee
+
+Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing, but it also depends on how much it is correlated with the already used set of tracers.
+\relook{
+\reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
+Note that the cross correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
+CIB and 21 cm surveys probes the high redshift structures and, independently from the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
+On the other end, galaxy clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
+However, given the low noise of this measurement and their small overlap with other probes they can still play an important role in delensing even if their overlap with the CMB lensing potential is not optimal.
+}
+
+% ===================
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.2,clip]{../images/compare_kernel.pdf}
+\caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
+The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential will be leading to a higher delensing efficiency.
+}
+\label{fig:kernel}
+\end{center}
+\end{figure}
+% ===================
+
+\section{Improving efficiency with tomographic binning}\label{sec:tomo-bin}
+The delensing efficiency of galaxy surveys can be improved by taking into account redshift information.
+When we weight a tracer with $\frac{C^{\kappa I}}{C^{II}}$ in \refeq{fweight} in order to maximize its ability to reconstruct the lensing potential we are only using redshift averaged information about the survey.
+However, as can be seen in \reffig{kernel}, the kernel overlap of a tracer with the CMB lensing varies as a function of redshifts.
+For this reason, the optimal approach is to weight galaxies at different redshifts with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
+We can see this with a simple example. Let's split a single survey I into two non-overlapping redshift bins $I_{1}$ and ${I_{2}}$ with $I=I_{1}+I_{2}$.
+For the full survey the effective cross-correlation is equal to 
+\be
+\rho_{\rm{full}}^{2}= \frac{(\cl^{\kappa I})^2}{\cl^{\kappa \kappa} \cl^{I I}}=\frac{(\cl^{\kappa I_{1}}+\cl^{\kappa I_{2}})^2}{\cl^{\kappa \kappa} (\cl^{I_{1} I_{1}}+\cl^{I_{2} I_{2}})}
+\ee
+while for the split survey it will be 
+\be
+\rho_{\rm{split}}^{2}= \frac{(\cl^{\kappa I_{1}})^2}{\cl^{\kappa \kappa} \cl^{I_{1} I_{1}}}+\frac{(\cl^{\kappa I_{2}})^2}{\cl^{\kappa \kappa} \cl^{I_{2} I_{2}}}.
+\ee
+Now it can be show that $\rho_{\rm{split}}\ge \rho_{\rm{full}}$ being:
+\be
+\rho_{\rm{split}}^{2} - \rho_{\rm{full}}^{2}\propto  (\cl^{I_{1} I_{1}}\cl^{\kappa I_{2}}-\cl^{I_{2} I_{2}}\cl^{\kappa I_{1}})^2
+\ee
+Then $\rho^2$ is always bigger in the tomographic case, and the two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the same for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
+Of course, binning will improve the efficiency of galaxy surveys, but it is not applicable to tracers with poor redshift information like the CIB or radio continuum surveys.
+
+In this work, we bin both photometric and spectroscopic galaxy surveys by splitting the window function \refeq{wg} into different slices such as all the bins contain the same number of galaxies. 
+For photometric surveys like DES and LSST, we assume a photometric redshift estimation gaussianly distributed around the true value with an rms fluctuation $\sigma(z)$.
+
+In that case the i$^{th}$ slice has a galaxy distribution \cite{hu:2004}:
+\begin{eqnarray}
+\label{eqn:zbin}
+W_{i}(z) \propto b(z) {\frac{dN(z)}{dz}} 
+\Bigl[{\,\rm erfc}\left({\Delta(i-1) -z \frac{\sigma(z)}{\sqrt{2}}}\right) 
+\\ \nonumber - {\,\rm erfc}
+\left({\Delta i  - \frac{z}{\sigma(z)}\sqrt{2}}\right)\Bigr].
+\end{eqnarray}
+
+% ===================
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.2,clip]{../images/compare_kernel_tomo.pdf}
+\caption{\textbf{Kernels Comparison with tomographic bins}: Comparison of the 10 LSST  tomographic bins together with the full LSST survey and the CMB lensing kernel. 
+Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross correlation with the CMB lensing.
+This leads to a better delensing efficiency.
+}
+\label{fig:kerneltomo}
+\end{center}
+\end{figure}
+% ===================
+
+For photometric surveys, the maximum number of bins is dictated by the fact that the bin width can not be smaller than the photo-z accuracy.
+We used 10 and 4 photometric bins for LSST and DES respectively with a common photo-z accuracy $\sigma(z) = 0.05(1+z)$. 
+\relook{In spectroscopic surveys, there are no limitations in increasing the number of bins. We split DESI into 4 spectroscopic bins with no overlap among each others.
+This number of bins is close to the saturation point where adding more bins does not improve delensing significantly while adding complexity to the analysis.} As an example, we show in \reffig{kerneltomo} the 10 bins and the full LSST redshift distributions together with the CMB lensing kernel.
+\relook{The improvement obtained by tomographic binning is illustrated in \reffig{rho-tomo}.
+In particular on large scales, binning can increase the value of $\rho$ by almost $30\%$ significantly improving the delensing efficiency of galaxy surveys.}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.2,clip]{../images/B_res_bin.pdf}
+\caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing.
+Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of angular scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line).}
+\label{fig:rho-tomo}
+\end{center}
+\end{figure}
 
@@
 \label{sec:for}
-In this section, we forecast the expected delensing efficiency and the relative importance of galaxies tracers for delensing in current and future experiments.
-We will use the Fisher information matrix to quantify the delensign efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
-We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction efficiency, together with a low-noise, low-resolution experiment which B-modes will be delensed and used to constrain the inflationary parameters.
-We focus on three different scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
 
-\section{Fisher method}
-In this framework, the error on a cosmological parameter constrained by a B-mode spectrum measurement is given as usual by:
-\begin{equation}
-\label{eqn:fisher}
-\sigma(p)= \left[\sum_l^{\ell_{\rm{max}}} \left(\frac{\partial {C}_{\ell}^{BB,r,n_{t}}}{\partial p}\right)^2/\mathrm{Cov}^2( C_{\ell}^{BB} ) \right]^{-\frac{1}{2}}
-\end{equation}
-where we consider the parameters $p =\{r,n_{t}\}$. 
-We will assume a gaussian covariance:
-\beq
-\sigma( C_{l}^{BB} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_l^{BB,r} + C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
+In this section, we forecast the expected delensing efficiency and the relative importance of galaxy tracers for delensing in current and future experiments.
+We will use the Fisher information matrix to quantify the delensing efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
+We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction performance, together with a low-noise, low-resolution experiment the B-modes of which \textcolor{red}{(?)} will be delensed and used to constrain the inflationary parameters.
+We focus on three distinct scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
+
+%\begin{figure}[htbp]
+%\begin{center}
+%\includegraphics[scale=1.2,clip]{../images/cmb_internal.pdf}
+%\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
+%CMB measurements will rapidly improve providing the most efficient (even if biased) way to reconstruct the structures lensing the CMB.}
+%\end{center}
+%\label{fig:corr-internal}
+%\end{figure}
+
+\section{Fisher Information Matrix}
+In the Fisher Information Matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtained from the inverse of the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
+We constrain the inflationary parameters $p =\{r,n_{t}\}$ with a CMB B-mode spectrum measurement so the Fisher matrix is:
+\beq\label{eqn:fisher}
+F_{ij} = \sum^{\ell_{max}}_{\ell=\ell_{min}}\frac{1}{\sigma( C_{\ell}^{BB} )^2}\frac{\partial {C}_{\ell}^{BB}}{\partial p_{i}}\frac{\partial {C}_{\ell}^{BB}}{\partial p_{j}}
 \eeq
-Here the B-modes noise spectrum is given by:
-\beq
-N_l^{BB} = \left({\Delta_P}/{T_{\mathrm{CMB}}}\right)^2  e^{ {l^2 \theta_\mathrm{FWHM}^2}/({8 \ln 2})}
+where we assume a Gaussian covariance:
+\beq\label{eqn:cova}
+\sigma( C_{\ell}^{BB} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_{\ell}^{BB,r} + C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
 \eeq
-where $\theta_\mathrm{FWHM}$ is the full half width of the telescope beam and $\Delta_P$ is the instrumental noise of the experiment.
-Even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012} this approximation is good enough to show the improvement due to delensing.
+The B-modes noise spectrum is given by \cite{knox:1995}:
+\beq \label{eqn:noise}
+N_{\ell}^{BB} = \left({\Delta_P}/{T_{\mathrm{CMB}}}\right)^2  e^{ {l^2 \theta_\mathrm{FWHM}^2}/({8 \ln 2})}
+\eeq
+where $\theta_\mathrm{FWHM}$ is the full half width of the telescope beam, and $\Delta_P$ is the instrumental noise of the experiment.
 
-The reason why removing the lensing contribution will improve parameters constraints can be seen from \refeq{fisher}.The parameter uncertainties are inversely proportional to the covariance of the measurement.
-Since lensing B-mode $C_{{l}}^{BB,\mathrm{lens}}$ is a substantial component of the covariance, removing part of them trough delensing will improve parameter constraints.
+It can be seen in \refeq{fisher} that removing the lensing contribution will improve parameters constraints.
+The parameter uncertainties are inversely proportional to the covariance of the measurement.
+Since the lensing B-modes $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them trough delensing will reduce the parameter statistical error.
+After delensing the error $\sigma^{\rm{del}}(p)$ can still be obtained from \refeq{fisher} with $C_{{l}}^{BB,\mathrm{res}}<C_{{l}}^{BB,\mathrm{lens}}$ in the covariance defined in \refeq{cova}.
+We define the improvement as the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$.
 
-Note that we are fixing the other cosmological parameters in \refeq{fisher}.
-Uncertainties in those will however propagate to uncertainties in $\{r,n_{t}\}$.
-While the constraints will be slightly optimistic, this has no real impact on the estimate of the improvement due to delensing which is the main focus of this work.
-We define that improvement as the ration of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$
+In \refeq{fisher} we are making a few important assumptions. 
+Firstly, we are fixing all the cosmological parameters except $\{r,n_{t}\}$.
+Uncertainties in those will, however, propagate to uncertainties in $\{r,n_{t}\}$.
+While neglecting this will lead to slightly optimistic constraints, it has no significant impact on the estimate of the improvement due to delensing.
+
+We are also neglecting an important contribution to the measured CMB B-modes: galactic polarized foregrounds. 
+The amplitude of these has been constrained in \cite{bicep2/keck-collaboration:2015, planck-collaboration:2015} and it strongly varies in different parts of the sky.
+Future experiments will use multi-band data to exploit the frequency dependence of these contaminants to remove them from the data.
+The amount of residual foregrounds depends both on uncertain foreground properties and experimental choices (see a review in \cite{abazajian:2016}).
+For this reason, accurately treating foreground requires the use of simulations and the knowledge of several experimental details. 
+We decided to focus on an ideal situation assuming no foregrounds or perfect cleaning even if the importance of delensing will be slightly overestimated.
+
+Furthermore, we are not considering the uncertainties on galaxy survey internal parameters such as biases, source distributions and photometric redshift uncertainties.
+These uncertaintes can significantly degrade inflationary constraints \cite{sherwin:2015,namikawa:2016a}.
+However, as shown in \cite{sherwin:2015,namikawa:2016a}, these can be auto-calibrated i.e.
+they can be tightly constrained using galaxy survey auto- and cross-correlation spectra.
+We checked this for a few of the tracer combinations used here and find it particularly true once several tracers are jointly taken into account.
+Finally, we have assumed a Gaussian covariance even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
+This approximation is good enough to show the improvement due to delensing.
+
 %===================
@@
 
-Very recently delensing has been proven possible on real data using as tracers both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as CIB data \cite{larsen:2016a,manzotti:2017}.
-Here we discuss the improvement due to combining these and other available surveys (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
+Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} as well as the internal CMB lensing potential reconstruction \cite{carron:2017} as large scale structure tracers.
+Here we discuss the improvement that can be obtained by combining these and other currently available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
 
-On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin} and a beam of \textcolor{red}{30 arcmin}.
-The CMB lensing reconstruction will be performed by a higher resolution CMB experiment. We explore two possibilities.
-First, we test the efficiency with an internal reconstruction performed by Planck \cite{2016A&A...594A..15P} using the CMB lensing noise curves publicly available
-\footnote{\textit{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
-Then we will assume a ground-based experiment with the final noise level of SPT-Pol which corresponds to \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
+On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment.
+For the deep experiment we assume an instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin}, a beam of \textcolor{red}{30 arcmin} and an angular scale range of $50<\ell<500$.
+We assume the CMB lensing reconstruction is  performed by the higher resolution CMB experiment. We explore two possibilities.
+First, we test the efficiency with an internal reconstruction of CMB lensing performed by Planck collaboration \cite{planck-collaboration:2016a}.
+We compute the noise in the CMB lensing map in \refeq{noisekappa} using the noise curves publicly available
+\footnote{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}.
+Then we assume a ground-based experiment with noise levels consistent with SPT-Pol: \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
+In this case we set the largest achievable scale at $\ell_{min}=300$.
+However, the results are robust against this choice since the neglected scales contribute only moderately to the reconstruction of the CMB lensing potential.
 
-Furthermore, we test how much not only CIB maps but also current low redshift galaxies surveys like DES and WISE can improve the delensing efficiency.
+We combined these CMB experiments with the CIB and current low redshift galaxy surveys like DES and WISE and we compute the improvement in the delensing efficiency.
+Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals.
+Optical surveys galaxies like DES are less affected by dust and can be used at larger scales.
+The achievable correlation is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
+In these figures we include a dashed curve that corresponds to (with arbitrary scale) $<C_{\ell}^{\kappa \kappa} \times \frac{\partial C^{BB}_{\ell'}}{\partial{C^{\kappa \kappa}_{\ell}}}>_{\ell ' <100}$, in order to show which scales in $\kappa$ contribute to the $\ell<100$ B-mode power.
 
-The correlation attainable using current experiments is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
-In general, we observe that CMB lensing reconstruction correlate very well at low $\ell$  and it then falls rapidly because of the raise of the reconstruction noise.
-For the noise levels of Planck, the internal reconstruction is at most $70\%$ correlated at very large scales and falls rapidly to $40\%$ at $\ell=200$.
-Its correlation is comparable to the on of LSS at almost all scales.
-On the other end, the CMB lensing reconstruction \cite{} from SPT-Pol will be more than 80\% correlated with the true field at large scales $\ell<300$, where the correlation become comparable with the current LSS surveys.
+In both cases the CMB lensing reconstruction correlates very well at low $\ell$ and it then falls rapidly at smaller scales because of the raise of the reconstruction noise.
+For Planck, the internal reconstruction is at most $70\%$ correlated at very large scales and then it falls rapidly to $40\%$ at $\ell=200$.
+Its correlation is comparable to the one of LSS at almost all scales.
+On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 80\% correlated with the true field at $\ell<300$, and only then the correlation becomes comparable with the current LSS surveys.
 
-Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals.
-Optical surveys galaxies like DES are less affected by that systematics and can be used at larger scales.
-Furthermore, in \reffig{corrnow} DES contribution to delensing is higher than WISE because the redshift distributions peaks at slightly higher redshift thus overlapping better with lensing.
+\reffig{corrnow} and \reffig{corrnow-planck} show that the DES galaxies are effective tracers of the LSS and can, at least in the near future, be used to improve delensing for CMB experiments that overlap with it.
+For example, DES delensing efficiency is higher than WISE firstly because of the lower level of noise and furthermore because DES galaxies are located at slightly higher redshift thus they better overlap with CMB lensing.
 
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/actual_scenario_planck.pdf}
-\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the angular scale $\ell$. 
+\includegraphics[scale=1.2,clip]{../images/actual_scenario_planck.pdf}
+\caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the angular scale $\ell$. 
 The dashed curve corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$, showing which scales contribute to the $\ell<100$ B-mode power.}
@@
 
-%At large angular scale, the CMB internal reconstruction is clearly the best tracers for delensing.
-Because of the large level of noise in current experiments, its efficiency became comparable to the CIB one at few degrees ($\ell \simeq 400$).
-
 \begin{figure}[htbp]
 \begin{center}
-\includegraphics[scale=1.]{../images/actual_scenario.pdf}
-\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
+\includegraphics[scale=1.2,clip]{../images/actual_scenario.pdf}
+\caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
 \label{fig:corrnow}
@@
 
-Using this correlation level we can, using \refeq{Bres} compute the residual B-mode power after delensing and test the consequent improvement on parameter constraints with \refeq{fisher}. 
-For current surveys we only consider $r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
+Using these correlation levels we can  compute the residual B-mode power after delensing using \refeq{Bres}, and test the consequent improvement on parameter constraints with \refeq{fisher}. 
+For current surveys we only consider $p=r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
+Note that we still let $n_{t}$ to vary in our fisher calculation even if it is not very degenerate with $r$.
 We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
-The results are summarized in \reftab{current}
+The results are summarized in \reftab{current} and \reftab{current-spt} for the Planck and SPTPol case respectively.
 
@@
 \centering
-\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to ration of the error before and after delensing.}
+
+\caption{$\alpha(r)$: Improvements on $\sigma(r)$ due to delensing for current generation and Planck lensing reconstruction.
+The values in parenthesis in the first column correspond to the fraction of lensing B-mode power removed using each LSS tracer. 
+The values in the other columns correspond to ratio of the error before and after delensing.}
   \vspace{0.2cm}
-  \begin{tabular}{ | c | c | c|}
+  \begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
-WISE & 1.10 & 1.04 \\ \hline
-DES & 1.19  & 1.07 \\ \hline
-CIB & 1.46  & 1.18\\ \hline
-LSS & 1.64  & 1.24\\ \hline
-CMB & 1.14  & 1.04 \\ \hline
-LSS+CMB & 1.71& 1.26  \\ \hline
-
+Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+WISE (6\%)& 1.09& 1.05 & 1.01 \\ \hline
+DES (14\%) &1.2 & 1.10  & 1.03 \\ \hline
+CIB (27\%) &1.45& 1.21  & 1.07\\ \hline
+LSS (36\%) &1.7& 1.29  & 1.09\\ \hline
+CMB Planck (8\%) &1.06& 1.05  & 1.01 \\ \hline
+LSS+CMB (41\%)& 1.77& 1.31& 1.10  \\ \hline
 \end{tabular}
@@
 
-When the same tracers are used, the results are in agreement with \cite{yu:2017,larsen:2016a}, with the CIB being the predominant source of delensing with an improvement of almost 50\%.
-Planck and DES are essential to reconstruct the largest scales of the lensing potential and as such the play a major role in delensing.
-The improvement due to WISE is quite marginal, but given the big sky coverage, it might be useful to delens large CMB polarization experiments.
-For reference we also test a scenario where GW are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.
+We also test a scenario where primordial gravitational waves are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.
 As expected the importance of delensing itself is now reduced given that the lensing component constitutes a smaller portion of the total B-modes variance.
 
+\reftab{current-spt} that uses SPTPol instead of Planck for the internal reconstruction shows consistent results.
 As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS. 
 This can be seen in \reftab{current-spt}. 
-In particular optical surveys will rapidly lose the role of filling in large-scale modes and they start, as the CIB, to supplement information about lensing at higher multipoles.
+In particular optical surveys will rapidly lose the role of filling in large-scale modes and they will start to supplement information at higher multipoles.
 
@@
 \vspace{0.2cm}
-  \begin{tabular}{ | c | c | c|}
+\begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
-WISE & 1.10 &  \\ \hline
-DES & 1.21  & \\ \hline
-CIB & 1.52  & \\ \hline
-LSS & 1.78  & \\ \hline
-CMB & 1.62  &  \\ \hline
-LSS+CMB & 2.35 &   \\ \hline
-
+Surveys &$\alpha(r=0), N^{B}_{\ell}=0$& $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+CMB SPTPol (35\%) && 1.62  &  \\ \hline
+LSS+CMB (56\%) && 2.35 &   \\ \hline
 \end{tabular}
@@
 
+%%%========================
+%%%========================
+
 \section{CMB-S3 Era}
-CMB measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
-To model the CMB at the beginning of the next decade, we will assume a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
-This is a reasonable level for a combination of the next generation of BK and SPT3G.
-We will assume this level of noise for the internal noise reconstruction.
+CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
+We model the CMB S3 stage with a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam. 
+We assume here that a deep CMB experiment and an overlapping high resolution one will be able to combine their measured modes.
+For this reason, we will also assume this level of noise for the internal noise reconstruction.
+With this experimental setup we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
+Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers.
 
-The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
-Not only will CMB experiments improved shortly: eventually  D.E.S.I will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
-DESI will be slightly more efficient than DES-like survey because of the slightly higher redshift that it can probe.
-However, the CIB will still play the major role among LSS.
+The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
+An interesting finding is that DESI will be less efficient (removing 10\% of the power) than a DES-like survey (14\%) despite the fact that it can probe slightly higher redshift.
+The reason is that, for the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES play a significant role in the delensing efficiency.
+Unfortunately, adding DESI will only bring the power removed using LSS fro 36 to 41\%.
+Indeed, even in the near future, CIB will still play the dominant role among LSS.
+Contrary to current experiments, th CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 550$.
 
-This generation of experiments will start exploring the small scales B-modes thus constraining, even if loosely, the value of $n_{t}$. 
-The improvement in cosmological parameters are summarized in \reftab{S3}
+%This generation of experiments will start exploring the small scales B-modes thus constraining, even if loosely, the value of $n_{t}$.
 
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/S3_scenario.pdf}
+\includegraphics[scale=1.2,clip]{../images/S3_scenario.pdf}
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
@@
 
+The improvement in the constraint on $r$ is summarized in \reftab{S3}. 
+Given the improvement in the noise of the high resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
+Despite this, adding galaxy survey will still lead to a decent improvement with a fairly low effort.
+Indeed galaxy surveys will still be able to remove an additional $20\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by $15\%$.
+Note that apart from the addition of DESI, the tracers used here are already available for the current generation.
+The ongoing effort can then lead to a significant improvement even for the next generation of experiments.
+Finally, the combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor 1.8 or to improve the constraint of a possible detection (r=0.12) by 30\%.
+As expected the improvement on $n_{t}$ is similar to the one in $r$ since they are both proportional to the variance of the measured B-modes.
+We tested this in the scenario $r=0.12$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}/8}$.
+It is important to note that, independently from the level of delensing, for CMB 3G experiments this statistical error will still be several times bigger than the fiducial value.
+
 \begin{table}
 \centering
+
 \caption{$\alpha$: S3 improvements. Note that the absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is $\sigma(r) =$  }
-  \begin{tabular}{ | c | c  |  c |}
+  \begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r)$ & $\alpha(n_{t})$\\ \hline \hline
-DESI & 1.34 & \\ \hline
-LSS & 1.71  &\\ \hline
-CMB & 1.71  &\\ \hline
+Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+DESI (10\%) &1.12 & 1.08  & 1.03 \\ \hline
+LSS (40\%) &1.73& 1.42  & 1.2\\ \hline
+CMB 3G (56\%) &2& 1.57  & 1.21 \\ \hline
+LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
 \end{tabular}
@@
 
-An ambitious program for a generation 4 ground CMB experiment is currently under planning \cite{abazajian:2016}. Moreover, satellite experiments have been proposed.
-The correlation attainable using current experiments is shown in \reffig{corrs4},
-In this futuristic but reasonable scenario, we add to the available large scale structures tracers  both 21 cm surveys like SKA and large scale structures like LSST and Euclid.
+An ambitious program for a generation 4 ground CMB experiment is currently under planning \cite{abazajian:2016}. 
+Moreover, satellite and balloon experiments have been proposed and have the potential to extend the accessible B-mode measurement to the largest scales.
+Here we follow \cite{abazajian:2016} and assume a CMB-S4 ground experiment with  with a level of noise of \textcolor{red}{$1\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
+By the mid 2020s several next generation of LSS survey will be online.
+Here, as an example for future optical galaxy surveys we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers.
+We test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} will have similar delensing performance.
+Even if they will observe similar part of the sky, because of the different  strength and weakness of this experiments
+it will still important to combined them efficiently.
+We also consider a radio continuum survey modeled following the SKA planned experiment.
+Radio continuum observation of the 21 cm line is in its early stage and several experimental and analysis challanges
+need to be overcome. However this technique has the potential to map LSS with relatively low noise up to redshift $z=6$.
+
+The correlation factor with CMB lensing of stage 4 experiments is shown in \reffig{corrs4}.
 
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/S4_scenario.pdf}
+\includegraphics[scale=1.2,clip]{../images/S4_scenario.pdf}
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
@@
 %\begin{center}
-%\includegraphics[scale=1.]{../images/S4_scenario_no_ska.pdf}
+%\includegraphics[scale=1.2,clip]{../images/S4_scenario_no_ska.pdf}
 %\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments without SKA. }
 %\label{fig:corrs4}
-%
 %\end{center}
 %\end{figure}
+
+A tomographically binned LSST-like experiment will be a very efficient CMB lensing tracer.
+Indeed it will be more than 70\% correlated with CMB lensing for all the scales $\ell<600$ a performance similar to a Stage 3 CMB internal reconstruction.
+SKA very new but potentially amazing
+With SKA and LSST, for the first we will have LSS tracers way more efficient than the CIB.
+Even if galaxy survey will improve dramatically the CMB internal reconstruction will still be the main source of delensing.
+This, having a perfect kernel overlap with the true lensing potential we will benefit for the very low level of noise of CMB S4 experiments.
+CMB S4 delensing will probably be limited by small secondary effect like foregrounds contamination, missing modes due to filtering etc.
+
+\reftab{S4} shows the improvement on the inflationary constraint.
 
@@
 \centering
+
+\caption{$\alpha$: Gen 4 experiments}
+  \begin{tabular}{|c | c | c | c|}
+\hline
+Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+LSST (10\%) &1.12 & 1.08  & 1.03 \\ \hline
+SKA (40\%) &1.73& 1.42  & 1.2\\ \hline
+CMB S4 (56\%) &2& 1.57  & 1.21 \\ \hline
+LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
+\end{tabular}
+\label{tab:S4}
+\end{table}
+
+%=======================================
+
+\begin{table}
+\centering
+
 \caption{$\alpha$: Gen 4 experiments}
@@
 LSS+CMB & 6.3  &\\ \hline
-
 \end{tabular}
@@
 
-\section{Bias uncertainties degradation}
-The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of the inflationary constraint of delensed spectra.
-In this section, we quantify this effect.
-We will now marginalize over unknown galaxies parameters, but we will use a full dataset of CMB and galaxies data.
-The idea is that as shown in the low level of noise in galaxies surveys might allow us to calibrate them internally.
-We will use a Fisher approach the Fisher matrix is:
-\beqn
-F_{pq} &=& \sum_{l_{a} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}  \sum_{l_{b} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}   \frac{\partial {C}_{l_a}^{BB,\mathrm{del}}}{\partial \theta_p } \left[ \mathrm{Cov}^{BB,BB}\right]_{l_a, l_b} ^{-1} \frac{\partial {C}_{l_b}^{BB,\mathrm{del}}}{\partial \theta_q }\nonumber \\&+&  \sum_j \frac{\frac{\partial {C}_j^{\kappa I}}{\partial \theta_p } \frac{\partial {C}_j^{\kappa I}}{\partial \theta_q }}{(\Delta C_j^{\kappa I})^2}
-+\sum_j \frac{\frac{\partial {C}_j^{I I}}{\partial \theta_p } \frac{\partial {C}_j^{I I}}{\partial \theta_q }}{(\Delta C_j^{I  I})^2}
-\eeqn
-
-\beq
-\alpha_{\mathrm{marginalized}} = \sigma_0(r) / \sigma_{\mathrm{marginalized/delensed}}(r)
-\eeq
-
-where the parameter array contains both the tensor to scalar rate $\theta=r$, and the galaxies surveys parameters like the bias $b_i$ or $p_i$ \footnote{In our analysis, since the fiducial value of $r$ is zero,
-the derivative of $C_\ell^{\rm BB,res}$ is non-zero if $\theta_i=r$.}.
-
-We compute the derivatives of the power spectra as described
+%\section{Bias uncertainties degradation}
+%The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of the inflationary constraint of delensed spectra.
+%In this section, we quantify this effect.
+%We will now marginalize over unknown galaxies parameters, but we will use a full dataset of CMB and galaxies data.
+The idea is that as shown in the low level of noise in galaxy surveys might allow us to calibrate them internally.
+%We will use a Fisher approach the Fisher matrix is:
+%\beqn
+%F_{pq} &=& \sum_{l_{a} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}  \sum_{l_{b} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}   \frac{\partial {C}_{l_a}^{BB,\mathrm{del}}}{\partial \theta_p } \left[ \mathrm{Cov}^{BB,BB}\right]_{l_a, l_b} ^{-1} \frac{\partial {C}_{l_b}^{BB,\mathrm{del}}}{\partial \theta_q }\nonumber \\&+&  \sum_j \frac{\frac{\partial {C}_j^{\kappa I}}{\partial \theta_p } \frac{\partial {C}_j^{\kappa I}}{\partial \theta_q }}{(\Delta C_j^{\kappa I})^2}
+%+\sum_j \frac{\frac{\partial {C}_j^{I I}}{\partial \theta_p } \frac{\partial {C}_j^{I I}}{\partial \theta_q }}{(\Delta C_j^{I  I})^2}
+%\eeqn
+%
+%\beq
+%\alpha_{\mathrm{marginalized}} = \sigma_0(r) / \sigma_{\mathrm{marginalized/delensed}}(r)
+%\eeq
+%
+%
+%where the parameter array contains both the tensor to scalar rate $\theta=r$, and the galaxy surveys parameters like the bias $b_i$ or $p_i$ \footnote{In our analysis, since the fiducial value of $r$ is zero,
+%the derivative of $C_\ell^{\rm BB,res}$ is non-zero if $\theta_i=r$.}.
+%
+%We compute the derivatives of the power spectra as described
 
@@
 
-Delensing or the ability to separate the lensing component of the CMB B-mode from a possible primordial inflationary signal is important to exploit the capabilities of future experiments thoroughly.
-In this paper, we studied the potential impact of large-scale structure surveys in this important endeavor.
-For current experiment, CIB data had already proven to be very efficient in delensing \cite{larsen:2016,manzotti:2017}. 
-If no CMB experiment is available with higher resolution then Planck is used to reconstruct the lensing potential $\phi$, a low-z optical survey like D.E.S will improve the delensing efficiency by $~10\%$ when combined with Planck lensing and a CIB tracers.
-As expected the lower the noise in the CMB experiment the more the delensign efficiency will be predominately come from the internal CMB reconstruction of the lensing potential.
-Indeed using the CMB itself, we reconstruct only the lenses that lens the CMB obtaining, as a result, an almost perfect cleaning in the absence of noise.
-This approach has already been applied to data \cite{carron:2017}.
-However it requires a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}.
-For this reason, the relatively good efficiency of galaxies tracers might be useful to cross-check these internal biases.
-This will probably be needed to confirm a possible detection of gravitation waves if this relies heavily on delensing.
+The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal ("delensing") is necessary to test the inflationary paradigm using the next generation of CMB polarization experiments.
+To do so, we need to accurately reconstruct the large scale structures that lens the CMB in order to build a template of the lensing B-modes.
+In this paper, we studied the potential impact of large-scale structure galaxy surveys in this important endeavor.
+%For current experiment, the cosmic infrared background had already proven to be very efficient in delensing \cite{larsen:2016,manzotti:2017}. 
+For ongoing experiments, we find that an optical survey like DES alone is able to remove $14\%$ of the lensing component and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the contaminant signal using only LSS survey.
+Depending on the CMB noise and the amount of galactic foreground cleaning these can correspond to an improvement of 30\% in the constrain on the tensor to scalar ratio $r$.
 
-Another possible application of LSS delensing, which we will explore in future works, is the possibility of removing just the low-z lensing of the CMB which will otherwise possibly adds the complexity of modeling nonlinearities in the study of CMB lensing.
+In the future, the improved level of noise of CMB experiments will allow to internally reconstruct the structures lensing the CMB. 
+The fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level. 
+Already in the 3G era, the CMB internal reconstruction will be the main source of delensing.
+However it will still be less efficient than galaxies at small scales$\ell>500$.
+For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $57\%$ to $68\%$.
+
+Furthermore internal delensing will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}.
+For this reason, efficient galaxies tracers are not only useful in the short term but will also play a role in cross-checking these internal biases. 
+For a future galaxy survey like LSST we find that using tomographic binning can improve its correlation with lensing by $30\%$.
+
+Giving the high level of correlation of future galaxy surveys and CMB lensing it is worth looking for possible application beside the detection of inflationary B-modes.
+For example LSS delensing allow to remove just the low-z lensing component from the CMB lensing.
+Removing this component can avoid the necessity of modelling nonlinearities in the study of CMB lensing power spectrum.
+
+% ===================
+%\onecolumngrid
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.2,clip]{../images/errors_summary.pdf}
+\caption{\relook{to be filled with final results}}
+\label{fig}
+\end{center}
+\end{figure}
+\twocolumngrid
 

### 2017-07-05 Notes/galaxy_delensing.tex
typos

@@
 For example, DES delensing efficiency is higher than WISE firstly because of the lower level of noise and furthermore because DES galaxies are located at slightly higher redshift thus they better overlap with CMB lensing.
+As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS.
+In particular it can be seen in \reffig{corrnow} that optical surveys will rapidly lose the role of filling in large-scale modes and they will start to supplement information at higher multipoles.
 
@@
 Indeed current experiments are still significantly limited by foregrounds and instrumental noise together and not only by lensing.
-If on the other side we consider an high-resolution ground base experiment the internal lensing reconstruction improves significantly.
-With an instrumental noise at the expected level for SPTPol the CMB will soon be able to remove an amount of power (35\%) at the level of the CIB.
+
+If, on the other side, we consider an high-resolution ground base experiment the internal lensing reconstruction improves significantly. 
+\reftab{current-spt} shows that, with an instrumental noise at the expected level for SPTPol, the CMB will soon be able to remove an amount of power (35\%) at the level of the CIB.
 Once combined with LSS tracer this will lead to removing 56\% of the power.
-The improvement of 1.57 in $\sigma(r)$ that will follow even when realistic noise is considered is such that \relook{here}
+The improvement of 1.57 in $\sigma(r)$ that will follow, even when realistic noise is considered, is such that it will make delensing a  needed step for CMB polarization data analysis.
 
@@
 We also test a scenario where primordial gravitational waves are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.
-As expected the importance of delensing itself is now reduced given that the lensing component constitutes a smaller portion of the total B-modes variance.
-
-\reftab{current-spt} that uses SPTPol instead of Planck for the internal reconstruction shows consistent results.
-As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS. 
-This can be seen in \reftab{current-spt}. 
-In particular optical surveys will rapidly lose the role of filling in large-scale modes and they will start to supplement information at higher multipoles.
+As expected the importance of delensing itself is now reduced given that the lensing component constitutes a smaller portion of the total B-modes variance
 
@@
 CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
-We model the CMB S3 stage with a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam. 
-We assume here that a deep CMB experiment and an overlapping high resolution one will be able to combine their measured modes.
+We model the CMB S3 stage with a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam. 
+We assume here that a deep CMB experiment and an overlapping high-resolution one will be able to combine their measured modes.
 For this reason, we will also assume this level of noise for the internal noise reconstruction.
-With this experimental setup we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
-Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers.
+With this experimental setup, we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
+Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the LSS tracers used in \refsec{del_res}.
 
@@
 An interesting finding is that DESI will be less efficient (removing 10\% of the power) than a DES-like survey (14\%) despite the fact that it can probe slightly higher redshift.
-The reason is that, for the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES play a significant role in the delensing efficiency.
-Unfortunately, adding DESI will only bring the power removed using LSS fro 36 to 41\%.
+The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES the delensing efficiency significantly.
+Unfortunately, adding DESI will only bring the power removed using LSS from 36 to 41\%.
 Indeed, even in the near future, CIB will still play the dominant role among LSS.
-Contrary to current experiments, th CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 550$.
+Contrary to current experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 550$.
 
@@
 
-The improvement in the constraint on $r$ is summarized in \reftab{S3}. 
-Given the improvement in the noise of the high resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
-Despite this, adding galaxy survey will still lead to a decent improvement with a fairly low effort.
+We summarize the improvement in the constraint on $r$ in \reftab{S3}. 
+Given the improvement in the noise of the high-resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
+Despite this, adding galaxy survey will still lead to a decent improvement with a relatively low effort.
 Indeed galaxy surveys will still be able to remove an additional $20\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by $15\%$.
@@
 As expected the improvement on $n_{t}$ is similar to the one in $r$ since they are both proportional to the variance of the measured B-modes.
-We tested this in the scenario $r=0.12$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}/8}$.
+We tested this in the scenario $r=0.12$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}}/8$.
 It is important to note that, independently from the level of delensing, for CMB 3G experiments this statistical error will still be several times bigger than the fiducial value.

### 2017-07-05 Notes/galaxy_delensing.tex
typos

@@
 Moreover, satellite and balloon experiments have been proposed and have the potential to extend the accessible B-mode measurement to the largest scales.
-Here we follow \cite{abazajian:2016} and assume a CMB-S4 ground experiment with  with a level of noise of \textcolor{red}{$1\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
-By the mid 2020s several next generation of LSS survey will be online.
-Here, as an example for future optical galaxy surveys we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers.
-We test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} will have similar delensing performance.
-Even if they will observe similar part of the sky, because of the different  strength and weakness of this experiments
-it will still important to combined them efficiently.
+Following \cite{abazajian:2016} we assume a CMB-S4 ground experiment with a level of noise of \textcolor{red}{$1\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam.
+Furthermore, by the mid 2020s, several next generation of LSS survey will be online.
+Here, as an example for future optical galaxy surveys, we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers.
+We test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance.
+Even if they will observe similar part of the sky, because of the different strength and weakness of this experiments it will still be important to consider all of them and combine them efficiently.
 We also consider a radio continuum survey modeled following the SKA planned experiment.
-Radio continuum observation of the 21 cm line is in its early stage and several experimental and analysis challanges
-need to be overcome. However this technique has the potential to map LSS with relatively low noise up to redshift $z=6$.
+Radio continuum observation of the 21 cm line is in its early stage, and several experimental and data-analysis challenges need to be overcome.
+However, this technique has the potential to map LSS with relatively low noise up to redshift $z=6$.
 
@@
 A tomographically binned LSST-like experiment will be a very efficient CMB lensing tracer.
-Indeed it will be more than 70\% correlated with CMB lensing for all the scales $\ell<600$ a performance similar to a Stage 3 CMB internal reconstruction.
+Indeed it will be more than 70\% correlated with CMB lensing for all the scales $\ell<600$, a performance similar to a Stage-3 CMB internal reconstruction.
 SKA very new but potentially amazing
-With SKA and LSST, for the first we will have LSS tracers way more efficient than the CIB.
+With SKA and LSST, for the first we will have LSS tracers with a higher delensing efficiency than the CIB.
 Even if galaxy survey will improve dramatically the CMB internal reconstruction will still be the main source of delensing.

### 2017-07-05 Code/Plot_figures.py
S4 images

@@ -133,13 +133,6 @@ labels = ['des_bin0', 'des_bin1', 'des_bin2', 'des_bin3']
 cmb = 'now'
 lbins, _, _, rho_gals_des, _ = multiple_survey_delens.main(labels, cmb)
 
-np.savetxt('rho_cib.txt', rho['cib'])
-np.savetxt('rho_des.txt', rho_gals_des)
-np.savetxt('rho_cmb_current.txt', rho_cmb)
-np.savetxt('rho_gals_current.txt', rho_gals)
-np.savetxt('rho_comb_current.txt', rho_comb)
-np.savetxt('lbins.txt', lbins)
-
 
 # In[24]:
 plt.close()
@@ -181,11 +174,6 @@ lbins, _, _, rho_gals_desi, _ = multiple_survey_delens.main(labels, cmb)
 # In[27]:
 
 
-np.savetxt('rho_cmb_S3.txt', rho_cmb)
-np.savetxt('rho_gals_S3.txt', rho_gals)
-np.savetxt('rho_comb_S3.txt', rho_comb)
-
-
 # In[28]:
 plt.close()
 
@@ -209,18 +197,16 @@ fg.tight_layout()
 plt.savefig('../images/S3_scenario.pdf', dpi=600, papertype='Letter')
 plt.savefig('../images/S3_scenario.png')
 
-sys.exit()
+
 # ## CMB S4 scenario
 
 # In[29]:
 
 # labels = ['cib', 'desi', 'des']
-labels = ['wise', 'lsst', 'ska10', 'ska01', 'ska5', 'ska1', 'cib', 'desi', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3']
+labels = ['wise', 'ska10', 'ska01', 'ska5', 'ska1', 'cib', 'desi', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3', 'lsst_bin0', 'lsst_bin1', 'lsst_bin2',
+          'lsst_bin3', 'lsst_bin4', 'lsst_bin5', 'lsst_bin6', 'lsst_bin7', 'lsst_bin8', 'lsst_bin9', 'desi_bin0', 'desi_bin1', 'desi_bin2', 'desi_bin3']
 cmb = 'S4'
 lbins, rho, rho_comb, rho_gals, rho_cmb = multiple_survey_delens.main(labels, cmb)
-np.savetxt('rho_cmb_S4.txt', rho_cmb)
-np.savetxt('rho_gals_S4.txt', rho_gals)
[…]

### 2017-07-05 Code/multiple_survey_delens.py

@@ -28,14 +28,8 @@ def nl(noise_uK_arcmin, fwhm_arcmin, lmax):
 
 def main(labels, cmb):
 
-    # cosmosis_dir = '/home/manzotti/cosmosis/'
-    # inifile = '/home/manzotti/cosmosis/modules/limber/galaxies_delens.ini'
-    # Config_ini = ConfigParser.ConfigParser()
-    # values = ConfigParser.ConfigParser()
-    # Config_ini.read(inifile)
-    # values_file = Config_ini.get('pipeline', 'values')
-    output_dir = '../Data/'
-    # values.read(cosmosis_dir + values_file)
+
+    output_dir =  '/home/manzotti/galaxies_delensing/Data/'
     cls = pickle.load(open('../Data/limber_spectra_delens.pkl', 'rb'))
     lbins = np.load('../Data/ells.npy')
 
@@ -63,7 +57,7 @@ def main(labels, cmb):
 
         if label == 'cib':
             cl_auto[label] = np.array([3500. * (1. * l / 3000.)**(-1.25) + 725 for l in lbins])
-            cl_auto[label] += 1000*cl_auto[label][100] * (lbins / lbins[100])**-4.6
+            cl_auto[label] += 1000 * cl_auto[label][100] * (lbins / lbins[100])**-4.6
 
         rho[label] = cl_cross_k[label] / np.sqrt(ckk[:] * cl_auto[label])
         if label == 'wise':
@@ -87,7 +81,7 @@ def main(labels, cmb):
 
             if (labels[i] == 'cib' and labels[j] == 'cib'):
                 cgg[i, j, :] = np.array([3500. * (1. * l / 3000.)**(-1.25) + 725 for l in lbins])
-                cgg[i, j, :] += 1000*cgg[i, j, 100] * (lbins / lbins[100])**-4.6
+                cgg[i, j, :] += 1000 * cgg[i, j, 100] * (lbins / lbins[100])**-4.6
 
             cgg[j, i, :] = cgg[i, j, :]
 
@@ -122,7 +116,7 @@ def main(labels, cmb):
         ckk_noise = noise_fun(lbins)
 
     elif cmb == 'S4':
-        noise = 1.0
+        noise = 1.
         beam = 1
         noise_phi = np.loadtxt(
             './quicklens_data/min_var_noise_{}muk_{}beam.txt'.format(noise, beam))

### 2017-07-05 Code/rho_to_Bres.py

@@ -232,8 +232,8 @@ def main(rho_names, nle):
 
     Config_ini = ConfigParser.ConfigParser()
     Config_ini.read(inifile)
-    output_dir = Config_ini.get('test', 'save_dir')
-
+    # output_dir = Config_ini.get('test', 'save_dir')
+    output_dir =  '/home/manzotti/galaxies_delensing/Data/'
     datadir = output_dir
 
     clpp = np.loadtxt(datadir + 'cmb_cl/pp.txt')
@@ -242,8 +242,6 @@ def main(rho_names, nle):
 
     clee *= 2. * np.pi / (ells_cmb.astype(float) * (ells_cmb.astype(float) + 1.))
     clpp = clpp * 2. * np.pi / (ells_cmb.astype(float) * (ells_cmb.astype(float) + 1.))
-
-    lbins = np.logspace(1, 3.5, 190)
 
     # clbb_th = np.loadtxt(
     #     output_dir + 'cmb_cl/bb.txt')

### 2017-07-05 Code/run_analysis.py

@@ -170,11 +170,11 @@ print('')
 # =====================================
 
 
-labels = ['wise', 'cib', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3']
+labels = ['wise', 'cib', 'des']
 cmb = 'Planck'
 multiple_survey_delens.main(labels, cmb)
 ells_cmb = np.loadtxt(output_dir + 'cmb_cl/ell.txt')
-rho_names = ['rho_cib.txt', 'rho_gals.txt',
+rho_names = ['rho_cib.txt', 'rho_des.txt', 'rho_gals.txt',
              'rho_wise.txt', 'rho_comb.txt', 'rho_cmb_' + cmb + '.txt']
 
 
@@ -330,12 +330,11 @@ for i, label in enumerate(rho_names):
     print('After delensing % errors', sigma_r_1, sigma_nt)
     print(probe, 'gain = ', sigma_r_1 / sigma_r)
 
-sys.exit()
 
 print(Fore.RED + 'Actual scenario High res SPT-pol')
 # In[274]:
 
-labels = ['wise', 'cib', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3']
+labels = ['wise', 'cib', 'des']
 cmb = 'now'
 multiple_survey_delens.main(labels, cmb)
 rho_names = ['rho_cib.txt', 'rho_des.txt', 'rho_gals.txt',
@@ -503,7 +502,7 @@ print(Fore.RED + 'CMB S3')
 print('')
 # possible names orders matters = ['k', 'euclid', 'des_weak', 'lsst',
 # 'ska10',            'ska01', 'ska5', 'ska1', 'cib', 'desi', 'des']
-labels = ['wise', 'cib', 'desi_bin0', 'desi_bin1', 'desi_bin2', 'desi_bin3', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3']
+labels = ['wise', 'cib', 'desi', 'des']
 cmb = 'S3'
 multiple_survey_delens.main(labels, cmb)
 rho_names = ['rho_cib.txt', 'rho_des.txt', 'rho_desi.txt',
@@ -549,7 +548,7 @@ print('')
 
 # noise_uK_arcmin=4.5,
 # fwhm_arcmin=4.,
-lmax = 3000
+lmax = 2000
 # This needs to be Bicep like, the value of the deep exp
 noise_uK_arcmin = 3
 fwhm_arcmin = 1.
[…]

### 2017-07-05 Code/Plot_figures.py
S4 images

@@ -1,17 +1,18 @@
-
-
 import numpy as np
 import matplotlib.pyplot as plt
 import multiple_survey_delens
-
+import sys
 import rho_to_Bres
 import configparser as ConfigParser
 import camb
-from camb import model, initialpower
+from camb import initialpower
 try:
     import functools32
 except ImportError:
     import functools as functools32
+
+
+
 # ============================================
 # SIZE OF THE PICTURE
 # ============================================
@@ -51,7 +52,6 @@ plt.rcParams['ytick.major.width'] = font_size / 10.
 # plt.rcParams['xtick.labelsize'] = font_size / 1.2
 # plt.rcParams['ytick.labelsize'] = font_size / 1.2
 
-# plt.rcParams['axes.prop_cycles'] = '#e41a1c,#377eb8,#4daf4a,#984ea3,#ff7f00,#ffff33,#a65628'
 plt.rcParams['lines.linewidth'] = font_size / 8.
 # ============================================
 
@@ -159,8 +159,6 @@ plt.savefig('../images/actual_scenario.png')
 
 # ## CMB S3 scenario
 
-# In[25]:
-
 labels = ['wise', 'cib', 'des_bin0', 'des_bin1', 'des_bin2',
           'des_bin3', 'desi_bin0', 'desi_bin1', 'desi_bin2', 'desi_bin3']
 cmb = 'S3'
@@ -170,11 +168,6 @@ labels = ['desi_bin0', 'desi_bin1', 'desi_bin2', 'desi_bin3']
 cmb = 'S3'
 lbins, _, _, rho_gals_desi, _ = multiple_survey_delens.main(labels, cmb)
 
-
-# In[27]:
-
-
-# In[28]:
 plt.close()
 
 fg = plt.figure(figsize=fig_dims)
@@ -202,28 +195,17 @@ plt.savefig('../images/S3_scenario.png')
 
 # In[29]:
 
-# labels = ['cib', 'desi', 'des']
[…]

### 2017-07-05 Code/Plot_figures.py
S4 images

@@ -12,7 +12,6 @@ except ImportError:
     import functools as functools32
 
 
-
 # ============================================
 # SIZE OF THE PICTURE
 # ============================================
@@ -193,9 +192,9 @@ plt.savefig('../images/S3_scenario.png')
 
 # ## CMB S4 scenario
 
-# In[29]:
+# In[29]: 'ska10', 'ska01', 'ska5', 'ska1'
 
-labels = ['wise', 'ska10', 'ska01', 'ska5', 'ska1', 'cib', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3', 'lsst_bin0', 'lsst_bin1', 'lsst_bin2',
+labels = ['wise', 'ska10', 'cib', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3', 'lsst_bin0', 'lsst_bin1', 'lsst_bin2',
           'lsst_bin3', 'lsst_bin4', 'lsst_bin5', 'lsst_bin6', 'lsst_bin7', 'lsst_bin8', 'lsst_bin9', 'desi_bin0', 'desi_bin1', 'desi_bin2', 'desi_bin3']
 cmb = 'S4'
 lbins, rho, rho_comb, rho_gals, rho_cmb = multiple_survey_delens.main(labels, cmb)
@@ -214,7 +213,7 @@ plt.plot(lbins, rho_gals, label='LSST')
 # plt.plot(BB_contr[:, 0], BB_contr[:, 1] - np.min(BB_contr[:, 1]),
 #          '--', alpha=0.6, linewidth=font_size / 14., color='k')
 
-labels = ['ska10', 'ska01', 'ska5', 'ska1']
+labels = ['ska10']
 cmb = 'S4'
 lbins, rho, rho_comb, rho_gals, rho_cmb = multiple_survey_delens.main(labels, cmb)
 plt.plot(lbins, rho_gals, label='SKA')
@@ -235,8 +234,42 @@ plt.xlabel(r'$\ell$')
 plt.ylabel(r'$\rho$')
 fg.tight_layout()
 
-plt.savefig('../images/S4_scenario.pdf', dpi=600, papertype='Letter')
-plt.savefig('../images/S4_scenario.png')
+plt.savefig('../images/S4_scenario_ska10.pdf', dpi=600, papertype='Letter')
[…]

### 2017-07-05 Code/run_analysis.py

@@ -652,25 +652,25 @@ for i, label in enumerate(rho_names):
 
     print('After delensing % errors', sigma_r_1, sigma_nt, sigma_nt_1)
     print(probe, 'gain = ', sigma_r_1 / sigma_r, sigma_nt_1 / sigma_nt)
-
-
-sys.exit()
+# ==============================
+# ==============================
+# CMB-S4
+# ==============================
+# ==============================
 
 print('')
-print(Fore.RED + 'CMB S4')
+print(Fore.RED + 'CMB S4 + SKA01')
 print('')
-# ## CMB S4
 
-# In[310]:
-
-#  'wise', 'euclid', 'des_weak', 'lsst', 'ska10',
-# #              'ska01', 'ska5', 'ska1', 'cib', 'desi', 'des']
-labels = ['wise', 'euclid', 'lsst', 'ska10',
-          'ska01', 'ska5', 'ska1', 'cib', 'desi', 'des']
+labels = ['wise', 'ska01', 'cib', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3', 'lsst_bin0', 'lsst_bin1', 'lsst_bin2',
+          'lsst_bin3', 'lsst_bin4', 'lsst_bin5', 'lsst_bin6', 'lsst_bin7', 'lsst_bin8', 'lsst_bin9', 'desi_bin0', 'desi_bin1', 'desi_bin2', 'desi_bin3']
 cmb = 'S4'
+
+print(Fore.RED + 'Tracers:' + '-'.join(labels))
+
 multiple_survey_delens.main(labels, cmb)
-rho_names = ['rho_euclid.txt', 'rho_lsst.txt', 'rho_ska10.txt', 'rho_ska5.txt',
-             'rho_ska1.txt', 'rho_ska01.txt', 'rho_comb.txt', 'rho_cmb_' + cmb + '.txt']
+rho_names = ['rho_ska01.txt', 'rho_gals.txt', 'rho_comb.txt', 'rho_cmb_' + cmb + '.txt']
+
 # deep survey to delens or what is giving you E-mode
 nle = nl(1, 1, lmax=ells_cmb[-1])[2:]
 B_res3 = rho_to_Bres.main(rho_names, nle)
@@ -770,17 +770,133 @@ for i, label in enumerate(rho_names):
[…]

### 2017-07-05 Code/Plot_figures.py
S4 images

@@ -216,7 +216,7 @@ plt.plot(lbins, rho_gals, label='LSST')
 labels = ['ska10']
 cmb = 'S4'
 lbins, rho, rho_comb, rho_gals, rho_cmb = multiple_survey_delens.main(labels, cmb)
-plt.plot(lbins, rho_gals, label='SKA')
+plt.plot(lbins, rho_gals, label=r'SKA $0.1\mu$Jy')
 
 
 labels = ['wise', 'cib', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3', 'lsst_bin0', 'lsst_bin1', 'lsst_bin2',
@@ -256,7 +256,7 @@ plt.plot(lbins, rho_comb, label='Galaxies + CMB S4')
 labels = ['ska01']
 cmb = 'S4'
 lbins, rho, rho_comb, rho_gals, rho_cmb = multiple_survey_delens.main(labels, cmb)
-plt.plot(lbins, rho_gals, label='SKA')
+plt.plot(lbins, rho_gals, label=r'SKA $0.1\mu$Jy')
 plt.plot(lbins, rho_cmb, label='CMB S4')
 
 plt.legend(loc=0, ncol=2)

### 2017-07-05 Code/run_analysis.py

@@ -524,7 +524,7 @@ for i, probe in enumerate(rho_names):
 
 print('')
 print(Fore.YELLOW + 'Fraction of removed Bmode power')
-for probe in rho_names[1:]:
+for probe in rho_names[0:]:
     probe = probe.split('.txt')[0].split('rho_')[1]
     print(probe)
     print('ell<100=', 1. - np.mean(clbb_res[probe](np.arange(4, 100, 25)) / clbb_lensed(np.arange(4, 100, 25))),
@@ -684,7 +684,7 @@ for i, probe in enumerate(rho_names):
 
 print('')
 print(Fore.YELLOW + 'Fraction of removed Bmode power')
-for probe in rho_names[1:]:
+for probe in rho_names[0:]:
     probe = probe.split('.txt')[0].split('rho_')[1]
     print(probe)
     print('ell<100=', 1. - np.mean(clbb_res[probe](np.arange(4, 100, 25)) / clbb_lensed(np.arange(4, 100, 25))),
@@ -910,7 +910,7 @@ for i, probe in enumerate(rho_names):
 
 print('')
 print(Fore.YELLOW + 'Fraction of removed Bmode power')
-for probe in rho_names[1:]:
+for probe in rho_names[0:]:
     probe = probe.split('.txt')[0].split('rho_')[1]
     print(probe)
     print('ell<100=', 1. - np.mean(clbb_res[probe](np.arange(4, 100, 25)) / clbb_lensed(np.arange(4, 100, 25))),
@@ -976,7 +976,7 @@ for i, probe in enumerate(rho_names):
 
 print('')
 print(Fore.YELLOW + 'Fraction of removed Bmode power')
-for probe in rho_names[1:]:
+for probe in rho_names[0:]:
     probe = probe.split('.txt')[0].split('rho_')[1]
     print(probe)
     print('ell<100=', 1. - np.mean(clbb_res[probe](np.arange(4, 100, 25)) / clbb_lensed(np.arange(4, 100, 25))),

### 2017-07-05 Notes/galaxy_delensing.tex

@@
 However, this technique has the potential to map LSS with relatively low noise up to redshift $z=6$.
+Here we considered just two possible  detection thresholds at 1 GHz (flux cut): 10 $\mu$Jy and 0.1 $\mu$Jy, which are representatives of the upper and lower limit levels for a SKA phase 1 (SKA1) and a SKA phase 2 (SKA2) survey.
 
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/S4_scenario.pdf}
+\includegraphics[scale=1.]{../images/S4_scenario_ska10.pdf}
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/S4_scenario_no_ska.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments without SKA. }
+\includegraphics[scale=1.]{../images/S4_scenario_ska01.pdf}
+\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
 \label{fig:corrs4}

### 2017-07-05 Code/Plot_figures.py
S4 images

@@ -216,7 +216,7 @@ plt.plot(lbins, rho_gals, label='LSST')
 labels = ['ska10']
 cmb = 'S4'
 lbins, rho, rho_comb, rho_gals, rho_cmb = multiple_survey_delens.main(labels, cmb)
-plt.plot(lbins, rho_gals, label=r'SKA $0.1\mu$Jy')
+plt.plot(lbins, rho_gals, label=r'SKA $10\mu$Jy')
 
 
 labels = ['wise', 'cib', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3', 'lsst_bin0', 'lsst_bin1', 'lsst_bin2',

### 2017-07-06 Notes/galaxy_delensing.tex
typos

@@
 Following \cite{abazajian:2016} we assume a CMB-S4 ground experiment with a level of noise of \textcolor{red}{$1\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam.
-Furthermore, by the mid 2020s, several next generation of LSS survey will be online.
+Furthermore, by the mid 2020s, several next-generation LSS surveys will be online.
 Here, as an example for future optical galaxy surveys, we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers.
 We test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance.
-Even if they will observe similar part of the sky, because of the different strength and weakness of this experiments it will still be important to consider all of them and combine them efficiently.
+Even if they will observe similar part of the sky and similar modes, because of the different strengths and weaknesses of these experiments it will still be important to combine them efficiently.
 We also consider a radio continuum survey modeled following the SKA planned experiment.
-Radio continuum observation of the 21 cm line is in its early stage, and several experimental and data-analysis challenges need to be overcome.
+Radio continuum observations of the 21 cm line are in theri early stage, and several experimental and data-analysis challenges need to be overcome.
 However, this technique has the potential to map LSS with relatively low noise up to redshift $z=6$.
@@
 
-The correlation factor with CMB lensing of stage 4 experiments is shown in \reffig{corrs4}.
+The correlation factor with CMB lensing of stage 4 experiments is shown in \reffig{corrs4_10}  and in \reffig{corrs4_01} for SKA1 and SKA2 respectively.
 
@@
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
-\label{fig:corrs4}
+\label{fig:corrs4_10}
 \end{center}
@@
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
-\label{fig:corrs4}
+\label{fig:corrs4_01}
 \end{center}
@@
 Indeed it will be more than 70\% correlated with CMB lensing for all the scales $\ell<600$, a performance similar to a Stage-3 CMB internal reconstruction.
-SKA very new but potentially amazing
+A very similar level of efficiency will be possible with a SKA1-like experiment.
 With SKA and LSST, for the first we will have LSS tracers with a higher delensing efficiency than the CIB.
 Even if galaxy survey will improve dramatically the CMB internal reconstruction will still be the main source of delensing.
-This, having a perfect kernel overlap with the true lensing potential we will benefit for the very low level of noise of CMB S4 experiments.
-CMB S4 delensing will probably be limited by small secondary effect like foregrounds contamination, missing modes due to filtering etc.
+This, having a perfect kernel overlap with the true lensing potential will benefit from the very low level of noise of CMB S4 experiments.
+\reffig{corrs4_10} shows that a CMB-S4 experiment will be able to internally reconstruct the lensing potential at more than 90\% up to $\ell=400$
+CMB S4 delensing will probably be limited by small secondary effects like foregrounds contamination, missing modes due to filtering etc.
+\reffig{corrs4_01} shows that a futuristic SKA-2 experiment will show a very high correlation with CMB lensing being more than 80\% correlated at all the scales.
+This performance is comparable with a CMB-S4 internal reconstruction.
 
 \reftab{S4} shows the improvement on the inflationary constraint.
+
+\begin{table}
+\caption{$\alpha$: Gen 4 experiments}
+  \begin{tabular}{|c | c | c | c|}
+\hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
+LSST (48\%) &1.12 & 1.08  & 1.03 \\ \hline
+LSS  (60\%) &1.12 & 1.08  & 1.03 \\ \hline
+SKA (10$\mu$Jy) (53\%) &1.73& 1.42  & 1.2\\ \hline
+SKA (0.1$\mu$Jy) (77\%) &1.73& 1.42  & 1.2\\ \hline
+LSS + SKA (10$\mu$Jy)  (62\%) &1.12 & 1.08  & 1.03 \\ \hline
+LSS + SKA (0.1$\mu$Jy)  (80\%) &1.12 & 1.08  & 1.03 \\ \hline
+CMB S4 (81\%) &2& 1.57  & 1.21 \\ \hline
+LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
+\end{tabular}
+\label{tab:S4}
+\end{table}
 
@@
 LSST (10\%) &1.12 & 1.08  & 1.03 \\ \hline
-SKA (40\%) &1.73& 1.42  & 1.2\\ \hline
+SKA (0.1$\mu$Jy) (40\%) &1.73& 1.42  & 1.2\\ \hline
 CMB S4 (56\%) &2& 1.57  & 1.21 \\ \hline
@@
 %=======================================
-
-\begin{table}
-\caption{$\alpha$: Gen 4 experiments}
-  \begin{tabular}{ | c | c  |  c |}
-\hline
-Surveys & $\alpha(r)$ & $\alpha(n_{t})$\\ \hline \hline
-Euclid & 1.53 & \\ \hline
-LSST & 1.65  &\\ \hline
-SKA & 4.35  &\\ \hline
-CMB S4 & 3.7  &\\ \hline
-LSS+CMB & 6.3  &\\ \hline
-\end{tabular}
-\label{tab:S4}
-\end{table}
 
@@
 %We compute the derivatives of the power spectra as described
-
-%\begin{table}
-%\caption{$\alpha$: improvement on r constraint}
-%  \begin{tabular}{ | c | c  | }
-%\hline
-%Surveys & $\alpha$\\ \hline \hline
-%des & 1.34  \\ \hline
-%cib & 1.71  \\ \hline
-%cmb current & 1.79  \\ \hline
-%gals current & 2.03  \\ \hline
-%cmb S3 & 2.15  \\ \hline
-%gals S3 & 2.16  \\ \hline
-%gals S4 & 2.16  \\ \hline
-%comb current & 2.66  \\ \hline
-%comb S3 & 3.14  \\ \hline
-%cmb S4 & 4.36  \\ \hline
-%comb S4 & 5.27  \\ \hline
-%  \end{tabular}
-%\end{table}
 
@@
 
-This work made use of computing resources and support provided by the Re- search Computing Center at the University of Chicago.
+This work made use of computing resources and support provided by the Research Computing Center at the University of Chicago.
 This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.

### 2017-07-06 Notes/galaxy_delensing.tex
ska description

@@
 The red solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed green line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
-The residual power left after delensing for some of the LSS tracers described in \refsec{model}. 
-The rapid improvement in the level of instrumental noise (dashed curves for SPTPol and CMB S4) will require a high level of delensing to exploit these experiments fully.}
+These are compared with the residual power left after delensing using some of the LSS tracers described in \refsec{model}. 
+The rapid improvement in the level of instrumental noise (dashed curves for SPTPol and CMB S4) will require a high delensing efficiency to exploit these experiments fully.}
 \label{fig:bbplot}
@@
 
-An ambitious program for a generation 4 ground CMB experiment is currently under planning \cite{abazajian:2016}. 
-Moreover, satellite and balloon experiments have been proposed and have the potential to extend the accessible B-mode measurement to the largest scales.
+An ambitious program for a stage-4 ground CMB experiment is currently under planning \cite{abazajian:2016}. 
+Moreover, satellite and balloon CMB experiments have been proposed and have the potential to extend the accessible B-mode measurement to the largest scales.
 Following \cite{abazajian:2016} we assume a CMB-S4 ground experiment with a level of noise of \textcolor{red}{$1\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam.
-Furthermore, by the mid 2020s, several next-generation LSS surveys will be online.
+Furthermore, by the mid-2020s, several next-generation LSS surveys will be online.
 Here, as an example for future optical galaxy surveys, we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers.
 We test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance. 
-Even if they will observe similar part of the sky and similar modes, because of the different strengths and weaknesses of these experiments it will still be important to combine them efficiently.
-We also consider a radio continuum survey modeled following the SKA planned experiment.
-Radio continuum observations of the 21 cm line are in theri early stage, and several experimental and data-analysis challenges need to be overcome.
-However, this technique has the potential to map LSS with relatively low noise up to redshift $z=6$.
+Note that even if they will observe an overlapping part of the sky and similar modes, because of their different strengths and weaknesses it will still be important to combine them efficiently.
+We also consider a radio continuum survey modeled following the SKA specifications.
+Radio continuum observations of the 21 cm line are in their early stage, and several experimental and data-analysis challenges need to be overcome.
+However, this technique has the potential to map the LSS with relatively low noise up to redshift $z=6$.
 Here we considered just two possible  detection thresholds at 1 GHz (flux cut): 10 $\mu$Jy and 0.1 $\mu$Jy, which are representatives of the upper and lower limit levels for a SKA phase 1 (SKA1) and a SKA phase 2 (SKA2) survey.
 
-The correlation factor with CMB lensing of stage 4 experiments is shown in \reffig{corrs4_10}  and in \reffig{corrs4_01} for SKA1 and SKA2 respectively.
+The correlation factor with CMB lensing for stage-4 experiments is shown in \reffig{corrs4_10}  and in \reffig{corrs4_01} for SKA1 and SKA2 respectively.
 
@@
 A very similar level of efficiency will be possible with a SKA1-like experiment.
-With SKA and LSST, for the first we will have LSS tracers with a higher delensing efficiency than the CIB.
-Even if galaxy survey will improve dramatically the CMB internal reconstruction will still be the main source of delensing.
+With SKA and LSST, for the first time we will have LSS tracers with a higher delensing efficiency than the CIB.
+Despite the improvement in galaxy surveys, the CMB internal reconstruction will still be the main source of delensing.
 This, having a perfect kernel overlap with the true lensing potential will benefit from the very low level of noise of CMB S4 experiments.
 \reffig{corrs4_10} shows that a CMB-S4 experiment will be able to internally reconstruct the lensing potential at more than 90\% up to $\ell=400$
-CMB S4 delensing will probably be limited by small secondary effects like foregrounds contamination, missing modes due to filtering etc.
-\reffig{corrs4_01} shows that a futuristic SKA-2 experiment will show a very high correlation with CMB lensing being more than 80\% correlated at all the scales.
+At this level, CMB S4 internal delensing will probably be limited by small secondary effects like foregrounds contamination, missing modes due to filtering etc.
+Finally, \reffig{corrs4_01} shows that a futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
 This performance is comparable with a CMB-S4 internal reconstruction.

### 2017-07-07 Notes/galaxy_delensing.tex
keep it up to data

@@
 \ee
-%As usual we define the power spectrum as:
-%\be
-%\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l .
-%\ee
-
 From this we get the power spectrum of the lensing component of the B-modes:
@@
 Indeed it will be more than 70\% correlated with CMB lensing for all the scales $\ell<600$, a performance similar to a Stage-3 CMB internal reconstruction.
-A very similar level of efficiency will be possible with a SKA1-like experiment.
+A very similar level of efficiency will be achieved by a SKA1-like experiment.
 With SKA and LSST, for the first time we will have LSS tracers with a higher delensing efficiency than the CIB.
@@
 
-\reftab{S4} shows the improvement on the inflationary constraint.
+\reftab{S4} shows the improvement on the inflationary constraint given the cross correlation factors described above. 
+LSST alone will be able to remove almost half of the B-mode lensing power.
+This will lead to an improvement in $\sigma(r)$ of a factor of 1.8.
+If we add all the other LSS tracers previously included (i.e.
+excluding SKA) the improvement factor increase to 2.3.
+SKA in its phase 1 will remove 53\% of the lensing power increasing
 
@@
 Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-LSST (48\%) &1.12 & 1.08  & 1.03 \\ \hline
-LSS  (60\%) &1.12 & 1.08  & 1.03 \\ \hline
-SKA (10$\mu$Jy) (53\%) &1.73& 1.42  & 1.2\\ \hline
-SKA (0.1$\mu$Jy) (77\%) &1.73& 1.42  & 1.2\\ \hline
-LSS + SKA (10$\mu$Jy)  (62\%) &1.12 & 1.08  & 1.03 \\ \hline
-LSS + SKA (0.1$\mu$Jy)  (80\%) &1.12 & 1.08  & 1.03 \\ \hline
-CMB S4 (81\%) &2& 1.57  & 1.21 \\ \hline
-LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
+LSST (48\%) &1.88 & 1.81  &  \\ \hline
+LSS  (60\%) &2.47 & 2.31  & \\ \hline
+SKA (10$\mu$Jy) (53\%) &2.18& 2.06  & \\ \hline
+SKA (0.1$\mu$Jy) (77\%) &4.6& 3.96  & \\ \hline
+LSS + SKA (10$\mu$Jy)  (62\%) &2.74 & 2.53  &  \\ \hline
+LSS + SKA (0.1$\mu$Jy)  (80\%) &4.66 & 3.96  &  \\ \hline
+CMB S4 (81\%) & 4.63 & 3.96  &  \\ \hline
+LSS+CMB (69\%)& 2.8 & 1.8&  \\ \hline
 \end{tabular}

### 2017-07-07 Notes/galaxy_delensing.tex
typos

@@
 We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
-We find that in general splitting galaxy surveys into tomographic binning can improve the correlation factor with CMB lensing by 30\%. 
-Considering available survey, a galaxy survey like DES can remove about $14\%$ of the lensing signal. current generation CMB experiments will benefit from 
+We find that, in general, splitting galaxy surveys into tomographic bins can improve the correlation factor with CMB lensing by 30\%. 
+Considering available survey, a galaxy survey like DES can remove about $14\%$ of the lensing signal. current generation CMB experiments will benefit from ...
 With the improvement of the noise CMB internal reconstruction will become more important and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level.

### 2017-07-07 Notes/galaxy_delensing.tex
Scott comments up to page 9

@@
 In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$. 
-Furthermore $W^{i}(z)$ is the kernel function of the field $i$ introduced in \refeq{wkernel}.
-We will now describe the kernels for each of the tracers used in this work.
+%Furthermore $W^{i}(z)$ is the kernel function of the field $i$ introduced in \refeq{wkernel}.
+We will now describe the kernels $W^{i}(z)$ for each of the tracers used in this work.
 
@@
 
-The CMB lensing potential is \textit{the} field that we need to reconstruct to reverse the effect of large scale structure and delens the CMB.
+The CMB lensing potential is \textit{the} field that we need in order to reverse the effect of large scale structure and delens the CMB.
 However, the lensing potential can also be reconstructed using the CMB itself.
@@
 We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$ and we set $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
-Note that there are several available CIB observations that have already been used to delens the CMB.
+There are several available CIB observations that have already been used to delens the CMB.
 Given its full-sky coverage, a promising one is the CIB map derived in \cite{2016A&A...596A.109P} using multi-frequency Planck data and the Generalized Needlet Internal Linear Combination (GNILC) component separation algorithm.
@@
 \label{sec:th}
-The large scale structures described in \refsec{model} have an important impact on the search of primordial CMB B-modes: they lens the primordial E-modes generating non inflationary B-modes that constitues an important source of noise.
+The large scale structures described in \refsec{model} have an important impact on the search of primordial CMB B-modes: they lens the primordial E-modes generating non inflationary B-modes that, then constitute an important source of noise.
 
@@
 \ee
+Here $\varphi_{\bl,\bl'}$ is the angle between the two different modes $\bl$ and $\bl'$.
 From this we get the power spectrum of the lensing component of the B-modes:
@@
 where $f(\bl,\bl')$ can be determined by minimizing the difference with the true $B^{\mathrm{lens}}(\bl)$ defined in \refeq{blens}.
-We introduce the instrumental noise in the CMB E-modes ($E^N$) that will also limit the ability to fully reconstruct the lensing B-modes.
+We include the instrumental noise in the CMB E-modes ($E^N$) that will also limit the ability to fully reconstruct the lensing B-modes.
 
@@
 
-\refeq{Bres} directly relate the residual power after delensing with the cross correlation coefficients with CMB lensing of the tracers used.
+\refeq{Bres} directly relates the residual power after delensing with the cross correlation coefficients with CMB lensing of the tracers used.
 The bigger the $\rho_{\ell}^{2}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
-An higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$ leading to a smaller residual power $C_{{l}}^{BB,\mathrm{res}}$.
+A higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$ leading to a smaller residual power $C_{{l}}^{BB,\mathrm{res}}$.
 We conclude this section showing in \reffig{bbplot} the expected residual lensing B-modes power spectrum for some of the tracers used in this work together with the primordial B-modes component and the instrumental noise for current and future experiments.
@@
 We start by assuming that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
-We can optimally combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ with:
+We can optimally combine them to estimate $\phi$ or, in other word, to maximize the correlation factor $\rho$ with:
 \beqn
 I         &=&    \sum_{i}c^{i}I^{i} \nonumber \\
-c_{i}     &=& (C_{II}^{-1})_{ij}C^{\kappa I^{j}}
+c_{i}     &=& (C_{II}^{-1})_{ij}C^{{\kappa I}^{j}}
 \label{eqn:combined}
@@
 \be
-\rho^{2}_{\ell} = \sum_{i,j}\frac{C^{\kappa i}_{\ell}~(C^{-1}_{\ell})_{ij}~C^{\kappa j}_{\ell}}{C^{\kappa\kappa}_{\ell}}.
+\rho^{2}_{\ell} = \sum_{i,j}\frac{C^{\kappa ^{i}}_{\ell}~(C^{-1}_{\ell})_{ij}~C^{\kappa ^{j}}_{\ell}}{C^{\kappa\kappa}_{\ell}}.
 \label{eqn:rho-combined}
@@
 \reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
-Note that the cross-correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
+The cross-correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
 CIB and 21 cm surveys probe the high redshift structures and, independently from the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
 On the other end, galaxy clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
-However, given the low noise of this measurement and their small overlap with other probes they can still play a major role in delensing even if their overlap with the CMB lensing potential is not optimal.
+However, given the low noise of these measurements and their small overlap with other probes, they can still play a major role in delensing even if their overlap with the CMB lensing potential is not optimal.
 
@@
 When we weight a tracer with $\frac{C^{\kappa I}}{C^{II}}$ in \refeq{fweight} in order to maximize its ability to reconstruct the lensing potential we are only using redshift averaged information about the survey.
-However, as can be seen in \reffig{kernel}, the kernel overlap of a tracer with the CMB lensing varies as a function of redshifts.
+However, as can be seen in \reffig{kernel}, the kernel overlap of a tracer with the CMB lensing varies as a function of redshift.
 For this reason, the optimal approach is to weight galaxies at different redshifts with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
@@
 \ee
-Now it can be show that $\rho_{\rm{split}}\ge \rho_{\rm{full}}$ being:
+Now it can be show that $\rho_{\rm{split}}\ge \rho_{\rm{full}}$ since:
 \be
@@
 \ee
-Then $\rho^2$ is always bigger in the tomographic case, and the two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the same for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
-Of course, binning will improve the efficiency of galaxy surveys, but it is not applicable to tracers with poor redshift information like the CIB or radio continuum surveys.
+Then $\rho^2$ is always larger in the tomographic case, and the two are equal only when $\frac{C^{\kappa I_{i}}}{C^{I_{i}I_{i}}}$ is the same for all the redshift bins in which case a single optimal weight is sufficient for the entire survey.
+Binning will improve the efficiency of galaxy surveys, but it is not very effective for tracers with poor redshift information like the CIB or radio continuum surveys.
 
@@
 We will use the Fisher information matrix to quantify the delensing efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
-We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction performance, together with a low-noise, low-resolution experiment the B-modes of which \textcolor{red}{(?)} will be delensed and used to constrain the inflationary parameters.
+We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the ability to internally reconstruct the potential $\kappa$, together with a low-noise, low-resolution experiment that measured the B-modes that will be delensed and used to constrain the inflationary parameters.
 We focus on three distinct scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
@@
 
-It can be seen in \refeq{fisher} that removing the lensing contribution will improve parameters constraints.
+It can be seen in \refeq{fisher} that removing the lensing contribution will improve parameter constraints.
 The parameter uncertainties are inversely proportional to the covariance of the measurement.
-Since the lensing B-modes $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them trough delensing will reduce the parameter statistical error.
-After delensing the error $\sigma^{\rm{del}}(p)$ can still be obtained from \refeq{fisher} with $C_{{l}}^{BB,\mathrm{res}}<C_{{l}}^{BB,\mathrm{lens}}$ in the covariance defined in \refeq{cova}.
+Since the lensing B-modes $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them through delensing will reduce the parameter statistical error.
+After delensing, the error $\sigma^{\rm{del}}(p)$ can still be obtained from \refeq{fisher} with $C_{{l}}^{BB,\mathrm{res}}<C_{{l}}^{BB,\mathrm{lens}}$ in the covariance defined in \refeq{cova}.
 We define the improvement as the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$.
@@
 In \refeq{fisher} we are making a few important assumptions. 
-Firstly, we are fixing all the cosmological parameters except $\{r,n_{t}\}$.
-Uncertainties in those will, however, propagate to uncertainties in $\{r,n_{t}\}$.
+First, we are fixing all the cosmological parameters except $\{r,n_{t}\}$.
+Uncertainties in those will propagate to larger uncertainties in $\{r,n_{t}\}$.
 While neglecting this will lead to slightly optimistic constraints, it has no significant impact on the estimate of the improvement due to delensing.
@@
 We checked this for a few of the tracer combinations used here and find it particularly true once several tracers are jointly taken into account.
-Finally, we have assumed a Gaussian covariance even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
+Finally, we have assumed a Gaussian covariance $\sigma( C_{\ell}^{BB} )$ even if has small non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
 This approximation is good enough to show the improvement due to delensing.
@@
 
-Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} as well as the internal CMB lensing potential reconstruction \cite{carron:2017} as large scale structure tracers.
+Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} and internal CMB lensing maps \cite{carron:2017} as large scale structure tracers.
 Here we discuss the improvement that can be obtained by combining these and other currently available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
@@
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
-Furthermore, we test a second scenario with a ground-based experiment with noise levels consistent with SPT-Pol: \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
+Then, we test a second scenario with a ground-based experiment with noise levels consistent with SPT-Pol: \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
 In this case we set the largest achievable scale at $\ell_{min}=300$.
@@
 
-We combined these CMB experiments with the CIB and current low redshift galaxy surveys like DES and WISE and we compute the improvement in the delensing efficiency.
+We combine these CMB experiments with the CIB and current low redshift galaxy surveys like DES and WISE and we compute the improvement in the delensing efficiency.
 Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals.
-Optical surveys galaxies like DES are less affected by dust and can be used at larger scales.
+Optical surveys galaxies like DES are less affected by dust and can be used on larger scales.
 The achievable correlation is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
@@
 \reffig{corrnow} and \reffig{corrnow-planck} show that the DES galaxies are effective tracers of the LSS and can, at least in the near future, be used to improve delensing for CMB experiments that overlap with it.
-For example, DES delensing efficiency is higher than WISE firstly because of the lower level of noise and furthermore because DES galaxies are located at slightly higher redshift thus they better overlap with CMB lensing.
+For example, DES delensing efficiency is higher than WISE first because of the lower level of noise and also because DES galaxies are located at slightly higher redshift so they better overlap with CMB lensing.
 As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS.
@@
 For current surveys we only consider $p=r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
-Note that we still let $n_{t}$ to vary in our fisher calculation even if it is not very degenerate with $r$.
+We still let $n_{t}$ to vary in our fisher calculation even if it is not very degenerate with $r$.
 We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
@@
 As shown in previous works \cite{sherwin:2015}, the best tracer is the CIB with a 27\% reduction in power.
-Nonetheless, it is still worth to consider all the tracers.
+Nonetheless, it is still worthwhile considering all the tracers.
 Indeed a multi tracer-approach can bring the removed power on the overlapping area from 27\% of the CIB alone to 36\% using all the LSS and 41\% when Planck lensing reconstruction is added to the set.
 However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 30\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
-Furthermore, as explained in \refsec{fisher}, this is only an upper limit and will be even smaller when, for example, a non-perfect foreground cleaning is considered.
+Furthermore, as explained in \refsec{fisher}, this is only an upper limit and will be even smaller when, for example, non-perfect foreground cleaning is considered.
 Indeed current experiments are still significantly limited by foregrounds and instrumental noise together and not only by lensing.
 
-If, on the other side, we consider an high-resolution ground base experiment the internal lensing reconstruction improves significantly. 
+If, on the other hand, we consider a high-resolution ground base experiment the internal lensing reconstruction improves significantly. 
 \reftab{current-spt} shows that, with an instrumental noise at the expected level for SPTPol, the CMB will soon be able to remove an amount of power (35\%) at the level of the CIB.
@@
 \begin{table}
-\caption{$\alpha$: S3 improvements. Note that the absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is $\sigma(r) =$  }
+\caption{$\alpha$: S3 improvements. The absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is $\sigma(r) =$  }
   \begin{tabular}{|c | c | c | c|}
@@
 We test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance. 
-Note that even if they will observe an overlapping part of the sky and similar modes, because of their different strengths and weaknesses it will still be important to combine them efficiently.
+Even if they will observe an overlapping part of the sky and similar modes, because of their different strengths and weaknesses it will still be important to combine them efficiently.
 We also consider a radio continuum survey modeled following the SKA specifications.

### 2017-07-07 Notes/galaxy_delensing.tex
Scott comments up to page 9

@@
 CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
-We model the CMB S3 stage with a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam. 
+We model the CMB S3 with a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam. 
 We assume here that a deep CMB experiment and an overlapping high-resolution one will be able to combine their measured modes.
@@
 An interesting finding is that DESI will be less efficient (removing 10\% of the power) than a DES-like survey (14\%) despite the fact that it can probe slightly higher redshift.
-The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES the delensing efficiency significantly.
+The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency significantly.
 Unfortunately, adding DESI will only bring the power removed using LSS from 36 to 41\%.
@@
 Indeed galaxy surveys will still be able to remove an additional $20\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by $15\%$.
-Note that apart from the addition of DESI, the tracers used here are already available for the current generation.
+Note that apart from the addition of DESI, the tracers used here are already available today.
 The ongoing effort can then lead to a significant improvement even for the next generation of experiments.
@@
 We tested this in the scenario $r=0.12$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}}/8$.
-It is important to note that, independently from the level of delensing, for CMB 3G experiments this statistical error will still be several times bigger than the fiducial value.
+It is important to note that, independent of the level of delensing, for CMB 3G experiments the statistical error on $n_{t}$ will still be several times bigger than the fiducial value.
 

### 2017-07-08 Notes/galaxy_delensing.tex
Scott comments up to page 9

@@
 
-Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational waves background, improving their measurement is a major objective of current and future CMB experiments.
-Even if inflationary B-modes have not been detected yet, a natural value $r \gtrsim 10^{-3}$ should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
+Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational wave background, improving their measurement is a major objective of current and future CMB experiments.
+Even if inflationary B-modes have not been detected yet, a natural value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
 However, just reducing the level of noise will not be enough to attain this goal.
@@
 
-We organize this article as follow: we describe the LSS tracers used in this analysis in \refsec{model}.
+We organize this article as follows: we describe the LSS tracers used in this analysis in \refsec{model}.
 In \refsec{th} we define the lensing B-mode component and the residual power after delensing with tracers of the lensing potential.
-The main result of this work is \refsec{for} : the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
+The main result of this work is presented in \refsec{for} : the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
 We conclude in \refsec{concl}.
@@
 
-Large scale structure surveys usually probe the 3D dark matter overdensities as a 2D field projected along the line of sight:
-
+Large scale structure surveys usually probe the 3D matter overdensities as a 2D field projected along the line of sight:
 \begin{equation}
@@
 \end{equation}
-
 where $\delta(\chi(z)\nver,z)$ corresponds to the dark matter overdensity field at a comoving distance $\chi(z)$ and at a redshift $z$ in the angular direction $\nver$.
@@
 
-In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$. 
+In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$. 
 %Furthermore $W^{i}(z)$ is the kernel function of the field $i$ introduced in \refeq{wkernel}.
@@
 where $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$, $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
+\relook{As can be seen in \reffig{kernel} the lensing kernel $W^{\kappa}(z)$ is quite a broad function of redshift peaking around $z=1.8$ but extending to high redshift.}
 
@@
  Different galaxy surveys in this work are then fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxy density.
-We test the delensing efficiency taking into account both current surveys like WISE or DES as well as future galaxy surveys like DESI and LSST together with 21 cm measurement like SKA.
+We test the delensing efficiency taking into account both current surveys like WISE or DES as well as future galaxy surveys like DESI and LSST as well as 21 cm measurement like SKA.
 
@@
 DES is modeled after the DES Science Verification public data release.
-For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy denisty of 0.63 galaxies per arcmin squared. 
+For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy denisty of 0.63 galaxies per squared arcmin. 
 %As in \cite{passaglia:2017} we use a linear bias $b$
@@
 We model both the redshift distribution and  bias of radio sources following \cite{namikawa:2016a}.
+
+\relook{In \reffig{kernel}, we compare the kernels $W^{g}(z)$ of all the tracers introduced here to the CMB lensing kernel $W^{\kappa}(z)$.}
 
@@
 We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$ and we set $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
+\relook{\reffig{kernel} shows how the CIB kernel peaks at higher redshift compared to other galaxy survey kernels, with a better overlap with the lensing one.}
+
 There are several available CIB observations that have already been used to delens the CMB.
@@
 However, this promising sign of primordial gravitational waves is partially obscured by gravitational interactions with large scale structures that generate CMB B-modes by distorting primordial E-modes.
-At first order, given the convergence field $\kappa= -\frac{1}{2}\nabla^2\phi$ introduced in \refsec{kappaCMB} the B-modes resulting from the lensing of primordial E-modes are:
+At first order, given the convergence field $\kappa= -\frac{1}{2}\nabla^2\phi$ introduced in \refsec{kappaCMB} the B-modes resulting from the lensing of the primordial E-modes are:
 \be
@@
 \be
-C_{\ell}^{BB,\mathrm{full}} = C_{\ell}^{BB,r} + C_{\ell}^{BB,\mathrm{lens}} + N_{\ell}^{BB}.
+C_{\ell}^{BB,\mathrm{measured}} = C_{\ell}^{BB,r} + C_{\ell}^{BB,\mathrm{lens}} + N_{\ell}^{BB}.
 \ee
-The lensing component is a significant source of B-modes that, at large scales, correspond to a white noise source of roughly $5 \mu K$-arcmin.
-This means that it is not only bigger than the allowed inflationary component at scales smaller than several degrees, but it is also comparable with current levels of instrumental noise.
+The lensing component is a significant source of B-modes that, at large scales, corresponds to a white noise source of roughly $5 \mu K$-arcmin independent of the angular scale.
+This means that it is not only larger than the allowed inflationary component at scales smaller than several degrees, but it is also comparable to current levels of instrumental noise.
 For this reason, it is critical to characterize and eventually remove it from the data.
 To do so, we build a template \refeq{blens} of the lensing B-modes in the observed patch given a measurement of the E-mode field and the lensing potential $\phi$.
-While E is measured directly, we can estimate $\phi$ using "tracers" of the dark matter distribution that sources the potential.
+While E is measured directly, we can estimate $\phi$ using "tracers" of the matter distribution that sources the potential.
+
+\subsection{Single tracer of the lensing potential}\label{sec:single-tracer}
 
@@
 
-The optimal weights $f(\bl,\bl')$ such that the residual lensing B mode power is minimized are:
+The optimal weights $f(\bl,\bl')$, chosen such that the residual lensing B mode power is minimized, are:
 \be
@@
 
-Notice that the first term consists of an inverse variance filter applied to the measured E-mode and the second minimizes the difference between the reconstructed $\phi$ and the CMB lensing potential.
+Notice that the first term consists of an inverse variance filter applied to the measured E-mode and the second captures the difference between the reconstructed $\phi$ and the CMB lensing potential.
 
-With this choice of $f(\bl,\bl')$ we finally have that the residual power is:
+With this choice of $f(\bl,\bl')$ we find that the residual power is:
 \beqn
@@
 \refeq{Bres} directly relates the residual power after delensing with the cross correlation coefficients with CMB lensing of the tracers used.
-The bigger the $\rho_{\ell}^{2}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
+The larger the $\rho_{\ell}^{2}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
 A higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$ leading to a smaller residual power $C_{{l}}^{BB,\mathrm{res}}$.
@@
 We start by assuming that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
-We can optimally combine them to estimate $\phi$ or, in other word, to maximize the correlation factor $\rho$ with:
+We can optimally combine them to estimate $\phi$ or, in other words, to maximize the correlation factor $\rho$ with:
 \beqn
@@
 The cross-correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
-CIB and 21 cm surveys probe the high redshift structures and, independently from the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
+The cosmic infrared background and 21 cm surveys probe the high redshift structures and, independent of the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
 On the other end, galaxy clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
@@
 However, as can be seen in \reffig{kernel}, the kernel overlap of a tracer with the CMB lensing varies as a function of redshift.
-For this reason, the optimal approach is to weight galaxies at different redshifts with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
+For this reason, the optimal approach is to weight galaxies at different redshift with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
 We can see this with a simple example. Let's split a single survey I into two non-overlapping redshift bins $I_{1}$ and ${I_{2}}$ with $I=I_{1}+I_{2}$.
@@
 The parameter uncertainties are inversely proportional to the covariance of the measurement.
-Since the lensing B-modes $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them through delensing will reduce the parameter statistical error.
-After delensing, the error $\sigma^{\rm{del}}(p)$ can still be obtained from \refeq{fisher} with $C_{{l}}^{BB,\mathrm{res}}<C_{{l}}^{BB,\mathrm{lens}}$ in the covariance defined in \refeq{cova}.
+Since the lensing B-modes power $C_{{l}}^{BB,\mathrm{lens}}$ is a substantial component of the covariance, removing part of it through delensing will reduce the parameter statistical error.
+We define the error after delensing as $\sigma^{\rm{del}}(p)$.
+This is computed from \refeq{fisher} substituting $C_{{l}}^{BB,\mathrm{lens}}$ with the residual power after delensing $C_{{l}}^{BB,\mathrm{res}}$ in the covariance matrix defined in \refeq{cova}.
 We define the improvement as the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$.
@@
 We checked this for a few of the tracer combinations used here and find it particularly true once several tracers are jointly taken into account.
-Finally, we have assumed a Gaussian covariance $\sigma( C_{\ell}^{BB} )$ even if has small non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
+Finally, although we have assumed a Gaussian covariance $\sigma( C_{\ell}^{BB} )$ it actually has small non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
 This approximation is good enough to show the improvement due to delensing.
@@
 
-In both cases the CMB lensing reconstruction correlates very well at low $\ell$ and it then falls rapidly at smaller scales because of the raise of the reconstruction noise.
 For Planck, the internal reconstruction is at most $70\%$ correlated at very large scales and then it falls rapidly to $40\%$ at $\ell=200$.
 Its correlation is comparable to the one of LSS at almost all scales.
-On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 80\% correlated with the true field at $\ell<300$, and only then the correlation becomes comparable with the current LSS surveys.
+On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 80\% correlated with the true field at $\ell<300$, and only then the correlation becomes comparable to the current LSS surveys.
+In both cases the CMB lensing reconstruction correlates very well at low $\ell$ and it then falls rapidly at smaller scales because of the raise of the reconstruction noise.
 
@@
 Finally, \reffig{corrs4_01} shows that a futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
-This performance is comparable with a CMB-S4 internal reconstruction.
+This performance is comparable to a CMB-S4 internal reconstruction.
 

### 2017-07-08 Notes/galaxy_delensing.tex
Scott comments up to page 9

@@
 \includegraphics[scale=1.]{../images/actual_scenario_planck.pdf}
-\caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the angular scale $\ell$. 
-The dashed curve corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$, showing which scales contribute to the $\ell<100$ B-mode power.}
+\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$ for current generation tracers.
+Here we show both current galaxy survey and internally reconstructed $\phi$ CMB lensing potential. 
+The dashed curve highlights which scales contribute to the $\ell<100$ B-mode power and are thus most useful to delens.
+It corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$}
 \label{fig:corrnow-planck}
@@
 \includegraphics[scale=1.]{../images/actual_scenario.pdf}
-\caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
+\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$ for current generation tracers. 
+This figure is the same as \reffig{corrnow-planck} but using SPTPol instead of Planck as the high-res CMB experiment performing the internal reconstruction.
+}
 \label{fig:corrnow}
@@
 Nonetheless, it is still worthwhile considering all the tracers.
-Indeed a multi tracer-approach can bring the removed power on the overlapping area from 27\% of the CIB alone to 36\% using all the LSS and 41\% when Planck lensing reconstruction is added to the set.
+Indeed a multi tracer-approach can bring the removed power on the overlapping area from 27\% of the CIB alone to 36\% using all the LSS (LSS-S2) and 41\% when Planck lensing reconstruction is added to the set.
 However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 30\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
@@
 \caption{$\alpha(r)$: Improvements on $\sigma(r)$ due to delensing for current generation and Planck lensing reconstruction.
+LSS-S2 corresponds to the combination of all the available LSS tracers.
 The values in parenthesis in the first column correspond to the fraction of lensing B-mode power removed using each LSS tracer. 
@@
 \hline
-Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
 WISE (6\%)& 1.09& 1.05 & 1.01 \\ \hline
@@
 CIB (27\%) &1.45& 1.21  & 1.07\\ \hline
-LSS (36\%) &1.7& 1.29  & 1.09\\ \hline
+LSS-S2 (36\%) &1.7& 1.29  & 1.09\\ \hline
 CMB Planck (8\%) &1.06& 1.05  & 1.01 \\ \hline
-LSS+CMB (41\%)& 1.77& 1.31& 1.10  \\ \hline
+LSS-S2 +CMB (41\%)& 1.77& 1.31& 1.10  \\ \hline
 \end{tabular}
@@
  \begin{table}
-\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to the ratio of the error before and after delensing.}
+\caption{$\alpha(r)$: Current generation improvements on $\sigma(r)$ due to delensing.
+We use the same LSS tracers as \reftab{current} but with the internal lensing reconstruction performed by SPTPol.}
   \vspace{0.2cm}
@@
 \hline
-Surveys &$\alpha(r=0), N^{B}_{\ell}=0$& $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
-CMB SPTPol (35\%) &1.46& 1.27  & 1.08 \\ \hline
-LSS+CMB (56\%) & 2.18 & 1.57 & 1.16  \\ \hline
+Surveys &$\alpha_{r=0}, N^{B}_{\ell}=0$& $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
+SPTPol (35\%) &1.46& 1.27  & 1.08 \\ \hline
+LSS-S2 +SPTPol (56\%) & 2.18 & 1.57 & 1.16  \\ \hline
 
@@
 
-\subsection{CMB-S3 Era}
+\subsection{CMB-S3 Era}\label{sec:S3}
 CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
@@
 \includegraphics[scale=1.]{../images/S3_scenario.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
+\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$.
+This figure is the same as \reffig{corrnow} but with a stage-3 CMB experiment performing the internal reconstrcution and with DESI added to the stage-2 galaxy surveys (LSS-S2).
+}
 \label{fig:corrS3}
@@
 Given the improvement in the noise of the high-resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
-Despite this, adding galaxy survey will still lead to a decent improvement with a relatively low effort.
+Adding galaxy surveys will lead to a further improvement without additional effort.
 Indeed galaxy surveys will still be able to remove an additional $20\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by $15\%$.
@@
 \begin{table}
-\caption{$\alpha$: S3 improvements. The absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is $\sigma(r) =$  }
+\caption{$\alpha$: Improvements on $\sigma(r)$ due to delensing for S3 experiments.
+Here we define LSS-S3 as the combination of S2 LSS tracers (LSS-S2) and DESI.
+We also use a CMB S3 experiment for the lensing internal reconstruction.
+See \refsec{S3} for details.
+%The absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is $\sigma(r) =$  
+}
   \begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
 DESI (10\%) &1.12 & 1.08  & 1.03 \\ \hline
-LSS (40\%) &1.73& 1.42  & 1.2\\ \hline
+LSS-S3 (40\%) &1.73& 1.42  & 1.2\\ \hline
 CMB 3G (56\%) &2& 1.57  & 1.21 \\ \hline

### 2017-07-08 Notes/galaxy_delensing.tex
text: abstract adnconcl

@@
 We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
-We find that, in general, splitting galaxy surveys into tomographic bins can improve the correlation factor with CMB lensing by 30\%. 
-Considering available survey, a galaxy survey like DES can remove about $14\%$ of the lensing signal. current generation CMB experiments will benefit from ...
-With the improvement of the noise CMB internal reconstruction will become more important and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level.
-Nevertheless optical galaxy surveys, in particular if split in different tomographic bin will still play an important role even for CMB S4.
+We find that, in general, galaxy surveys should be splitted into tomographic bins as this can improve the correlation with CMB lensing by 30\%.
+Considering currently available survey, a DES-like galaxy survey can remove about $14\%$ of the lensing signal.
+The CMB experiments of the current generation will particularly benefit from large scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
+With the decrease of the instrumental noise, the CMB internal reconstruction will increase its efficiency and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level.
+Nevertheless optical galaxy surveys will still play an important role even for CMB S4.
+In particular LSST can a achieve a delensing performance comparable to a 3G CMB experiment but and completely different systematics.
+This will be important to prove the robustness against systematics of an eventual detection of primordial B-modes.
+
+%
 
@@
 
-Notice that the first term consists of an inverse variance filter applied to the measured E-mode and the second captures the difference between the reconstructed $\phi$ and the CMB lensing potential.
-
 With this choice of $f(\bl,\bl')$ we find that the residual power is:
@@
 
-\refeq{Bres} directly relates the residual power after delensing with the cross correlation coefficients with CMB lensing of the tracers used.
+\refeq{Bres} highligts the different factors that control the delensing efficiency.
+The first part of the second term in the parenthesis consists of an inverse variance filter applied to the measured E-mode.
+The smaller the noise in the E-modes ($N^{EE}$) is the closer this term is to one: a less noisy measurement improves the template of the lensing B-modes.
+The second captures the difference between the reconstructed $\phi$ and the CMB lensing potential, and it directly relates the residual power after delensing with the cross-correlation coefficients with CMB lensing of the tracers used.
 The larger the $\rho_{\ell}^{2}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
@@
 \includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
+\caption{Kernels Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
 The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential will be leading to a higher delensing efficiency.
@@
 \includegraphics[scale=1.]{../images/compare_kernel_tomo.pdf}
-\caption{\textbf{Kernels Comparison with tomographic bins}: Comparison of the 10 LSST  tomographic bins together with the full LSST survey and the CMB lensing kernel. 
-Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross correlation with the CMB lensing.
+\caption{Kernels Comparison with tomographic bins: Comparison of the 10 LSST  tomographic bins together with the full LSST survey and the CMB lensing kernel. 
+Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross-correlation with the CMB lensing.
 This leads to a better delensing efficiency.
@@
 For photometric surveys, the maximum number of bins is dictated by the fact that the bin width can not be smaller than the photo-z accuracy.
-We used 10 and 4 photometric bins for LSST and DES respectively with a common photo-z accuracy $\sigma(z) = 0.05(1+z)$. 
+We used 10 and 4 photometric bins for LSST and DES respectively with a photo-z accuracy of $\sigma(z)_{\textrm{DES}} = 0.05(1+z)$ for DES and $\sigma(z)_{\textrm{LSST}} = 0.01(1+z)$ for LSST. 
 In spectroscopic surveys, there are no limitations in increasing the number of bins. We split DESI into 4 spectroscopic bins with no overlap among each others.
@@
 CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
-We model the CMB S3 with a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam. 
-We assume here that a deep CMB experiment and an overlapping high-resolution one will be able to combine their measured modes.
-For this reason, we will also assume this level of noise for the internal noise reconstruction.
+As we did for Stage-2, we model the CMB S3 as two different overlapping experiments that will be able to combine their measurement.
+The two experiments combined will have a level of noise of \textcolor{red}{$3\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam.
+This level of noise is also assumed for the internal noise reconstruction.
 With this experimental setup, we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
@@
 We also use a CMB S3 experiment for the lensing internal reconstruction.
-See \refsec{S3} for details.
+See \refsec{S3} for details.}
 %The absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is $\sigma(r) =$ 
-}
+\hspace{2.2cm}
 \begin{tabular}{|c | c | c | c|}
@@
 
-\reftab{S4} shows the improvement on the inflationary constraint given the cross correlation factors described above. 
+\reftab{S4} shows the improvement on the inflationary constraint given the cross-correlation factors described above. 
 LSST alone will be able to remove almost half of the B-mode lensing power.
@@
 %For current experiment, the cosmic infrared background had already proven to be very efficient in delensing \cite{larsen:2016,manzotti:2017}.
-For ongoing experiments, we find that an optical survey like DES alone is able to remove $14\%$ of the lensing component and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the contaminant signal using only LSS survey.
+
+For ongoing experiments, we find that combining LSS tracers with CMB internal reconstruction will be particularly beneficial.
+An optical survey like DES alone is able to remove $14\%$ of the lensing component and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the contaminant signal using only LSS survey.
 Depending on the CMB noise and the amount of galactic foreground cleaning these can correspond to an improvement of 30\% in the constrain on the tensor to scalar ratio $r$.
-
-In the future, the improved level of noise of CMB experiments will allow to internally reconstruct the structures lensing the CMB. 
-The fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level. 
+In the future, the decreasing level of instrumental noise of CMB experiments will dramatically improve the internal reconstruction of the structures lensing the CMB. 
+The fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level. 
 Already in the 3G era, the CMB internal reconstruction will be the main source of delensing.
@@
 
-Furthermore internal delensing will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}.
+Even for CMB S4, galaxy surveys will still play an important role.
+For example, a tomographically binned LSST-survey will remove 56\% of the lensing power.
+While this performance is comparable to a Stage-3 CMB it will allow to probe the robustness against systematics of an eventual detection.
+Indeed delensing with just CMB data will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}.
 For this reason, efficient galaxies tracers are not only useful in the short term but will also play a role in cross-checking these internal biases.
-For a future galaxy survey like LSST we find that using tomographic binning can improve its correlation with lensing by $30\%$.
 
-Giving the high level of correlation of future galaxy surveys and CMB lensing it is worth looking for possible application beside the detection of inflationary B-modes.
-For example LSS delensing allow to remove just the low-z lensing component from the CMB lensing.
-Removing this component can avoid the necessity of modelling nonlinearities in the study of CMB lensing power spectrum.
+Given the high level of correlation of future galaxy surveys and CMB lensing, it is also worth looking for potential application besides the detection of inflationary B-modes.
+For example, LSS delensing allows to selectively remove from the CMB  the gravitational effect coming from the non-linear structures in the low-redshift universe. 
+Removing this component will reduce the level of non-linearities both in the CMB and in the CMB lensing power spectrum providing an alternative route to exploit all the measurable modes without having to model the non-linear component at small angular scales.
 

### 2017-07-08 Code/Plot_figures.py
images change labels

@@ -518,9 +518,9 @@ plt_func(ell, ell * (ell + 1) * np.array(B_res3[-1]) /
 
 # plt_func(clbb(r=0.01, lmax=3000), label=r'$C^{BB^{\rm{tot}}}_{\ell}$')
 fact = np.arange(0, 4001) * (np.arange(0, 4001) + 1.) / 2. / np.pi
-plt.plot(np.arange(0, 4001), fact * nl(9, 1, 4000), 'b', linestyle='--',
+plt.plot(np.arange(0, 4001), fact * nl(9, 1, 4000), color = 'black', linestyle='--',
          linewidth=font_size / 14.5, alpha=0.7)
-plt.plot(np.arange(0, 4001), fact * nl(1, 1, 4000), 'b', linestyle='--',
+plt.plot(np.arange(0, 4001), fact * nl(1, 1, 4000), color = 'black', linestyle='--',
          linewidth=font_size / 14.5, alpha=0.7)
 plt.ylim(1e-5, 1.5e-1)
 plt.xlim(10, 1700)

### 2017-07-08 Code/limber_integrals.pyx

@@ -105,7 +105,7 @@ def  cl_limber_z(chi_z, hspline, rbs, l, kernel_1, kernel_2=None,  zmin=0.0,  zm
     # print('')
     # print(func(zmax-(zmax-zmin)/2.))
     # print(scipy.integrate.quad(integrand, zmin, zmax, limit=300, epsrel=1.49e-06)[0])
-    return scipy.integrate.quad(integrand, zmin, zmax, limit=600, epsabs=0., epsrel=0.49e-03)[0]
+    return scipy.integrate.quad(integrand, zmin, zmax, limit=600, epsabs=0., epsrel=5e-04)[0]
 
 
 

### 2017-07-08 Code/plots/plot_kernel.py

@@ -141,21 +141,21 @@ print(fig_dims)
 ax1 = fg.add_subplot(111)
 
 plt.plot(kappakernel_l30[:, 0], kappakernel_l100[:, 1] /
-         np.max(kappakernel_l100[:, 1]), label=r'$CMB$', alpha=0.8)
+         np.max(kappakernel_l100[:, 1]), label=r'CMB', alpha=0.8)
 # plt.plot(kappakernel_l30[:, 0], kappakernel_l100[:, 1] /
 #          np.max(kappakernel_l100[:, 1]), color='#e41a1c', linestyle=':')
 # plt.plot(kappakernel_l30[:, 0], kappakernel_l500[:, 1] /
 #          np.max(kappakernel_l500[:, 1]), color='#e41a1c', linestyle='--')
 
 plt.plot(cibkernel_l100[:, 0], cibkernel_l100[:, 1] /
-         np.max(cibkernel_l100[:, 1]), label=r'$CIB$')
+         np.max(cibkernel_l100[:, 1]), label=r'CIB')
 # plt.plot(cibkernel_l30[:, 0], cibkernel_l100[:, 1] /
 #          np.max(cibkernel_l100[:, 1]), color='#377eb8', linestyle=':')
 # plt.plot(cibkernel_l30[:, 0], cibkernel_l500[:, 1] /
 #          np.max(cibkernel_l500[:, 1]), color='#377eb8', linestyle='--')
 
 plt.plot(deskernel_l30[:, 0], deskernel_l100[:, 1] /
-         np.max(deskernel_l100[:, 1]), label=r'$DES$', alpha=0.8)
+         np.max(deskernel_l100[:, 1]), label=r'DES', alpha=0.8)
 # plt.plot(deskernel_l30[:, 0], deskernel_l100[:, 1] /
 #          np.max(deskernel_l100[:, 1]), color='#4daf4a', linestyle=':')
 # plt.plot(deskernel_l30[:, 0], deskernel_l500[:, 1] /
@@ -163,14 +163,14 @@ plt.plot(deskernel_l30[:, 0], deskernel_l100[:, 1] /
 
 
 plt.plot(skakernel_l100[:, 0], skakernel_l100[:, 1] /
-         np.max(skakernel_l100[:, 1]), label=r'$SKA$', alpha=0.8)
+         np.max(skakernel_l100[:, 1]), label=r'SKA', alpha=0.8)
 
 
[…]

### 2017-07-08 Code/Plot_figures.py
images change labels

@@ -273,7 +273,7 @@ plt.savefig('../images/S4_scenario_ska01.pdf', dpi=600, papertype='Letter')
 plt.savefig('../images/S4_scenario_ska01.png')
 
 
-sys.exit()
+# sys.exit()
 
 
 labels = ['lsst_bin0', 'lsst_bin1', 'lsst_bin2', 'lsst_bin3', 'lsst_bin4',

### 2017-07-09 Code/limber_integrals.pyx
images change labels

@@ -98,13 +98,6 @@ def  cl_limber_z(chi_z, hspline, rbs, l, kernel_1, kernel_2=None,  zmin=0.0,  zm
             pk=rbs.ev((l + 0.5) / x, z)
             return 1. / x /x * h * k1 * k2 * pk
 
-    # print 'here' ,integrand(0.5)
-
-    # sys.exit()
-    # func = InterpolatedUnivariateSpline(np.linspace(zmin,zmax,100), np.vectorize(integrand)(np.linspace(zmin,zmax,100)), ext='zeros')
-    # print('')
-    # print(func(zmax-(zmax-zmin)/2.))
-    # print(scipy.integrate.quad(integrand, zmin, zmax, limit=300, epsrel=1.49e-06)[0])
     return scipy.integrate.quad(integrand, zmin, zmax, limit=600, epsabs=0., epsrel=5e-04)[0]
 
 

### 2017-07-09 Code/run_analysis.py

@@ -493,8 +493,18 @@ for i, label in enumerate(rho_names):
 
     print('After delensing % errors', sigma_r_1, sigma_nt)
     print(probe, 'gain = ', sigma_r_1 / sigma_r)
+# ====================================
+# ====================================
 
-# ## CMB S3
+## CMB S3
+
+# ====================================
+# ====================================
+
+
+
+
+
 
 # In[278]:
 print('')
@@ -607,17 +617,15 @@ for i, label in enumerate(rho_names):
         noise_uK_arcmin=noise_uK_arcmin,
         fwhm_arcmin=fwhm_arcmin)
 
-    print('After delensing % errors', sigma_r, sigma_nt)
+    print('After delensing % errors', sigma_r, sigma_r_1, sigma_nt)
     print(probe, 'gain = ', sigma_r_1 / sigma_r, sigma_nt_1 / sigma_nt)
 
 
 # ### r=0.12
+clbb.cache_clear()
+clbb_tens.cache_clear()
 
-# In[307]:
-
-# noise_uK_arcmin=4.5,
-# fwhm_arcmin=4.,
-lmax = 2000
+lmax = 4000
 noise_uK_arcmin = 3.
 fwhm_arcmin = 1.
 r_fid = 0.12

### 2017-07-09 Notes/galaxy_delensing.tex
typos

@@
 
-\title{Future Cosmic Microwave Background delensing with galaxy surveys.}         % Enter your title between curly braces
-\author{A. Manzotti}        % Enter your name between curly braces
+\title{Future Cosmic Microwave Background delensing with galaxy surveys.}       
+\author{A. Manzotti} 
 \email{manzotti.alessandro@gmail.com}
@@
 \date{\today}
-
-        % Enter your date or \today between curly braces
 \begin{abstract}
@@
 This approximation is good enough to show the improvement due to delensing.
+The approximations made here go in the direction of a slightly optimistic \textit{absolute} value for the parameters statistical uncertainties both for the standard and the delensed case.
+The delensing improvement is defined as the \textit{relative} value of these uncertainties $\alpha_{i} = \sigma^{\rm{del}}(i)/\sigma(i)$ and, at first order, it is not affected by this assumptions.
 
@@
 For this reason, efficient galaxies tracers are not only useful in the short term but will also play a role in cross-checking these internal biases. 
-
-Given the high level of correlation of future galaxy surveys and CMB lensing, it is also worth looking for potential application besides the detection of inflationary B-modes.
-For example, LSS delensing allows to selectively remove from the CMB  the gravitational effect coming from the non-linear structures in the low-redshift universe. 
-Removing this component will reduce the level of non-linearities both in the CMB and in the CMB lensing power spectrum providing an alternative route to exploit all the measurable modes without having to model the non-linear component at small angular scales.
-
 % ===================
-%\onecolumngrid
+\onecolumngrid
 \begin{figure}[htbp]
 \begin{center}
-\includegraphics[scale=1.]{../images/errors_summary.pdf}
+\includegraphics[scale=1]{../images/errors_summary.pdf}
 \caption{\relook{to be filled with final results}}
@@
 % ===================
+
+Given the high level of correlation of future galaxy surveys and CMB lensing, it is also worth looking for potential application besides the detection of inflationary B-modes.
+For example, LSS delensing allows to selectively remove from the CMB  the gravitational effect coming from the non-linear structures in the low-redshift universe. 
+Removing this component will reduce the level of non-linearities both in the CMB and in the CMB lensing power spectrum providing an alternative route to exploit all the measurable modes without having to model the non-linear component at small angular scales.
 

### 2017-07-10 Notes/galaxy_delensing.tex
refs

@@
 This will be important to prove the robustness against systematics of an eventual detection of primordial B-modes.
-
-%
-
+% maybe non linear scales?
 \end{abstract}
@@
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
-This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,polarbear2014c,vanengelen14b,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
+This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
 
@@
 Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
-Furthermore, it has recently been performed on CMB temperature data using the cosmic infrared background as LSS tracer \cite{larsen:2016} and on CMB temperature and polarization data using CMB data to internally reconstruct the LSS lensing potential \cite{ carron17}.
+Furthermore, it has recently been performed on CMB temperature data using the cosmic infrared background as LSS tracer \cite{larsen:2016a} and on CMB temperature and polarization data using CMB data to internally reconstruct the LSS lensing potential \cite{carron:2017}.
 Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in which $~28\%$ of the lensing power was removed.
@@
 where $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$, $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
-\relook{As can be seen in \reffig{kernel} the lensing kernel $W^{\kappa}(z)$ is quite a broad function of redshift peaking around $z=1.8$ but extending to high redshift.}
+As can be seen in \reffig{kernel}, the lensing kernel $W^{\kappa}(z)$ is a broad function of redshift peaking around $z=1.8$ but extending to high redshift.
 
@@
 We model both the redshift distribution and  bias of radio sources following \cite{namikawa:2016a}. 
-
-\relook{In \reffig{kernel}, we compare the kernels $W^{g}(z)$ of all the tracers introduced here to the CMB lensing kernel $W^{\kappa}(z)$.}
+In \reffig{kernel}, we compare the CMB lensing kernel $W^{\kappa}(z)$ to the kernels $W^{g}(z)$ of all the tracers introduced here.
 
@@
 This model provides an accurate fit to several experimental results.
-For the cross-spectra with the CMB lensing or other galaxy tracers, $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall10}.
+For the cross-spectra with the CMB lensing or other galaxy tracers, $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall:2010}.
 It corresponds to the kernel:
@@
 We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$ and we set $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
-\relook{\reffig{kernel} shows how the CIB kernel peaks at higher redshift compared to other galaxy survey kernels, with a better overlap with the lensing one.}
+\reffig{kernel} shows how the CIB kernel peaks at higher redshift compared to other galaxy survey kernels, with a better overlap with the CMB lensing one.
 
@@
 SKA in its phase 1 will remove 53\% of the lensing power increasing 
+\relook{FINISH}
 
@@
 \hline
-Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
 LSST (10\%) &1.12 & 1.08  & 1.03 \\ \hline
@@
 \end{tabular}
-\label{tab:S4}
+\label{tab:S4_2}
 \end{table}
@@
 While this performance is comparable to a Stage-3 CMB it will allow to probe the robustness against systematics of an eventual detection.
-Indeed delensing with just CMB data will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}.
+Indeed delensing with just CMB data will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2017,namikawa:2017}.
 For this reason, efficient galaxies tracers are not only useful in the short term but will also play a role in cross-checking these internal biases.
-% ===================
-\onecolumngrid
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1]{../images/errors_summary.pdf}
-\caption{\relook{to be filled with final results}}
-\label{fig}
-\end{center}
-\end{figure}
-\twocolumngrid
-
-% ===================
 
@@
 
+% ===================
+\newpage
+\onecolumngrid
+\begin{figure}[htbp]
+%\begin{center}
+\includegraphics[scale=1]{../images/errors_summary.pdf}
+\caption{\relook{to be filled with final results}}
+\label{fig}
+%\end{center}
+\end{figure}
+\twocolumngrid
+% ===================
+
 \bibliographystyle{plain}
 \bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
+%/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
 

### 2017-07-10 Notes/galaxy_delensing.tex
refs

@@
 
-For LSST we follow \cite{schaan:2016}: $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
+For LSST we follow \cite{schaan:2017}: $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
 Furthermore we assume a density of 26 galaxies per arcmin squared.

### 2017-07-10 Notes/galaxy_delensing.tex
S4 resu

@@
 \subsection{CMB-S4 Era}
-
-An ambitious program for a stage-4 ground CMB experiment is currently under planning \cite{abazajian:2016}. 
-Moreover, satellite and balloon CMB experiments have been proposed and have the potential to extend the accessible B-mode measurement to the largest scales.
-Following \cite{abazajian:2016} we assume a CMB-S4 ground experiment with a level of noise of \textcolor{red}{$1\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam.
+Having as a major goal the detection of inflationary B-modes, CMB data will continue improving even after Stage-3.
+An ambitious program for a Stage-4 ground CMB experiment is in the planning phase \cite{abazajian:2016}. 
+Moreover, satellite and balloon CMB experiments have been proposed and have the potential to extend the accessible B-mode measurements to the largest scales.
+Given the unprecedented low level of noise of these experiment, delensing will be even more important than previous generations.
+Following \cite{abazajian:2016}, here we assume a CMB-S4 ground experiment with a level of noise of \textcolor{red}{$1\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam.
 Furthermore, by the mid-2020s, several next-generation LSS surveys will be online. 
-Here, as an example for future optical galaxy surveys, we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers.
-We test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance. 
-Even if they will observe an overlapping part of the sky and similar modes, because of their different strengths and weaknesses it will still be important to combine them efficiently.
+Here, as an example of future optical galaxy surveys, we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers. 
+We do not consider other surveys but we test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance. 
+Even if these LSS surveys will observe an overlapping part of the sky and similar modes, because of their different strengths and weaknesses, it will still be important to combine them efficiently.
 We also consider a radio continuum survey modeled following the SKA specifications.
@@
 However, this technique has the potential to map the LSS with relatively low noise up to redshift $z=6$.
-Here we considered just two possible  detection thresholds at 1 GHz (flux cut): 10 $\mu$Jy and 0.1 $\mu$Jy, which are representatives of the upper and lower limit levels for a SKA phase 1 (SKA1) and a SKA phase 2 (SKA2) survey.
+Here we considered just two possible detection thresholds at 1 GHz (flux cut): 10 $\mu$Jy and 0.1 $\mu$Jy, which are representatives of the upper and lower limit levels for a SKA phase 1 (SKA1) and a following SKA phase 2 (SKA2) survey.
+Given the importance of delensing for Stage-4 experiments, here we combine radio and optical survey to test their delensing efficiency.
 
-The correlation factor with CMB lensing for stage-4 experiments is shown in \reffig{corrs4_10}  and in \reffig{corrs4_01} for SKA1 and SKA2 respectively.
-
+The correlation factor with CMB lensing for Stage-4 experiments is shown in \reffig{corrs4_10}  and in \reffig{corrs4_01} for SKA1 and SKA2 respectively.
+%=======
 \begin{figure}[htbp]
@@
 \end{figure}
-
+%=======
 \begin{figure}[htbp]
@@
 \end{figure}
-
+%=======
 A tomographically binned LSST-like experiment will be a very efficient CMB lensing tracer.
@@
 Despite the improvement in galaxy surveys, the CMB internal reconstruction will still be the main source of delensing.
-This, having a perfect kernel overlap with the true lensing potential will benefit from the very low level of noise of CMB S4 experiments.
-\reffig{corrs4_10} shows that a CMB-S4 experiment will be able to internally reconstruct the lensing potential at more than 90\% up to $\ell=400$
-At this level, CMB S4 internal delensing will probably be limited by small secondary effects like foregrounds contamination, missing modes due to filtering etc.
-Finally, \reffig{corrs4_01} shows that a futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
+Having a perfect kernel overlap with the true lensing potential it will benefit from the very low level of noise of CMB S4 experiments.
+\reffig{corrs4_10} shows that a CMB-S4 experiment will be able to internally reconstruct the lensing potential at more than 90\% up to $\ell=400$.
+At this level, CMB S4 internal delensing will not be limited by noise but by small secondary effects like foregrounds contamination, filtered modes etc.
+Finally, \reffig{corrs4_01} shows that a very futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
 This performance is comparable to a CMB-S4 internal reconstruction.
@@
 This will lead to an improvement in $\sigma(r)$ of a factor of 1.8.
-If we add all the other LSS tracers previously included (i.e.
+If we add all the other infrared and optical LSS tracers previously included (i.e.
 excluding SKA) the improvement factor increase to 2.3.
-SKA in its phase 1 will remove 53\% of the lensing power increasing 
+These gives an important opportunity: for sufficiently high value of $r$ it will be possible to confirm a detection of primordial gravity waves using spectra delensed with LSS, completely independent of CMB internal data.
+Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of CMB S3 internal reconstruction.
+Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power.
+
 \relook{FINISH}
@@
 \end{table}
-
 %=======================================
-
 \begin{table}
@@
 \end{table}
-
 %=======================================
-
 %\subsection{Bias uncertainties degradation}
@@
 %We compute the derivatives of the power spectra as described
-
 \section{Conclusions}
 \label{sec:concl}
-
 The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal ("delensing") is necessary to test the inflationary paradigm using the next generation of CMB polarization experiments.

### 2017-07-10 Notes/galaxy_delensing.tex
S4 resu

@@
 Nevertheless optical galaxy surveys will still play an important role even for CMB S4.
-In particular LSST can a achieve a delensing performance comparable to a 3G CMB experiment but and completely different systematics.
+In particular an LSST-like survey can a achieve a delensing performance comparable to a 3G CMB experiment but with completely different systematics.
 This will be important to prove the robustness against systematics of an eventual detection of primordial B-modes.
@@
 These gives an important opportunity: for sufficiently high value of $r$ it will be possible to confirm a detection of primordial gravity waves using spectra delensed with LSS, completely independent of CMB internal data.
-Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of CMB S3 internal reconstruction.
+Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of the CMB S3 internal reconstruction.
 Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power.
+This will bring an almost factor of 4 improvement in the null test case even with realistic noise.
+Combining with LSS will only increase the removed power compared to CMB only by $\simeq 5\%$ but it will be highly beneficial to test the robustness of the final result.
 
-\relook{FINISH}
+We find the improvement in $\sigma(n_{t})$ to be very similar to the one in $\sigma(r)$ for all the CMB-S4 cases.
+This is reasonable because, given the level of noise in CMB-S4, the variance at angular scales $\ell<200$ are dominated by instrumental noise and not by the lensing component.
+Independent of the delensing level, the small value allowed for $r$ will not allow to test the consistency relations.
 
@@
 CMB S4 (81\%) & 4.63 & 3.96  &  \\ \hline
-LSS+CMB (69\%)& 2.8 & 1.8&  \\ \hline
+LSS+CMB (86\%)& 5.2 &  &  \\ \hline
 \end{tabular}
@@
 %=======================================
-\begin{table}
-\caption{$\alpha$: Gen 4 experiments}
-  \begin{tabular}{|c | c | c | c|}
-\hline
-Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-LSST (10\%) &1.12 & 1.08  & 1.03 \\ \hline
-SKA (0.1$\mu$Jy) (40\%) &1.73& 1.42  & 1.2\\ \hline
-CMB S4 (56\%) &2& 1.57  & 1.21 \\ \hline
-LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
-\end{tabular}
-\label{tab:S4_2}
-\end{table}
+%\begin{table}
+%\caption{$\alpha$: Gen 4 experiments}
+%  \begin{tabular}{|c | c | c | c|}
+%\hline
+%Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
+%LSST (10\%) &1.12 & 1.08  & 1.03 \\ \hline
+%SKA (0.1$\mu$Jy) (40\%) &1.73& 1.42  & 1.2\\ \hline
+%CMB S4 (56\%) &2& 1.57  & 1.21 \\ \hline
+%LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
+%\end{tabular}
+%\label{tab:S4_2}
+%\end{table}
 %=======================================
@@
 \label{sec:concl}
-The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal ("delensing") is necessary to test the inflationary paradigm using the next generation of CMB polarization experiments.
-To do so, we need to accurately reconstruct the large scale structures that lens the CMB in order to build a template of the lensing B-modes.
+The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal ("delensing") is necessary to test inflation using the next generation of CMB polarization experiments.
+To delens, we need to accurately reconstruct the large scale structures that lens the CMB in order to build a template of the expected lensing B-modes in the observed patch.
 In this paper, we studied the potential impact of large-scale structure galaxy surveys in this important endeavor.
@@
 For ongoing experiments, we find that combining LSS tracers with CMB internal reconstruction will be particularly beneficial.
-An optical survey like DES alone is able to remove $14\%$ of the lensing component and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the contaminant signal using only LSS survey.
-Depending on the CMB noise and the amount of galactic foreground cleaning these can correspond to an improvement of 30\% in the constrain on the tensor to scalar ratio $r$.
+An optical survey like DES alone is able to remove $14\%$ of the B-mode lensing power and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the power using only LSS survey.
+Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to an improvement of 30\% in the constrain on the tensor to scalar ratio $r$.
 In the future, the decreasing level of instrumental noise of CMB experiments will dramatically improve the internal reconstruction of the structures lensing the CMB. 
@@
 For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $57\%$ to $68\%$.
-
 Even for CMB S4, galaxy surveys will still play an important role.
@@
 While this performance is comparable to a Stage-3 CMB it will allow to probe the robustness against systematics of an eventual detection.
-Indeed delensing with just CMB data will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2017,namikawa:2017}.
-For this reason, efficient galaxies tracers are not only useful in the short term but will also play a role in cross-checking these internal biases.
+Indeed delensing with just CMB data will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) that we are trying to delens to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2017,namikawa:2017}.
+For this reason, efficient galaxies tracers are not only useful in the short term but will also play a role in cross-checking these internal biases in the future.
 
 Given the high level of correlation of future galaxy surveys and CMB lensing, it is also worth looking for potential application besides the detection of inflationary B-modes.
-For example, LSS delensing allows to selectively remove from the CMB  the gravitational effect coming from the non-linear structures in the low-redshift universe. 
+For example, delensing with galaxy surveys allows to selectively remove from the CMB only the gravitational effect coming from the non-linear structures in the low-redshift universe. 
 Removing this component will reduce the level of non-linearities both in the CMB and in the CMB lensing power spectrum providing an alternative route to exploit all the measurable modes without having to model the non-linear component at small angular scales.
@@
 \section{Acknowledgments}
-We thank S. Passaglia, S. Dodelson, K. Story and K. Wu for useful discussion and their review of the draft. 
+We thank S. Dodelson, S. Passaglia, K. Story and K. Wu for useful discussion and their review of the draft. 
 Furthermore we thank D. Alonso for his feedback on radio continuum surveys.

### 2017-07-10 Notes/galaxy_delensing.tex
S4 resu

@@
 For ongoing experiments, we find that combining LSS tracers with CMB internal reconstruction will be particularly beneficial.
-An optical survey like DES alone is able to remove $14\%$ of the B-mode lensing power and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the power using only LSS survey.
-Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to an improvement of 30\% in the constrain on the tensor to scalar ratio $r$.
-In the future, the decreasing level of instrumental noise of CMB experiments will dramatically improve the internal reconstruction of the structures lensing the CMB. 
+An optical survey like DES is able to remove $14\%$ of the B-mode lensing power alone, and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the power using only LSS surveys.
+Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to an improvement of 30\% in the constraint on the tensor to scalar ratio $r$.
+In the future, the decreasing level of instrumental noise of CMB experiments will dramatically improve its internal reconstruction of the structures lensing the CMB. 
 The fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level. 
-Already in the 3G era, the CMB internal reconstruction will be the main source of delensing.
-However it will still be less efficient than galaxies at small scales$\ell>500$.
-For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $57\%$ to $68\%$.
+Already for Stage-3, the CMB internal reconstruction will be the main source of delensing.
+However it will still be less efficient than galaxies at delensing small scales $\ell>500$.
+For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $57\%$ to $68\%$ for 3G.
 Even for CMB S4, galaxy surveys will still play an important role.
@@
 Given the high level of correlation of future galaxy surveys and CMB lensing, it is also worth looking for potential application besides the detection of inflationary B-modes.
-For example, delensing with galaxy surveys allows to selectively remove from the CMB only the gravitational effect coming from the non-linear structures in the low-redshift universe. 
+For example, delensing with galaxy surveys allows us to selectively remove from the CMB only the gravitational effect coming from the non-linear structures in the low-redshift universe. 
 Removing this component will reduce the level of non-linearities both in the CMB and in the CMB lensing power spectrum providing an alternative route to exploit all the measurable modes without having to model the non-linear component at small angular scales.

### 2017-07-11 Notes/galaxy_delensing.tex
text: conclusions

@@
 
-For ongoing experiments, we find that combining LSS tracers with CMB internal reconstruction will be particularly beneficial.
+For ongoing experiments, we find that LSS tracers will be particularly beneficial.
 An optical survey like DES is able to remove $14\%$ of the B-mode lensing power alone, and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the power using only LSS surveys.
-Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to an improvement of 30\% in the constraint on the tensor to scalar ratio $r$.
-In the future, the decreasing level of instrumental noise of CMB experiments will dramatically improve its internal reconstruction of the structures lensing the CMB. 
-The fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level. 
+Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to an improvement of 30\% in the constraint on the tensor to scalar ratio $r$ compared to the value before delensing.
+In the future, the decreasing level of instrumental noise in CMB experiments will dramatically improve its internal reconstruction of the structures lensing the CMB. 
+The fraction of removed lensing B-modes will rapidly improve from the current expected levels for Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) levels. 
 Already for Stage-3, the CMB internal reconstruction will be the main source of delensing.
-However it will still be less efficient than galaxies at delensing small scales $\ell>500$.
+However, it will still be less efficient than galaxies at tracing the lensing potential at small scales $\ell>500$.
 For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $57\%$ to $68\%$ for 3G.
 Even for CMB S4, galaxy surveys will still play an important role.
-For example, a tomographically binned LSST-survey will remove 56\% of the lensing power.
-While this performance is comparable to a Stage-3 CMB it will allow to probe the robustness against systematics of an eventual detection.
+For example, a tomographically binned LSST-like survey by itself will remove 56\% of the lensing power.
+This performance is lower than a S4 CMB internal reconstruction and comparable to a Stage-3 CMB.
+However it will allow us to probe the robustness against systematics of an eventual detection of primordial gravitational waves.
 Indeed delensing with just CMB data will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) that we are trying to delens to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2017,namikawa:2017}.
-For this reason, efficient galaxies tracers are not only useful in the short term but will also play a role in cross-checking these internal biases in the future.
+For this reason, efficient galaxies tracers are not only useful in the short term but, in the future, will also play a role in testing internal biases and performing consistency tests.
 

### 2017-07-11 Notes/galaxy_delensing.tex
text: conclusions

@@
 This performance is lower than a S4 CMB internal reconstruction and comparable to a Stage-3 CMB.
-However it will allow us to probe the robustness against systematics of an eventual detection of primordial gravitational waves.
-Indeed delensing with just CMB data will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) that we are trying to delens to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2017,namikawa:2017}.
+However, it will allow us to probe the robustness against systematics of an eventual detection of primordial gravitational waves.
+Indeed delensing with just CMB data will require a careful study of possible biases and systematic effects because we delens the same dataset (the CMB) that we also use to reconstruct the lensing potential \cite{carron:2017,sehgal:2017,namikawa:2017}.
 For this reason, efficient galaxies tracers are not only useful in the short term but, in the future, will also play a role in testing internal biases and performing consistency tests.

### 2017-07-25 Notes/galaxies_delensing_uchicago_thesis.tex
keep it up to data
text: conclusions

@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/BB_res_ell2.pdf}
+\includegraphics[scale=1.2,clip]{../images/BB_res_ell2_thesis.pdf}
 \caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/compare_kernel.pdf}
+\includegraphics[scale=1.2,clip]{../images/compare_kernel_thesis.pdf}
 \caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/compare_kernel_tomo.pdf}
+\includegraphics[scale=1.2,clip]{../images/compare_kernel_tomo_thesis.pdf}
 \caption{\textbf{Kernels Comparison with tomographic bins}: Comparison of the 10 LSST  tomographic bins together with the full LSST survey and the CMB lensing kernel. 
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/B_res_bin.pdf}
+\includegraphics[scale=1.2,clip]{../images/B_res_bin_thesis.pdf}
 \caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing.
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/actual_scenario_planck.pdf}
+\includegraphics[scale=1.2,clip]{../images/actual_scenario_planck_thesis.pdf}
 \caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the angular scale $\ell$. 
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/actual_scenario.pdf}
+\includegraphics[scale=1.2,clip]{../images/actual_scenario_thesis.pdf}
 \caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/S3_scenario.pdf}
+\includegraphics[scale=1.2,clip]{../images/S3_scenario_thesis.pdf}
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/S4_scenario.pdf}
+\includegraphics[scale=1.2,clip]{../images/S4_scenario_thesis.pdf}
 \caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/errors_summary.pdf}
+\includegraphics[scale=1.2,clip]{../images/errors_summary_thesis.pdf}
 \caption{\relook{to be filled with final results}}
@@
 \end{figure}
-\twocolumngrid
 

### 2017-07-25 Notes/galaxy_delensing.tex
text: conclusions

@@
 \usepackage{epsfig}
-\usepackage{natbib}
 \usepackage{graphicx,epsfig}
@@
 \usepackage{ifthen}
+\usepackage [english]{babel}
 \usepackage{xstring}
@@
 
-\title{Future Cosmic Microwave Background delensing with galaxy surveys.}       
+\title{Future cosmic microwave background delensing with galaxy surveys.}       
 \author{A. Manzotti} 
@@
 In particular, the gravitational interactions of CMB photons with large scale structures will distort the primordial E-modes adding a lensing B-mode component to the primordial signal.
-Removing the lensing part ('delensing') from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
+Removing the lensing part (``delensing'') from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
 Here we discuss the role of current and future large scale structure surveys in a multi-tracers approach to delensing that will improve the reconstruction of the lensing potential that lenses the CMB photons and, as a consequence, the delensing efficiency.
 We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
-We find that, in general, galaxy surveys should be splitted into tomographic bins as this can improve the correlation with CMB lensing by 30\%.
+We find that, in general, galaxy surveys should be split into tomographic bins as this can improve the correlation with CMB lensing by 30\%.
 Considering currently available survey, a DES-like galaxy survey can remove about $14\%$ of the lensing signal.
 The CMB experiments of the current generation will particularly benefit from large scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
-With the decrease of the instrumental noise, the CMB internal reconstruction will increase its efficiency and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level.
+With the decrease of instrumental noise, the CMB internal reconstruction will increase its efficiency and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level.
 Nevertheless optical galaxy surveys will still play an important role even for CMB S4.
@@
 \label{sec:intro}
-In the standard cosmological paradigm the early universe underwent a period of near-exponential expansion called "cosmic inflation." 
+In the standard cosmological paradigm the early universe underwent a period of near-exponential expansion called ``cosmic inflation''.
 All the cosmological observations agree with this picture, making it a compelling and elegant description of the Universe initial conditions.
@@
 Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski15}.
-This prediction set inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologists are looking for.
+This prediction sets inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologists are looking for.
 These primordial gravitational waves in the early universe would imprint a unique signature on the polarized anisotropies of the CMB.
@@
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
-This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
+This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
 
@@
 Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
-Furthermore, it has recently been performed on CMB temperature data using the cosmic infrared background as LSS tracer \cite{larsen:2016a} and on CMB temperature and polarization data using CMB data to internally reconstruct the LSS lensing potential \cite{carron:2017}.
-Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in which $~28\%$ of the lensing power was removed.
+Furthermore, it has recently been performed on CMB temperature data using the cosmic infrared background (CIB) as LSS tracer \cite{larsen:2016a} and on CMB temperature and polarization data using CMB data to internally reconstruct the LSS lensing potential \cite{carron:2017}.
+Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in which $~28\%$ of the lensing power was removed \cite{manzotti:2017}.
 
@@
 Here $\frac{dN}{dz}$ is the number of galaxies observed by the survey as a function of redshift while $b(z)$ is the galaxy bias that connects the amplitude of galaxy overdensities to the underlying dark matter density.
+We use a linear bias independent of the angular scale, which is a reasonable assumption for the relatively large scales relevant for delensing ($\ell<1000$).
 When computing the auto-spectrum of the galaxy density, a shot noise term needs to be taken into account. 
@@
 \caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
-The red solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed green line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
+The red solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed blue line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
 These are compared with the residual power left after delensing using some of the LSS tracers described in \refsec{model}. 
@@
 \includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{Kernels Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
+\caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
 The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential will be leading to a higher delensing efficiency.
@@
 \includegraphics[scale=1.]{../images/compare_kernel_tomo.pdf}
-\caption{Kernels Comparison with tomographic bins: Comparison of the 10 LSST  tomographic bins together with the full LSST survey and the CMB lensing kernel. 
+\caption{Kernel Comparison with tomographic bins: Comparison of the 10 LSST  tomographic bins together with the full LSST survey and the CMB lensing kernel. 
 Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross-correlation with the CMB lensing.
@@
 
-On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment.
+On the CMB side, we will combine a BICEP-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment.
 For the deep experiment we assume an instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin}, a beam of \textcolor{red}{30 arcmin} and an angular scale range of $50<\ell<500$.
@@
 excluding SKA) the improvement factor increase to 2.3.
-These gives an important opportunity: for sufficiently high value of $r$ it will be possible to confirm a detection of primordial gravity waves using spectra delensed with LSS, completely independent of CMB internal data.
+These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravity waves using spectra delensed with LSS, completely independent of CMB internal data.
 Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of the CMB S3 internal reconstruction.
@@
 \label{sec:concl}
-The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal ("delensing") is necessary to test inflation using the next generation of CMB polarization experiments.
+The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal (``delensing'') is necessary to test inflation using the next generation of CMB polarization experiments.
 To delens, we need to accurately reconstruct the large scale structures that lens the CMB in order to build a template of the expected lensing B-modes in the observed patch.
@@
 \includegraphics[scale=1]{../images/errors_summary.pdf}
-\caption{\relook{to be filled with final results}}
+\caption{Summary of the amount of lensing B-modes power removed by different tracers for different generation of experiments.}
 \label{fig}
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{apsrev4-1}
 \bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}

### 2017-07-25 Code/multiple_survey_delens.py
new images for Uchicago thesis

@@ -100,7 +100,7 @@ def main(labels, cmb):
         ckk_noise = noise_fun(lbins)
 
     if cmb == 'S3':
-        noise = 1.5
+        noise = 2.0
         beam = 2
         noise_phi = np.loadtxt(
             './quicklens_data/min_var_noise_{}muk_{}beam.txt'.format(noise, beam))

### 2017-07-25 Code/plot_kernel_tomo_thesis.py

@@ -1,0 +1,393 @@
+import scipy as sp
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+try:
+    import pyfits
+    import pdb
+
+except:
+    pass
+
+# try:
+#     import palettable
+
+# except Exception as e:
+#     pass
+import os
+import sys
+import time
+import subprocess
+import pyximport
+pyximport.install(reload_support=True)
+import numpy as np
+import kappa_cmb_kernel as kappa_kernel
+import gals_kernel
+# import kappa_gals_kernel
+import hall_CIB_kernel as cib_hall
+import scipy.integrate
+from scipy.interpolate import RectBivariateSpline, interp1d, InterpolatedUnivariateSpline
+import limber_integrals
+import pickle as pk
+import camb
+from camb import model
+import configparser
+from joblib import Parallel, delayed
+# from profiling.sampling import SamplingProfiler
+# profiler = SamplingProfiler()
+import DESI
+from pprint import pprint
+import builtins
+
+from matplotlib import cm
+
+# try:
+#     cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+#     plt.rcParams['image.cmap'] = cmap
+# except Exception as e:
+#     print('probably palettable not found')
+
+
+def find_bins(z, dndz, nbins):
+    '''
+    Function that, given a redshift distribution returns binning with equal number of bins.
+
+    It returns a list of the redshift of the bins.
+    '''
+    cum = np.cumsum(dndz)
+    args = np.hstack((np.array(0.), np.searchsorted(
+        cum, [cum[-1] / nbins * n for n in np.arange(1, nbins + 1)]))).astype(np.int)
+    args[-1] = len(cum) - 1
+    return [z[args[i]:args[i + 1]] for i in np.arange(0, len(args) - 1.)], args
+
+
+def make_tomo_bins(z, dndz, sigmaz, width, nbins, hspline, omegac, h, b=1.):
+    '''
[…]

### 2017-07-25 Code/plots/plot_kernel_thesis.py

@@ -1,0 +1,195 @@
+
+
+# from __future__ import print_function
+import scipy as sp
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+
+import matplotlib.pyplot as plt
+
+try:
+    import pyfits
+    import pdb
+
+except:
+    pass
+
+# try:
+#     import palettable
+
+# except Exception as e:
+#     pass
+import os
+import os
+import sys
+import time
+import subprocess
+
+from pprint import pprint
+import builtins
+import pyximport
+pyximport.install(reload_support=True)
+
+# try:
+#     cmap = palettable.colorbrewer.get_map('RdBu', 'diverging', 11, reverse=True).mpl_colormap
+#     plt.rcParams['image.cmap'] = cmap
+# except Exception as e:
+#     print('probably palettable not found')
+
+
+datadir = '../../Data/limber_spectra/'
+
+image_dir = '../../images/'
+
+
+# Import data
+kappakernel_l30 = np.loadtxt(datadir + 'kappa_kernel_l30.txt')
+kappakernel_l100 = np.loadtxt(datadir + 'kappa_kernel_l100.txt')
+kappakernel_l500 = np.loadtxt(datadir + 'kappa_kernel_l500.txt')
+
+lsstkernel_l100 = np.loadtxt(datadir + 'lsst_kernel_l100.txt')
+
+skakernel_l100 = np.loadtxt(datadir + 'ska_kernel_l100.txt')
+weakkernel_l100 = np.loadtxt(datadir + 'weak_kernel_l100.txt')
+
+
+# cibkernel_l30 = np.loadtxt(datadir + 'cib_kernel_l30.txt')
+cibkernel_l100 = np.loadtxt(datadir + 'cib_kernel_l100.txt')
+# cibkernel_l500 = np.loadtxt(datadir + 'cib_kernel_l500.txt')
+
+
+deskernel_l30 = np.loadtxt(datadir + 'des_kernel_l30.txt')
+deskernel_l100 = np.loadtxt(datadir + 'des_kernel_l100.txt')
+deskernel_l500 = np.loadtxt(datadir + 'des_kernel_l500.txt')
+
+
+desikernel_l30 = np.loadtxt(datadir + 'desi_kernel_l30.txt')
+desikernel_l100 = np.loadtxt(datadir + 'desi_kernel_l100.txt')
+desikernel_l500 = np.loadtxt(datadir + 'desi_kernel_l500.txt')
+
+
+# ============================================
+# SIZE OF THE PICTURE
[…]

### 2017-07-26 Notes/galaxy_delensing.tex
abby comments

@@
 We find that, in general, galaxy surveys should be split into tomographic bins as this can improve the correlation with CMB lensing by 30\%.
-Considering currently available survey, a DES-like galaxy survey can remove about $14\%$ of the lensing signal.
-The CMB experiments of the current generation will particularly benefit from large scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
+Among currently available surveys, a DES-like galaxy survey can remove about $14\%$ of the lensing signal.
+Ongoing CMB experiments (CMB-S2) will particularly benefit from large scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
 With the decrease of instrumental noise, the CMB internal reconstruction will increase its efficiency and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level.
@@
 Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational wave background, improving their measurement is a major objective of current and future CMB experiments.
-Even if inflationary B-modes have not been detected yet, a natural value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
+Even if inflationary B-modes have not been detected yet, the target value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15,abazajian:2016}.
 However, just reducing the level of noise will not be enough to attain this goal.
@@
 Indeed, the contamination from lensing B modes is already at the level of the instrumental noise of current experiments \citep{bk14}.
-Thus, together with the experimental effort to reduce the amount of noise, the effect of the spurious lensing component must be understood.
+Thus, together with the experimental effort to reduce the amount of noise, the effect of the lensing component must be understood.
 
-The optimal way to reduce the lensing contributions to the B-modes is to reconstruct the expected realization of lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
+The optimal way to reduce the lensing contributions to the B-modes is to reconstruct a template of the actual lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
 We can delens the observed B-modes by combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
@@
 The CMB polarization is usually decomposed into odd-parity Fourier modes E and B. 
-As shown in \cite{}, because of the symmetry of the problem, tensor perturbations are the principal source of the B-modes configuration.
+As shown in \cite{seljak:1997}, because of the symmetry of the problem, tensor perturbations are the principal source of the B-modes configuration.
 For this reason, B-modes are a promising signature of early universe tensor perturbations.

### 2017-07-26 Notes/galaxy_delensing.tex
abby comments

@@
 \newcommand{\apjl}{ApJL}
-\def\araa{\ref@jnl{ARA\&A}}
+\newcommand{\araa}{ARA\&A}
 \newcommand{\apjs}{ApJS}
@@
 Despite the experimental effort, other possible explanations are still valid, and a conclusive evidence of inflation is still to be found.
-Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski15}.
+Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski:2016}.
 This prediction sets inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologists are looking for.
@@
 Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational wave background, improving their measurement is a major objective of current and future CMB experiments.
-Even if inflationary B-modes have not been detected yet, the target value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15,abazajian:2016}.
+Even if inflationary B-modes have not been detected yet, the target value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski:2016,abazajian:2016}.
 However, just reducing the level of noise will not be enough to attain this goal.
@@
 CMB 3G (56\%) &2& 1.57  & 1.21 \\ \hline
-LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
+LSS-S3+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
 \end{tabular}
@@
 \includegraphics[scale=1.]{../images/S4_scenario_ska10.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
+\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$.
+This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstrcution and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
+}
 \label{fig:corrs4_10}
@@
 \includegraphics[scale=1.]{../images/S4_scenario_ska01.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
+\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$.
+This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstrcution and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
+}
 \label{fig:corrs4_01}
@@
 excluding SKA) the improvement factor increase to 2.3.
-These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravity waves using spectra delensed with LSS, completely independent of CMB internal data.
+These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSS, completely independent of CMB internal data.
 Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of the CMB S3 internal reconstruction.
@@
 Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-LSST (48\%) &1.88 & 1.81  &  \\ \hline
-LSS  (60\%) &2.47 & 2.31  & \\ \hline
-SKA (10$\mu$Jy) (53\%) &2.18& 2.06  & \\ \hline
-SKA (0.1$\mu$Jy) (77\%) &4.6& 3.96  & \\ \hline
-LSS + SKA (10$\mu$Jy)  (62\%) &2.74 & 2.53  &  \\ \hline
-LSS + SKA (0.1$\mu$Jy)  (80\%) &4.66 & 3.96  &  \\ \hline
-CMB S4 (81\%) & 4.63 & 3.96  &  \\ \hline
+LSST (48\%) &1.84 & 1.76  & 1.74  \\ \hline
+LSS-S4  (60\%) &2.47 & 2.31  & \\ \hline
+SKA (10$\mu$Jy) (53\%) &2.18& 2.06  & 2.02\\ \hline
+SKA (0.1$\mu$Jy) (77\%) &4.6& 3.96  & 3.9\\ \hline
+LSS-S4 + SKA (10$\mu$Jy)  (62\%) &2.74 & 2.53  &  \\ \hline
+LSS-S4 + SKA (0.1$\mu$Jy)  (80\%) &4.66 & 3.96  &  \\ \hline
+CMB S4 (81\%) & 5.16 & 4.32  & 4.2 \\ \hline
 LSS+CMB (86\%)& 5.2 &  &  \\ \hline
@@
 
-\bibliographystyle{apsrev4-1}
+\bibliographystyle{mnrasunsrt}
 \bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}

### 2017-09-28 Code/money_plot.py
add label last plot

@@ -88,7 +88,7 @@ plt.close()
 fg = plt.figure(figsize=fig_dims)
 N = 4
 ind = np.arange(N)    # the x locations for the groups
-ind= np.array([0,2,4,6])
+ind = np.array([0, 2, 4, 6])
 width = 0.55       # the width of the bars: can also be len(x) sequence
 
 # ==========================================
@@ -100,7 +100,8 @@ p2 = plt.barh(ind[0], -14, width, left=-6 - 8, alpha=0.8)
 p2 = plt.barh(ind[0], -27, width, left=-6 - 8 - 14, alpha=0.8)
 
 p1 = plt.barh(ind[0] + 1.5 * width, 36, width, color='#7f7f7f', alpha=0.8)
-p2 = plt.barh(ind[0] + 1.5 * width, 41 - 36, width, color='#bcbd22', left=36, alpha=0.8)
+p2 = plt.barh(ind[0] + 1.5 * width, 41 - 36, width,
+              color='#bcbd22', left=36, alpha=0.8)
 
 # ==========================================
 # ==========================================
@@ -113,8 +114,10 @@ p2 = plt.barh(ind[1], -35., width * 1.1, left=-6., alpha=0.8)
 p2 = plt.barh(ind[1], -14, width * 1.1, left=-6 - 35, alpha=0.8)
 p2 = plt.barh(ind[1], -27, width * 1.1, left=-6 - 35 - 14, alpha=0.8)
 
-p1 = plt.barh(ind[1] + 1.5 * width, 36, width * 0.9, color='#7f7f7f', alpha=0.8)
-p2 = plt.barh(ind[1] + 1.5 * width, 56 - 36, width * 0.9, color='#bcbd22', left=36, alpha=0.8)
+p1 = plt.barh(ind[1] + 1.5 * width, 36, width *
+              0.9, color='#7f7f7f', alpha=0.8)
+p2 = plt.barh(ind[1] + 1.5 * width, 56 - 36, width *
+              0.9, color='#bcbd22', left=36, alpha=0.8)
 # ==========================================
 
 
@@ -125,8 +128,10 @@ p1 = plt.barh(ind[2], -10., width * 1.1, alpha=0.8)
 p2 = plt.barh(ind[2], -40., width * 1.1, left=-10., alpha=0.8)
 p2 = plt.barh(ind[2], -56., width * 1.1, left=-10. - 40., alpha=0.8)
 
[…]

### 2017-09-28 Code/run_analysis.py

@@ -60,7 +60,7 @@ def clbb(r=0.1, nt=None, lmax=3000):
     inflation_params = initialpower.InitialPowerParams()
     if nt is None:
         nt = -r / 8.
-    inflation_params.set_params(As=2.1e-9, r=r, nt=nt)
+    inflation_params.set_params(As=2.1e-9, r=r, nt=nt, pivot_tensor=0.01)
     results.power_spectra_from_transfer(inflation_params)
     return results.get_total_cls(lmax)[:, 2] * 7.42835025e12
 
@@ -70,7 +70,7 @@ def clbb_tens(r=0.1, nt=None, lmax=3000):
     inflation_params = initialpower.InitialPowerParams()
     if nt is None:
         nt = -r / 8.
-    inflation_params.set_params(As=2.1e-9, r=r, nt=nt)
+    inflation_params.set_params(As=2.1e-9, r=r, nt=nt, pivot_tensor=0.01)
     results.power_spectra_from_transfer(inflation_params)
     return results.get_tensor_cls(lmax)[:, 2] * 7.42835025e12
 
@@ -249,6 +249,7 @@ print(Fore.YELLOW + 'r=0, no noise')
 print('')
 for i, label in enumerate(rho_names):
     probe = label.split('.txt')[0].split('rho_')[1]
+    # delensed
     sigma_r, sigma_nt, sigr, sigmant = fisher_r_nt(
         r_fid=r_fid,
         lmin=50,
@@ -258,11 +259,13 @@ for i, label in enumerate(rho_names):
                                  ) + clbb_tens(r_fid, lmax=lmax),
         noise_uK_arcmin=0.,
         fwhm_arcmin=fwhm_arcmin)
+
     sigma_r_1, sigma_nt_1, sigr_1, sigmant_1 = fisher_r_nt(
         r_fid=r_fid,
         lmin=50,
         lmax=lmax,
         fsky=fsky,
+        # clbb also contains r tensor contributions
         clbb_cov=clbb(r_fid, lmax=lmax),
         noise_uK_arcmin=0.,
         fwhm_arcmin=fwhm_arcmin)
@@ -339,6 +342,8 @@ for i, label in enumerate(rho_names):
 
[…]

### 2017-09-28 Notes/galaxy_delensing.tex

@@
 \date{\today}
+
 \begin{abstract}
@@
 \caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
-The red solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed blue line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
+The orange solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed blue line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
 These are compared with the residual power left after delensing using some of the LSS tracers described in \refsec{model}. 
@@
 \caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
-The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential will be leading to a higher delensing efficiency.
+The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency.
 }
@@
 \includegraphics[scale=1.]{../images/compare_kernel_tomo.pdf}
-\caption{Kernel Comparison with tomographic bins: Comparison of the 10 LSST  tomographic bins together with the full LSST survey and the CMB lensing kernel. 
+\caption{Comparison of kernels of the 10 LSST tomographic bins together with the full LSST survey and the CMB lensing kernel. 
 Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross-correlation with the CMB lensing.
@@
 \caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing.
-Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of angular scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line).}
+Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of angular scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line).
+For details about the binning, see \refsec{tomo-bin} and in particular \refeq{zbin}.
+}
 \label{fig:rho-tomo}
@@
 We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the ability to internally reconstruct the potential $\kappa$, together with a low-noise, low-resolution experiment that measured the B-modes that will be delensed and used to constrain the inflationary parameters.
-We focus on three distinct scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
+Even if there is not a definite distinction between different stages, we focus on three distinct scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
 
@@
 Given a fiducial value for $r$, the value of $n_{t}$ is fixed imposing the consistency relation $n_t=-r/8$ \cite{LIDDLE1992391}.
+In order to minimize the degeneracy between $r$ and $n_{t}$ \cite{abazajian:2016}, we choose as a pivot scale $k_{t}=0.01 $ Mpc$^{-1}$ .
 
@@
 Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} and internal CMB lensing maps \cite{carron:2017} as large scale structure tracers.
-Here we discuss the improvement that can be obtained by combining these and other currently available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
+Here we discuss the improvement that can be obtained by combining these and other currently Stage-2 available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
 
 On the CMB side, we will combine a BICEP-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment.
-For the deep experiment we assume an instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin}, a beam of \textcolor{red}{30 arcmin} and an angular scale range of $50<\ell<500$.
+For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin, a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
 We assume the CMB lensing reconstruction is  performed by the higher resolution CMB experiment. We explore two possibilities.
-First, we test the efficiency with an internal reconstruction of CMB lensing performed by the Planck collaboration \cite{planck-collaboration:2016a}.
+First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck collaboration \cite{planck-collaboration:2016a}.
 We compute the noise in the CMB lensing map in \refeq{noisekappa} using the noise curves publicly available
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
-Then, we test a second scenario with a ground-based experiment with noise levels consistent with SPT-Pol: \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
-In this case we set the largest achievable scale at $\ell_{min}=300$.
-However, the results are robust against this choice since the neglected scales contribute only moderately to the reconstruction of the CMB lensing potential.
+Then, we test a second scenario with a ground-based experiment characterized by noise levels consistent with SPT-Pol \cite{henning:2017}: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin.
+In this case we set the angular scale range at $50<\ell<4000$.
+However, the results are quite robust against this choice.
+%since the neglected scales contribute only moderately to the reconstruction of the CMB lensing potential.
 
@@
 \includegraphics[scale=1.]{../images/actual_scenario_planck.pdf}
-\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$ for current generation tracers.
-Here we show both current galaxy survey and internally reconstructed $\phi$ CMB lensing potential. 
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers.
+Here we show both current galaxy survey and internally reconstructed CMB lensing potential. 
 The dashed curve highlights which scales contribute to the $\ell<100$ B-mode power and are thus most useful to delens.
@@
 \includegraphics[scale=1.]{../images/actual_scenario.pdf}
-\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$ for current generation tracers. 
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers. 
 This figure is the same as \reffig{corrnow-planck} but using SPTPol instead of Planck as the high-res CMB experiment performing the internal reconstruction.
@@
 Using these correlation levels we can  compute the residual B-mode power after delensing using \refeq{Bres}, and test the consequent improvement on parameter constraints with \refeq{fisher}. 
-For current surveys we only consider $p=r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
+For current surveys we only report the improvements in $r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
 We still let $n_{t}$ to vary in our fisher calculation even if it is not very degenerate with $r$.
@@
 As shown in previous works \cite{sherwin:2015}, the best tracer is the CIB with a 27\% reduction in power.
-Nonetheless, it is still worthwhile considering all the tracers.
+Nonetheless, it is worthwhile combining all the tracers.
 Indeed a multi tracer-approach can bring the removed power on the overlapping area from 27\% of the CIB alone to 36\% using all the LSS (LSS-S2) and 41\% when Planck lensing reconstruction is added to the set.
@@
 \begin{table}
-\caption{$\alpha(r)$: Improvements on $\sigma(r)$ due to delensing for current generation and Planck lensing reconstruction.
+\caption{$\alpha(r)$: Improvements on $\sigma(r)$ due to delensing for completed and ongoing Stage-2 surveys and Planck lensing reconstruction.
 LSS-S2 corresponds to the combination of all the available LSS tracers.
 The values in parenthesis in the first column correspond to the fraction of lensing B-mode power removed using each LSS tracer. 
-The values in the other columns correspond to ratio of the error before and after delensing.}
+The values in the other columns correspond to ratio of the error before and after delensing for 3 cases: no instrumental noise in the B-mode measurement and realistic noise for two different values of $r$.}
   \vspace{0.2cm}
@@
  \begin{table}
-\caption{$\alpha(r)$: Current generation improvements on $\sigma(r)$ due to delensing.
-We use the same LSS tracers as \reftab{current} but with the internal lensing reconstruction performed by SPTPol.}
+\caption{$\alpha(r)$: Stage-2 improvements on $\sigma(r)$ due to delensing. We use the same LSS tracers as \reftab{current} but with the internal lensing reconstruction performed by SPTPol.}
   \vspace{0.2cm}
@@
 As we did for Stage-2, we model the CMB S3 as two different overlapping experiments that will be able to combine their measurement.
-The two experiments combined will have a level of noise of \textcolor{red}{$3\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam.
-This level of noise is also assumed for the internal noise reconstruction.
+The two experiments combined will have a level of noise of $3\mu$K-arcmin and a 1 arcmin beam. This level of noise is also assumed for the internal noise reconstruction.
 With this experimental setup, we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
@@
 \includegraphics[scale=1.]{../images/S3_scenario.pdf}
-\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$.
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
 This figure is the same as \reffig{corrnow} but with a stage-3 CMB experiment performing the internal reconstrcution and with DESI added to the stage-2 galaxy surveys (LSS-S2).
@@
 
-\subsection{CMB-S4 Era}
+\subsection{CMB-S4 Era}\label{sec:S4}
 Having as a major goal the detection of inflationary B-modes, CMB data will continue improving even after Stage-3.
@@
 Given the unprecedented low level of noise of these experiment, delensing will be even more important than previous generations.
-Following \cite{abazajian:2016}, here we assume a CMB-S4 ground experiment with a level of noise of \textcolor{red}{$1\mu$K-arcmin} and a \textcolor{red}{1 arcmin} beam.
+Following \cite{abazajian:2016}, here we assume a CMB-S4 ground experiment with a level of noise of $1\mu$K-arcmin and a 1 arcmin beam.
 Furthermore, by the mid-2020s, several next-generation LSS surveys will be online. 
@@
 \includegraphics[scale=1.]{../images/S4_scenario_ska10.pdf}
-\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$.
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
 This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstrcution and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
@@
 \includegraphics[scale=1.]{../images/S4_scenario_ska01.pdf}
-\caption{Correlation factor $\rho$ with the lensing potential $\kappa$ as a function of the angular scale $\ell$.
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
 This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstrcution and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
@@
 \begin{table}
-\caption{$\alpha$: Gen 4 experiments}
+\caption{$\alpha$: Improvements on $\sigma(r)$ due to delensing for S4 experiments.
+Here we define LSS-S4 as the combination of S3 LSS tracers (LSS-S3) and LSST.
+We also consider SKA-like radio-continuum surveys.
+We also use a CMB S4 experiment for the lensing internal reconstruction.
+See \refsec{S4} for details.}
 \begin{tabular}{|c | c | c | c|}
@@
 Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-LSST (48\%) &1.84 & 1.76  & 1.74  \\ \hline
-LSS-S4  (60\%) &2.47 & 2.31  & \\ \hline
-SKA (10$\mu$Jy) (53\%) &2.18& 2.06  & 2.02\\ \hline
-SKA (0.1$\mu$Jy) (77\%) &4.6& 3.96  & 3.9\\ \hline
-LSS-S4 + SKA (10$\mu$Jy)  (62\%) &2.74 & 2.53  &  \\ \hline
-LSS-S4 + SKA (0.1$\mu$Jy)  (80\%) &4.66 & 3.96  &  \\ \hline
-CMB S4 (81\%) & 5.16 & 4.32  & 4.2 \\ \hline
+LSST (48\%) &1.8 & 1.7  & 1.7  \\ \hline
+LSS-S4  (60\%) &2.5 & 2.3  & 2.3 \\ \hline
+SKA (10$\mu$Jy) (53\%) &2.1& 2  & 2\\ \hline
+SKA (0.1$\mu$Jy) (77\%) &4.6& 3.9  & 3.9\\ \hline
+LSS-S4 + SKA (10$\mu$Jy)  (65\%) &2.7 & 2.5  & 2.5  \\ \hline
+LSS-S4 + SKA (0.1$\mu$Jy)  (80\%) &4.6 & 3.9  & 3.9 \\ \hline
+CMB S4 (81\%) & 5.1 & 4.3  & 4.2 \\ \hline
 LSS+CMB (86\%)& 5.2 &  &  \\ \hline
@@
 An optical survey like DES is able to remove $14\%$ of the B-mode lensing power alone, and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the power using only LSS surveys.
-Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to an improvement of 30\% in the constraint on the tensor to scalar ratio $r$ compared to the value before delensing.
-In the future, the decreasing level of instrumental noise in CMB experiments will dramatically improve its internal reconstruction of the structures lensing the CMB. 
+Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to a maximum improvement of 30\% in the constraint on the tensor to scalar ratio $r$ compared to the value before delensing.
+In the future, the decreasing level of instrumental noise in CMB experiments will dramatically improve the internal reconstruction of the structures lensing the CMB. 
 The fraction of removed lensing B-modes will rapidly improve from the current expected levels for Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) levels. 
-Already for Stage-3, the CMB internal reconstruction will be the main source of delensing.
+For Stage-3 experiments, the CMB internal reconstruction will be the main source of delensing.
 However, it will still be less efficient than galaxies at tracing the lensing potential at small scales $\ell>500$.
@@
 \section{Acknowledgments}
-We thank S. Dodelson, S. Passaglia, K. Story and K. Wu for useful discussion and their review of the draft. 
+We thank S. Dodelson, S. Passaglia, K. Story, A. Vieregg and K. Wu for useful discussion and their review of the draft. 
 Furthermore we thank D. Alonso for his feedback on radio continuum surveys.

### 2017-10-03 Notes/galaxy_delensing.tex
results and conclusion text

@@
 We will use the Fisher information matrix to quantify the delensing efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
-We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the ability to internally reconstruct the potential $\kappa$, together with a low-noise, low-resolution experiment that measured the B-modes that will be delensed and used to constrain the inflationary parameters.
+We mainly focus on $r$ since these two parameters show similar improvement when delensing is moderate (does not remove more than $\sim 80\%$) and since, at least for single field inflation, only $r$ would be detectable by future experiments.
+
+We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the ability to internally reconstruct the lensing potential $\kappa$, together with a low-noise, low-resolution experiment that measures the B-modes that will be delensed and used to constrain the inflationary parameters.
+As in \cite{abazajian:2016} and other works in the literature, we do not combine the observations from the two experiments to obtain a single CMB map with optimally low noise.
+Depending on the experiments, this might be possible and, once used for both lensing reconstruction and B-mode measurements, it can improve the level of delensing and the cosmological constraints.
+
 Even if there is not a definite distinction between different stages, we focus on three distinct scenarios: the current stage, a third generation stage (3G) and finally the futuristic CMB Stage 4.
@@
 \subsection{Fisher Information Matrix}\label{sec:fisher}
-In the Fisher Information Matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtained from the inverse of the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
+In the Fisher information matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtained from the inverse of the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
 We constrain the inflationary parameters $p =\{r,n_{t}\}$ with a CMB B-mode spectrum measurement so the Fisher matrix is:
@@
 Since the lensing B-modes power $C_{{l}}^{BB,\mathrm{lens}}$ is a substantial component of the covariance, removing part of it through delensing will reduce the parameter statistical error.
-We define the error after delensing as $\sigma^{\rm{del}}(p)$.
+In the following, we sometimes refer to the fraction of lensing B-mode power removed by delensing.
+This quantity is defined as: $\left< C_{\ell}^{BB,\mathrm{lens}}/C_{{l}}^{BB,\mathrm{res}}\right>_{4<\ell<100}.$
+
+To quantify the improvement after delensing we introduce the error on a parameter $p$ constrained with \textit{delensed} spectra as $\sigma^{\rm{del}}(p)$.
 This is computed from \refeq{fisher} substituting $C_{{l}}^{BB,\mathrm{lens}}$ with the residual power after delensing $C_{{l}}^{BB,\mathrm{res}}$ in the covariance matrix defined in \refeq{cova}.
-We define the improvement as the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$.
-The fiducial values used are described in \refsec{del_res}.
+The improvement is then the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$.
+
+\refeq{fisher} requires fiducial values for the parameters $p =\{r,n_{t}\}$. We explore different value for $r$ in \refsec{del_res}.
 Given a fiducial value for $r$, the value of $n_{t}$ is fixed imposing the consistency relation $n_t=-r/8$ \cite{LIDDLE1992391}.
+Furthermore, these results depend mildly on the choice of the pivot scale for the tensor and scalar primordial perturbation spectra.
 In order to minimize the degeneracy between $r$ and $n_{t}$ \cite{abazajian:2016}, we choose as a pivot scale $k_{t}=0.01 $ Mpc$^{-1}$ .
@@
 We assume the CMB lensing reconstruction is  performed by the higher resolution CMB experiment. We explore two possibilities.
-First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck collaboration \cite{planck-collaboration:2016a}.
-We compute the noise in the CMB lensing map in \refeq{noisekappa} using the noise curves publicly available
+First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck satellite \cite{planck-collaboration:2016a}.
+For the noise in the CMB lensing map in \refeq{noisekappa} we use the actual noise curves publicly available
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
@@
 
-For Planck, the internal reconstruction is at most $70\%$ correlated at very large scales and then it falls rapidly to $40\%$ at $\ell=200$.
+For Planck, the internal reconstruction is at most $\sim 65\%$ correlated at very large scales and then it falls rapidly to $40\%$ at $\ell=200$.
 Its correlation is comparable to the one of LSS at almost all scales.
-On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 80\% correlated with the true field at $\ell<300$, and only then the correlation becomes comparable to the current LSS surveys.
+On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 85\% correlated with the true field at $\ell<300$, and only then the correlation goes below the level of current LSS surveys.
 In both cases the CMB lensing reconstruction correlates very well at low $\ell$ and it then falls rapidly at smaller scales because of the raise of the reconstruction noise.
@@
 \reffig{corrnow} and \reffig{corrnow-planck} show that the DES galaxies are effective tracers of the LSS and can, at least in the near future, be used to improve delensing for CMB experiments that overlap with it.
-For example, DES delensing efficiency is higher than WISE first because of the lower level of noise and also because DES galaxies are located at slightly higher redshift so they better overlap with CMB lensing.
+For example, DES delensing efficiency is higher than WISE first because of the lower level of noise and secondly because DES galaxies are located at slightly higher redshift so they better overlap with CMB lensing.
 As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS.
@@
 Using these correlation levels we can  compute the residual B-mode power after delensing using \refeq{Bres}, and test the consequent improvement on parameter constraints with \refeq{fisher}. 
-For current surveys we only report the improvements in $r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
-We still let $n_{t}$ to vary in our fisher calculation even if it is not very degenerate with $r$.
+We only report the improvements in $r$ but we still let $n_{t}$ to vary in our fisher calculation even if it is not very degenerate with $r$.
 We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
@@
 
-The lensing reconstruction noise in Planck is such that internal delensing can only remove $8\%$ of the power, improving the constraints on $r$ by a factor of 1.06 even in the ideal scenario of no instrumental noise in the B-modes.
+The lensing reconstruction noise in Planck is such that internal delensing can only remove \textcolor{blue}{$8\%$} of the power, improving the constraints on $r$ by a factor of \textcolor{blue}{1.06} even in the ideal scenario of no instrumental noise in the B-modes.
 A similar performance is achieved with WISE.
-DES can do better, removing a level of 14\% of power.
+DES can do better, removing a level of \textcolor{blue}{11\%} of power.
 However, it covers a smaller fraction of the sky than the two previous probes.
-As shown in previous works \cite{sherwin:2015}, the best tracer is the CIB with a 27\% reduction in power.
+As shown in previous works \cite{sherwin:2015}, the best tracer is the CIB with a \textcolor{blue}{27\%} reduction in power.
 Nonetheless, it is worthwhile combining all the tracers.
-Indeed a multi tracer-approach can bring the removed power on the overlapping area from 27\% of the CIB alone to 36\% using all the LSS (LSS-S2) and 41\% when Planck lensing reconstruction is added to the set.
+Indeed a multi tracer-approach can bring the removed power on the overlapping area from \textcolor{blue}{27\%} of the CIB alone to \textcolor{blue}{36\%} using all the LSS (LSS-S2) and \textcolor{blue}{41\%} when Planck lensing reconstruction is added to the set.
 However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 30\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
@@
 If, on the other hand, we consider a high-resolution ground base experiment the internal lensing reconstruction improves significantly. 
-\reftab{current-spt} shows that, with an instrumental noise at the expected level for SPTPol, the CMB will soon be able to remove an amount of power (35\%) at the level of the CIB.
-Once combined with LSS tracer this will lead to removing 56\% of the power.
-The improvement of 1.57 in $\sigma(r)$ that will follow, even when realistic noise is considered, is such that it will make delensing a  needed step for CMB polarization data analysis.
+\reftab{current-spt} shows that, with an instrumental noise at the expected level for SPTPol, the CMB will soon be able to remove an amount of power \textcolor{blue}{35\%} at the level of the CIB.
+Once combined with LSS tracer this will lead to removing \textcolor{blue}{56\%} of the power.
+The improvement of \textcolor{blue}{1.57} in $\sigma(r)$ that will follow, even when realistic noise is considered, is such that it will make delensing a needed step for CMB polarization data analysis.
 
+%===========================
+%===========================
 \begin{table}
@@
 The values in parenthesis in the first column correspond to the fraction of lensing B-mode power removed using each LSS tracer. 
-The values in the other columns correspond to ratio of the error before and after delensing for 3 cases: no instrumental noise in the B-mode measurement and realistic noise for two different values of $r$.}
+It is defined as $\left< \frac{C_{\ell}^{BB,\mathrm{lens}}}{C_{{l}}^{BB,\mathrm{res}}}\right>_{4<\ell<100}.$
+\newline
+The values in the other columns correspond to ratio of the error before and after delensing for 3 cases: no instrumental noise in the B-mode measurement (but with instrumental noise in the E-mode used to delens) and realistic noise for two different values of $r$.}
   \vspace{0.2cm}
@@
 Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-WISE (6\%)& 1.09& 1.05 & 1.01 \\ \hline
-DES (14\%) &1.2 & 1.10  & 1.03 \\ \hline
-CIB (27\%) &1.45& 1.21  & 1.07\\ \hline
-LSS-S2 (36\%) &1.7& 1.29  & 1.09\\ \hline
-CMB Planck (8\%) &1.06& 1.05  & 1.01 \\ \hline
-LSS-S2 +CMB (41\%)& 1.77& 1.31& 1.10  \\ \hline
+WISE (6\%)& 1.09& 1.06 & 1.02 \\ \hline
+DES (11\%) &1.16 & 1.10  & 1.04 \\ \hline
+CIB (25\%) &1.37& 1.21  & 1.09\\ \hline
+LSS-S2 (34\%) &1.53& 1.29  & 1.11\\ \hline
+CMB Planck (8\%) &1.08& 1.05  & 1.02 \\ \hline
+LSS-S2 +CMB (37\%)& 1.58& 1.33& 1.13  \\ \hline
 \end{tabular}
@@
 \end{table}
+%===========================
+%===========================
 
@@
 SPTPol (35\%) &1.46& 1.27  & 1.08 \\ \hline
-LSS-S2 +SPTPol (56\%) & 2.18 & 1.57 & 1.16  \\ \hline
+LSS-S2 +SPTPol (58\%) & 2.18 & 1.57 & 1.16  \\ \hline
 
@@
 The two experiments combined will have a level of noise of $3\mu$K-arcmin and a 1 arcmin beam. This level of noise is also assumed for the internal noise reconstruction.
-With this experimental setup, we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
-Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the LSS tracers used in \refsec{del_res}.
+We still use an  angular scale range in the Fisher matrix of $50<\ell<4000$ even if most of the high $\ell$ scale do not contribute to the constraints.
+Furthermore, around 2019, DESI will start taking data. For this reason, we add DESI to the LSS tracers used in \refsec{del_res}.
 
 The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
-An interesting finding is that DESI will be less efficient (removing 10\% of the power) than a DES-like survey (14\%) despite the fact that it can probe slightly higher redshift.
-The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency significantly.
-Unfortunately, adding DESI will only bring the power removed using LSS from 36 to 41\%.
+An interesting finding is that DESI will be less efficient (removing \textcolor{blue}{9\%} of the power) than a DES-like survey (\textcolor{blue}{11\%}) despite the fact that it can probe slightly higher redshift.
+The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency.
+Unfortunately, adding DESI will only bring the power removed using LSS from \textcolor{blue}{36 to 41\%}.
 Indeed, even in the near future, CIB will still play the dominant role among LSS.
 Contrary to current experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 550$.
-
-%This generation of experiments will start exploring the small scales B-modes thus constraining, even if loosely, the value of $n_{t}$.
 
@@
 We summarize the improvement in the constraint on $r$ in \reftab{S3}. 
-Given the improvement in the noise of the high-resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
+Given the improvement in the noise of the high-resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of \textcolor{blue}{2} through internal delensing in the ideal case of no instrumental noise in the B-modes \textcolor{red}{(but with instrumental noise in the E-mode used to delens)}.
 Adding galaxy surveys will lead to a further improvement without additional effort.
-Indeed galaxy surveys will still be able to remove an additional $20\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by $15\%$.
+Indeed galaxy surveys will still be able to remove an additional \textcolor{blue}{$20\%$} of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by \textcolor{blue}{$15\%$}.
 Note that apart from the addition of DESI, the tracers used here are already available today.
 The ongoing effort can then lead to a significant improvement even for the next generation of experiments.
-Finally, the combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor 1.8 or to improve the constraint of a possible detection (r=0.12) by 30\%.
+Finally, the combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor \textcolor{blue}{1.8} or to improve the constraint of a possible detection (r=0.12) by \textcolor{blue}{30\%}.
 As expected the improvement on $n_{t}$ is similar to the one in $r$ since they are both proportional to the variance of the measured B-modes.
@@
 Moreover, satellite and balloon CMB experiments have been proposed and have the potential to extend the accessible B-mode measurements to the largest scales.
-Given the unprecedented low level of noise of these experiment, delensing will be even more important than previous generations.
+Given the unprecedented low level of noise of these experiment, delensing will be even more important than in previous generations.
 Following \cite{abazajian:2016}, here we assume a CMB-S4 ground experiment with a level of noise of $1\mu$K-arcmin and a 1 arcmin beam.
-Furthermore, by the mid-2020s, several next-generation LSS surveys will be online. 
+Furthermore, in the next decade, several next-generation LSS surveys will be online. 
+Even if these LSS surveys will observe an overlapping part of the sky and similar modes, because of their different strengths and weaknesses, it will still be important to combine them efficiently.
 Here, as an example of future optical galaxy surveys, we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers. 
 We do not consider other surveys but we test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance. 
-Even if these LSS surveys will observe an overlapping part of the sky and similar modes, because of their different strengths and weaknesses, it will still be important to combine them efficiently.
 We also consider a radio continuum survey modeled following the SKA specifications.
@@
 
-The correlation factor with CMB lensing for Stage-4 experiments is shown in \reffig{corrs4_10}  and in \reffig{corrs4_01} for SKA1 and SKA2 respectively.
+The correlation factor with CMB lensing for Stage-4 experiments is shown in \reffig{corrs4_10}.
 %=======
@@
 \end{figure}
-%=======
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/S4_scenario_ska01.pdf}
-\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
+%%=======
+%\begin{figure}[htbp]
+%\begin{center}
+%\includegraphics[scale=1.]{../images/S4_scenario_ska01.pdf}
+%\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
 This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstrcution and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
 }
-\label{fig:corrs4_01}
-\end{center}
-\end{figure}
-%=======
+%\label{fig:corrs4_01}
+%\end{center}
+%\end{figure}
+%%=======
 A tomographically binned LSST-like experiment will be a very efficient CMB lensing tracer.
@@
 At this level, CMB S4 internal delensing will not be limited by noise but by small secondary effects like foregrounds contamination, filtered modes etc.
-Finally, \reffig{corrs4_01} shows that a very futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
+Finally, \reftab{S4} shows that a very futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
 This performance is comparable to a CMB-S4 internal reconstruction.
@@
 \reftab{S4} shows the improvement on the inflationary constraint given the cross-correlation factors described above. 
-LSST alone will be able to remove almost half of the B-mode lensing power.
-This will lead to an improvement in $\sigma(r)$ of a factor of 1.8.
+LSST alone will be able to remove almost \textcolor{blue}{half} of the B-mode lensing power.
+This will lead to an improvement in $\sigma(r)$ of a factor of \textcolor{blue}{1.8}.
 If we add all the other infrared and optical LSS tracers previously included (i.e.
-excluding SKA) the improvement factor increase to 2.3.
+excluding SKA) the improvement factor increase to \textcolor{blue}{2.3}.
 These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSS, completely independent of CMB internal data.
-Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of the CMB S3 internal reconstruction.
+Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove \textcolor{blue}{53\%} of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS \textcolor{blue}{(62\%)} to surpass the level of the CMB S3 internal reconstruction.
 Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power.
 This will bring an almost factor of 4 improvement in the null test case even with realistic noise.
-Combining with LSS will only increase the removed power compared to CMB only by $\simeq 5\%$ but it will be highly beneficial to test the robustness of the final result.
+Combining with LSS will only increase the removed power compared to CMB only by \textcolor{blue}{$\simeq 5\%$} but it will be highly beneficial to test the robustness of the final result.
 
@@
 This is reasonable because, given the level of noise in CMB-S4, the variance at angular scales $\ell<200$ are dominated by instrumental noise and not by the lensing component.
-Independent of the delensing level, the small value allowed for $r$ will not allow to test the consistency relations.
+Independent of the delensing level, given the small value allowed for $r$ will not allow to test the consistency relations.
 
@@
 %=======================================
-%\begin{table}
-%\caption{$\alpha$: Gen 4 experiments}
-%  \begin{tabular}{|c | c | c | c|}
-%\hline
-%Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-%LSST (10\%) &1.12 & 1.08  & 1.03 \\ \hline
-%SKA (0.1$\mu$Jy) (40\%) &1.73& 1.42  & 1.2\\ \hline
-%CMB S4 (56\%) &2& 1.57  & 1.21 \\ \hline
-%LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
-%\end{tabular}
-%\label{tab:S4_2}
-%\end{table}
-%=======================================
-%\subsection{Bias uncertainties degradation}
-%The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of the inflationary constraint of delensed spectra.
-%In this section, we quantify this effect.
-%We will now marginalize over unknown galaxies parameters, but we will use a full dataset of CMB and galaxies data.
-The idea is that as shown in the low level of noise in galaxy surveys might allow us to calibrate them internally.
-%We will use a Fisher approach the Fisher matrix is:
-%\beqn
-%F_{pq} &=& \sum_{l_{a} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}  \sum_{l_{b} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}   \frac{\partial {C}_{l_a}^{BB,\mathrm{del}}}{\partial \theta_p } \left[ \mathrm{Cov}^{BB,BB}\right]_{l_a, l_b} ^{-1} \frac{\partial {C}_{l_b}^{BB,\mathrm{del}}}{\partial \theta_q }\nonumber \\&+&  \sum_j \frac{\frac{\partial {C}_j^{\kappa I}}{\partial \theta_p } \frac{\partial {C}_j^{\kappa I}}{\partial \theta_q }}{(\Delta C_j^{\kappa I})^2}
-%+\sum_j \frac{\frac{\partial {C}_j^{I I}}{\partial \theta_p } \frac{\partial {C}_j^{I I}}{\partial \theta_q }}{(\Delta C_j^{I  I})^2}
-%\eeqn
-%
-%\beq
-%\alpha_{\mathrm{marginalized}} = \sigma_0(r) / \sigma_{\mathrm{marginalized/delensed}}(r)
-%\eeq
-%
-%
-%where the parameter array contains both the tensor to scalar rate $\theta=r$, and the galaxy surveys parameters like the bias $b_i$ or $p_i$ \footnote{In our analysis, since the fiducial value of $r$ is zero,
-%the derivative of $C_\ell^{\rm BB,res}$ is non-zero if $\theta_i=r$.}.
-%
-%We compute the derivatives of the power spectra as described
+
 \section{Conclusions}
@@
 In this paper, we studied the potential impact of large-scale structure galaxy surveys in this important endeavor. 
-%For current experiment, the cosmic infrared background had already proven to be very efficient in delensing \cite{larsen:2016,manzotti:2017}.
+We find that, to improve the delensing efficiency of galaxy surveys, tomographic delensing will be very important: using several tomographic galaxy bins can improve the correlation of galaxies with CMB lensing by 10\%-30\% on a wide range of angular scales.
 
 For ongoing experiments, we find that LSS tracers will be particularly beneficial. 
-An optical survey like DES is able to remove $14\%$ of the B-mode lensing power alone, and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the power using only LSS surveys.
-Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to a maximum improvement of 30\% in the constraint on the tensor to scalar ratio $r$ compared to the value before delensing.
+An optical survey like DES is able to remove \textcolor{blue}{$14\%$} of the B-mode lensing power alone, and together with WISE(\textcolor{blue}{$8\%$}) and the CIB (\textcolor{blue}{$27\%$}) will allow to remove $36\%$ of the power using only LSS surveys.
+Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to a maximum improvement of \textcolor{blue}{30\%} in the constraint on the tensor to scalar ratio $r$ compared to the value before delensing.
 In the future, the decreasing level of instrumental noise in CMB experiments will dramatically improve the internal reconstruction of the structures lensing the CMB. 
-The fraction of removed lensing B-modes will rapidly improve from the current expected levels for Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) levels. 
-For Stage-3 experiments, the CMB internal reconstruction will be the main source of delensing.
+The fraction of removed lensing B-modes will rapidly improve from the current expected levels for Planck (\textcolor{blue}{$8\%$}) and SPTPol (\textcolor{blue}{$35\%$}) to 3G (\textcolor{blue}{56\%}) and CMB S4 (\textcolor{blue}{85\%}) levels.
+Indeed for Stage-3 experiments, the CMB internal reconstruction will be the main source of delensing.
 However, it will still be less efficient than galaxies at tracing the lensing potential at small scales $\ell>500$.
-For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $57\%$ to $68\%$ for 3G.
+For this reason, combining galaxy survey with the CMB will push the fraction of removed power from \textcolor{blue}{$57\%$} to \textcolor{blue}{$68\%$} for 3G.
 Even for CMB S4, galaxy surveys will still play an important role.
-For example, a tomographically binned LSST-like survey by itself will remove 56\% of the lensing power.
+For example, a tomographically binned LSST-like survey by itself will remove \textcolor{blue}{56\%} of the lensing power.
 This performance is lower than a S4 CMB internal reconstruction and comparable to a Stage-3 CMB.

### 2017-10-04 Code/money_plot.py
label wrong summary plit

@@ -199,7 +199,7 @@ plt.text(-44, 4 - 0.2, 'LSS-S3', horizontalalignment='left',
 plt.text(-6.9, 4 + 0.4, 'DESI', horizontalalignment='left',
          verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', rotation=90.)
 
-plt.text(-160, 6 - 0.2, 'SPT-S4', horizontalalignment='left',
+plt.text(-160, 6 - 0.2, 'CMB-S4', horizontalalignment='left',
          verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', )
 
 plt.text(-98, 6 - 0.2, r'SKA (10 $\mu$Jy)', horizontalalignment='left',

### 2017-10-04 Code/run_analysis.py

@@ -1,8 +1,6 @@
 
 # # Fisher Analysis
 
-# !module load gcc/6.1
-
 import numpy as np
 import matplotlib.pyplot as plt
 try:
@@ -11,8 +9,7 @@ except ImportError:
     import functools as functools32
 
 import sys
-import os
-import numpy as np
+# import os
 import camb
 from camb import model, initialpower
 import multiple_survey_delens
@@ -20,7 +17,7 @@ import configparser as ConfigParser
 import rho_to_Bres
 from scipy.interpolate import InterpolatedUnivariateSpline
 from colorama import init
-from colorama import Fore, Back, Style
+from colorama import Fore
 
 init(autoreset=True)
 np.seterr(divide='ignore', invalid='ignore')
@@ -30,7 +27,7 @@ plt.rc('text', usetex=True)
 plt.rc('font', **{'family': 'serif', 'serif': ['Computer Modern Roman']})
 # ============================================
 
-# COSMOSIS VALUES!!!!!!
+# cosmology values!!!!!!
 
 pars = camb.CAMBparams()
 # This function sets up CosmoMC-like settings, with one massive neutrino
@@ -50,10 +47,9 @@ pars.max_eta_k_tensor = 3000.
 
 # print(pars) # if you want to test parasm
 results = camb.get_results(pars)
-tens_cl = results.get_tensor_cls(3000)
 
+# Remember there is a 7.4e12 missing and CAMB always give you $ \ell (\ell +1)/2\pi  $
 
-# ### Remember there is a 7.4e12 missing and CAMB always give you $ \ell (\ell +1)/2\pi  $
 
 @functools32.lru_cache(maxsize=64)
 def clbb(r=0.1, nt=None, lmax=3000):
@@ -154,7 +150,6 @@ def fisher_r_nt(r_fid=0.2, fid=None,
     return (sigma_r, sigma_nt, np.sqrt(1 / Frr), np.sqrt(1 / Fnn))
 
 
-cosmosis_dir = '/home/manzotti/cosmosis/'
 inifile = '/home/manzotti/cosmosis/modules/limber/galaxies_delens.ini'
 Config_ini = ConfigParser.ConfigParser()
 values = ConfigParser.ConfigParser()
[…]

### 2017-10-04 Code/run_galaxies_delens.py

@@ -23,8 +23,8 @@ import DESI
 
 
 def setup(ini_file='./gal_delens_values.ini'):
+    #  read possible setup options from ini file
     config = configparser.ConfigParser()
-
     # use configparser for this
     config.read('gal_delens_values.ini')
     # # L BINS
@@ -37,6 +37,8 @@ def setup(ini_file='./gal_delens_values.ini'):
     lbins = 10. ** lbins
     ini_pars = {}
     ini_pars['lbins'] = lbins
+    ini_pars['noisy'] = noisy
+
     return ini_pars
 
 
@@ -85,6 +87,7 @@ def make_tomo_bins_equal_gals(z, dndz, sigmaz, nbins, hspline, omegac, h, b=1.):
 
     z_bins = find_bins(z, dndz, nbins)
     # print('z_bins', z_bins)
+
     def p_z_ph_z(z_ph, z, sigma_z):
         return np.exp(-(z_ph - z)**2 / (2. * sigma_z**2)) / np.sqrt(2 * np.pi * sigma_z**2)
 
@@ -119,6 +122,7 @@ def make_tomo_bins_equal_gals_lsst(z, dndz, sigmaz, nbins, hspline, omegac, h, b
 
     z_bins = find_bins(z, dndz, nbins)
     # print('z_bins', z_bins)
+
     def p_z_ph_z(z_ph, z, sigma_z):
         return np.exp(-(z_ph - z)**2 / (2. * sigma_z**2)) / np.sqrt(2 * np.pi * sigma_z**2)
 
@@ -152,7 +156,8 @@ def make_spec_bins(z, dndz_fun, nbins, hspline, omegac, h, b=1.):
     '''
     spec_bins = []
     galaxies_fraction = []
-    z_bins = [z[i:i + int(len(z) / nbins) + 1] for i in range(0, len(z), int(len(z) / nbins))]
+    z_bins = [z[i:i + int(len(z) / nbins) + 1]
+              for i in range(0, len(z), int(len(z) / nbins))]
     # +1 is inserted not to have gaps
     print(z_bins)
     for z in z_bins:
@@ -166,7 +171,8 @@ def make_spec_bins(z, dndz_fun, nbins, hspline, omegac, h, b=1.):
         dndz_bin = InterpolatedUnivariateSpline(
             z, dndz / norm, ext='zeros')
         # print('here', dndz_bin.integral(z[0], z[-1]))
[…]

### 2017-10-04 Notes/galaxies_delensing_uchicago_thesis.tex
Kimmy and Toshiya input

--- 2017-07-05 Notes/galaxies_delensing_uchicago_thesis.tex
@@
 %% Use these commands to set biographic information for the title page:
-\title{Future Cosmic Microwave Background delensing with galaxies surveys}
+\title{Future Cosmic Microwave Background delensing with galaxy surveys}
 \author{Alessandro Manzotti}
@@
 \division{Physical Sciences}
-\degree{Doctor of Phylosophy}
+\degree{Doctor of Philosophy}
 \date{August 2017}
@@
 \makecopyright
-%\makededication
-%\makeepigraph
 \tableofcontents
@@
 A particular component, the so-called B-modes, is indeed a direct signature of a prediction of inflation: the presence of gravitational waves in the early universe.
-However, improving the level of noise in future experiments will not be enough.
+However, reducing the instrumental noise in future experiments will not be enough to detect this signal.
 Secondary effects in the low redshift universe will also produce non-primordial B-modes adding confusion to the inflationary signal.
 In particular, the gravitational interactions of CMB photons with large scale structures will distort the primordial E-modes adding a lensing B-mode component to the primordial signal.
-Removing the lensing component from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
-Here we discuss the role of current and future large scale structure surveys in multi-tracers approach that will improve the reconstruction of the lensing potential that lenses the CMB photons and and, as a consequence, the delensing efficiency.
+Removing the lensing part (``delensing'') from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
+Here we discuss the role of current and future large scale structure surveys in a multi-tracers approach to delensing that will improve the reconstruction of the lensing potential that lenses the CMB photons and, as a consequence, the delensing efficiency.
 We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
-We find that a galaxy survey like DES can remove about $14\%$ of the lensing signal. current generation CMB experiments will benefit from 
-With the improvement of the noise CMB internal reconstruction will become more important and the fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level.
-Nevertheless optical galaxy surveys, in particular if split in different tomographic bin will still play an important role even for CMB S4. 
+We find that, in general, galaxy surveys should be split into tomographic bins as this can improve the correlation with CMB lensing by 30\%.
+Among currently available surveys, a DES-like galaxy survey can remove about $14\%$ of the lensing signal.
+Ongoing CMB experiments (CMB-S2) will particularly benefit from large scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
+With the decrease of instrumental noise, the CMB internal reconstruction will increase its efficiency and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level.
+Nevertheless optical galaxy surveys will still play an important role even for CMB S4.
+In particular an LSST-like survey can a achieve a delensing performance comparable to a 3G CMB experiment but with completely different systematics.
+This will be important to prove the robustness against systematics of an eventual detection of primordial B-modes.
+% maybe non linear scales?
+
+\vspace{1cm}
 \mainmatter
-
 \chapter{Introduction}
 \label{sec:intro}
-In the standard cosmological paradigm the early universe underwent a period of near-exponential expansion called "cosmic inflation." 
+In the standard cosmological paradigm the early universe underwent a period of near-exponential expansion called ``cosmic inflation''.
 All the cosmological observations agree with this picture, making it a compelling and elegant description of the Universe initial conditions.
 Despite the experimental effort, other possible explanations are still valid, and a conclusive evidence of inflation is still to be found.
-Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski15}.
-This prediction set inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologists are looking for.
+Inflation generically predicts a stochastic background of gravitational waves \cite[see e.g.][for a review]{kamionkowski:2016}.
+This prediction sets inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologists are looking for.
 These primordial gravitational waves in the early universe would imprint a unique signature on the polarized anisotropies of the CMB.
@@
 
-Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational waves background, improving their measurement is a major objective of current and future CMB experiments.
-Even if inflationary B-modes have not been detected yet, a natural value $r \gtrsim 10^{-3}$ should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski15}.
+Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational wave background, improving their measurement is a major objective of current and future CMB experiments.
+Even if inflationary B-modes have not been detected yet, the target value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski:2016,abazajian:2016}.
 However, just reducing the level of noise will not be enough to attain this goal.
@@
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
-This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,polarbear2014c,vanengelen14b,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
+This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
 
@@
 Indeed, the contamination from lensing B modes is already at the level of the instrumental noise of current experiments \citep{bk14}.
-Thus, together with the experimental effort to reduce the amount of noise, the effect of the spurious lensing component must be understood.
+Thus, together with the experimental effort to reduce the amount of noise, the effect of the lensing component must be understood.
 
-The optimal way to reduce the lensing contributions to the B-modes is to reconstruct the expected realization of lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
+The optimal way to reduce the lensing contributions to the B-modes is to reconstruct a template of the actual lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
 We can delens the observed B-modes by combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
 Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
-Furthermore, it has recently been performed on CMB temperature data using the cosmic infrared background as LSS tracer \cite{larsen:2016} and on CMB temperature and polarization data using CMB data to internally reconstruct the LSS lensing potential \cite{ carron17}.
-Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in which $~28\%$ of the lensing power was removed.
+Furthermore, it has recently been performed on CMB temperature data using the cosmic infrared background (CIB) as LSS tracer \cite{larsen:2016a} and on CMB temperature and polarization data using CMB data to internally reconstruct the LSS lensing potential \cite{carron:2017}.
+Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in which $~28\%$ of the lensing power was removed \cite{manzotti:2017}.
 
@@
 
-In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$. 
-Furthermore $W^{i}(z)$ is the kernel function of the field $i$ introduced in \refeq{wkernel}.
-We will now describe the kernels for each of the tracers used in this work.
+In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$. 
+%Furthermore $W^{i}(z)$ is the kernel function of the field $i$ introduced in \refeq{wkernel}.
+We will now describe the kernels $W^{i}(z)$ for each of the tracers used in this work.
 
-\section{CMB lensing potential}
+\subsection{CMB lensing potential}
 \label{sec:kappaCMB}
@@
 where $\chi_*$ is the comoving distance to the last-scattering surface at $z_*\simeq 1090$, $\Omega_{\rm m}$ and $H_0$ are the present day values of the Hubble and matter density parameters, respectively.
+As can be seen in \reffig{kernel}, the lensing kernel $W^{\kappa}(z)$ is a broad function of redshift peaking around $z=1.8$ but extending to high redshift.
 
-The CMB lensing potential is \textit{the} field that we need to reconstruct to reverse the effect of large scale structure and delens the CMB.
+The CMB lensing potential is \textit{the} field that we need in order to reverse the effect of large scale structure and delens the CMB.
 However, the lensing potential can also be reconstructed using the CMB itself.
@@
 
-\section{Galaxies}
+\subsection{Galaxies}
 
@@
 DES is modeled after the DES Science Verification public data release.
-For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy denisty of 0.63 galaxies per arcmin squared. 
+For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy denisty of 0.63 galaxies per squared arcmin. 
 %As in \cite{passaglia:2017} we use a linear bias $b$
 
-For LSST we follow \cite{schaan:2016}: $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
+For LSST we follow \cite{schaan:2017}: $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
 Furthermore we assume a density of 26 galaxies per arcmin squared.
@@
 We model both the redshift distribution and  bias of radio sources following \cite{namikawa:2016a}. 
+In \reffig{kernel}, we compare the CMB lensing kernel $W^{\kappa}(z)$ to the kernels $W^{g}(z)$ of all the tracers introduced here.
 
@@
 
-\section{Cosmic infrared Background (CIB)}
+\subsection{Cosmic infrared Background (CIB)}
 \label{sec:cib}
@@
 We place the peak of the CIB emissivity at redshift $z_c = 2$ with a broad redshift kernel of width $\sigma_z = 2$ and we set $T$ = 34K and $\nu^{\prime} \approx$ 4955 GHz.
+\reffig{kernel} shows how the CIB kernel peaks at higher redshift compared to other galaxy survey kernels, with a better overlap with the CMB lensing one.
+
+There are several available CIB observations that have already been used to delens the CMB.
+Given its full-sky coverage, a promising one is the CIB map derived in \cite{2016A&A...596A.109P} using multi-frequency Planck data and the Generalized Needlet Internal Linear Combination (GNILC) component separation algorithm.
 
@@
 \ee
-%As usual we define the power spectrum as:
-%\be
-%\langle B^{\mathrm{lens}}(\bl) ~ B^{\mathrm{lens}^*}(\tilde{\bl}) \rangle \equiv (2\pi)^2 \delta^D(\bl - \tilde{\bl}) C^{BB,\mathrm{lens}}_l .
-%\ee
-
+Here $\varphi_{\bl,\bl'}$ is the angle between the two different modes $\bl$ and $\bl'$.
 From this we get the power spectrum of the lensing component of the B-modes:
@@
 
-Notice that the first term consists of an inverse variance filter applied to the measured E-mode and the second minimizes the difference between the reconstructed $\phi$ and the CMB lensing potential.
-
-With this choice of $f(\bl,\bl')$ we finally have that the residual power is:
+With this choice of $f(\bl,\bl')$ we find that the residual power is:
 \beqn
@@
 
-\refeq{Bres} directly relate the residual power after delensing with the cross correlation coefficients with CMB lensing of the tracers used.
-The bigger the $\rho_{\ell}^{2}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
-An higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$ leading to a smaller residual power $C_{{l}}^{BB,\mathrm{res}}$.
-\relook{We conclude this section showing in \reffig{bbplot} the expected residual lensing B-modes power spectrum for some of the tracers used in this work together with the primordial B-modes component and the instrumental noise for current and future experiments.
-}
+\refeq{Bres} highligts the different factors that control the delensing efficiency.
+The first part of the second term in the parenthesis consists of an inverse variance filter applied to the measured E-mode.
+The smaller the noise in the E-modes ($N^{EE}$) is the closer this term is to one: a less noisy measurement improves the template of the lensing B-modes.
+The second captures the difference between the reconstructed $\phi$ and the CMB lensing potential, and it directly relates the residual power after delensing with the cross-correlation coefficients with CMB lensing of the tracers used.
+The larger the $\rho_{\ell}^{2}$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons.
+A higher correlation allows for a better reconstruction of $\phi$ and, as a consequence, of $B^{\rm{lens}}$ leading to a smaller residual power $C_{{l}}^{BB,\mathrm{res}}$.
+We conclude this section showing in \reffig{bbplot} the expected residual lensing B-modes power spectrum for some of the tracers used in this work together with the primordial B-modes component and the instrumental noise for current and future experiments.
 
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/BB_res_ell2.pdf}
+\includegraphics[scale=1.]{../images/BB_res_ell2_thesis.pdf}
 \caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
-The red solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed green line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
-The residual power left after delensing for some of the LSS tracers described in \refsec{model}. 
-The rapid improvement in the level of instrumental noise (dashed curves for SPTPol and CMB S4) will require a high level of delensing to exploit these experiments fully.}
+The orange solid line corresponds to the fiducial lensing B-mode component of the signal while the dashed blue line corresponds to the inflationary one for the higher amplitude allowed by current experiments.
+These are compared with the residual power left after delensing using some of the LSS tracers described in \refsec{model}. 
+The rapid improvement in the level of instrumental noise (dashed curves for SPTPol and CMB S4) will require a high delensing efficiency to exploit these experiments fully.}
 \label{fig:bbplot}
@@
 
-\section{Multiple tracers of the lensing potential}\label{sec:galcontrib}
+\subsection{Multiple tracers of the lensing potential}\label{sec:galcontrib}
 
@@
 We start by assuming that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
-We can optimally combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ with:
+We can optimally combine them to estimate $\phi$ or, in other words, to maximize the correlation factor $\rho$ with:
 \beqn
 I         &=&    \sum_{i}c^{i}I^{i} \nonumber \\
-c_{i}     &=& (C_{II}^{-1})_{ij}C^{\kappa I^{j}}
+c_{i}     &=& (C_{II}^{-1})_{ij}C^{{\kappa I}^{j}}
 \label{eqn:combined}
@@
 Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing, but it also depends on how much it is correlated with the already used set of tracers.
-\relook{
 \reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
-Note that the cross correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
-CIB and 21 cm surveys probes the high redshift structures and, independently from the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
+The cross-correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
+The cosmic infrared background and 21 cm surveys probe the high redshift structures and, independent of the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
 On the other end, galaxy clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
-However, given the low noise of this measurement and their small overlap with other probes they can still play an important role in delensing even if their overlap with the CMB lensing potential is not optimal.
-}
+However, given the low noise of these measurements and their small overlap with other probes, they can still play a major role in delensing even if their overlap with the CMB lensing potential is not optimal.
 
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/compare_kernel.pdf}
-\caption{\textbf{Kernels Comparison}: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
-The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential will be leading to a higher delensing efficiency.
+\includegraphics[scale=1.]{../images/compare_kernel_thesis.pdf}
+\caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
+The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency.
 }
@@
 
-\section{Improving efficiency with tomographic binning}\label{sec:tomo-bin}
+\subsection{Improving efficiency with tomographic binning}\label{sec:tomo-bin}
 The delensing efficiency of galaxy surveys can be improved by taking into account redshift information.
 When we weight a tracer with $\frac{C^{\kappa I}}{C^{II}}$ in \refeq{fweight} in order to maximize its ability to reconstruct the lensing potential we are only using redshift averaged information about the survey.
-However, as can be seen in \reffig{kernel}, the kernel overlap of a tracer with the CMB lensing varies as a function of redshifts.
-For this reason, the optimal approach is to weight galaxies at different redshifts with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
+However, as can be seen in \reffig{kernel}, the kernel overlap of a tracer with the CMB lensing varies as a function of redshift.
+For this reason, the optimal approach is to weight galaxies at different redshift with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
 We can see this with a simple example. Let's split a single survey I into two non-overlapping redshift bins $I_{1}$ and ${I_{2}}$ with $I=I_{1}+I_{2}$.
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/compare_kernel_tomo.pdf}
-\caption{\textbf{Kernels Comparison with tomographic bins}: Comparison of the 10 LSST  tomographic bins together with the full LSST survey and the CMB lensing kernel. 
-Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross correlation with the CMB lensing.
+\includegraphics[scale=1.]{../images/compare_kernel_tomo_thesis.pdf}
+\caption{Comparison of kernels of the 10 LSST tomographic bins together with the full LSST survey and the CMB lensing kernel. 
+Compared to a full survey approach, tomographic binning allows to optimally weight different bins according to their cross-correlation with the CMB lensing.
 This leads to a better delensing efficiency.
@@
 For photometric surveys, the maximum number of bins is dictated by the fact that the bin width can not be smaller than the photo-z accuracy.
-We used 10 and 4 photometric bins for LSST and DES respectively with a common photo-z accuracy $\sigma(z) = 0.05(1+z)$. 
-\relook{In spectroscopic surveys, there are no limitations in increasing the number of bins. We split DESI into 4 spectroscopic bins with no overlap among each others.
-This number of bins is close to the saturation point where adding more bins does not improve delensing significantly while adding complexity to the analysis.} As an example, we show in \reffig{kerneltomo} the 10 bins and the full LSST redshift distributions together with the CMB lensing kernel.
-\relook{The improvement obtained by tomographic binning is illustrated in \reffig{rho-tomo}.
-In particular on large scales, binning can increase the value of $\rho$ by almost $30\%$ significantly improving the delensing efficiency of galaxy surveys.}
+We used 10 and 4 photometric bins for LSST and DES respectively with a photo-z accuracy of $\sigma(z)_{\textrm{DES}} = 0.05(1+z)$ for DES and $\sigma(z)_{\textrm{LSST}} = 0.01(1+z)$ for LSST. 
+In spectroscopic surveys, there are no limitations in increasing the number of bins. We split DESI into 4 spectroscopic bins with no overlap among each others.
+This number of bins is close to the saturation point where adding more bins does not improve delensing significantly while adding complexity to the analysis.
+As an example, we show in \reffig{kerneltomo} the 10 bins and the full LSST redshift distributions together with the CMB lensing kernel.
+\reffig{rho-tomo} illustrate the improvement obtained by tomographic binning.
+In particular on large scales, binning can increase the value of $\rho$ by almost $30\%$ significantly improving the delensing efficiency of galaxy surveys.
 
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/B_res_bin.pdf}
+\includegraphics[scale=1.]{../images/B_res_bin_thesis.pdf}
 \caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing.
-Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of angular scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line).}
+Here we show the cross-correlation coefficient (\refeq{rho1}) as a function of angular scale for full surveys (solid lines) and tomographically binned surveys (dot-dashed line).
+For details about the binning, see \refsec{tomo-bin} and in particular \refeq{zbin}.
+}
 \label{fig:rho-tomo}
@@
 We will use the Fisher information matrix to quantify the delensing efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
-We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction performance, together with a low-noise, low-resolution experiment the B-modes of which \textcolor{red}{(?)} will be delensed and used to constrain the inflationary parameters.
-We focus on three distinct scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
+We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the ability to internally reconstruct the potential $\kappa$, together with a low-noise, low-resolution experiment that measured the B-modes that will be delensed and used to constrain the inflationary parameters.
+Even if there is not a definite distinction between different stages, we focus on three distinct scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.
 
@@
 %\begin{center}
-%\includegraphics[scale=1.2,clip]{../images/cmb_internal.pdf}
+%\includegraphics[scale=1.]{../images/cmb_internal.pdf}
 %\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
@@
 
-\section{Fisher Information Matrix}
+\subsection{Fisher Information Matrix}\label{sec:fisher}
 In the Fisher Information Matrix formalism \cite{10.2307/2342435}, the statistical uncertainty on a cosmological parameter $p$ can be obtained from the inverse of the Fisher matrix $F_{ij}$ as $\sigma(p) =\sqrt{(\mathbf{F})_{pp}^{-1}}$.
@@
 
-It can be seen in \refeq{fisher} that removing the lensing contribution will improve parameters constraints.
+It can be seen in \refeq{fisher} that removing the lensing contribution will improve parameter constraints.
 The parameter uncertainties are inversely proportional to the covariance of the measurement.
-Since the lensing B-modes $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them trough delensing will reduce the parameter statistical error.
-After delensing the error $\sigma^{\rm{del}}(p)$ can still be obtained from \refeq{fisher} with $C_{{l}}^{BB,\mathrm{res}}<C_{{l}}^{BB,\mathrm{lens}}$ in the covariance defined in \refeq{cova}.
+Since the lensing B-modes power $C_{{l}}^{BB,\mathrm{lens}}$ is a substantial component of the covariance, removing part of it through delensing will reduce the parameter statistical error.
+We define the error after delensing as $\sigma^{\rm{del}}(p)$.
+This is computed from \refeq{fisher} substituting $C_{{l}}^{BB,\mathrm{lens}}$ with the residual power after delensing $C_{{l}}^{BB,\mathrm{res}}$ in the covariance matrix defined in \refeq{cova}.
 We define the improvement as the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$.
+The fiducial values used are described in \refsec{del_res}.
+Given a fiducial value for $r$, the value of $n_{t}$ is fixed imposing the consistency relation $n_t=-r/8$ \cite{LIDDLE1992391}.
 
 In \refeq{fisher} we are making a few important assumptions. 
-Firstly, we are fixing all the cosmological parameters except $\{r,n_{t}\}$.
-Uncertainties in those will, however, propagate to uncertainties in $\{r,n_{t}\}$.
+First, we are fixing all the cosmological parameters except $\{r,n_{t}\}$.
+Uncertainties in those will propagate to larger uncertainties in $\{r,n_{t}\}$.
 While neglecting this will lead to slightly optimistic constraints, it has no significant impact on the estimate of the improvement due to delensing.
@@
 We checked this for a few of the tracer combinations used here and find it particularly true once several tracers are jointly taken into account.
-Finally, we have assumed a Gaussian covariance even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
+Finally, although we have assumed a Gaussian covariance $\sigma( C_{\ell}^{BB} )$ it actually has small non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
 This approximation is good enough to show the improvement due to delensing.
+The approximations made here go in the direction of a slightly optimistic \textit{absolute} value for the parameters statistical uncertainties both for the standard and the delensed case.
+The delensing improvement is defined as the \textit{relative} value of these uncertainties $\alpha_{i} = \sigma^{\rm{del}}(i)/\sigma(i)$ and, at first order, it is not affected by this assumptions.
 
 %===================
-\section{Delensing with current CMB and LSS}
+\subsection{Delensing with current CMB and LSS}
+\label{sec:del_res}
 
@@
 
-Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} as well as the internal CMB lensing potential reconstruction \cite{carron:2017} as large scale structure tracers.
-Here we discuss the improvement that can be obtained by combining these and other currently available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
+Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} and internal CMB lensing maps \cite{carron:2017} as large scale structure tracers.
+Here we discuss the improvement that can be obtained by combining these and other currently Stage-2 available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
 
-On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment.
-For the deep experiment we assume an instrumental noise equal to \textcolor{red}{$3\mu$K-arcmin}, a beam of \textcolor{red}{30 arcmin} and an angular scale range of $50<\ell<500$.
+On the CMB side, we will combine a BICEP-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment.
+For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin, a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
 We assume the CMB lensing reconstruction is  performed by the higher resolution CMB experiment. We explore two possibilities.
-First, we test the efficiency with an internal reconstruction of CMB lensing performed by Planck collaboration \cite{planck-collaboration:2016a}.
+First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck collaboration \cite{planck-collaboration:2016a}.
 We compute the noise in the CMB lensing map in \refeq{noisekappa} using the noise curves publicly available
 \footnote{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}.
-Then we assume a ground-based experiment with noise levels consistent with SPT-Pol: \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.
+Then, we test a second scenario with a ground-based experiment characterized by noise levels consistent with SPT-Pol: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin.
 In this case we set the largest achievable scale at $\ell_{min}=300$.
@@
 
-In both cases the CMB lensing reconstruction correlates very well at low $\ell$ and it then falls rapidly at smaller scales because of the raise of the reconstruction noise.
 For Planck, the internal reconstruction is at most $70\%$ correlated at very large scales and then it falls rapidly to $40\%$ at $\ell=200$.
 Its correlation is comparable to the one of LSS at almost all scales.
-On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 80\% correlated with the true field at $\ell<300$, and only then the correlation becomes comparable with the current LSS surveys.
+On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 80\% correlated with the true field at $\ell<300$, and only then the correlation becomes comparable to the current LSS surveys.
+In both cases the CMB lensing reconstruction correlates very well at low $\ell$ and it then falls rapidly at smaller scales because of the raise of the reconstruction noise.
 
 \reffig{corrnow} and \reffig{corrnow-planck} show that the DES galaxies are effective tracers of the LSS and can, at least in the near future, be used to improve delensing for CMB experiments that overlap with it.
-For example, DES delensing efficiency is higher than WISE firstly because of the lower level of noise and furthermore because DES galaxies are located at slightly higher redshift thus they better overlap with CMB lensing.
+For example, DES delensing efficiency is higher than WISE first because of the lower level of noise and also because DES galaxies are located at slightly higher redshift so they better overlap with CMB lensing.
+As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS.
+In particular it can be seen in \reffig{corrnow} that optical surveys will rapidly lose the role of filling in large-scale modes and they will start to supplement information at higher multipoles.
 
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/actual_scenario_planck.pdf}
-\caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the angular scale $\ell$. 
-The dashed curve corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$, showing which scales contribute to the $\ell<100$ B-mode power.}
+\includegraphics[scale=1.]{../images/actual_scenario_planck_thesis.pdf}
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers.
+Here we show both current galaxy survey and internally reconstructed CMB lensing potential. 
+The dashed curve highlights which scales contribute to the $\ell<100$ B-mode power and are thus most useful to delens.
+It corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$}
 \label{fig:corrnow-planck}
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/actual_scenario.pdf}
-\caption{Correlation factor between current galaxy survey and internally reconstructed $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
+\includegraphics[scale=1.]{../images/actual_scenario_thesis.pdf}
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers. 
+This figure is the same as \reffig{corrnow-planck} but using SPTPol instead of Planck as the high-res CMB experiment performing the internal reconstruction.
+}
 \label{fig:corrnow}
@@
 
+The lensing reconstruction noise in Planck is such that internal delensing can only remove $8\%$ of the power, improving the constraints on $r$ by a factor of 1.06 even in the ideal scenario of no instrumental noise in the B-modes.
+A similar performance is achieved with WISE.
+DES can do better, removing a level of 14\% of power.
+However, it covers a smaller fraction of the sky than the two previous probes.
+As shown in previous works \cite{sherwin:2015}, the best tracer is the CIB with a 27\% reduction in power.
+Nonetheless, it is worthwhile combining all the tracers.
+Indeed a multi tracer-approach can bring the removed power on the overlapping area from 27\% of the CIB alone to 36\% using all the LSS (LSS-S2) and 41\% when Planck lensing reconstruction is added to the set.
+However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 30\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
+Furthermore, as explained in \refsec{fisher}, this is only an upper limit and will be even smaller when, for example, non-perfect foreground cleaning is considered.
+Indeed current experiments are still significantly limited by foregrounds and instrumental noise together and not only by lensing.
+
+If, on the other hand, we consider a high-resolution ground base experiment the internal lensing reconstruction improves significantly. 
+\reftab{current-spt} shows that, with an instrumental noise at the expected level for SPTPol, the CMB will soon be able to remove an amount of power (35\%) at the level of the CIB.
+Once combined with LSS tracer this will lead to removing 56\% of the power.
+The improvement of 1.57 in $\sigma(r)$ that will follow, even when realistic noise is considered, is such that it will make delensing a  needed step for CMB polarization data analysis.
+
 \begin{table}
 \centering
-
-\caption{$\alpha(r)$: Improvements on $\sigma(r)$ due to delensing for current generation and Planck lensing reconstruction.
+\caption{$\alpha(r)$: Improvements on $\sigma(r)$ due to delensing for completed and ongoing Stage-2 surveys and Planck lensing reconstruction.
+LSS-S2 corresponds to the combination of all the available LSS tracers.
 The values in parenthesis in the first column correspond to the fraction of lensing B-mode power removed using each LSS tracer. 
-The values in the other columns correspond to ratio of the error before and after delensing.}
+The values in the other columns correspond to ratio of the error before and after delensing for 3 cases: no instrumental noise in the B-mode measurement and realistic noise for two different values of $r$.}
   \vspace{0.2cm}
@@
  \centering
-\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to the ratio of the error before and after delensing.}
+\caption{$\alpha(r)$: Stage-2 improvements on $\sigma(r)$ due to delensing. We use the same LSS tracers as \reftab{current} but with the internal lensing reconstruction performed by SPTPol.}
   \vspace{0.2cm}
@@
 \hline
-Surveys &$\alpha(r=0), N^{B}_{\ell}=0$& $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
-CMB SPTPol (35\%) && 1.62  &  \\ \hline
-LSS+CMB (56\%) && 2.35 &   \\ \hline
+Surveys &$\alpha_{r=0}, N^{B}_{\ell}=0$& $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
+SPTPol (35\%) &1.46& 1.27  & 1.08 \\ \hline
+LSS-S2 +SPTPol (56\%) & 2.18 & 1.57 & 1.16  \\ \hline
+
   \end{tabular}
@@
 
-\section{CMB-S3 Era}
+\subsection{CMB-S3 Era}\label{sec:S3}
 CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
-We model the CMB S3 stage with a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam. 
-We assume here that a deep CMB experiment and an overlapping high resolution one will be able to combine their measured modes.
-For this reason, we will also assume this level of noise for the internal noise reconstruction.
-With this experimental setup we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
-Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers.
+As we did for Stage-2, we model the CMB S3 as two different overlapping experiments that will be able to combine their measurement.
+The two experiments combined will have a level of noise of $3\mu$K-arcmin and a 1 arcmin beam. This level of noise is also assumed for the internal noise reconstruction.
+With this experimental setup, we can push the angular scale range used to calculate the Fisher matrix to $50<\ell<3000$ even if most of the high $\ell$ scale do not contribute to the constraints.
+Not only will CMB experiments improve shortly: eventually DESI will start taking data. For this reason, we add DESI to the LSS tracers used in \refsec{del_res}.
 
@@
 An interesting finding is that DESI will be less efficient (removing 10\% of the power) than a DES-like survey (14\%) despite the fact that it can probe slightly higher redshift.
-The reason is that, for the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES play a significant role in the delensing efficiency.
-Unfortunately, adding DESI will only bring the power removed using LSS fro 36 to 41\%.
+The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency significantly.
+Unfortunately, adding DESI will only bring the power removed using LSS from 36 to 41\%.
 Indeed, even in the near future, CIB will still play the dominant role among LSS.
-Contrary to current experiments, th CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 550$.
+Contrary to current experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 550$.
 
@@
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/S3_scenario.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 3 experiments. }
+\includegraphics[scale=1.]{../images/S3_scenario_thesis.pdf}
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
+This figure is the same as \reffig{corrnow} but with a stage-3 CMB experiment performing the internal reconstrcution and with DESI added to the stage-2 galaxy surveys (LSS-S2).
+}
 \label{fig:corrS3}
@@
 
-The improvement in the constraint on $r$ is summarized in \reftab{S3}. 
-Given the improvement in the noise of the high resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
-Despite this, adding galaxy survey will still lead to a decent improvement with a fairly low effort.
+We summarize the improvement in the constraint on $r$ in \reftab{S3}. 
+Given the improvement in the noise of the high-resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of 2 through internal delensing in the ideal case of no instrumental noise.
+Adding galaxy surveys will lead to a further improvement without additional effort.
 Indeed galaxy surveys will still be able to remove an additional $20\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by $15\%$.
-Note that apart from the addition of DESI, the tracers used here are already available for the current generation.
+Note that apart from the addition of DESI, the tracers used here are already available today.
 The ongoing effort can then lead to a significant improvement even for the next generation of experiments.
@@
 As expected the improvement on $n_{t}$ is similar to the one in $r$ since they are both proportional to the variance of the measured B-modes.
-We tested this in the scenario $r=0.12$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}/8}$.
-It is important to note that, independently from the level of delensing, for CMB 3G experiments this statistical error will still be several times bigger than the fiducial value.
+We tested this in the scenario $r=0.12$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}}/8$.
+It is important to note that, independent of the level of delensing, for CMB 3G experiments the statistical error on $n_{t}$ will still be several times bigger than the fiducial value.
 
@@
 \centering
-
-\caption{$\alpha$: S3 improvements. Note that the absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is $\sigma(r) =$  }
+\caption{$\alpha$: Improvements on $\sigma(r)$ due to delensing for S3 experiments.
+Here we define LSS-S3 as the combination of S2 LSS tracers (LSS-S2) and DESI.
+We also use a CMB S3 experiment for the lensing internal reconstruction.
+See \refsec{S3} for details.}
+%\hspace{2.2cm}
 \begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
 DESI (10\%) &1.12 & 1.08  & 1.03 \\ \hline
-LSS (40\%) &1.73& 1.42  & 1.2\\ \hline
+LSS-S3 (40\%) &1.73& 1.42  & 1.2\\ \hline
 CMB 3G (56\%) &2& 1.57  & 1.21 \\ \hline
-LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
+LSS-S3+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
 \end{tabular}
@@
 
-\section{CMB-S4 Era}
+\subsection{CMB-S4 Era}\label{sec:S4}
+Having as a major goal the detection of inflationary B-modes, CMB data will continue improving even after Stage-3.
+An ambitious program for a Stage-4 ground CMB experiment is in the planning phase \cite{abazajian:2016}. 
+Moreover, satellite and balloon CMB experiments have been proposed and have the potential to extend the accessible B-mode measurements to the largest scales.
+Given the unprecedented low level of noise of these experiment, delensing will be even more important than previous generations.
+Following \cite{abazajian:2016}, here we assume a CMB-S4 ground experiment with a level of noise of $1\mu$K-arcmin and a 1 arcmin beam.
+Furthermore, by the mid-2020s, several next-generation LSS surveys will be online. 
+Here, as an example of future optical galaxy surveys, we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers. 
+We do not consider other surveys but we test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance. 
+Even if these LSS surveys will observe an overlapping part of the sky and similar modes, because of their different strengths and weaknesses, it will still be important to combine them efficiently.
+We also consider a radio continuum survey modeled following the SKA specifications.
+Radio continuum observations of the 21 cm line are in their early stage, and several experimental and data-analysis challenges need to be overcome.
+However, this technique has the potential to map the LSS with relatively low noise up to redshift $z=6$.
+Here we considered just two possible detection thresholds at 1 GHz (flux cut): 10 $\mu$Jy and 0.1 $\mu$Jy, which are representatives of the upper and lower limit levels for a SKA phase 1 (SKA1) and a following SKA phase 2 (SKA2) survey.
+Given the importance of delensing for Stage-4 experiments, here we combine radio and optical survey to test their delensing efficiency.
 
-An ambitious program for a generation 4 ground CMB experiment is currently under planning \cite{abazajian:2016}. 
-Moreover, satellite and balloon experiments have been proposed and have the potential to extend the accessible B-mode measurement to the largest scales.
-Here we follow \cite{abazajian:2016} and assume a CMB-S4 ground experiment with  with a level of noise of \textcolor{red}{$1\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam.
-By the mid 2020s several next generation of LSS survey will be online.
-Here, as an example for future optical galaxy surveys we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers.
-We test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} will have similar delensing performance.
-Even if they will observe similar part of the sky, because of the different  strength and weakness of this experiments
-it will still important to combined them efficiently.
-We also consider a radio continuum survey modeled following the SKA planned experiment.
-Radio continuum observation of the 21 cm line is in its early stage and several experimental and analysis challanges
-need to be overcome. However this technique has the potential to map LSS with relatively low noise up to redshift $z=6$.
-
-The correlation factor with CMB lensing of stage 4 experiments is shown in \reffig{corrs4}.
-
+The correlation factor with CMB lensing for Stage-4 experiments is shown in \reffig{corrs4_10}  and in \reffig{corrs4_01} for SKA1 and SKA2 respectively.
+%=======
 \begin{figure}[htbp]
 \begin{center}
-\includegraphics[scale=1.2,clip]{../images/S4_scenario.pdf}
-\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments. }
-\label{fig:corrs4}
+\includegraphics[scale=1.]{../images/S4_scenario_ska10_thesis.pdf}
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
+This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstrcution and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
+}
+\label{fig:corrs4_10}
 \end{center}
 \end{figure}
+%=======
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/S4_scenario_ska01_thesis.pdf}
+\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
+This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstrcution and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
+}
+\label{fig:corrs4_01}
+\end{center}
+\end{figure}
+%=======
+A tomographically binned LSST-like experiment will be a very efficient CMB lensing tracer.
+Indeed it will be more than 70\% correlated with CMB lensing for all the scales $\ell<600$, a performance similar to a Stage-3 CMB internal reconstruction.
+A very similar level of efficiency will be achieved by a SKA1-like experiment.
+With SKA and LSST, for the first time we will have LSS tracers with a higher delensing efficiency than the CIB.
+Despite the improvement in galaxy surveys, the CMB internal reconstruction will still be the main source of delensing.
+Having a perfect kernel overlap with the true lensing potential it will benefit from the very low level of noise of CMB S4 experiments.
+\reffig{corrs4_10} shows that a CMB-S4 experiment will be able to internally reconstruct the lensing potential at more than 90\% up to $\ell=400$.
+At this level, CMB S4 internal delensing will not be limited by noise but by small secondary effects like foregrounds contamination, filtered modes etc.
+Finally, \reffig{corrs4_01} shows that a very futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
+This performance is comparable to a CMB-S4 internal reconstruction.
 
-%\begin{figure}[htbp]
-%\begin{center}
-%\includegraphics[scale=1.2,clip]{../images/S4_scenario_no_ska.pdf}
-%\caption{Correlation factor. Same as \reffig{corrnow} but for stage 4 experiments without SKA. }
-%\label{fig:corrs4}
-%\end{center}
-%\end{figure}
+\reftab{S4} shows the improvement on the inflationary constraint given the cross-correlation factors described above. 
+LSST alone will be able to remove almost half of the B-mode lensing power.
+This will lead to an improvement in $\sigma(r)$ of a factor of 1.8.
+If we add all the other infrared and optical LSS tracers previously included (i.e.
+excluding SKA) the improvement factor increase to 2.3.
+These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSS, completely independent of CMB internal data.
+Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of the CMB S3 internal reconstruction.
+Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power.
+This will bring an almost factor of 4 improvement in the null test case even with realistic noise.
+Combining with LSS will only increase the removed power compared to CMB only by $\simeq 5\%$ but it will be highly beneficial to test the robustness of the final result.
 
-A tomographically binned LSST-like experiment will be a very efficient CMB lensing tracer.
-Indeed it will be more than 70\% correlated with CMB lensing for all the scales $\ell<600$ a performance similar to a Stage 3 CMB internal reconstruction.
-SKA very new but potentially amazing
-With SKA and LSST, for the first we will have LSS tracers way more efficient than the CIB.
-Even if galaxy survey will improve dramatically the CMB internal reconstruction will still be the main source of delensing.
-This, having a perfect kernel overlap with the true lensing potential we will benefit for the very low level of noise of CMB S4 experiments.
-CMB S4 delensing will probably be limited by small secondary effect like foregrounds contamination, missing modes due to filtering etc.
-
-\reftab{S4} shows the improvement on the inflationary constraint.
-
-%=======================================
+We find the improvement in $\sigma(n_{t})$ to be very similar to the one in $\sigma(r)$ for all the CMB-S4 cases.
+This is reasonable because, given the level of noise in CMB-S4, the variance at angular scales $\ell<200$ are dominated by instrumental noise and not by the lensing component.
+Independent of the delensing level, the small value allowed for $r$ will not allow to test the consistency relations.
 
@@
 \centering
-
-\caption{$\alpha$: Gen 4 experiments}
+\caption{$\alpha$: Improvements on $\sigma(r)$ due to delensing for S4 experiments.
+Here we define LSS-S4 as the combination of S3 LSS tracers (LSS-S3) and LSST.
+We also consider SKA-like radio-continuum surveys.
+We also use a CMB S4 experiment for the lensing internal reconstruction.
+See \refsec{S4} for details.}
 \begin{tabular}{|c | c | c | c|}
 \hline
-Surveys & $\alpha(r=0), N^{B}_{\ell}=0$ & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
-LSST (10\%) &1.12 & 1.08  & 1.03 \\ \hline
-SKA (40\%) &1.73& 1.42  & 1.2\\ \hline
-CMB S4 (56\%) &2& 1.57  & 1.21 \\ \hline
-LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
+LSST (48\%) &1.8 & 1.7  & 1.7  \\ \hline
+LSS-S4  (60\%) &2.5 & 2.3  & 2.3 \\ \hline
+SKA (10$\mu$Jy) (53\%) &2.1& 2  & 2\\ \hline
+SKA (0.1$\mu$Jy) (77\%) &4.6& 3.9  & 3.9\\ \hline
+LSS-S4 + SKA (10$\mu$Jy)  (65\%) &2.7 & 2.5  & 2.5  \\ \hline
+LSS-S4 + SKA (0.1$\mu$Jy)  (80\%) &4.6 & 3.9  & 3.9 \\ \hline
+CMB S4 (81\%) & 5.1 & 4.3  & 4.2 \\ \hline
+LSS+CMB (86\%)& 5.2 & 4.4  & 4.3 \\ \hline
 \end{tabular}
@@
 \end{table}
-
 %=======================================
-
-\begin{table}
-\centering
-
-\caption{$\alpha$: Gen 4 experiments}
-  \begin{tabular}{ | c | c  |  c |}
-\hline
-Surveys & $\alpha(r)$ & $\alpha(n_{t})$\\ \hline \hline
-Euclid & 1.53 & \\ \hline
-LSST & 1.65  &\\ \hline
-SKA & 4.35  &\\ \hline
-CMB S4 & 3.7  &\\ \hline
-LSS+CMB & 6.3  &\\ \hline
-\end{tabular}
-\label{tab:S4}
-\end{table}
-
-%\section{Bias uncertainties degradation}
+%\begin{table}
+%\caption{$\alpha$: Gen 4 experiments}
+%  \begin{tabular}{|c | c | c | c|}
+%\hline
+%Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
+%LSST (10\%) &1.12 & 1.08  & 1.03 \\ \hline
+%SKA (0.1$\mu$Jy) (40\%) &1.73& 1.42  & 1.2\\ \hline
+%CMB S4 (56\%) &2& 1.57  & 1.21 \\ \hline
+%LSS+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
+%\end{tabular}
+%\label{tab:S4_2}
+%\end{table}
+%=======================================
+%\subsection{Bias uncertainties degradation}
 %The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of the inflationary constraint of delensed spectra.
@@
 %We compute the derivatives of the power spectra as described
-
-%\begin{table}
-%\caption{$\alpha$: improvement on r constraint}
-%  \begin{tabular}{ | c | c  | }
-%\hline
-%Surveys & $\alpha$\\ \hline \hline
-%des & 1.34  \\ \hline
-%cib & 1.71  \\ \hline
-%cmb current & 1.79  \\ \hline
-%gals current & 2.03  \\ \hline
-%cmb S3 & 2.15  \\ \hline
-%gals S3 & 2.16  \\ \hline
-%gals S4 & 2.16  \\ \hline
-%comb current & 2.66  \\ \hline
-%comb S3 & 3.14  \\ \hline
-%cmb S4 & 4.36  \\ \hline
-%comb S4 & 5.27  \\ \hline
-%  \end{tabular}
-%\end{table}
-
 \chapter{Conclusions}
 \label{sec:concl}
-
-The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal ("delensing") is necessary to test the inflationary paradigm using the next generation of CMB polarization experiments.
-To do so, we need to accurately reconstruct the large scale structures that lens the CMB in order to build a template of the lensing B-modes.
+The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal (``delensing'') is necessary to test inflation using the next generation of CMB polarization experiments.
+To delens, we need to accurately reconstruct the large scale structures that lens the CMB in order to build a template of the expected lensing B-modes in the observed patch.
 In this paper, we studied the potential impact of large-scale structure galaxy surveys in this important endeavor.
 %For current experiment, the cosmic infrared background had already proven to be very efficient in delensing \cite{larsen:2016,manzotti:2017}.
-For ongoing experiments, we find that an optical survey like DES alone is able to remove $14\%$ of the lensing component and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the contaminant signal using only LSS survey.
-Depending on the CMB noise and the amount of galactic foreground cleaning these can correspond to an improvement of 30\% in the constrain on the tensor to scalar ratio $r$.
 
-In the future, the improved level of noise of CMB experiments will allow to internally reconstruct the structures lensing the CMB. 
-The fraction of removed lensing B-modes will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (76\%) level. 
-Already in the 3G era, the CMB internal reconstruction will be the main source of delensing.
-However it will still be less efficient than galaxies at small scales$\ell>500$.
-For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $57\%$ to $68\%$.
+For ongoing experiments, we find that LSS tracers will be particularly beneficial.
+An optical survey like DES is able to remove $14\%$ of the B-mode lensing power alone, and together with WISE($8\%$) and the CIB ($27\%$) will allow to remove $36\%$ of the power using only LSS surveys.
+Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to a maximum improvement of 30\% in the constraint on the tensor to scalar ratio $r$ compared to the value before delensing.
+In the future, the decreasing level of instrumental noise in CMB experiments will dramatically improve the internal reconstruction of the structures lensing the CMB. 
+The fraction of removed lensing B-modes will rapidly improve from the current expected levels for Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) levels. 
+For Stage-3 experiments, the CMB internal reconstruction will be the main source of delensing.
+However, it will still be less efficient than galaxies at tracing the lensing potential at small scales $\ell>500$.
+For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $57\%$ to $68\%$ for 3G.
+Even for CMB S4, galaxy surveys will still play an important role.
+For example, a tomographically binned LSST-like survey by itself will remove 56\% of the lensing power.
+This performance is lower than a S4 CMB internal reconstruction and comparable to a Stage-3 CMB.
+However, it will allow us to probe the robustness against systematics of an eventual detection of primordial gravitational waves.
+Indeed delensing with just CMB data will require a careful study of possible biases and systematic effects because we delens the same dataset (the CMB) that we also use to reconstruct the lensing potential \cite{carron:2017,sehgal:2017,namikawa:2017}.
+For this reason, efficient galaxies tracers are not only useful in the short term but, in the future, will also play a role in testing internal biases and performing consistency tests.
 
-Furthermore internal delensing will require a careful study of possible biases and systematics effects due to the use of the same source (the CMB) we are delensing to reconstruct the lensing effect itself \cite{carron:2017,sehgal:2016,namikawa:2017}.
-For this reason, efficient galaxies tracers are not only useful in the short term but will also play a role in cross-checking these internal biases. 
-For a future galaxy survey like LSST we find that using tomographic binning can improve its correlation with lensing by $30\%$.
-
-Giving the high level of correlation of future galaxy surveys and CMB lensing it is worth looking for possible application beside the detection of inflationary B-modes.
-For example LSS delensing allow to remove just the low-z lensing component from the CMB lensing.
-Removing this component can avoid the necessity of modelling nonlinearities in the study of CMB lensing power spectrum.
+Given the high level of correlation of future galaxy surveys and CMB lensing, it is also worth looking for potential application besides the detection of inflationary B-modes.
+For example, delensing with galaxy surveys allows us to selectively remove from the CMB only the gravitational effect coming from the non-linear structures in the low-redshift universe. 
+Removing this component will reduce the level of non-linearities both in the CMB and in the CMB lensing power spectrum providing an alternative route to exploit all the measurable modes without having to model the non-linear component at small angular scales.
 
 % ===================
-%\onecolumngrid
+\newpage
 \begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.2,clip]{../images/errors_summary.pdf}
-\caption{\relook{to be filled with final results}}
+%\begin{center}
+\includegraphics[scale=0.5]{../images/errors_summary_thesis_noted.pdf}
+\caption{Summary of the amount of lensing B-modes power removed by different tracers for different generations of experiments.
+The bars on the left represent the individual contribution of different galaxy tracers.
+On the right, the gray bar represents the contribution from LSS only and the yellow one the total removed power once LSS and CMB internal reconstruction are combined.
+The values are taken from \reftab{current}, \reftab{current-spt}, \reftab{S3} and \reftab{S4}. }
 \label{fig}
-\end{center}
+%\end{center}
 \end{figure}
-\twocolumngrid
+% ===================
+
+%=================================================================
+\chapter{Acknowledgments}
+We thank S. Dodelson, S. Passaglia, K. Story, A. Vieregg and K. Wu for useful discussion and their review of the draft. 
+Furthermore we thank D. Alonso for his feedback on radio continuum surveys.
+
+This work made use of computing resources and support provided by the Research Computing Center at the University of Chicago.
+This work was partially supported by the Kavli Institute for Cosmological Physics at the University of Chicago through grants NSF PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.
+%%%=================================================================
 

### 2017-10-04 Notes/galaxy_delensing.tex
Kimmy and Toshiya input

@@
 \usepackage{xstring}
-% \usepackage{graphicx}\usepackage[top=2.5 cm, bottom=2.5 cm, left=1.05in, right=1.05in]{geometry}
 \usepackage[applemac]{inputenc}
@@
 \usepackage{epstopdf}
-%\usepackage[caption=false]{subcaption}
-%\epstopdfDeclareGraphicsRule{.gif}{png}{.png}{convert gif:#1 png:\OutputFile}
-% \AppendGraphicsExtensions{.gif}
+\usepackage{caption}
+\captionsetup{justification=centerlast,singlelinecheck=false,font=small,skip=2pt}
 % notes to make comment is non-zero if
@@
 This will be important to prove the robustness against systematics of an eventual detection of primordial B-modes.
-% maybe non linear scales?
 \end{abstract}
@@
 
-The optimal way to reduce the lensing contributions to the B-modes is to reconstruct a template of the actual lensing B modes on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
+The optimal way to reduce the lensing contributions to the B-modes is to reconstruct a template of the actual lensing B modes (or the actual large scale structure lensing potential) on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
 We can delens the observed B-modes by combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
@@
 For LSST we follow \cite{schaan:2017}: $\frac{dN}{dz}\propto z^{\alpha}\exp^{-(z/z_{0})^{\beta}}$ with $\alpha = 1.27$, $\beta = 1.02$, and $z_{0}= 0.5$.
+These correspond to a median redshift for LSST galaxies around $z_{\text{median}}\sim1.1$.
 Furthermore we assume a density of 26 galaxies per arcmin squared.
@@
 
+% ===================
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=1.]{../images/compare_kernel.pdf}
+\caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
+The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency.
+}
+\label{fig:kernel}
+\end{center}
+\end{figure}
+% ===================
+
+% ===================
+
 \begin{figure}[htbp]
@@
 \end{figure}
+% ===================
 
@@
 \be
-\rho^{2}_{\ell} = \sum_{i,j}\frac{C^{\kappa ^{i}}_{\ell}~(C^{-1}_{\ell})_{ij}~C^{\kappa ^{j}}_{\ell}}{C^{\kappa\kappa}_{\ell}}.
+\rho^{2}_{\ell} = \sum_{i,j}\frac{C^{\kappa I^{i}}_{\ell}~(C^{II}_{\ell})^{-1}_{ij}~C^{\kappa I^{j}}_{\ell}}{C^{\kappa\kappa}_{\ell}}.
 \label{eqn:rho-combined}
@@
 However, given the low noise of these measurements and their small overlap with other probes, they can still play a major role in delensing even if their overlap with the CMB lensing potential is not optimal.
-
-% ===================
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=1.]{../images/compare_kernel.pdf}
-\caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
-The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency.
-}
-\label{fig:kernel}
-\end{center}
-\end{figure}
-% ===================
 
@@
 
-For photometric surveys, the maximum number of bins is dictated by the fact that the bin width can not be smaller than the photo-z accuracy.
+For photometric surveys, the maximum number of bins is dictated by the fact that the bin width cannot be smaller than the photo-z accuracy.
+Since we are not considering any possible photo-z bias, this is the only way photo-z uncertainties affect the delensing efficiency.
 We used 10 and 4 photometric bins for LSST and DES respectively with a photo-z accuracy of $\sigma(z)_{\textrm{DES}} = 0.05(1+z)$ for DES and $\sigma(z)_{\textrm{LSST}} = 0.01(1+z)$ for LSST.
+
 In spectroscopic surveys, there are no limitations in increasing the number of bins. We split DESI into 4 spectroscopic bins with no overlap among each others.
@@
 In the following, we sometimes refer to the fraction of lensing B-mode power removed by delensing.
-This quantity is defined as: $\left< C_{\ell}^{BB,\mathrm{lens}}/C_{{l}}^{BB,\mathrm{res}}\right>_{4<\ell<100}.$
+This quantity is defined as: $\left< C_{\ell}^{BB,\mathrm{res}}/C_{{l}}^{BB,\mathrm{lens}}\right>_{4<\ell<100}.$
 
@@
 The values in parenthesis in the first column correspond to the fraction of lensing B-mode power removed using each LSS tracer. 
-It is defined as $\left< \frac{C_{\ell}^{BB,\mathrm{lens}}}{C_{{l}}^{BB,\mathrm{res}}}\right>_{4<\ell<100}.$
-\newline
+It is defined as $\left<C_{\ell}^{BB,\mathrm{res}}/C_{l}^{BB,\mathrm{lens}}\right>_{4<\ell<100}$.
+%\newline
 The values in the other columns correspond to ratio of the error before and after delensing for 3 cases: no instrumental noise in the B-mode measurement (but with instrumental noise in the E-mode used to delens) and realistic noise for two different values of $r$.}
@@
 \end{table}
+
+% ===================
+\begin{figure*}[ht]
+\includegraphics[scale=1]{../images/errors_summary.pdf}
+\caption[]{A summary of the amount of lensing B-modes power removed by different tracers for different generations of experiments.
+
+On the \textit{left} we show the contribution from each lensing tracers.
+The cross-correlation among tracers is not considered and for this reason, the sum of all the contributions can be bigger than one.
+The purpose is to highlight the relative importance of each tracer.
+
+On the \textit{right} we show the power removed by LSS alone and final delensing efficiency once CMB internal delensing is added.
+Here the double-counting information is taken into account and the bars correspond to final delensing efficiency levels.
+}
+\label{fig}
+\end{figure*}
+%% ===================
+
 %=======================================
@@
 
-% ===================
-\newpage
-\onecolumngrid
-\begin{figure}[htbp]
-%\begin{center}
-\includegraphics[scale=1]{../images/errors_summary.pdf}
-\caption{Summary of the amount of lensing B-modes power removed by different tracers for different generation of experiments.}
-\label{fig}
-%\end{center}
-\end{figure}
-\twocolumngrid
-% ===================
-
 \bibliographystyle{mnrasunsrt}

### 2017-10-05 Notes/galaxy_delensing.tex
remove intensity mapping

@@
 The SKA is a planned radio array that will survey large scale structure primarily by detecting the redshifted neutral hydrogen (HI) 21cm emission line from a large number of galaxies out to high redshift.
-We will assume an intensity mapping survey mode where SKA will measure the large-scale fluctuations of the \textit{integrated} 21cm intensity from many unresolved galaxies.  
+We will assume a radio continuum survey mode where SKA will detect radio galaxies through their total emission out to very high redshift.  
 We model both the redshift distribution and  bias of radio sources following \cite{namikawa:2016a}. 
@@
 We checked this for a few of the tracer combinations used here and find it particularly true once several tracers are jointly taken into account.
-Finally, although we have assumed a Gaussian covariance $\sigma( C_{\ell}^{BB} )$ it actually has small non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012}.
+Finally, although we have assumed a Gaussian covariance $\sigma( C_{\ell}^{BB} )$ it actually has a non-Gaussian contribution \cite{motloch:2017,benoit-levy:2012}.
 This approximation is good enough to show the improvement due to delensing.

### 2017-10-12 Code/Plot_figures.py
new figures same colors

@@ -81,6 +81,15 @@ plt.rcParams['legend.handlelength'] = font_size / 9.
 
 BB_contr = np.genfromtxt('../Data/BB_contribution.csv', delimiter=',')
 
+# dictionary of colors
+colors = {}
+colors['CIB'] = plt.cm.Set1(0)
+colors['DES'] = plt.cm.Set1(1)
+colors['WISE'] = plt.cm.Set1(2)
+colors['DESI'] = plt.cm.Set1(3)
+colors['LSST'] = plt.cm.Set1(4)
+colors['SKA'] = plt.cm.Set1(5)
+
 
 plt.rcParams["figure.figsize"] = fig_dims
 
@@ -97,12 +106,12 @@ lbins, _, _, rho_gals_des, _ = multiple_survey_delens.main(labels, cmb)
 fg = plt.figure(figsize=fig_dims)
 ax1 = fg.add_subplot(111)
 
-plt.plot(lbins, rho['cib'], label='CIB')
-plt.plot(lbins, rho['wise'], label='WISE')
-plt.plot(lbins, rho_gals_des, label='DES')
-plt.plot(lbins, rho_cmb, label='Planck')
-plt.plot(lbins, rho_gals, label='DES+CIB+WISE (LSS S2)')
-plt.plot(lbins, rho_comb, label='LSS S2 + Planck')
+plt.plot(lbins, rho['cib'], color = colors['CIB'], label='CIB')
+plt.plot(lbins, rho['wise'], color = colors['WISE'],label='WISE')
+plt.plot(lbins, rho_gals_des,color = colors['DES'], label='DES')
+plt.plot(lbins, rho_cmb, label='Planck', color = plt.cm.Set1(6))
+plt.plot(lbins, rho_gals, label='DES+CIB+WISE (LSS S2)', color = plt.cm.Set1(7))
+plt.plot(lbins, rho_comb, label='LSS S2 + Planck', color = plt.cm.Set1(8))
 plt.plot(BB_contr[:, 0], BB_contr[:, 1] - np.min(BB_contr[:, 1]),
          '--', alpha=0.6, linewidth=font_size / 14., color='k')
 
@@ -138,12 +147,12 @@ plt.close()
 
 fg = plt.figure(figsize=fig_dims)
 
-plt.plot(lbins, rho['cib'], label='CIB')
[…]

### 2017-10-12 Code/compute_kernel_to_compare.py

@@ -136,9 +136,15 @@ dndzfun = interp1d(z_ska, dndzska01)
 norm = scipy.integrate.quad(dndzfun, z_ska[0], z_ska[-1], limit=100, epsrel=1.49e-03)[0]
 # print(norm)
 # normalize
+# dndzska01 = InterpolatedUnivariateSpline(
+#     z_ska, dndzska01 / norm * gals_kernel.dNdZ_SKA_bias(z_ska, mujk=0.1))
+
 dndzska01 = InterpolatedUnivariateSpline(
-    z_ska, dndzska01 / norm * gals_kernel.dNdZ_SKA_bias(z_ska, mujk=0.1))
+    z_ska, dndzska01 / norm)
+
+
 ska01 = gals_kernel.kern(z_ska, dndzska01, hspline, omega_m, h0, b=1.)
+# ska10 = gals_kernel.kern(z_ska, dndzska10, hspline, omega_m, h0, b=1.)
 
 
 for i, z in enumerate(z_ska):

### 2017-10-12 Code/money_plot.py

@@ -208,11 +208,11 @@ plt.text(-98, 6 - 0.2, r'SKA (10 $\mu$Jy)', horizontalalignment='left',
 plt.text(-35, 6 - 0.2, r'LSST', horizontalalignment='left',
          verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold')
 
-fg.tight_layout()
+# fg.tight_layout()
 
-# ============================================
-# FINALLY SAVE
-plt.savefig('../images/' + 'errors_summary.pdf', papertype='Letter', dpi=800)
-plt.savefig('../images/' + 'errors_summary.png', papertype='Letter', dpi=800)
-# plt.clf()
-# plt.close()
+# # ============================================
+# # FINALLY SAVE
+# plt.savefig('../images/' + 'errors_summary.pdf', papertype='Letter', dpi=800)
+# plt.savefig('../images/' + 'errors_summary.png', papertype='Letter', dpi=800)
+# # plt.clf()
+# # plt.close()

### 2017-10-12 Code/multiple_survey_delens.py

@@ -26,6 +26,7 @@ def nl(noise_uK_arcmin, fwhm_arcmin, lmax):
     return (noise_uK_arcmin * np.pi / 180. / 60.)**2 / bl(fwhm_arcmin, lmax)**2
 
 
+
 def main(labels, cmb):
 
     output_dir = '/home/manzotti/galaxies_delensing/Data/'
@@ -51,8 +52,10 @@ def main(labels, cmb):
             # noise is already where you compute spectra
 
         if label == 'cib':
-            cl_auto[label] = np.array([3400. * (1. * l / 3000.)**(-1.25) + 525 for l in lbins])
-            cl_auto[label] += 300 * cl_auto[label][100] * (lbins / lbins[100])**-4.6
+            cl_auto[label] = np.array(
+                [3400. * (1. * l / 3000.)**(-1.25) + 525 for l in lbins])
+            cl_auto[label] += 300 * cl_auto[label][100] * \
+                (lbins / lbins[100])**-4.6
 
         rho[label] = cl_cross_k[label] / np.sqrt(ckk[:] * cl_auto[label])
         if label == 'wise':
@@ -75,8 +78,10 @@ def main(labels, cmb):
             cgg[i, j, :] = np.array(cls[labels[i] + labels[j]])
 
             if (labels[i] == 'cib' and labels[j] == 'cib'):
-                cgg[i, j, :] = np.array([3500. * (1. * l / 3000.)**(-1.25) + 525 for l in lbins])
-                cgg[i, j, :] += 300 * cgg[i, j, 100] * (lbins / lbins[100])**-4.6
+                cgg[i, j, :] = np.array(
+                    [3500. * (1. * l / 3000.)**(-1.25) + 525 for l in lbins])
+                cgg[i, j, :] += 300 * cgg[i, j, 100] * \
+                    (lbins / lbins[100])**-4.6
 
             cgg[j, i, :] = cgg[i, j, :]
 
@@ -84,7 +89,8 @@ def main(labels, cmb):
 
         noise_phi = np.loadtxt('./quicklens_data/nlkk.dat')
         # noise_cmb = nl(noise, beam, lmax=4000)
-        noise_fun = interp1d(noise_phi[:, 0], noise_phi[:, 1], bounds_error=False, fill_value=1e10)
+        noise_fun = interp1d(
+            noise_phi[:, 0], noise_phi[:, 1], bounds_error=False, fill_value=1e10)
[…]

### 2017-10-12 Code/plots/plot_corr_fac.py

@@ -108,6 +108,23 @@ fg = plt.figure(figsize=fig_dims)
 
 ax1 = fg.add_subplot(111)
 
+
+# build a dictionary of colors
+
+
+colors = {}
+
+color['CIB'] = plt.cm.Set1(0)
+color['DES'] = plt.cm.Set1(1)
+color['CIB'] = plt.cm.Set1(0)
+color['CIB'] = plt.cm.Set1(0)
+color['CIB'] = plt.cm.Set1(0)
+
+
+
+
+
+
 # ax1.set_color_cycle(palettable.colorbrewer.qualitative.Dark2_8.mpl_colors)
 
 # plt.plot(ell, clkappacib / np.sqrt(clkappa * (clcib + 225.)), linestyle='--')

### 2017-10-12 Code/rho_to_Bres.py

@@ -3,7 +3,7 @@ from scipy.interpolate import RectBivariateSpline, interp1d, InterpolatedUnivari
 import sys
 import configparser as ConfigParser
 import numpy as np
-from joblib import Parallel, delayed
+from joblib import Parallel, delayed, cpu_count
 import scipy.integrate as integrate
 import profiling
 from profiling.sampling import SamplingProfiler
@@ -83,6 +83,24 @@ def compute_res_parallel(rho_filename, output_dir, clee_fun, clpp_fun, nle_fun):
     return clbb_res_ell
 
 
+def load_res(rho_filenames):
+    output_dir = '/home/manzotti/galaxies_delensing/Data/'
+    print(output_dir)
+    clbb_res_ell = []
+    lbins_int = np.loadtxt(output_dir + 'limber_spectra/cbb_res_ls.txt')
+
+    for rho_filename in rho_filenames:
+        # print('start integration')
+        if rho_filename is not 'test':
+            rho_filename = output_dir + 'limber_spectra/' + rho_filename
+            print(rho_filename)
+
+        clbb_res_ell.append(np.loadtxt(
+            rho_filename.split('.txt')[0] + 'Cbb_res.txt'))
+
+    return clbb_res_ell, lbins_int
+
+
 def compute_deriv_2(ells, clee_fun, clpp, l_phi, l_bb):
 
     clpp_mask = np.where(np.logical_and(
@@ -105,7 +123,8 @@ def compute_deriv(ells, clee_fun, clpp, l_phi, l_bb):
     # dx = 2.75
     clpp_fun_test = InterpolatedUnivariateSpline(
         ells[:5000], clpp_mask, ext='zeros')
-    clbb_der = np.array(compute_BB(clee_fun, clpp_fun_test, l_bb, l_phi))  # / dx
+    clbb_der = np.array(compute_BB(
+        clee_fun, clpp_fun_test, l_bb, l_phi))  # / dx
     return clbb_der
 
 
@@ -184,12 +203,14 @@ def compute_deriv_grid_CEE(delta_e, delta_b, n_jobs=15):
         clbb_der[i, :] = np.array(Parallel(n_jobs=n_jobs, verbose=50)(delayed(compute_deriv_CEE)(
[…]

### 2017-10-12 Code/multiple_z_med_delens.py
add Z_median plot

@@ -1,0 +1,178 @@
+'''
+This take the power spectra cross with CMB lensign and auto of different surveys ans spits out their rho as in Shwerwin Smithful
+'''
+
+import numpy as np
+# import scipy.integrate
+from scipy.interpolate import interp1d
+import pickle
+
+
+def bl(fwhm_arcmin, lmax):
+    """ returns the map-level transfer function for a symmetric Gaussian beam.
+         * fwhm_arcmin      - beam full-width-at-half-maximum (fwhm) in arcmin.
+         * lmax             - maximum multipole.
+    """
+    ls = np.arange(0, lmax + 1)
+    return np.exp(-(fwhm_arcmin * np.pi / 180. / 60.)**2 / (16. * np.log(2.)) * ls * (ls + 1.))
+
+
+def nl(noise_uK_arcmin, fwhm_arcmin, lmax):
+    """ returns the beam-deconvolved noise power spectrum in units of uK^2 for
+          * noise_uK_arcmin - map noise level in uK.arcmin
+          * fwhm_arcmin     - beam full-width-at-half-maximum (fwhm) in arcmin.
+          * lmax            - maximum multipole.
+    """
+    return (noise_uK_arcmin * np.pi / 180. / 60.)**2 / bl(fwhm_arcmin, lmax)**2
+
+
+def compute_rho_sigle_tracer(surveys, lbins, cls, ckk, output_dir, galaxies_steradians=1e15):
+    cl_cross_k = {}
+    cl_auto = {}
+    rho = {}
+
+    for label in surveys:
+        cl_cross_k[label] = np.array(cls['k' + label])
+        cl_auto[label] = np.array(cls[label + label]) + \
+            (galaxies_steradians / (0.000290888)**2)**(-1)
+        rho[label] = cl_cross_k[label] / np.sqrt(ckk[:] * cl_auto[label])
+        np.savetxt(output_dir + '/limber_spectra/rho_test_' + label + '_ngal_{:.1f}.txt'.format(galaxies_steradians),
+                   np.vstack((lbins, rho[label])).T)
+    return rho
+
+
+def compute_rho_single_tracer_plus_CMB(surveys, lbins, cls, ckk, ckk_noise, output_dir, galaxies_steradians=1e15):
+    rho_comb_dict = {}
+    # return lbins, rho
+    for label in surveys:
[…]

### 2017-10-12 Code/run_analysis_median_z_gal_num.py

@@ -1,0 +1,296 @@
+
+# # Fisher Analysis
+
+import numpy as np
+import matplotlib.pyplot as plt
+try:
+    import functools32
+except ImportError:
+    import functools as functools32
+
+import sys
+# import os
+import camb
+from camb import model, initialpower
+import multiple_z_med_delens
+import configparser as ConfigParser
+import rho_to_Bres
+from scipy.interpolate import InterpolatedUnivariateSpline
+from colorama import init
+from colorama import Fore
+
+init(autoreset=True)
+np.seterr(divide='ignore', invalid='ignore')
+# ============================================
+# SET LATEX
+plt.rc('text', usetex=True)
+plt.rc('font', **{'family': 'serif', 'serif': ['Computer Modern Roman']})
+# ============================================
+
+# cosmology values!!!!!!
+
+pars = camb.CAMBparams()
+# This function sets up CosmoMC-like settings, with one massive neutrino
+# and helium set using BBN consistency
+pars.set_cosmology(H0=70, ombh2=0.0226, omch2=0.112,
+                   mnu=0.029, omk=0, tau=0.079)
+pars.InitPower.set_params(ns=0.96, r=0., nt=0)
+pars.set_for_lmax(5000, lens_potential_accuracy=3)
+# pars.set_for_lmax?
+
+pars.AccurateBB = True
+pars.OutputNormalization = False
+pars.WantTensors = True
+pars.DoLensing = True
+pars.max_l_tensor = 3000
+pars.max_eta_k_tensor = 3000.
+
+# print(pars) # if you want to test parasm
+results = camb.get_results(pars)
+
+# Remember there is a 7.4e12 missing and CAMB always give you $ \ell (\ell +1)/2\pi  $
+
+
+@functools32.lru_cache(maxsize=64)
+def clbb(r=0.1, nt=None, lmax=3000):
+    inflation_params = initialpower.InitialPowerParams()
+    if nt is None:
+        nt = -r / 8.
+    inflation_params.set_params(As=2.1e-9, r=r, nt=nt, pivot_tensor=0.01)
[…]

### 2017-10-12 Code/Plot_figures.py
new figures same colors

@@ -7,9 +7,9 @@ import configparser as ConfigParser
 import camb
 from camb import initialpower
 try:
-    import functools32
+  import functools32
 except ImportError:
-    import functools as functools32
+  import functools as functools32
 
 
 # ============================================
@@ -43,7 +43,7 @@ plt.rcParams['font.size'] = font_size
 plt.rcParams['axes.labelsize'] = font_size * 1.3
 plt.rcParams['axes.linewidth'] = font_size / 10.
 plt.rcParams['axes.titlesize'] = font_size * 1.3
-plt.rcParams['legend.fontsize'] = font_size/ 1.2
+plt.rcParams['legend.fontsize'] = font_size / 1.2
 plt.rcParams['xtick.labelsize'] = font_size / 1.2
 plt.rcParams['ytick.labelsize'] = font_size / 1.2
 plt.rcParams['xtick.major.width'] = font_size / 10.
@@ -80,22 +80,24 @@ plt.rcParams['legend.handlelength'] = font_size / 9.
 # =============================================
 
 BB_contr = np.genfromtxt('../Data/BB_contribution.csv', delimiter=',')
-
+# colormap = plt.cm.Set1
+colormap = plt.cm.Vega10
 # dictionary of colors
 colors = {}
-colors['CIB'] = plt.cm.Set1(0)
-colors['DES'] = plt.cm.Set1(1)
-colors['WISE'] = plt.cm.Set1(2)
-colors['DESI'] = plt.cm.Set1(3)
-colors['LSST'] = plt.cm.Set1(4)
-colors['SKA'] = plt.cm.Set1(5)
+colors['CIB'] = colormap(0)
+colors['DES'] = colormap(1)
+colors['WISE'] = colormap(2)
+colors['DESI'] = colormap(3)
+colors['LSST'] = colormap(4)
+colors['SKA'] = colormap(5)
 
 
 plt.rcParams["figure.figsize"] = fig_dims
 
 labels = ['wise', 'cib', 'des_bin0', 'des_bin1', 'des_bin2', 'des_bin3']
 cmb = 'Planck'
-lbins, rho, rho_comb, rho_gals, rho_cmb = multiple_survey_delens.main(labels, cmb)
[…]

### 2017-10-12 Code/Plot_figures.py
new figures same colors

@@ -80,8 +80,8 @@ plt.rcParams['legend.handlelength'] = font_size / 9.
 # =============================================
 
 BB_contr = np.genfromtxt('../Data/BB_contribution.csv', delimiter=',')
-# colormap = plt.cm.Set1
-colormap = plt.cm.Vega10
+colormap = plt.cm.Set1
+# colormap = plt.cm.Vega10
 # dictionary of colors
 colors = {}
 colors['CIB'] = colormap(0)

### 2017-10-12 Code/money_plot.py

@@ -163,50 +163,61 @@ plt.gca().yaxis.grid(False)
 
 plt.gca().set_axisbelow(True)
 
+colormap = plt.cm.Set1
+# colormap = plt.cm.Vega10
+# dictionary of colors
+colors = {}
+colors['CIB'] = colormap(0)
+colors['DES'] = colormap(1)
+colors['WISE'] = colormap(2)
+colors['DESI'] = colormap(3)
+colors['LSST'] = colormap(4)
+colors['SKA'] = colormap(5)
+
 
 # start annotation
 
 plt.text(-47, -0.2, 'CIB', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold')
+         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', color=colors['CIB'])
 
 plt.text(-28, - 0.18, 'DES', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.8, fontweight='bold')
+         verticalalignment='bottom', fontsize=font_size / 1.8, fontweight='bold', color=colors['DES'])
 
 plt.text(-12, 0.3, 'Planck', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', rotation=90.)
+         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', rotation=90., color=colormap(6))
 
 plt.text(-5, 0.35, 'WISE', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', rotation=90.)
+         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', rotation=90., color=colors['WISE'])
 
 plt.text(-75, 2 - 0.2, 'CIB', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold')
+         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', color=colors['CIB'])
 
 plt.text(-54.5, 2 - 0.18, 'DES', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.8, fontweight='bold')
+         verticalalignment='bottom', fontsize=font_size / 1.8, fontweight='bold', color=colors['DES'])
 
 plt.text(-35, 2 - 0.2, 'SPTPol', horizontalalignment='left',
[…]

### 2017-10-12 Code/plots/plot_kernel.py

@@ -37,6 +37,16 @@ pyximport.install(reload_support=True)
 # except Exception as e:
 #     print('probably palettable not found')
 
+colormap = plt.cm.Set1
+# colormap = plt.cm.Vega10
+# dictionary of colors
+colors = {}
+colors['CIB'] = colormap(0)
+colors['DES'] = colormap(1)
+colors['WISE'] = colormap(2)
+colors['DESI'] = colormap(3)
+colors['LSST'] = colormap(4)
+colors['SKA'] = colormap(5)
 
 datadir = '../../Data/limber_spectra/'
 
@@ -141,21 +151,21 @@ print(fig_dims)
 ax1 = fg.add_subplot(111)
 
 plt.plot(kappakernel_l30[:, 0], kappakernel_l100[:, 1] /
-         np.max(kappakernel_l100[:, 1]), label=r'CMB', alpha=0.8)
+         np.max(kappakernel_l100[:, 1]), label=r'CMB', alpha=0.8, color=colormap(6))
 # plt.plot(kappakernel_l30[:, 0], kappakernel_l100[:, 1] /
 #          np.max(kappakernel_l100[:, 1]), color='#e41a1c', linestyle=':')
 # plt.plot(kappakernel_l30[:, 0], kappakernel_l500[:, 1] /
 #          np.max(kappakernel_l500[:, 1]), color='#e41a1c', linestyle='--')
 
 plt.plot(cibkernel_l100[:, 0], cibkernel_l100[:, 1] /
-         np.max(cibkernel_l100[:, 1]), label=r'CIB')
+         np.max(cibkernel_l100[:, 1]), label=r'CIB', color=colors['CIB'])
 # plt.plot(cibkernel_l30[:, 0], cibkernel_l100[:, 1] /
 #          np.max(cibkernel_l100[:, 1]), color='#377eb8', linestyle=':')
 # plt.plot(cibkernel_l30[:, 0], cibkernel_l500[:, 1] /
 #          np.max(cibkernel_l500[:, 1]), color='#377eb8', linestyle='--')
 
 plt.plot(deskernel_l30[:, 0], deskernel_l100[:, 1] /
-         np.max(deskernel_l100[:, 1]), label=r'DES', alpha=0.8)
+         np.max(deskernel_l100[:, 1]), label=r'DES', alpha=0.8, color=colors['DES'])
[…]

### 2017-10-12 Code/money_plot.py
new figures same colors

@@ -223,7 +223,7 @@ plt.text(-35, 6 - 0.2, r'LSST', horizontalalignment='left',
 
 # # ============================================
 # # FINALLY SAVE
-# plt.savefig('../images/' + 'errors_summary.pdf', papertype='Letter', dpi=800)
-# plt.savefig('../images/' + 'errors_summary.png', papertype='Letter', dpi=800)
-# # plt.clf()
-# # plt.close()
+plt.savefig('../images/' + 'errors_summary.pdf', papertype='Letter', dpi=800)
+plt.savefig('../images/' + 'errors_summary.png', papertype='Letter', dpi=800)
+# plt.clf()
+# plt.close()

### 2017-10-12 Code/money_plot.py
new figures same colors

@@ -81,6 +81,19 @@ plt.rcParams['legend.handletextpad'] = 0.3
 plt.rcParams['legend.handlelength'] = font_size / 9.
 # ============================================
 
+colormap = plt.cm.Set1
+# colormap = plt.cm.Vega10
+# dictionary of colors
+colors = {}
+colors['CIB'] = colormap(0)
+colors['DES'] = colormap(1)
+colors['WISE'] = colormap(2)
+colors['DESI'] = colormap(3)
+colors['LSST'] = colormap(4)
+colors['SKA'] = colormap(5)
+
+
+
 plt.close()
 plt.close()
 
@@ -163,66 +176,55 @@ plt.gca().yaxis.grid(False)
 
 plt.gca().set_axisbelow(True)
 
-colormap = plt.cm.Set1
-# colormap = plt.cm.Vega10
-# dictionary of colors
-colors = {}
-colors['CIB'] = colormap(0)
-colors['DES'] = colormap(1)
-colors['WISE'] = colormap(2)
-colors['DESI'] = colormap(3)
-colors['LSST'] = colormap(4)
-colors['SKA'] = colormap(5)
-
 
 # start annotation
 
 plt.text(-47, -0.2, 'CIB', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', color=colors['CIB'])
+         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold')
 
 plt.text(-28, - 0.18, 'DES', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.8, fontweight='bold', color=colors['DES'])
+         verticalalignment='bottom', fontsize=font_size / 1.8, fontweight='bold')
 
 plt.text(-12, 0.3, 'Planck', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', rotation=90., color=colormap(6))
+         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', rotation=90.)
 
 plt.text(-5, 0.35, 'WISE', horizontalalignment='left',
-         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', rotation=90., color=colors['WISE'])
+         verticalalignment='bottom', fontsize=font_size / 1.5, fontweight='bold', rotation=90.)
 
 plt.text(-75, 2 - 0.2, 'CIB', horizontalalignment='left',
[…]

### 2017-10-12 Notes/galaxy_delensing.tex
explicitly say about f_sky
better definition for current

@@
 Among currently available surveys, a DES-like galaxy survey can remove about $14\%$ of the lensing signal.
-Ongoing CMB experiments (CMB-S2) will particularly benefit from large scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
+Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
 With the decrease of instrumental noise, the CMB internal reconstruction will increase its efficiency and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level.
@@
 These will translate into a better reconstruction of the B-modes in the measured patch, and, as a consequence will improve the constraints on inflation through delensing.
-We model several actual and future surveys, and after computing the residual B-modes, we forecast the resulting statistical uncertainties on the amplitude and the shape of the inflationary tensor perturbations for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).
+We model several actual and future surveys, and after computing the residual B-modes, we forecast the resulting statistical uncertainties on the amplitude and the shape of the inflationary tensor perturbations for ongoing or recently completed CMB experiments (S2) as well as the next generation (S3) and the planned fourth generation (S4).
 
@@
 
-Even if there is not a definite distinction between different stages, we focus on three distinct scenarios: the current stage, a third generation stage (3G) and finally the futuristic CMB Stage 4.
+Even if there is not a definite distinction between different stages, we focus on three distinct scenarios: the ongoing or recently concluded stage (S2), a third generation stage (3G) of experiments in an advanced building phase or that have just started taking data and finally the futuristic CMB Stage 4.
 
@@
 The improvement is then the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$. 
+The ratio $\alpha_{r,n_{t}}$ is insensitive to the fraction of the sky observed by the experiment: while the absolute parameter constraints depend on $\fsky$ the relative improvement does not.
+For this reason, we do not quote $\fsky$ values for the CMB experiments investigated here.
 
@@
 %===================
-\subsection{Delensing with current CMB and LSS}
+\subsection{Delensing with CMB Stage-2 and current LSS}
 \label{sec:del_res}
@@
 Its correlation is comparable to the one of LSS at almost all scales.
-On the other end, the CMB lensing reconstruction from SPT-Pol will be more than 85\% correlated with the true field at $\ell<300$, and only then the correlation goes below the level of current LSS surveys.
+On the other end, the CMB lensing reconstruction from SPT-Pol will be more than \textcolor{blue}{70\%} correlated with the true field at $\ell<300$, and only then the correlation goes below the level of current LSS surveys.
 In both cases the CMB lensing reconstruction correlates very well at low $\ell$ and it then falls rapidly at smaller scales because of the raise of the reconstruction noise.
@@
 }
-\label{fig}
+\label{fig:recap}
 \end{figure*}
@@
 
+A summary of the delensing efficiency for different lensing tracers is shown in \reffig{recap}.
 For ongoing experiments, we find that LSS tracers will be particularly beneficial. 

### 2017-10-18 Code/rho_to_Bres.py
new figures same colors

@@ -7,6 +7,7 @@ from joblib import Parallel, delayed, cpu_count
 import scipy.integrate as integrate
 import profiling
 from profiling.sampling import SamplingProfiler
+import time
 
 
 def bl(fwhm_arcmin, lmax):
@@ -51,10 +52,12 @@ def compute_BB(clee_fun, clpp_fun, ell_b_bin, ell_phi_bin):
 
 
 def compute_res_parallel(rho_filename, output_dir, clee_fun, clpp_fun, nle_fun):
+
     print(output_dir)
     # print('start integration')
+
     if rho_filename == 'test':
-        lmax = 3000
+        lmax = 2000
 
         def integrand(theta, ell, L):
             clee = clee_fun(ell)
@@ -72,11 +75,13 @@ def compute_res_parallel(rho_filename, output_dir, clee_fun, clpp_fun, nle_fun):
             clee = clee_fun(ell)
             return (ell / (2. * np.pi)**2 * (L * ell * np.cos(theta) - ell**2)**2 * clpp_fun(np.sqrt(L**2 + ell**2 - 2. * ell * L * np.cos(theta))) * clee * (np.sin(2. * theta))**2) * (1. - (clee / (clee + nle_fun(ell))) * rho_fun(ell) ** 2)
 
-    lbins_int = np.linspace(10, 2500, 80)
-
-    clbb_res_ell = [integrate.dblquad(
-        integrand, 8, lmax, lambda x: 0, lambda x: 2. * np.pi, args=(L,), epsabs=0., epsrel=1.49e-05)[0] for L in lbins_int]
-
+    lbins_int = np.linspace(10, 2000, 50)
+    options1 = {'limit': 500, 'epsabs': 0., 'epsrel': 1.e-3}
+    options2 = {'limit': 40, 'epsabs': 0., 'epsrel': 1.e-3}
+    tic =time.time()
+    clbb_res_ell = [integrate.nquad(
+        integrand, [[0., 2. * np.pi], [8, lmax]], args=(L,), opts=[options1, options2])[0] for L in lbins_int]
+    print('time for the integral total',tic -time.time())
     np.savetxt(rho_filename.split('.txt')[0] + 'Cbb_res.txt', clbb_res_ell)
     np.savetxt(output_dir + 'limber_spectra/cbb_res_ls.txt', lbins_int)
 

### 2017-10-18 Code/run_analysis.py

@@ -34,7 +34,7 @@ pars = camb.CAMBparams()
 # and helium set using BBN consistency
 pars.set_cosmology(H0=70, ombh2=0.0226, omch2=0.112,
                    mnu=0.029, omk=0, tau=0.079)
-pars.InitPower.set_params(ns=0.96, r=0., nt=0)
+pars.InitPower.set_params(ns=0.96, r=0., nt=0, pivot_tensor=0.01, As=2.1e-9)
 pars.set_for_lmax(5000, lens_potential_accuracy=3)
 # pars.set_for_lmax?
 
@@ -47,9 +47,12 @@ pars.max_eta_k_tensor = 3000.
 
 # print(pars) # if you want to test parasm
 results = camb.get_results(pars)
-
+powers = results.get_cmb_power_spectra(pars)
 # Remember there is a 7.4e12 missing and CAMB always give you $ \ell (\ell +1)/2\pi  $
-
+totCL = powers['total']
+cle = totCL[:, 1]
+clp = powers['lens_potential'][:, 0]
+ells_cmb = np.arange(0,len(clp))
 
 @functools32.lru_cache(maxsize=64)
 def clbb(r=0.1, nt=None, lmax=3000):
@@ -106,6 +109,7 @@ def fisher_r_nt(r_fid=0.2, fid=None,
                 fsky=0.5
                 ):
 
+    print('noise', noise_uK_arcmin, 'beam=', fwhm_arcmin)
     nlb = nl(noise_uK_arcmin, fwhm_arcmin, lmax=lmax)
     ell_nlb = np.arange(0, len(nlb))
     nlb = nlb * ell_nlb * (ell_nlb + 1.) / 2. / np.pi
@@ -189,10 +193,10 @@ rho_names = ['rho_cib.txt', 'rho_des.txt', 'rho_gals.txt',
 noise_uK_arcmin = 3.
 fwhm_arcmin = 30.
 # not used right now
-ell_range_deep = [20, 400]
-ell_range_high = [200, ells_cmb[-1]]
-nle_deep = nl(noise_uK_arcmin, fwhm_arcmin, lmax=ells_cmb[-1])[2:]
-nle_high = nl(9., 1., lmax=ells_cmb[-1])[2:]
+# ell_range_deep = [20, 400]
+# ell_range_high = [200, ells_cmb[-1]]
[…]

### 2017-10-18 Notes/galaxy_delensing.tex

@@
 On the CMB side, we will combine a BICEP-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment.
-For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin, a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
+For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt(2)$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
 We assume the CMB lensing reconstruction is  performed by the higher resolution CMB experiment. We explore two possibilities.
@@
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
-Then, we test a second scenario with a ground-based experiment characterized by noise levels consistent with SPT-Pol \cite{henning:2017}: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin.
+Then, we test a second scenario with a ground-based experiment characterized by noise levels in polarization consistent with SPT-Pol \cite{henning:2017}: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin.
 In this case we set the angular scale range at $50<\ell<4000$.
@@
 CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
-As we did for Stage-2, we model the CMB S3 as two different overlapping experiments that will be able to combine their measurement.
-The two experiments combined will have a level of noise of $3\mu$K-arcmin and a 1 arcmin beam. This level of noise is also assumed for the internal noise reconstruction.
+As we did for Stage-2, we model the CMB S3 as two different overlapping experiments.
+For the deep experiment we assume an instrumental noise equal to $2\mu$K-arcmin in polarization and a beam of 30 arcmin.
+The high resolution experiment will have a level of noise in polarization of $3\mu$K-arcmin and a 1 arcmin beam. This level of noise is also assumed for the internal noise reconstruction.
 We still use an angular scale range in the Fisher matrix of $50<\ell<4000$ even if most of the high $\ell$ scale do not contribute to the constraints.
@@
 Given the unprecedented low level of noise of these experiment, delensing will be even more important than in previous generations.
-Following \cite{abazajian:2016}, here we assume a CMB-S4 ground experiment with a level of noise of $1\mu$K-arcmin and a 1 arcmin beam.
+Following the B-mode constraints section in \cite{abazajian:2016}, here we assume a CMB-S4 ground experiment composed of two different telescopes.
+For the deep experiment, we assume an instrumental noise equal to $1\mu$K-arcmin in polarization and a beam of 15 arcmin.
+The high-resolution experiment will have a level of noise in polarization of $1.5\mu$K-arcmin and a 1 arcmin beam.
+
 Furthermore, in the next decade, several next-generation LSS surveys will be online. 
@@
 \section{Acknowledgments}
-We thank S. Dodelson, S. Passaglia, K. Story, A. Vieregg and K. Wu for useful discussion and their review of the draft. 
+We thank T. Crawford, S. Dodelson, Gil Holder, T. Namikawa, S. Passaglia, K. Story, A. Vieregg and K. Wu for useful discussion and their review of the draft. 
 Furthermore we thank D. Alonso for his feedback on radio continuum surveys.

### 2017-10-26 Notes/galaxy_delensing.tex
more text review also after Giulio input

@@
 Even if inflationary B-modes have not been detected yet, the target value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski:2016,abazajian:2016}.
-However, just reducing the level of noise will not be enough to attain this goal.
+However, just reducing the level of noise will not be enough to reach this target.
 Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe. 
@@
 
-The optimal way to reduce the lensing contributions to the B-modes is to reconstruct a template of the actual lensing B modes (or the actual large scale structure lensing potential) on the observed part of the sky and then use it to clean the data in a process called ``delensing''.
+The optimal way to reduce the lensing contributions to the B-modes is to reconstruct a template of the actual lensing B modes (or the actual large scale structure lensing potential) on the observed part of the sky and then use it to marginalize the lensing contribution in the data in a process called ``delensing''.
 We can delens the observed B-modes by combining CMB polarization data (what is \textit{lensed}) with tracers of the large scale structure (what is \textit{lensing}) to reconstruct a template of the expected lensing B-modes.
@@
 In this section, we introduce the different large scale structure tracers considered in this work to reconstruct the lensing potential.
-Gravitational lensing distorts primordial E-modes generating a non-primordial B-mode component.
+We will see how large scale structures distorts primordial E-modes generating a non-primordial B-mode component in the next section.
 Also, we define the power spectra that we will use later in \refsec{th}.
@@
 \begin{equation}
-C_{\ell}^{ij}= \int_0^{\infty} \frac{dz}{c} \frac{H(z)}{\chi(z)^2}\, W^{i}(z)W^{j}(z)P(k,z).
+C_{\ell}^{ij}= \int_0^{\infty} \frac{dz}{c} \frac{H(z)}{\chi(z)^2}\, W^{i}(z)W^{j}(z)P(k = \ell/\chi(z),z).
 \label{eqn:wkappa}{}
@@
 
-In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, and $P(k,z)$ is the matter power spectrum evaluated at wavenumber $k = \ell/\chi(z)$ and redshift $z$. 
+In this equation, $H(z)$ is the Hubble factor at redshift $z$, $c$ is the speed of light, and $P(k,z)$ is the matter power spectrum evaluated at redshift $z$. 
 %Furthermore $W^{i}(z)$ is the kernel function of the field $i$ introduced in \refeq{wkernel}.
@@
 
-The galaxy clustering kernel is:
+The normalized galaxy clustering kernel is:
 \begin{equation}
@@
 To do so, we add a constant term to the power spectrum equal to the inverse of the number of galaxies per steradians.
- Different galaxy surveys in this work are then fully characterized by their b(z), $\frac{dN}{dz}$ and the observed galaxy density.
+ Different galaxy surveys in this work are then fully characterized by their $b(z)$, $\frac{dN}{dz}$ and the observed galaxy density.
 We test the delensing efficiency taking into account both current surveys like WISE or DES as well as future galaxy surveys like DESI and LSST as well as 21 cm measurement like SKA.
@@
 There are several available CIB observations that have already been used to delens the CMB.
-Given its full-sky coverage, a promising one is the CIB map derived in \cite{2016A&A...596A.109P} using multi-frequency Planck data and the Generalized Needlet Internal Linear Combination (GNILC) component separation algorithm.
+Given its full-sky coverage and the small contribution from foregrounds contamination, a promising CIB map for future experiments is the one derived in \cite{2016A&A...596A.109P} using multi-frequency Planck data and the Generalized Needlet Internal Linear Combination (GNILC) component separation algorithm.
 
@@
 As shown in \cite{seljak:1997}, because of the symmetry of the problem, tensor perturbations are the principal source of the B-modes configuration.
-For this reason, B-modes are a promising signature of early universe tensor perturbations.
+For this reason, primordial B-modes are a promising signature of early universe tensor perturbations.
 
-However, this promising sign of primordial gravitational waves is partially obscured by gravitational interactions with large scale structures that generate CMB B-modes by distorting primordial E-modes.
+However, primordial B-modes are obscured by gravitational interactions between the large scale structures and the CMB that generate CMB B-modes by distorting primordial E-modes.
 At first order, given the convergence field $\kappa= -\frac{1}{2}\nabla^2\phi$ introduced in \refsec{kappaCMB} the B-modes resulting from the lensing of the primordial E-modes are:
@@
 The lensing component is a significant source of B-modes that, at large scales, corresponds to a white noise source of roughly $5 \mu K$-arcmin independent of the angular scale.
-This means that it is not only larger than the allowed inflationary component at scales smaller than several degrees, but it is also comparable to current levels of instrumental noise.
+This means that it is not only larger than the allowed inflationary component at scales smaller than several degrees ($r_{0.05} < 0.12$ from \cite{bicep2/keck-collaboration:2015}), but it is also comparable to current levels of instrumental noise.
 For this reason, it is critical to characterize and eventually remove it from the data.
-To do so, we build a template \refeq{blens} of the lensing B-modes in the observed patch given a measurement of the E-mode field and the lensing potential $\phi$.
+To do so, while other approaches are possible (\cite{larsen:2016a,carron:2017,millea:2017}), here we assume a "template approach": we build a template \refeq{blens} of the lensing B-modes in the observed patch given a measurement of the E-mode field and the lensing potential $\phi$.
 While E is measured directly, we can estimate $\phi$ using "tracers" of the matter distribution that sources the potential.
@@
 
-The optimal weights $f(\bl,\bl')$, chosen such that the residual lensing B mode power is minimized, are:
+The optimal weights $f(\bl,\bl')$, chosen such that the residual lensing B mode power is minimized, are \cite{sherwin:2015}:
 \be
@@
 The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency.
-}
+Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not simply to the redshift distribution of galaxies.}
 \label{fig:kernel}
@@
 We start by assuming that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
-We can optimally combine them to estimate $\phi$ or, in other words, to maximize the correlation factor $\rho$ with:
+We can optimally combine them to estimate $\phi$ or, in other words, to maximize the correlation factor $\rho$ with \cite{sherwin:2015}:
 \beqn
@@
 \eeqn
-where $C_{II}$ is the covariance matrix of the LSS tracers.
+where $C_{II}$ is the covariance matrix of the LSS tracers. This is assumed to be gaussian and it is computed using \refeq{wkappa}.
 The residual B-mode power can be derived from \refeq{Bres} using an ``effective'' correlation $\rho^{2}$ of these combined tracers with gravitational lensing:
@@
 The cross-correlation of a tracer with the CMB lensing is directly proportional to the overlap of their kernels.
-The cosmic infrared background and 21 cm surveys probe the high redshift structures and, independent of the model assumed, they show a relatively good overlap with the CMB lensing kernel. 
+It can be seen, for example, that the cosmic infrared background and 21 cm surveys probe the high redshift structures and, independent of the model assumed, they show a relatively good overlap with the CMB lensing kernel.
 On the other end, galaxy clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
@@
 When we weight a tracer with $\frac{C^{\kappa I}}{C^{II}}$ in \refeq{fweight} in order to maximize its ability to reconstruct the lensing potential we are only using redshift averaged information about the survey.
-However, as can be seen in \reffig{kernel}, the kernel overlap of a tracer with the CMB lensing varies as a function of redshift.
+However, as can be seen in \reffig{kernel}, the kernel overlap of a tracer with the CMB lensing varies as a function of redshift and the latter decreases steeply at low redshift.
 For this reason, the optimal approach is to weight galaxies at different redshift with different weights according to both their cross-correlation with $\kappa$ and their auto-spectrum.
@@
 In this work, we bin both photometric and spectroscopic galaxy surveys by splitting the window function \refeq{wg} into different slices such as all the bins contain the same number of galaxies.
-For photometric surveys like DES and LSST, we assume a photometric redshift estimation gaussianly distributed around the true value with an rms fluctuation $\sigma(z)$.
+For photometric surveys like DES and LSST, we assume a photometric redshift estimation Gaussianly distributed around the true value with an rms fluctuation $\sigma(z)$.
 
@@
 To quantify the improvement after delensing we introduce the error on a parameter $p$ constrained with \textit{delensed} spectra as $\sigma^{\rm{del}}(p)$.
-This is computed from \refeq{fisher} substituting $C_{{l}}^{BB,\mathrm{lens}}$ with the residual power after delensing $C_{{l}}^{BB,\mathrm{res}}$ in the covariance matrix defined in \refeq{cova}.
+Motivated by \cite{namikawa:2015} we assume the likelihood after delensing is well approximated by a multivariate gaussian.
+Then we compute $\sigma^{\rm{del}}(p)$ from \refeq{fisher} substituting $C_{{l}}^{BB,\mathrm{lens}}$ with the residual power after delensing $C_{{l}}^{BB,\mathrm{res}}$ in the covariance matrix defined in \refeq{cova}.
 The improvement is then the ratio of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$. 
@@
 Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-WISE (6\%)& 1.09& 1.06 & 1.02 \\ \hline
-DES (11\%) &1.16 & 1.10  & 1.04 \\ \hline
-CIB (25\%) &1.37& 1.21  & 1.09\\ \hline
-LSS-S2 (34\%) &1.53& 1.29  & 1.11\\ \hline
-CMB Planck (8\%) &1.08& 1.05  & 1.02 \\ \hline
-LSS-S2 +CMB (37\%)& 1.58& 1.33& 1.13  \\ \hline
+WISE (7\%)& 1.09& 1.06 & 1.02 \\ \hline
+DES (15\%) &1.19 & 1.12  & 1.04 \\ \hline
+CIB (31\%) &1.48& 1.28  & 1.08\\ \hline
+LSS-S2 (41\%) &1.76& 1.41  & 1.11\\ \hline
+CMB Planck (7\%) &1.08 & 1.05  & 1.02 \\ \hline
+LSS-S2 +CMB (44\%)& 1.83& 1.44& 1.13  \\ \hline
 \end{tabular}
@@
 Surveys &$\alpha_{r=0}, N^{B}_{\ell}=0$& $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-SPTPol (35\%) &1.46& 1.27  & 1.08 \\ \hline
-LSS-S2 +SPTPol (58\%) & 2.18 & 1.57 & 1.16  \\ \hline
+SPTPol (31\%) &1.46& 1.27  & 1.08 \\ \hline
+LSS-S2 +SPTPol (54\%) & 2.24 & 1.59 & 1.14  \\ \hline
 
@@
 DESI (10\%) &1.12 & 1.08  & 1.03 \\ \hline
-LSS-S3 (40\%) &1.73& 1.42  & 1.2\\ \hline
-CMB 3G (56\%) &2& 1.57  & 1.21 \\ \hline
-LSS-S3+CMB (69\%)& 2.8 & 1.8& 1.31  \\ \hline
+LSS-S3 (43\%) &1.82& 1.42  & 1.12\\ \hline
+CMB 3G (56\%) &2.38& 1.9  & 1.16 \\ \hline
+LSS-S3+CMB (69\%)& 3.2 & 2.3& 1.21  \\ \hline
 \end{tabular}
@@
 LSS-S4  (60\%) &2.5 & 2.3  & 2.3 \\ \hline
-SKA (10$\mu$Jy) (53\%) &2.1& 2  & 2\\ \hline
-SKA (0.1$\mu$Jy) (77\%) &4.6& 3.9  & 3.9\\ \hline
+SKA (10$\mu$Jy) (53\%) &2.1& 1.9  & 1.16\\ \hline
 LSS-S4 + SKA (10$\mu$Jy)  (65\%) &2.7 & 2.5  & 2.5  \\ \hline
-LSS-S4 + SKA (0.1$\mu$Jy)  (80\%) &4.6 & 3.9  & 3.9 \\ \hline
 CMB S4 (81\%) & 5.1 & 4.3  & 4.2 \\ \hline
@@
 \section{Acknowledgments}
-We thank T. Crawford, S. Dodelson, Gil Holder, T. Namikawa, S. Passaglia, K. Story, A. Vieregg and K. Wu for useful discussion and their review of the draft. 
+We thank T. Crawford, S. Dodelson, G. Fabbian, G. Holder, T. Namikawa, S. Passaglia, K. Story, A. Vieregg and K. Wu for useful discussions. 
 Furthermore we thank D. Alonso for his feedback on radio continuum surveys.

### 2017-10-27 Notes/galaxy_delensing.tex

@@
 
-Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} and internal CMB lensing maps \cite{carron:2017} as large scale structure tracers.
+Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} and internally reconstructed CMB lensing maps \cite{carron:2017} as large scale structure tracers.
 Here we discuss the improvement that can be obtained by combining these and other currently Stage-2 available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
@@
 For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt(2)$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
-We assume the CMB lensing reconstruction is performed by the higher resolution CMB experiment. We explore two possibilities.
+We assume that the CMB lensing reconstruction is performed by the higher resolution CMB experiment. We explore two possibilities.
 First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck satellite \cite{planck-collaboration:2016a}.
-For the noise in the CMB lensing map in \refeq{noisekappa} we use the actual noise curves publicly available
+In this case, for the noise in the CMB lensing map in \refeq{noisekappa} we use the actual noise curves publicly available
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
@@
 However, the results are quite robust against this choice.
-%since the neglected scales contribute only moderately to the reconstruction of the CMB lensing potential.
 
@@
 The achievable correlation is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
-In these figures we include a dashed curve that corresponds to (with arbitrary scale) $<C_{\ell}^{\kappa \kappa} \times \frac{\partial C^{BB}_{\ell'}}{\partial{C^{\kappa \kappa}_{\ell}}}>_{\ell ' <100}$, in order to show which scales in $\kappa$ contribute to the $\ell<100$ B-mode power.
+In these figures, in order to show which scales in $\kappa$ contribute to the $\ell<100$ B-mode power, we include a dashed curve that corresponds to (with arbitrary scale) $<C_{\ell}^{\kappa \kappa} \times \frac{\partial C^{BB}_{\ell'}}{\partial{C^{\kappa \kappa}_{\ell}}}>_{\ell ' <100}$.
 
@@
 
-The lensing reconstruction noise in Planck is such that internal delensing can only remove \textcolor{blue}{$8\%$} of the power, improving the constraints on $r$ by a factor of \textcolor{blue}{1.06} even in the ideal scenario of no instrumental noise in the B-modes.
-A similar performance is achieved with WISE.
-DES can do better, removing a level of \textcolor{blue}{11\%} of power.
+The lensing reconstruction noise in Planck is such that internal delensing can only remove $7\%$ of the power, improving the constraints on $r$ by a factor of 1.06 even in the ideal scenario of no instrumental noise in the B-modes.
+A slightly better performance is achieved with WISE.
+DES can do better, once binned in redshift, removing a level of $17\%$ of power.
 However, it covers a smaller fraction of the sky than the two previous probes.
-As shown in previous works \cite{sherwin:2015}, the best tracer is the CIB with a \textcolor{blue}{27\%} reduction in power.
+As shown in previous works \cite{sherwin:2015}, the best tracer is the CIB with a $30\%$ reduction in power.
 Nonetheless, it is worthwhile combining all the tracers.
-Indeed a multi tracer-approach can bring the removed power on the overlapping area from \textcolor{blue}{27\%} of the CIB alone to \textcolor{blue}{36\%} using all the LSS (LSS-S2) and \textcolor{blue}{41\%} when Planck lensing reconstruction is added to the set.
-However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 30\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
+Indeed a multi tracer-approach can bring the removed power on the overlapping area from $30\%$ of the CIB alone to $43\%$ using all the LSS (LSS-S2) and $45\%$ when Planck lensing reconstruction is added to the set.
+However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 45\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
 Furthermore, as explained in \refsec{fisher}, this is only an upper limit and will be even smaller when, for example, non-perfect foreground cleaning is considered.
@@
 If, on the other hand, we consider a high-resolution ground base experiment the internal lensing reconstruction improves significantly. 
-\reftab{current-spt} shows that, with an instrumental noise at the expected level for SPTPol, the CMB will soon be able to remove an amount of power \textcolor{blue}{35\%} at the level of the CIB.
-Once combined with LSS tracer this will lead to removing \textcolor{blue}{56\%} of the power.
-The improvement of \textcolor{blue}{1.57} in $\sigma(r)$ that will follow, even when realistic noise is considered, is such that it will make delensing a needed step for CMB polarization data analysis.
+\reftab{current-spt} shows that, with an instrumental noise at the expected level for SPTPol, the CMB will soon be able to remove an amount of power $31\%$ at the level of the CIB.
+Once combined with LSS tracer this will lead to removing $55\%$ of the power.
+The improvement of 1.6 in $\sigma(r)$ that will follow, even when realistic noise is considered, is such that it will make delensing a needed step for CMB polarization data analysis.
 
@@
 %\newline
-The values in the other columns correspond to ratio of the error before and after delensing for 3 cases: no instrumental noise in the B-mode measurement (but with instrumental noise in the E-mode used to delens) and realistic noise for two different values of $r$.}
+The values in the other columns correspond to ratio of the error before and after delensing for 3 cases: no instrumental noise in the B-mode measurement (but with instrumental noise in the E-mode used to delens) and with instrumental noise for two different values of $r$.}
   \vspace{0.2cm}
@@
 Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-WISE (7\%)& 1.09& 1.06 & 1.02 \\ \hline
-DES (15\%) &1.19 & 1.12  & 1.04 \\ \hline
-CIB (31\%) &1.48& 1.28  & 1.08\\ \hline
-LSS-S2 (41\%) &1.76& 1.41  & 1.11\\ \hline
-CMB Planck (7\%) &1.08 & 1.05  & 1.02 \\ \hline
-LSS-S2 +CMB (44\%)& 1.83& 1.44& 1.13  \\ \hline
+WISE (8\%)& 1.08& 1.05 & 1.02 \\ \hline
+DES (17\%) &1.20 & 1.13  & 1.05 \\ \hline
+CIB (30\%) &1.44& 1.26  & 1.08\\ \hline
+LSS-S2 (42\%) &1.75& 1.42  & 1.11\\ \hline
+CMB Planck (7\%) &1.06 & 1.04  & 1.02 \\ \hline
+LSS-S2 +CMB (45\%)& 1.81& 1.45& 1.13  \\ \hline
 \end{tabular}
@@
 Surveys &$\alpha_{r=0}, N^{B}_{\ell}=0$& $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-SPTPol (31\%) &1.46& 1.27  & 1.08 \\ \hline
-LSS-S2 +SPTPol (54\%) & 2.24 & 1.59 & 1.14  \\ \hline
+SPTPol (31\%) &1.43& 1.26  & 1.08 \\ \hline
+LSS-S2 +SPTPol (55\%) & 2.2 & 1.6 & 1.14  \\ \hline
 
@@
 The high resolution experiment will have a level of noise in polarization of $3\mu$K-arcmin and a 1 arcmin beam. This level of noise is also assumed for the internal noise reconstruction.
-We still use an angular scale range in the Fisher matrix of $50<\ell<4000$ even if most of the high $\ell$ scale do not contribute to the constraints.
+We use an angular scale range in the Fisher matrix of $50<\ell<800$ even if most of the high $\ell$ scale do not contribute to the constraints.
 Furthermore, around 2019, DESI will start taking data. For this reason, we add DESI to the LSS tracers used in \refsec{del_res}.
@@
 The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
-An interesting finding is that DESI will be less efficient (removing \textcolor{blue}{9\%} of the power) than a DES-like survey (\textcolor{blue}{11\%}) despite the fact that it can probe slightly higher redshift.
+An interesting finding is that DESI will be less efficient (removing $11\%$ of the power) than a DES-like survey ($17\%$) despite the fact that it can probe slightly higher redshift.
 The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency.
-Unfortunately, adding DESI will only bring the power removed using LSS from \textcolor{blue}{36 to 41\%}.
-Indeed, even in the near future, CIB will still play the dominant role among LSS.
-Contrary to current experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 550$.
+Unfortunately, adding DESI will only bring the power removed using LSS from 41 to 45\%.
+Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
 
@@
 We summarize the improvement in the constraint on $r$ in \reftab{S3}. 
-Given the improvement in the noise of the high-resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of \textcolor{blue}{2} through internal delensing in the ideal case of no instrumental noise in the B-modes \textcolor{red}{(but with instrumental noise in the E-mode used to delens)}.
+Given the improvement in the noise of the high-resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of $\sim 2.4$ through internal delensing in the ideal case of no instrumental noise in the B-modes (but with instrumental noise in the E-mode used to delens).
 Adding galaxy surveys will lead to a further improvement without additional effort.
-Indeed galaxy surveys will still be able to remove an additional \textcolor{blue}{$20\%$} of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by \textcolor{blue}{$15\%$}.
+Indeed galaxy surveys will still be able to remove an additional $11\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by $34\%$.
 Note that apart from the addition of DESI, the tracers used here are already available today.
 The ongoing effort can then lead to a significant improvement even for the next generation of experiments.
-Finally, the combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor \textcolor{blue}{1.8} or to improve the constraint of a possible detection (r=0.12) by \textcolor{blue}{30\%}.
+Finally, the combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor \textcolor{blue}{1.8} or to improve the constraint of a possible detection (r=0.07) by 34\%.
 As expected the improvement on $n_{t}$ is similar to the one in $r$ since they are both proportional to the variance of the measured B-modes. 
 We tested this in the scenario $r=0.12$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}}/8$. 
-It is important to note that, independent of the level of delensing, for CMB 3G experiments the statistical error on $n_{t}$ will still be several times bigger than the fiducial value.
+It is important to note that, independent of the level of delensing, for CMB 3G experiments the statistical error on $n_{t}$ will still be several times bigger than the fiducial value but it can still be tight enough to constrain more exotic inflationary models.
 
@@
 Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-DESI (10\%) &1.12 & 1.08  & 1.03 \\ \hline
-LSS-S3 (43\%) &1.82& 1.42  & 1.12\\ \hline
-CMB 3G (56\%) &2.38& 1.9  & 1.16 \\ \hline
-LSS-S3+CMB (69\%)& 3.2 & 2.3& 1.21  \\ \hline
+DESI (11\%) &1.13 & 1.08  & 1.03 \\ \hline
+LSS-S3 (45\%) &1.82& 1.42  & 1.12\\ \hline
+CMB 3G (56\%) &2.38& 1.7  & 1.16 \\ \hline
+LSS-S3+CMB (68\%)& 3.2 & 2.3& 1.21  \\ \hline
 \end{tabular}
@@
 Here, as an example of future optical galaxy surveys, we add LSST \cite{lsst-science-collaboration:2009} to the CMB lensing tracers. 
-We do not consider other surveys but we test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance. 
+We do not show results for other surveys here but we test that other experiments like Euclid \cite{laureijs:2011} and WFIRST \cite{spergel:2013} have similar delensing performance. 
 We also consider a radio continuum survey modeled following the SKA specifications.
@@
 However, this technique has the potential to map the LSS with relatively low noise up to redshift $z=6$.
-Here we considered just two possible detection thresholds at 1 GHz (flux cut): 10 $\mu$Jy and 0.1 $\mu$Jy, which are representatives of the upper and lower limit levels for a SKA phase 1 (SKA1) and a following SKA phase 2 (SKA2) survey.
+Here we considered a detection thresholds at 1 GHz (flux cut) of 10 $\mu$Jy which should be representative of SKA phase 1 (SKA1) \cite{namikawa:2016a}.
 Given the importance of delensing for Stage-4 experiments, here we combine radio and optical survey to test their delensing efficiency.
@@
 \end{figure}
-%%=======
-%\begin{figure}[htbp]
-%\begin{center}
-%\includegraphics[scale=1.]{../images/S4_scenario_ska01.pdf}
-%\caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
-This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstrcution and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
-}
-%\label{fig:corrs4_01}
-%\end{center}
-%\end{figure}
-%%=======
+
 A tomographically binned LSST-like experiment will be a very efficient CMB lensing tracer.
@@
 Having a perfect kernel overlap with the true lensing potential it will benefit from the very low level of noise of CMB S4 experiments.
-\reffig{corrs4_10} shows that a CMB-S4 experiment will be able to internally reconstruct the lensing potential at more than 90\% up to $\ell=400$.
+\reffig{corrs4_10} shows that a CMB-S4 experiment will be able to internally reconstruct the lensing potential at more than 90\% up to $\ell=450$.
 At this level, CMB S4 internal delensing will not be limited by noise but by small secondary effects like foregrounds contamination, filtered modes etc.
-Finally, \reftab{S4} shows that a very futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
+%Finally, \reftab{S4} shows that a very futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
 This performance is comparable to a CMB-S4 internal reconstruction.
@@
 \reftab{S4} shows the improvement on the inflationary constraint given the cross-correlation factors described above. 
-LSST alone will be able to remove almost \textcolor{blue}{half} of the B-mode lensing power.
-This will lead to an improvement in $\sigma(r)$ of a factor of \textcolor{blue}{1.8}.
+LSST alone will be able to remove almost half of the B-mode lensing power.
+This will lead to an improvement in $\sigma(r)$ of a factor of 2.
 If we add all the other infrared and optical LSS tracers previously included (i.e.
-excluding SKA) the improvement factor increase to \textcolor{blue}{2.3}.
+excluding SKA) the improvement factor increase to 2.5.
 These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSS, completely independent of CMB internal data.
-Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove \textcolor{blue}{53\%} of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS \textcolor{blue}{(62\%)} to surpass the level of the CMB S3 internal reconstruction.
+Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of the CMB S3 internal reconstruction.
 Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power.
-This will bring an almost factor of 4 improvement in the null test case even with realistic noise.
-Combining with LSS will only increase the removed power compared to CMB only by \textcolor{blue}{$\simeq 5\%$} but it will be highly beneficial to test the robustness of the final result.
+This will bring an almost factor of 3.7 improvement in the null test case even with realistic noise.
+Combining with LSS will only increase the removed power compared to CMB only by $\simeq 4.3\%$ but it will be highly beneficial to test the robustness of the final result.
 
@@
 Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-LSST (48\%) &1.8 & 1.7  & 1.7  \\ \hline
-LSS-S4  (60\%) &2.5 & 2.3  & 2.3 \\ \hline
+LSST (51\%) &2.0 & 1.9  & 1.7  \\ \hline
+LSS-S4  (59\%) &2.5 & 2.3  & 2.3 \\ \hline
 SKA (10$\mu$Jy) (53\%) &2.1& 1.9  & 1.16\\ \hline
-LSS-S4 + SKA (10$\mu$Jy)  (65\%) &2.7 & 2.5  & 2.5  \\ \hline
-CMB S4 (81\%) & 5.1 & 4.3  & 4.2 \\ \hline
-LSS+CMB (86\%)& 5.2 &  &  \\ \hline
+LSS-S4 + SKA (10$\mu$Jy)  (62\%) &2.6 & 2.3  & 2.5  \\ \hline
+CMB S4 (81\%) & 5.0 & 3.7  & 4.2 \\ \hline
+LSS-S4+CMB (86\%)& 6.6 & 4.3  &  \\ \hline
 \end{tabular}

### 2017-10-29 Notes/galaxy_delensing.tex
closer to final

@@
 We find that, in general, galaxy surveys should be split into tomographic bins as this can improve the correlation with CMB lensing by 30\%.
-Among currently available surveys, a DES-like galaxy survey can remove about $14\%$ of the lensing signal.
+Among currently available surveys, a DES-like galaxy survey can remove about $17\%$ of the lensing signal.
 Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
-With the decrease of instrumental noise, the CMB internal reconstruction will increase its efficiency and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($8\%$) and SPTPol ($35\%$) to 3G (56\%) and CMB S4 (85\%) level.
+With the decrease of instrumental noise, the CMB internal reconstruction will increase its efficiency and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($7\%$) and SPTPol ($31\%$) to 3G (56\%) and CMB S4 (81\%) level.
 Nevertheless optical galaxy surveys will still play an important role even for CMB S4.
@@
 The lensing component is a significant source of B-modes that, at large scales, corresponds to a white noise source of roughly $5 \mu K$-arcmin independent of the angular scale.
-This means that it is not only larger than the allowed inflationary component at scales smaller than several degrees ($r_{0.05} < 0.12$ from \cite{bicep2/keck-collaboration:2015}), but it is also comparable to current levels of instrumental noise.
+This means that it is not only larger than the allowed inflationary component at scales smaller than several degrees ($r_{0.05} < 0.07$ from \cite{bicep2-collaboration:2016a}), but it is also comparable to current levels of instrumental noise.
 For this reason, it is critical to characterize and eventually remove it from the data.
@@
 \hline
-Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.07}$ \\ \hline \hline
 WISE (8\%)& 1.08& 1.05 & 1.02 \\ \hline
 DES (17\%) &1.20 & 1.13  & 1.05 \\ \hline
-CIB (30\%) &1.44& 1.26  & 1.08\\ \hline
-LSS-S2 (42\%) &1.75& 1.42  & 1.11\\ \hline
+CIB (30\%) &1.44& 1.26  & 1.10\\ \hline
+LSS-S2 (42\%) &1.75& 1.42  & 1.15\\ \hline
 CMB Planck (7\%) &1.06 & 1.04  & 1.02 \\ \hline
-LSS-S2 +CMB (45\%)& 1.81& 1.45& 1.13  \\ \hline
+LSS-S2 +CMB (45\%)& 1.81& 1.45& 1.16  \\ \hline
 \end{tabular}
@@
 
-We also test a scenario where primordial gravitational waves are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}.
+We also test a scenario where primordial gravitational waves are present at their highest possible value of $r_{\rm{fid}}=0.07$ \cite{bicep2-collaboration:2016a}.
 As expected the importance of delensing itself is now reduced given that the lensing component constitutes a smaller portion of the total B-modes variance
@@
 \hline
-Surveys &$\alpha_{r=0}, N^{B}_{\ell}=0$& $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-SPTPol (31\%) &1.43& 1.26  & 1.08 \\ \hline
-LSS-S2 +SPTPol (55\%) & 2.2 & 1.6 & 1.14  \\ \hline
+Surveys &$\alpha_{r=0}, N^{B}_{\ell}=0$& $\alpha_{r=0}$ & $\alpha_{r=0.07}$ \\ \hline \hline
+SPTPol (31\%) &1.43& 1.26  & 1.10 \\ \hline
+LSS-S2 +SPTPol (55\%) & 2.2 & 1.6 & 1.20  \\ \hline
 
@@
 As expected the improvement on $n_{t}$ is similar to the one in $r$ since they are both proportional to the variance of the measured B-modes. 
-We tested this in the scenario $r=0.12$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}}/8$. 
+We tested this in the scenario $r=0.07$ where we additionally impose the consistency relation $n_{t}^{\rm{fid}}=-r^{\rm{fid}}/8$. 
 It is important to note that, independent of the level of delensing, for CMB 3G experiments the statistical error on $n_{t}$ will still be several times bigger than the fiducial value but it can still be tight enough to constrain more exotic inflationary models.
@@
 \hline
-Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.07}$ \\ \hline \hline
 DESI (11\%) &1.13 & 1.08  & 1.03 \\ \hline
-LSS-S3 (45\%) &1.82& 1.42  & 1.12\\ \hline
-CMB 3G (56\%) &2.38& 1.7  & 1.16 \\ \hline
+LSS-S3 (45\%) &1.82& 1.59  & 1.18\\ \hline
+CMB 3G (56\%) &2.31& 1.65  & 1.20 \\ \hline
 LSS-S3+CMB (68\%)& 3.2 & 2.3& 1.21  \\ \hline
@@
 \hline
-Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.12}$ \\ \hline \hline
-LSST (51\%) &2.0 & 1.9  & 1.7  \\ \hline
-LSS-S4  (59\%) &2.5 & 2.3  & 2.3 \\ \hline
-SKA (10$\mu$Jy) (53\%) &2.1& 1.9  & 1.16\\ \hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.07}$ \\ \hline \hline
+LSST (51\%) &2.0 & 1.9  & 1.21  \\ \hline
+LSS-S4  (59\%) &2.4 & 2.1  & 1.25 \\ \hline
+SKA (10$\mu$Jy) (53\%) &2.1& 1.9  & 1.22\\ \hline
 LSS-S4 + SKA (10$\mu$Jy)  (62\%) &2.6 & 2.3  & 2.5  \\ \hline
-CMB S4 (81\%) & 5.0 & 3.7  & 4.2 \\ \hline
+CMB S4 (81\%) & 5.0 & 3.7  & 1.43 \\ \hline
 LSS-S4+CMB (86\%)& 6.6 & 4.3  &  \\ \hline
@@
 In this paper, we studied the potential impact of large-scale structure galaxy surveys in this important endeavor. 
+We focus on galaxy clustering here, leaving other probes like weak lensing, for future works.
+However,  weak lensing is usually lower redshift probes than clustering resulting in a worse delensing efficiency.
 We find that, to improve the delensing efficiency of galaxy surveys, tomographic delensing will be very important: using several tomographic galaxy bins can improve the correlation of galaxies with CMB lensing by 10\%-30\% on a wide range of angular scales.

### 2017-10-30 Code/compute_chi2_lensingiswpowerspectrum/compute_chi2_lensingiswpowerspectrum.f90
rearranging code

@@ -1,82 +1,0 @@
-PROGRAM compute_chi2_lensingiswpowerspectrum
-! A code for computing the full-sky, cosmic-variance only chi-squared of the 
-! lensing-ISW power spectrum, including variance from the lensing signal itself.
-!
-! Chi-squared is computed as
-!
-! Chi^2 = sum_l [Q(l)]^2/{C(l)*(N(l)+Ctheta(l))+[Q(l)]^2}
-!
-! where
-! - C(l) is the lensed CMB temperature power spectrum
-! - Q(l) is the lensing-ISW cross-power spectrum
-!   [more precisely, this is the lensing-temperature power spectrum, as it 
-!    includes all the linear effects (not just the late-time ISW)]
-! - Ctheta(l) is the power spectrum of lensing potential
-!
-! all of which are computed from CAMB, and
-!
-! - N(l) is the noisebias of the reconstructed lensing potential 
-!   (computed by "compute_noise" in this package)
-!
-! REF: Lewis, Challinor & Hanson, JCAP 03, 018 (2011), arXiv:1101.2234
-! April 25, 2012: E.Komatsu 
-  IMPLICIT none
-  character(len=128) :: filename
-  integer :: lmax=1500 ! maximum multipole
-  integer :: lmin=2    ! minimum multipole
-  integer :: i,l,info,dum
-  double precision :: twopi,dummy
-  double precision :: chi2a,chi2b,chi2c,chi2d
-  double precision, allocatable, dimension(:) :: cl,nl,ctheta,ql
-  print*,'lmin=',lmin
-  print*,'lmax=',lmax
-  twopi=2d0*3.1415926535d0
-! read in noisebias, N(l)
-  filename='multipole_noisebias.txt'
-  print*,'noisebias file= ',trim(filename)
-  open(1,file=filename,status='old')
-  allocate(nl(lmax))
-  do i=2,lmax
-     read(1,*)dum,nl(i)
-  enddo
-  close(1)
-! read in lens-temperature cross power spectrum and lensing potential power
-! spectrum
[…]

### 2017-10-30 Code/compute_chi2_lensingiswpowerspectrum/compute_noisebias.f90

@@ -1,73 +1,0 @@
-PROGRAM compute_noisebias
-! A for computing "noisebias," Nl, of the reconstructed lensing potential:
-!
-! 1/Nl = 1/(2l+1) * sum_{l2=l}^{lmax} * sum_{l3=l2}^{lmax}
-!                 * A_{l1l2l3}^2/[C(l2)C(l3)Delta(l1,l2,l3)].
-!
-! REF: Lewis, Challinor & Hanson, JCAP 03, 018 (2011), arXiv:1101.2234
-! April 25, 2012: E.Komatsu
-
-  IMPLICIT none
-  character(len=128) :: filename
-  integer :: lmax=3000 ! maximum multipole
-  integer :: lmin=2    ! minimum multipole
-  integer :: i,l,l1,l2,l3,info,dum
-  double precision :: l3min,l3max,gaunt,fourpi,twopi,delta
-  double precision :: Al1l2l3,noisebias
-  double precision, allocatable, dimension(:) :: cl,wigner3j
-  print*,'lmin=',lmin
-  print*,'lmax=',lmax
-  twopi=2d0*3.1415926535d0
-  fourpi=4d0*3.1415926535d0
-! read in lensed Cl
-  allocate(cl(lmax))
-  filename='wmap5baosn_max_likelihood_lensedCls.dat'
-  print*,'cl file (lensed)= ',trim(filename)
-  open(2,file=filename,status='old')
-  do l=2,lmax
-     read(2,*)dum,cl(l)
-     cl(l)=cl(l)/dble(l)/(dble(l)+1d0)*twopi/(2.726d6)**2d0
-  enddo
-  close(2)
-! begin computing noisebias
-  allocate(wigner3j(2*lmax+1))
-  open(3,file='multipole_noisebias.txt',status='unknown')
-  do l1=lmin,lmax
-     noisebias=0d0
-     do l2=l1,lmax
-        call DRC3JJ(dble(l1),dble(l2),0d0,0d0, &
-             l3min,l3max,wigner3j,2*lmax+1,info)
-        if ((int(l3max).ge.l2)) then
-           do l3=l2,int(l3max)
-              if ((l3.le.lmax).and.(mod(l1+l2+l3,2).eq.0))then
-                 gaunt=wigner3j(l3-int(l3min)+1)/dsqrt(fourpi) &
-                      *dsqrt(dble(2*l1+1))*dsqrt(dble(2*l2+1)) &
[…]

### 2017-10-30 Code/multiple_survey_delens.py

@@ -126,7 +126,7 @@ def main(labels, cmb):
         ckk_noise = noise_fun(lbins)
 
     elif cmb == 'S4':
-        noise = 0.3
+        noise = 0.1
         beam = 2
         noise_phi = np.loadtxt(
             './quicklens_data/min_var_noise_{}muk_{}beam.txt'.format(noise, beam))

### 2017-10-30 Code/run_analysis.py

@@ -21,7 +21,7 @@ from colorama import Fore
 import datetime
 
 
-sys.stdout = open('run_analysis' + datetime.datetime.now().strftime("%Y-%m-%d %H:%M:%S"+'.txt'), 'w')
+# sys.stdout = open('run_analysis_S4' + datetime.datetime.now().strftime("%Y-%m-%d %H:%M:%S"+'.txt'), 'w')
 
 
 init(autoreset=True)
@@ -89,15 +89,15 @@ def run_fisher_cases(rho_names, lmin, lmax, deep, fsky=0.06, r_tens_case=0.07):
         print('gain', (probe, 'gain = ', sigma_r_1 / sigma_r,
                        sigma_nt_1 / sigma_nt))
 
-    print('After delensing % errors', sigma_r_1 * 1e2)
-
+    print('After delensing % errors sigma(r)*1e2', sigma_r * 1e2)
+    print('After delensing % errors sigma(nt)', sigma_nt)
     print('')
     print('')
 
     print('')
     print(Fore.YELLOW + 'r=0')
     print('')
-    r_fid = 0
+    r_fid = 1e-4
 
     for i, label in enumerate(rho_names):
         probe = label.split('.txt')[0].split('rho_')[1]
@@ -121,8 +121,8 @@ def run_fisher_cases(rho_names, lmin, lmax, deep, fsky=0.06, r_tens_case=0.07):
         print('gain', (probe, 'gain = ', sigma_r_1 / sigma_r,
                        sigma_nt_1 / sigma_nt))
 
-    print('After delensing % errors', sigma_r_1 * 1e2)
-
+    print('After delensing % errors sigma(r)*1e2', sigma_r * 1e2)
+    print('After delensing % errors sigma(nt)', sigma_nt)
     print('')
     print('')
 
@@ -161,7 +161,8 @@ def run_fisher_cases(rho_names, lmin, lmax, deep, fsky=0.06, r_tens_case=0.07):
         print('gain', (probe, 'gain = ', sigma_r_1 / sigma_r,
                        sigma_nt_1 / sigma_nt))
 
-    print('After delensing % errors sigma(r)*1e2', sigma_r_1 * 1e2)
+    print('After delensing % errors sigma(r)*1e2', sigma_r * 1e2)
[…]

### 2017-10-30 Code/res_to_alpha.py
rearranging code

@@ -1,112 +1,0 @@
-# coding: utf-8
-
-'''
-Testing different method to compute a 2d B_res.
-
-Use method 3 after the test for best accuracy
-
-'''
-
-
-
-from scipy.interpolate import RectBivariateSpline, interp1d
-# import lensing
-import sys
-import configparser as ConfigParser
-import numpy as np
-
-noise = '/home/manzotti/my_version_cosmosis/cosmosis-standard-library/my_projects/CIB_sherwin/compute_chi2_lensingiswpowerspectrum/multipole_noisebias.txt'
-
-
-# call external Eichiro code
-
-# def generate_CMB_rho():
-#     '''
-#     todo: to make it more realistic add beam here. an extenstion to more observable TE and mean variance would be good
-#     '''
-
-#     clpp_th = np.loadtxt(
-#         '/home/manzotti/cosmosis/modules/limber/galaxies_delens/cmb_cl/pp.txt')
-
-#     ells_cmb = np.loadtxt(
-#         '/home/manzotti/cosmosis/modules/limber/galaxies_delens/cmb_cl/ell.txt')
-
-#     # noise is a 2d array first column is ell second is cl
-#     noise_cl = np.loadtxt(
-#         '/home/manzotti/my_version_cosmosis/cosmosis-standard-library/my_projects/CIB_sherwin/compute_chi2_lensingiswpowerspectrum/multipole_noisebias.txt')
-#     # check the first ell is the same
-#     assert (ells_cmb[0] == noise_cl[0, 0])
-#     noise_cl[:, 1] = noise_cl[:, 1] * noise_cl[:, 0] ** 4
-#     # cut one of the 2 at the proper lmax if needed
-#     lmax = np.min((np.shape(noise_cl)[0], np.shape(clpp_th)[0]))
-#     # try to generate them at the same ells
-#     return np.sqrt(clpp_th[:lmax] / (clpp_th[:lmax] + noise_cl[:lmax, 1]))
-
-def bl(fwhm_arcmin, lmax):
-    """ returns the map-level transfer function for a symmetric Gaussian beam.
-         * fwhm_arcmin      - beam full-width-at-half-maximum (fwhm) in arcmin.
[…]

### 2017-10-30 Notes/galaxy_delensing.tex
more text review also after Giulio input

@@
 
-\title{Future cosmic microwave background delensing with galaxy surveys.}       
+\title{Future cosmic microwave background delensing with galaxy surveys}
 \author{A. Manzotti} 
@@
 Removing the lensing part (``delensing'') from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
-Here we discuss the role of current and future large scale structure surveys in a multi-tracers approach to delensing that will improve the reconstruction of the lensing potential that lenses the CMB photons and, as a consequence, the delensing efficiency.
+Here we discuss the role of current and future large scale structure surveys in a multi-tracers approach to delensing that will improve thesreconstruction of the lensing potential that lenses the CMB photons and, as a consequence, the delensing efficiency.
 We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
@@
 DES is modeled after the DES Science Verification public data release.
-For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy denisty of 0.63 galaxies per squared arcmin. 
+For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy density of 0.63 galaxies per squared arcmin.
 %As in \cite{passaglia:2017} we use a linear bias $b$
@@
 
-Finally, we consider the Square Kilometre Array (SKA).
+Finally, we consider the Square Kilometer Array (SKA).
 The SKA is a planned radio array that will survey large scale structure primarily by detecting the redshifted neutral hydrogen (HI) 21cm emission line from a large number of galaxies out to high redshift.
@@
 The CIB consists of diffuse extragalactic radiation generated by the unresolved emission from star-forming galaxies (see \cite{dole:2006} and references therein). 
-In these galaxies, the UV light from young stars heats the dust regions around them that then reradiates thermally in the infrared with a graybody spectrum of $T \simeq30$K.
+In these galaxies, the UV light from young stars heats the dust regions around them that then reradiates thermally in the infrared with a gray-body spectrum of $T \simeq30$K.
 
@@
 
-\refeq{Bres} highligts the different factors that control the delensing efficiency.
+\refeq{Bres} highlights the different factors that control the delensing efficiency.
 The first part of the second term in the parenthesis consists of an inverse variance filter applied to the measured E-mode.
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/compare_kernel.pdf}
+\includegraphics[scale=1.]{images/compare_kernel.pdf}
 \caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/BB_res_ell2.pdf}
+\includegraphics[scale=1.]{images/BB_res_ell2.pdf}
 \caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/compare_kernel_tomo.pdf}
+\includegraphics[scale=1.]{images/compare_kernel_tomo.pdf}
 \caption{Comparison of kernels of the 10 LSST tomographic bins together with the full LSST survey and the CMB lensing kernel. 
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/B_res_bin.pdf}
+\includegraphics[scale=1.]{images/B_res_bin.pdf}
 \caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing.
@@
 We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the ability to internally reconstruct the lensing potential $\kappa$, together with a low-noise, low-resolution experiment that measures the B-modes that will be delensed and used to constrain the inflationary parameters.
-As in \cite{abazajian:2016} and other works in the literature, we do not combine the observations from the two experiments to obtain a single CMB map with optimally low noise.
-Depending on the experiments, this might be possible and, once used for both lensing reconstruction and B-mode measurements, it can improve the level of delensing and the cosmological constraints.
+As in \cite{abazajian:2016} and other works in the literature, we only combine the observations from the two experiments to obtain a single CMB E-mode map with optimally low noise.
+We assume independent measurements for the lensing reconstruction and the B-modes measurement.
+Depending on the experiments, it might be possible to optimally combined the two experiments and, once used for both lensing reconstruction and B-mode measurements, it can improve the level of delensing and the cosmological constraints.
 
-Even if there is not a definite distinction between different stages, we focus on three distinct scenarios: the ongoing or recently concluded stage (S2), a third generation stage (3G) of experiments in an advanced building phase or that have just started taking data and finally the futuristic CMB Stage 4.
-
-%\begin{figure}[htbp]
-%\begin{center}
-%\includegraphics[scale=1.]{../images/cmb_internal.pdf}
-%\caption{\textbf{CMB internal delensing:} Correlation factor between CMB internal reconstructed potential and the actual lensing potential for different CMB experiments and corresponding noise reconstructions.
-%CMB measurements will rapidly improve providing the most efficient (even if biased) way to reconstruct the structures lensing the CMB.}
-%\end{center}
-%\label{fig:corr-internal}
-%\end{figure}
-
+Even if there is not a definite distinction between different experimental stages, we focus on three distinct scenarios: the ongoing or recently concluded stage (S2), a third generation stage (3G) of experiments in an advanced building phase or that have just started taking data and finally the futuristic CMB Stage 4.
+s
 \subsection{Fisher Information Matrix}\label{sec:fisher}
@@
 Furthermore, we are not considering the uncertainties on galaxy survey internal parameters such as biases, source distributions and photometric redshift uncertainties.
-These uncertaintes can significantly degrade inflationary constraints \cite{sherwin:2015,namikawa:2016a}.
+These uncertainties can degrade inflationary constraints \cite{sherwin:2015,namikawa:2016a}.
 However, as shown in \cite{sherwin:2015,namikawa:2016a}, these can be auto-calibrated i.e.
@@
 The approximations made here go in the direction of a slightly optimistic \textit{absolute} value for the parameters statistical uncertainties both for the standard and the delensed case.
-The delensing improvement is defined as the \textit{relative} value of these uncertainties $\alpha_{i} = \sigma^{\rm{del}}(i)/\sigma(i)$ and, at first order, it is not affected by this assumptions.
+The delensing improvement is defined as the \textit{relative} value of these uncertainties $\alpha_{i} = \sigma^{\rm{del}}(i)/\sigma(i)$ and, at first order, it is not affected by these assumptions.
 
@@
 Recently delensing has been performed for the first time on data using both CIB maps \cite{larsen:2016a,manzotti:2017} and internally reconstructed CMB lensing maps \cite{carron:2017} as large scale structure tracers.
-Here we discuss the improvement that can be obtained by combining these and other currently Stage-2 available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
+In this section we discuss the improvement that can be obtained by combining these and other currently Stage-2 available tracers (see \cite{yu:2017} for publicly accessible multi-tracers data-products).
 
 On the CMB side, we will combine a BICEP-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment.
-For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt(2)$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
-We assume that the CMB lensing reconstruction is performed by the higher resolution CMB experiment. We explore two possibilities.
+For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt{2}$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
+We assume that the CMB lensing reconstruction is performed by the higher resolution CMB experiment.
+We explore two possibilities.
 First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck satellite \cite{planck-collaboration:2016a}.
@@
 \footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
+For consistency, we also combine a Planck-like E-mode map of $60\mu$K-arcmin in polarization with a beam equal to 7 arcmin that however adds almost nothing to the \textit{deep} CMB experiment.
 Then, we test a second scenario with a ground-based experiment characterized by noise levels in polarization consistent with SPT-Pol \cite{henning:2017}: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin.
@@
 Its correlation is comparable to the one of LSS at almost all scales.
-On the other end, the CMB lensing reconstruction from SPT-Pol will be more than \textcolor{blue}{70\%} correlated with the true field at $\ell<300$, and only then the correlation goes below the level of current LSS surveys.
+On the other end, the CMB lensing reconstruction from SPT-Pol will be more than $70\%$ correlated with the true field at $\ell<300$, and only then the correlation goes below the level of current LSS surveys.
 In both cases the CMB lensing reconstruction correlates very well at low $\ell$ and it then falls rapidly at smaller scales because of the raise of the reconstruction noise.
@@
 \reffig{corrnow} and \reffig{corrnow-planck} show that the DES galaxies are effective tracers of the LSS and can, at least in the near future, be used to improve delensing for CMB experiments that overlap with it.
-For example, DES delensing efficiency is higher than WISE first because of the lower level of noise and secondly because DES galaxies are located at slightly higher redshift so they better overlap with CMB lensing.
+For example, DES delensing efficiency is higher than WISE first as a result of the lower level of noise and secondly because DES galaxies are located at slightly higher redshift so they better overlap with CMB lensing.
 As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS.
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/actual_scenario_planck.pdf}
+\includegraphics[scale=1.]{images/actual_scenario_planck.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers.
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/actual_scenario.pdf}
+\includegraphics[scale=1.]{images/actual_scenario.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers. 
@@
 A slightly better performance is achieved with WISE.
-DES can do better, once binned in redshift, removing a level of $17\%$ of power.
+Once binned in redshift, DES can do better, removing a level of $17\%$ of power.
 However, it covers a smaller fraction of the sky than the two previous probes.
@@
 Nonetheless, it is worthwhile combining all the tracers.
-Indeed a multi tracer-approach can bring the removed power on the overlapping area from $30\%$ of the CIB alone to $43\%$ using all the LSS (LSS-S2) and $45\%$ when Planck lensing reconstruction is added to the set.
+Indeed a multi tracer-approach can bring the removed power on the overlapping area from $30\%$ of the CIB alone to $42\%$ using all the LSS (LSS-S2) and $45\%$ when Planck lensing reconstruction is added to the set.
 However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 45\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/S3_scenario.pdf}
+\includegraphics[scale=1.]{images/S3_scenario.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
-This figure is the same as \reffig{corrnow} but with a stage-3 CMB experiment performing the internal reconstrcution and with DESI added to the stage-2 galaxy surveys (LSS-S2).
+This figure is the same as \reffig{corrnow} but with a stage-3 CMB experiment performing the internal reconstruction and with DESI added to the stage-2 galaxy surveys (LSS-S2).
 }
@@
 We summarize the improvement in the constraint on $r$ in \reftab{S3}. 
-Given the improvement in the noise of the high-resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of $\sim 2.4$ through internal delensing in the ideal case of no instrumental noise in the B-modes (but with instrumental noise in the E-mode used to delens).
+Given the improvement in the noise of the high-resolution experiment, CMB alone will be able to improve constraint on $r$ by a factor of $\sim 2.3$ through internal delensing in the ideal case of no instrumental noise in the B-modes (but with instrumental noise in the E-mode used to delens).
 Adding galaxy surveys will lead to a further improvement without additional effort.
-Indeed galaxy surveys will still be able to remove an additional $11\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0$) by $34\%$.
+Indeed galaxy surveys will still be able to remove an additional $12\%$ of power from $C_{\ell}^{BB,\mathrm{lens}}$ and to improve our constraint on the null hypothesis ($r=0, $) by $54\%$ in the noiseless case.
 Note that apart from the addition of DESI, the tracers used here are already available today.
 The ongoing effort can then lead to a significant improvement even for the next generation of experiments.
-Finally, the combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor \textcolor{blue}{1.8} or to improve the constraint of a possible detection (r=0.07) by 34\%.
+Finally, the combination of CMB and galaxy surveys will be able to improve the constrain on the null hypothesis of no primordial waves by a factor 2.3 or to improve the constraint of a possible detection (r=0.07) by 30\% even in the noisy case.
 As expected the improvement on $n_{t}$ is similar to the one in $r$ since they are both proportional to the variance of the measured B-modes. 
@@
 Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.07}$ \\ \hline \hline
-DESI (11\%) &1.13 & 1.08  & 1.03 \\ \hline
+DESI (11\%) &1.13 & 1.1  & 1.04 \\ \hline
 LSS-S3 (45\%) &1.82& 1.59  & 1.18\\ \hline
-CMB 3G (56\%) &2.31& 1.65  & 1.20 \\ \hline
-LSS-S3+CMB (68\%)& 3.2 & 2.3& 1.21  \\ \hline
+CMB 3G (56\%) &2.31& 1.9  & 1.23 \\ \hline
+LSS-S3+CMB (68\%)& 3.17 & 2.3& 1.30  \\ \hline
 \end{tabular}
@@
 \begin{center}
-\includegraphics[scale=1.]{../images/S4_scenario_ska10.pdf}
+\includegraphics[scale=1.]{images/S4_scenario_ska10.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
-This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstrcution and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
+This figure is the same as \reffig{corrnow} but with a stage-4 CMB experiment performing the internal reconstruction and with LSST and SKA added to the stage-2 and stage-3 galaxy surveys (LSS-S3).
 }
@@
 \reffig{corrs4_10} shows that a CMB-S4 experiment will be able to internally reconstruct the lensing potential at more than 90\% up to $\ell=450$.
-At this level, CMB S4 internal delensing will not be limited by noise but by small secondary effects like foregrounds contamination, filtered modes etc.
-%Finally, \reftab{S4} shows that a very futuristic SKA-2 experiment will have a very high correlation with CMB lensing, being more than 80\% correlated at all the scales.
-This performance is comparable to a CMB-S4 internal reconstruction.
+At this level, CMB S4 internal delensing will not be limited by noise but by small secondary effects like foregrounds contamination, filtered modes etc (see \cite{manzotti:2017}).
 
@@
 If we add all the other infrared and optical LSS tracers previously included (i.e.
-excluding SKA) the improvement factor increase to 2.5.
+excluding SKA) the improvement factor increase to 2.4.
 These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSS, completely independent of CMB internal data.
@@
 Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power.
-This will bring an almost factor of 3.7 improvement in the null test case even with realistic noise.
+This will bring a factor of $\sim 4$ improvement in the null test case even with realistic noise.
 Combining with LSS will only increase the removed power compared to CMB only by $\simeq 4.3\%$ but it will be highly beneficial to test the robustness of the final result.
 
-We find the improvement in $\sigma(n_{t})$ to be very similar to the one in $\sigma(r)$ for all the CMB-S4 cases.
+For the case $r=0.07$, we find the improvement in $\sigma(n_{t})$ to be just slightly higher to the one in $\sigma(r)$ for all the CMB-S4 cases.
 This is reasonable because, given the level of noise in CMB-S4, the variance at angular scales $\ell<200$ are dominated by instrumental noise and not by the lensing component.
-Independent of the delensing level, given the small value allowed for $r$ will not allow to test the consistency relations.
+The conclusions about the relative importance of LSS tracers we find for $\sigma(r)$ are still valid for $\sigma(n_{t})$.
+Delensing can then be useful also to constrain more exotic inflationary models.
 
 \begin{table}
-\caption{$\alpha$: Improvements on $\sigma(r)$ due to delensing for S4 experiments.
+\caption{$\alpha$: Improvements on $\sigma(r)$ and $\sigma(n_{t})$ after delensing for S4 experiments.
 Here we define LSS-S4 as the combination of S3 LSS tracers (LSS-S3) and LSST.
@@
 We also use a CMB S4 experiment for the lensing internal reconstruction.
+The improvement for $\sigma(n_{t})$ is computed for fiducial values $r=0.07$ and $n_t=-r/8$.
 See \refsec{S4} for details.}
-\begin{tabular}{|c | c | c | c|}
+\begin{tabular}{|c | c | c | c| c|}
 \hline
-Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.07}$ \\ \hline \hline
-LSST (51\%) &2.0 & 1.9  & 1.21  \\ \hline
-LSS-S4  (59\%) &2.4 & 2.1  & 1.25 \\ \hline
-SKA (10$\mu$Jy) (53\%) &2.1& 1.9  & 1.22\\ \hline
-LSS-S4 + SKA (10$\mu$Jy)  (62\%) &2.6 & 2.3  & 2.5  \\ \hline
-CMB S4 (81\%) & 5.0 & 3.7  & 1.43 \\ \hline
-LSS-S4+CMB (86\%)& 6.6 & 4.3  &  \\ \hline
+Surveys & $\alpha_{r=0}, N^{B}_{\ell}=0$ & $\alpha_{r=0}$ & $\alpha_{r=0.07}$ & $\alpha^{n_{t}}_{r=0.07}$ \\ \hline \hline
+LSST (51\%) &2.1 & 1.9  & 1.21 & 1.47 \\ \hline
+LSS-S4  (59\%) &2.4 & 2.1  & 1.25 & 1.57\\ \hline
+SKA (10$\mu$Jy) (53\%) &2.1& 1.9  & 1.22& 1.46\\ \hline
+LSS-S4 + SKA (62\%) &2.6 & 2.3  & 1.3& 1.6  \\ \hline
+CMB S4 (83\%) & 6.22 & 4.1  & 1.42& 2.3 \\ \hline
+LSS-S4+CMB (86\%)& 7.7 & 4.7  & 1.46 & 2.46\\ \hline
 \end{tabular}
@@
 \begin{figure*}[ht]
-\includegraphics[scale=1]{../images/errors_summary.pdf}
+\includegraphics[scale=0.8]{images/errors_summary2.pdf}
 \caption[]{A summary of the amount of lensing B-modes power removed by different tracers for different generations of experiments.
@@
 The ability to separate the lensing component of the CMB B-modes from a possible primordial inflationary signal (``delensing'') is necessary to test inflation using the next generation of CMB polarization experiments.
-To delens, we need to accurately reconstruct the large scale structures that lens the CMB in order to build a template of the expected lensing B-modes in the observed patch.
+To delens, we need to accurately reconstruct the large scale structures that lens the CMB in order to marginalize the expected lensing B-modes components in the observed patch.
 In this paper, we studied the potential impact of large-scale structure galaxy surveys in this important endeavor. 
 We focus on galaxy clustering here, leaving other probes like weak lensing, for future works.
-However,  weak lensing is usually lower redshift probes than clustering resulting in a worse delensing efficiency.
+However, weak lensing is a lower redshift probes than clustering resulting in general in a worse delensing efficiency.
 We find that, to improve the delensing efficiency of galaxy surveys, tomographic delensing will be very important: using several tomographic galaxy bins can improve the correlation of galaxies with CMB lensing by 10\%-30\% on a wide range of angular scales.
@@
 For ongoing experiments, we find that LSS tracers will be particularly beneficial. 
-An optical survey like DES is able to remove \textcolor{blue}{$14\%$} of the B-mode lensing power alone, and together with WISE(\textcolor{blue}{$8\%$}) and the CIB (\textcolor{blue}{$27\%$}) will allow to remove $36\%$ of the power using only LSS surveys.
-Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using LSS survey can correspond to a maximum improvement of \textcolor{blue}{30\%} in the constraint on the tensor to scalar ratio $r$ compared to the value before delensing.
+An optical survey like DES is able to remove $17\%$ of the B-mode lensing power alone, and together with WISE($8\%$) and the CIB ($30\%$) will allow to remove $42\%$ of the power using only LSS surveys.
+Depending on the CMB instrumental noise and the amount of galactic foreground cleaning, delensing using currently available LSS survey can correspond to a maximum improvement of $42\%$ in the constraint on the tensor to scalar ratio $r$ compared to the value before delensing.
 In the future, the decreasing level of instrumental noise in CMB experiments will dramatically improve the internal reconstruction of the structures lensing the CMB. 
-The fraction of removed lensing B-modes will rapidly improve from the current expected levels for Planck (\textcolor{blue}{$8\%$}) and SPTPol (\textcolor{blue}{$35\%$}) to 3G (\textcolor{blue}{56\%}) and CMB S4 (\textcolor{blue}{85\%}) levels.
+The fraction of removed lensing B-modes will rapidly improve from the current expected levels for Planck ($7\%$) and SPTPol ($31\%$) to 3G (56\%) and CMB S4 (81\%) levels.
 Indeed for Stage-3 experiments, the CMB internal reconstruction will be the main source of delensing.
 However, it will still be less efficient than galaxies at tracing the lensing potential at small scales $\ell>500$.
-For this reason, combining galaxy survey with the CMB will push the fraction of removed power from \textcolor{blue}{$57\%$} to \textcolor{blue}{$68\%$} for 3G.
+For this reason, combining galaxy survey with the CMB will push the fraction of removed power from $56\%$ to $68\%$ for 3G.
 Even for CMB S4, galaxy surveys will still play an important role.
-For example, a tomographically binned LSST-like survey by itself will remove \textcolor{blue}{56\%} of the lensing power.
-This performance is lower than a S4 CMB internal reconstruction and comparable to a Stage-3 CMB.
+For example, a tomographically binned LSST-like survey by itself will remove 51\% of the lensing power. This performance is lower than a S4 CMB internal reconstruction and comparable to a Stage-3 CMB.
 However, it will allow us to probe the robustness against systematics of an eventual detection of primordial gravitational waves.
@@
 
-Given the high level of correlation of future galaxy surveys and CMB lensing, it is also worth looking for potential application besides the detection of inflationary B-modes.
+Given the high level of correlation of future galaxy surveys and CMB lensing, it is also worth looking for potential delensing application besides the detection of inflationary B-modes.
 For example, delensing with galaxy surveys allows us to selectively remove from the CMB only the gravitational effect coming from the non-linear structures in the low-redshift universe. 

### 2017-11-23 Code/Plot_figures.py
working on referree

@@ -134,9 +134,9 @@ plt.xlim(10, 1400)
 plt.ylim(0, 1.15)
 fg.tight_layout()
 
-plt.savefig('../images/actual_scenario_planck.pdf',
+plt.savefig('../Notes/images/actual_scenario_planck.pdf',
             dpi=600, papertype='Letter')
-plt.savefig('../images/actual_scenario_planck.png')
+plt.savefig('../Notes/images/actual_scenario_planck.png')
 
 
 # ## CMB current scenario
@@ -165,6 +165,8 @@ plt.plot(lbins, rho['wise'], color=colors['WISE'], label='WISE')
 plt.plot(lbins, rho_cmb, label='SPT Pol', color=colormap(6))
 plt.plot(lbins, rho_gals, label='DES+CIB', color=colormap(7))
 plt.plot(lbins, rho_comb, label='LSS S2 +  SPTPol', color=colormap(8))
+plt.plot(BB_contr[:, 0], BB_contr[:, 1] - np.min(BB_contr[:, 1]),
+         '--', alpha=0.6, linewidth=font_size / 14., color='k')
 # plt.plot(BB_contr[:, 0], BB_contr[:, 1] - np.min(BB_contr[:, 1]), '--', alpha=0.6, color='k')
 plt.legend(loc=0, ncol=2)
 plt.ylim(0, 1.2)
@@ -175,8 +177,8 @@ plt.xlim(10, 1400)
 fg.tight_layout()
 
 
-plt.savefig('../images/actual_scenario.pdf', dpi=600, papertype='Letter')
-plt.savefig('../images/actual_scenario.png')
+plt.savefig('../Notes/images/actual_scenario.pdf', dpi=600, papertype='Letter')
+plt.savefig('../Notes/images/actual_scenario.png')
 
 # ## CMB S3 scenario
 
@@ -198,6 +200,9 @@ plt.plot(lbins, rho_gals_desi, color=colors['DESI'], label='DESI')
 plt.plot(lbins, rho_cmb, label='SPT 3G', color=colormap(6))
 plt.plot(lbins, rho_gals, label='LSS-S2+DESI (LSS-S3)', color=colormap(7))
 plt.plot(lbins, rho_comb, label='LSS-S3+SPT 3G', color=colormap(8))
+plt.plot(BB_contr[:, 0], BB_contr[:, 1] - np.min(BB_contr[:, 1]),
[…]

### 2017-11-23 Code/gals_kernel.pyx

@@ -298,7 +298,7 @@ class kern():
         self.dndzfun = dndzfun
         self.norm = scipy.integrate.quad(
             dndzfun, self.zmin, self.zmax, limit=100, epsrel=1.49e-03)[0]
-        print(self.norm, self.zmin, self.zmax)
+        # print(self.norm, self.zmin, self.zmax)
         self.hspline = hspline
 
     def w_lxz(self, float l, float x, float z):
@@ -314,7 +314,7 @@ class kern():
        '''
         cdef double dndz, b
         dndz = self.dndzfun(z)
-
+        # assert(dndz is not np.nan)
         b = self.b
         # print z,self.dndzfun(z),  self.norm  , self.b
         return dndz * self.b

### 2017-11-23 Code/multiple_survey_delens.py

@@ -26,10 +26,10 @@ def nl(noise_uK_arcmin, fwhm_arcmin, lmax):
     return (noise_uK_arcmin * np.pi / 180. / 60.)**2 / bl(fwhm_arcmin, lmax)**2
 
 
-def main(labels, cmb):
+def main(labels, cmb, spectra_file='../Data/limber_spectra_delens.pkl'):
 
     output_dir = '/home/manzotti/galaxies_delensing/Data/'
-    cls = pickle.load(open('../Data/limber_spectra_delens.pkl', 'rb'))
+    cls = pickle.load(open(spectra_file, 'rb'))
     lbins = np.load('../Data/ells.npy')
 
     ckk = cls['kk']
@@ -107,6 +107,17 @@ def main(labels, cmb):
     if cmb == 'S3':
         noise = 2.0
         beam = 2
+        noise_phi = np.loadtxt(
+            './quicklens_data/min_var_noise_{}muk_{}beam.txt'.format(noise, beam))
+        noise_phi *= np.arange(0, len(noise_phi))**4. / 4.
+        # noise_cmb = nl(noise, beam, lmax=4000)
+        noise_fun = interp1d(np.arange(0, len(noise_phi)), noise_phi)
+        ckk_noise = np.zeros_like(ckk)
+        ckk_noise = noise_fun(lbins)
+
+    if cmb == 'SPT_deep':
+        noise = 7.0
+        beam = 1
         noise_phi = np.loadtxt(
             './quicklens_data/min_var_noise_{}muk_{}beam.txt'.format(noise, beam))
         noise_phi *= np.arange(0, len(noise_phi))**4. / 4.

### 2017-11-23 Code/run_galaxies_delens.py

@@ -58,7 +58,7 @@ def find_bins(z, dndz, nbins):
     args = np.hstack((np.array(0.), np.searchsorted(
         cum, [cum[-1] / nbins * n for n in np.arange(1, nbins + 1)]))).astype(np.int)
     args[-1] = len(cum) - 1
-    return [z[args[i]:args[i + 1]] for i in np.arange(0, len(args) - 1.)], args
+    return [z[args[np.int(i)]:args[np.int(i) + 1]] for i in np.arange(0, len(args) - 1.)], args
 
 
 def make_tomo_bins(z, dndz, sigmaz, width, nbins, hspline, omegac, h, b=1.):
@@ -70,11 +70,11 @@ def make_tomo_bins(z, dndz, sigmaz, width, nbins, hspline, omegac, h, b=1.):
         dndz_win = dndz * (scipy.special.erfc((width * (n - 1) - z) / (sigmaz * np.sqrt(2))
                                               ) - scipy.special.erfc((width * n - z) / (sigmaz * np.sqrt(2))))
         dndzlsst_temp = InterpolatedUnivariateSpline(
-            z, dndz_win, ext='zeros')
+            z, dndz_win, ext='raise')
         norm = dndzlsst_temp.integral(z[0], z[-1])
         # print(norm, dndz_win, z)
         dndzlsst_temp_fun = InterpolatedUnivariateSpline(
-            z, dndz_win / norm * np.sqrt(1. + z), ext='zeros')
+            z, dndz_win / norm * np.sqrt(1. + z), ext='raise')
         lsst_tomo_bins.append(gals_kernel.kern(
             z, dndzlsst_temp_fun, hspline, omegac, h, b=1.))
     return lsst_tomo_bins
@@ -104,12 +104,12 @@ def make_tomo_bins_equal_gals(z, dndz, sigmaz, nbins, hspline, omegac, h, b=1.):
 
         dndz_win = dndz * photoz_confusion
         dndzlsst_temp = InterpolatedUnivariateSpline(
-            z, dndz_win, ext='zeros')
+            z, dndz_win, ext='raise')
         norm = dndzlsst_temp.integral(z[0], z[-1])
         # print(norm, dndz_win, z)
         galaxies_fraction.append(norm)
         dndzlsst_temp_fun = InterpolatedUnivariateSpline(
-            z, dndz_win / norm, ext='zeros')
[…]

### 2017-11-23 Code/compute_kernel_to_compare.py
find a bug in CIB kernel plot

@@ -89,7 +89,7 @@ norm = dndzfun.integral(dndz[0, 0], dndz[-2, 0])
 dndzfun_desi = InterpolatedUnivariateSpline(dndz[:, 0], dndz[:, 1] / norm)
 
 
-wise_dn_dz = np.loadtxt('/home/manzotti/galaxies_delensing/wise_dn_dz.txt')
+wise_dn_dz = np.loadtxt('/home/manzotti/galaxies_delensing/Code/wise_dn_dz.txt')
 dndzwise = InterpolatedUnivariateSpline(wise_dn_dz[:, 0], wise_dn_dz[:, 1], k=3, ext='zeros')
 norm = dndzwise.integral(0, 2)
 dndzwise = InterpolatedUnivariateSpline(
@@ -111,7 +111,7 @@ desi = gals_kernel.kern(dndz[:, 0], dndzfun_desi, hspline, omega_m, h0)
 
 des_weak = kappa_gals_kernel.kern(dndz_des[:, 0], dndzfun_des, chispline, hspline, omega_m, h0)
 
-l = 500
+l = 100
 
 
 z_kappa = np.linspace(0, 13, 500)
@@ -174,7 +174,7 @@ w_structure = np.zeros_like(z_kappa_gal)
 for i, z in enumerate(z_kappa_gal):
 
     x = chispline(z)
-    w_kappa_gal[i] = 1. / x * des_weak.w_lxz(l, x, z) * np.sqrt(rbs.ev((l + 0.5) / x, z))
+    # w_kappa_gal[i] = 1. / x * des_weak.w_lxz(l, x, z) * np.sqrt(rbs.ev((l + 0.5) / x, z))
     w_structure[i] = np.sqrt(rbs.ev((l + 0.5) / x, z))  # des_weak.w_lxz(l, x, z)
 
 z_cib = np.linspace(0, 13., 500)
@@ -182,7 +182,8 @@ w_cib = np.zeros_like(z_cib)
 for i, z in enumerate(z_cib):
 
     x = chispline(z)
-    w_cib[i] = 1. / x * cib.w_lxz(l, x, z) * np.sqrt(rbs.ev((l + 0.5) / x, z))
+    # w_cib[i] = 1. / x * cib.w_lxz(l, x, z) * np.sqrt(rbs.ev((l + 0.5) / x, z))
+    w_cib[i] = cib.w_lxz(l, x, z)
 
 # plt.plot(z_cib,w_cib,label = 'cib')
 

### 2017-11-23 Code/plots/plot_kernel.py

@@ -50,13 +50,13 @@ colors['SKA'] = colormap(5)
 
 datadir = '../../Data/limber_spectra/'
 
-image_dir = '../../images/'
+image_dir = '../../Notes/images/'
 
 
 # Import data
-kappakernel_l30 = np.loadtxt(datadir + 'kappa_kernel_l30.txt')
+# kappakernel_l30 = np.loadtxt(datadir + 'kappa_kernel_l30.txt')
 kappakernel_l100 = np.loadtxt(datadir + 'kappa_kernel_l100.txt')
-kappakernel_l500 = np.loadtxt(datadir + 'kappa_kernel_l500.txt')
+# kappakernel_l500 = np.loadtxt(datadir + 'kappa_kernel_l500.txt')
 
 lsstkernel_l100 = np.loadtxt(datadir + 'lsst_kernel_l100.txt')
 
@@ -69,14 +69,14 @@ cibkernel_l100 = np.loadtxt(datadir + 'cib_kernel_l100.txt')
 # cibkernel_l500 = np.loadtxt(datadir + 'cib_kernel_l500.txt')
 
 
-deskernel_l30 = np.loadtxt(datadir + 'des_kernel_l30.txt')
+# deskernel_l30 = np.loadtxt(datadir + 'des_kernel_l30.txt')
 deskernel_l100 = np.loadtxt(datadir + 'des_kernel_l100.txt')
-deskernel_l500 = np.loadtxt(datadir + 'des_kernel_l500.txt')
+# deskernel_l500 = np.loadtxt(datadir + 'des_kernel_l500.txt')
 
 
-desikernel_l30 = np.loadtxt(datadir + 'desi_kernel_l30.txt')
+# desikernel_l30 = np.loadtxt(datadir + 'desi_kernel_l30.txt')
 desikernel_l100 = np.loadtxt(datadir + 'desi_kernel_l100.txt')
-desikernel_l500 = np.loadtxt(datadir + 'desi_kernel_l500.txt')
+# desikernel_l500 = np.loadtxt(datadir + 'desi_kernel_l500.txt')
 
 
 # ============================================
@@ -150,7 +150,7 @@ fg = plt.figure(figsize=fig_dims)
 print(fig_dims)
 ax1 = fg.add_subplot(111)
 
-plt.plot(kappakernel_l30[:, 0], kappakernel_l100[:, 1] /
[…]

### 2017-11-23 Notes/referee_response.tex

\documentclass[12pt]{amsart}
\usepackage[top=2.5 cm, bottom=2.5 cm, left=1.05in, right=1.05in]{geometry}
%\geometry{a4paper} % or letter or a5paper or ... etc
% \geometry{landscape} % rotated page geometry
\usepackage[applemac]{inputenc}
\usepackage{amsmath,amssymb} 
% See the ``Article customise'' template for come common customisations

\title{Referee Response}
\author{A. Manzotti}
\date{} % delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle

\begin{enumerate}
\item The abstract is unnecessarily very long and could be considerably
shortened.

\vspace{0.4cm}
The abstract length is now reduced to $~80\%$ of its previous length, mainly removing redundancy and quantitative results that can be found easily in the tables but leaving the main general points.
\vspace{1cm}

\item Putting the lmin-lmax ranges in the tables as well as having the
dashed grey line of Fig. 5 onto Fig 6-7-8 would help the readability
of the paper. 
\vspace{0.4cm}

I agree with the referee on this point. I first thought this can cause a bit of clatter, but the gain in readability is worth the risk.
I added the dashed line to all the figures.

\vspace{1cm}
\item Twice in the paper the author writes that lensing is only one of
secondaries producing B-modes, or his writing gives that impression.
It is unclear to me what is meant by that, as far as I know, galactic
foregrounds should be basically the only expected source of worries
for the experiments considered there.

\vspace{0.4cm}

We agree that the phrasing of the sentences where secondary B-modes sources are introduced can be confusing. While it is well known in the field, it is definitely important to be clear to a wider audience.
For this reason we add this sentence in the introduction:

\textit{Two main effects produce non-primordial B-modes: the polarized foregrounds from our own galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
This two components needs to be treated with very different techniques and in this work we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}

\vspace{1cm}
\item The author introduces sharp cuts to CIB and WISE at l < 100. Is the
alternative approach of incorporating the foreground content in the
auto spectra of the tracer (E.g. Eq. 17) combination possible ? That
would retain some more of the signal presumably. 
\vspace{0.4cm}

\vspace{1cm}
\item Fig 1 might look confusing at first. The SKA kernel there shows
brilliant overlap with the CMB lensing kernel, but is not the most
powerful tracer. On the other hand, DES for instance has very moderate
overlap but can still achieve very high performance on large scales.
Can the author comment and help the reader make sense of that figure ?

\vspace{0.4cm}

\vspace{1cm}

\item Is it clear, as the author claims in IV A with no supporting
evidence, that fixing the cosmology has no impact on the {r,nt}
constraints ? The cosmology fixes the residual B-mode amplitude, which
is unequally important in the delensed vs nominal cases. An slightly
improved discussion would be welcome here.

\vspace{0.4cm}

\vspace{1cm}

\item The author makes a strong point that survey data should be split
into bins, as this makes the correlation increases by 30%. I am not
entirely convinced that this is as relevant as claimed. From Fig 4, it
seems that the improvement is fairly limited on the scales that are
relevant for B-mode delensing. Presumably, that is more important for
T or E delensing. The author might want to comment or adapt.

\vspace{0.4cm}

\vspace{1cm}

\end{enumerate}

\end{document}

### 2017-11-27 Notes/referee_response.tex
working on referee response

@@
 
-\title{Referee Response}
+\title{  \scriptsize{Response to Referee for: "Future cosmic microwave background delensing with galaxy surveys"}}
 \author{A. Manzotti}
@@
 
+Thank you for your very careful review of our paper, and for the comments, corrections and suggestions that ensued.
+
+We sequentially address all of the points raised below.
+While working on this answer, I found a minor bug in the part of the code that plots the CIB kernel in Fig 1 (the CIB kernel peaks is slightly shifted).
+This is fixed now and it has no effect on the results.
+\vspace{1.cm}
+
 \begin{enumerate}
@@
 \vspace{0.4cm}
-The abstract length is now reduced to $~80\%$ of its previous length, mainly removing redundancy and quantitative results that can be found easily in the tables but leaving the main general points.
+The abstract length is now reduced to $\sim 80\%$ of its previous length, mainly removing redundancy and quantitative results that can be found easily in the tables but leaving the main general points.
 \vspace{1cm}
@@
 
-I agree with the referee on this point. I first thought this can cause a bit of clatter, but the gain in readability is worth the risk.
+I agree with the referee on this point. I first thought this could cause a bit of clatter, but the gain in readability is worth the risk.
 I added the dashed line to all the figures.
@@
 
-We agree that the phrasing of the sentences where secondary B-modes sources are introduced can be confusing.
-While it is well known in the field, it is definitely important to be clear to a wider audience.
+We agree that the phrasing of the sentences where secondary B-modes sources are introduced can be confusing. While it is well known in the field, it is  important to be clear to a broader audience.
 For this reason we add this sentence in the introduction:
@@
 \textit{Two main effects produce non-primordial B-modes: the polarized foregrounds from our own galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
-This two components needs to be treated with very different techniques and in this work we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}
+This two components needs to be treated with very different techniques and in this work we will focus on the latter (see (CORE cleaning paper) for a review on the former).}
 
@@
 \vspace{0.4cm}
+Fig 1 can be confusing because it does not tell the entire story. 
+The redshift overlap is not the only ingredient for the delensing efficiency.
+For example, the noise is relevant as well, and this is why for example for S2 experiments CIB can be better than the CMB reconstruction even if the latter has a perfect overlap.
+Additionally, not all the redshift matters at the same way since they correspond to different angular scales (depending on the distance of the tracer) which, as shown by the dashed curves in the figures 5-8, contributes differently.
+Showing all of this in one plot is hard.
 
+However, I think Fig 1 tells most of the story. For example, the referee is right that DES is less efficient than SKA in delensing efficiency (16\%) vs (52\%).
+Furthermore, I think this figure highlights the main difference in the calculation of different tracers spectra (using Eq. 2) for people not familiar with it.
+
+I added to the caption that the efficiency is not proportional to the overlapping area of this plot because different redshifts contribute differently both because of noise and the geometric properties of the lensing kernel.
 \vspace{1cm}
@@
 \vspace{0.4cm}
+Note that indeed, as specified in the text, the chosen cosmology has an impact on the $r$,$n_{t}$ constraints.
+What I claim cosmology has almost no impact on is the delensing efficiency (how much these constraints improve), i.e.
+the ratio nominal/delensed.
+What matters for the efficiency is the \textit{fraction} of removed power.
+That depends only on the factor  $1-\rho^2$ where $\rho$ is the correlation factor.
+The correlation factor has a very little dependence on cosmology in particular given the tight constraints from Planck, and it is expected to be a secondary uncertainty here compared to, for example, the redshift distribution of tracers.
+Indeed $\rho$ contains $C^{II}$ $C^{I\phi}$, $C^{\phi\phi}$ that are integrals, with slightly different weights in redshift, of the power spectrum $P(k)$.
+For this reason, they should scale similarly with cosmology leading to a cancellation in the cosmological dependence of $\rho$.
+
+%Furthermore, the amplitude of the lensing B-modes (more so given that we do not use the reionization peak in the constraints) scale very simply with cosmology.
 
@@
 \item The author makes a strong point that survey data should be split
-into bins, as this makes the correlation increases by 30%. I am not
+into bins, as this makes the correlation increases by 30\%. I am not
 entirely convinced that this is as relevant as claimed. From Fig 4, it
@@
 \vspace{0.4cm}
+In general, I agree with the point that quoting the improvement in the correlation is not entirely fair because the scale at which that happen is also necessary to understand how much power will be removed.
+For this reason, I added in the abstract this quantity too.
+
+Furthermore, while I agree that the improvement would be even more significant for T-E delensing, it is nonetheless also quite important for BB too.
+As can be seen in Figure 4 in the paper LSST shows a significant improvement in the correlation with the lensing potential all the way to $\ell\sim 800$.
+Even for other probes, in general, less powerful like, there is still an improvement for example where the contribution to BB (the dashed line peaks at $\ell\sim350$).
+Below I am attaching a table that shows the improvement in the BB power removed and constraints on $\sigma(r)$.
+
+\begin{table}[htp]
+\caption{Improvement on $\sigma(r)$ constraints together with the BB power removed (in parenthesis) with and without binning.}
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+Survey & No Binning & Binning\\
+\hline
+DES & 1.17 (14\%)& 1.20 (16\%) \\
+\hline
+LSST & 1.67 (40\%) & 2.10 (52\%)\\
+\hline
+DESI & 1.10 (9\%) & 1.13 (12\%)  \\
+\hline
+\end{tabular}
+\end{center}
+\label{default}
+\end{table}%
 

### 2017-11-27 Notes/galaxy_delensing.tex
working on referee response

@@
 
-The cosmic microwave background (CMB) polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe. 
-A particular component, the so-called B-modes, is indeed a direct signature of a prediction of inflation: the presence of gravitational waves in the early universe.
-However, reducing the instrumental noise in future experiments will not be enough to detect this signal.
-Secondary effects in the low redshift universe will also produce non-primordial B-modes adding confusion to the inflationary signal.
-In particular, the gravitational interactions of CMB photons with large scale structures will distort the primordial E-modes adding a lensing B-mode component to the primordial signal.
-Removing the lensing part (``delensing'') from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
-Here we discuss the role of current and future large scale structure surveys in a multi-tracers approach to delensing that will improve thesreconstruction of the lensing potential that lenses the CMB photons and, as a consequence, the delensing efficiency.
-We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$).
-We find that, in general, galaxy surveys should be split into tomographic bins as this can improve the correlation with CMB lensing by 30\%.
-Among currently available surveys, a DES-like galaxy survey can remove about $17\%$ of the lensing signal.
-Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
-With the decrease of instrumental noise, the CMB internal reconstruction will increase its efficiency and the fraction of removed lensing B-modes with CMB alone will rapidly improve from the current level of Planck ($7\%$) and SPTPol ($31\%$) to 3G (56\%) and CMB S4 (81\%) level.
-Nevertheless optical galaxy surveys will still play an important role even for CMB S4.
-In particular an LSST-like survey can a achieve a delensing performance comparable to a 3G CMB experiment but with completely different systematics.
-This will be important to prove the robustness against systematics of an eventual detection of primordial B-modes.
+The primordial B-modes, a particular component of the cosmic microwave background (CMB) polarization are a promising experimental dataset to probe the inflationary paradigm.
+B-modes are indeed a direct consequence of the presence of gravitational waves in the early universe.
+However, several secondary effects in the low redshift universe will produce \textit{non-primordial} B-modes.
+In particular, the gravitational interactions of CMB photons with large-scale structures will distort the primordial E-modes, adding a lensing B-mode component to the primordial signal.
+Removing the lensing part ("delensing") from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
+Here we discuss the role of current and future large-scale structure surveys in a multi-tracers approach to CMB delensing.
+We find that, in general, galaxy surveys should be split into tomographic bins as this can improve the reconstruction of the lensing potential that lenses the CMB photons by 30\%.
+Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large-scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
+With the decrease of instrumental noise, the CMB lensing internal reconstruction will improve, and the fraction of removed lensing B-modes using CMB alone will rapidly improve for CMB S3 and S4.
+Nevertheless, optical galaxy surveys will still play an important role even for CMB S4.
+In particular, an LSST-like survey can a achieve a delensing performance comparable to a 3G CMB experiment but with entirely different systematics.
+This redundancy will be essential to prove the robustness against systematics of an eventual detection of primordial B-modes.
+
 \end{abstract}
@@
 Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe.
-In this work, we will focus on the effect of gravitational interactions with large scale structures (LSS).
+Two main effects produce non-primordial B-modes: the polarized foregrounds from our own galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
+This two components needs to be treated with very different techniques and in this work we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
@@
 \begin{center}
-\includegraphics[scale=1.]{images/compare_kernel.pdf}
+\includegraphics[scale=1.]{./images/compare_kernel.pdf}
 \caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
-The bigger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency.
-Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not simply to the redshift distribution of galaxies.}
+The larger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency.
+However, the delensing efficiency is not exactly proportional to the area of overlap here.
+Indeed each redshift contributes differently to CMB lensing because of both the noise in the measurement and the geometric properties of the lensing kernel.
+Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}
 \label{fig:kernel}
@@
 \begin{center}
-\includegraphics[scale=1.]{images/BB_res_ell2.pdf}
+\includegraphics[scale=1.]{./images/BB_res_ell2.pdf}
 \caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
@@
 \begin{center}
-\includegraphics[scale=1.]{images/compare_kernel_tomo.pdf}
+\includegraphics[scale=1.]{./images/compare_kernel_tomo.pdf}
 \caption{Comparison of kernels of the 10 LSST tomographic bins together with the full LSST survey and the CMB lensing kernel. 
@@
 \begin{center}
-\includegraphics[scale=1.]{images/B_res_bin.pdf}
+\includegraphics[scale=1.]{./images/B_res_bin.pdf}
 \caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing.
@@
 \begin{center}
-\includegraphics[scale=1.]{images/actual_scenario_planck.pdf}
+\includegraphics[scale=1.]{./images/actual_scenario_planck.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers.
@@
 \begin{center}
-\includegraphics[scale=1.]{images/actual_scenario.pdf}
+\includegraphics[scale=1.]{./images/actual_scenario.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers. 
@@
 \begin{center}
-\includegraphics[scale=1.]{images/S3_scenario.pdf}
+\includegraphics[scale=1.]{./images/S3_scenario.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
@@
 \begin{center}
-\includegraphics[scale=1.]{images/S4_scenario_ska10.pdf}
+\includegraphics[scale=1.]{./images/S4_scenario_ska10.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
@@
 The conclusions about the relative importance of LSS tracers we find for $\sigma(r)$ are still valid for $\sigma(n_{t})$.
-Delensing can then be useful also to constrain more exotic inflationary models.
+Even if a test of consistency relation is not possible even for S4 experiment, delensing can still be useful also to constrain more exotic inflationary models.
 
@@
 \begin{figure*}[ht]
-\includegraphics[scale=0.8]{images/errors_summary2.pdf}
+\includegraphics[scale=0.8]{./images/errors_summary2.pdf}
 \caption[]{A summary of the amount of lensing B-modes power removed by different tracers for different generations of experiments.

### 2017-11-27 Notes/referee_response.tex
working on referee response

@@
 \vspace{1cm}
-\item The author introduces sharp cuts to CIB and WISE at l < 100. Is the
+\item The author introduces sharp cuts to CIB and WISE at $\ell < 100$. Is the
 alternative approach of incorporating the foreground content in the
@@
 \vspace{0.4cm}
+This is a good idea.
+The real obstacle to that is our current knowledge of foreground spectra.
+There is not a trustworthy model for that.
+However one might measure that empirically by measuring $C^{II}$ and $C^{I\phi}$ directly from the data.
+Doing that will call everything in the tracer map that does not correlate with $\phi$, "noise".
+That noise will include foregrounds too.
+However..
+Given the small amount of delensing power for $\ell<100$ the field decided, so far, to cut those in forecasts.
+Given how much foregrounds are complicated I believed only data will answer this.
+An exploration of these on simulations is probably beyond the scope of this paper, but very soon future delensing work will need to explore this on real data.
 

### 2017-11-28 Notes/referee_response.tex
working on referee response

@@
 \vspace{0.4cm}
+
 This is a good idea.
@@
 
+\item
+The author uses iterative internal delensing noise levels for the
+CMB. I believe this could be discussed in more details. Can he specify
+what multipole range is considered for this calculation ? I suppose
+this multipole range should not overlap with the primordial B that is
+used for constraints on r, however, this range (unnecessarily) extends
+up to 800 in the S3 section for example. Most likely, 200 would be
+largely enough, does it imply a sizable reduction on the iterative
+reconstruction noise levels ? Besides, iterative delensing is
+complicated and not established yet, in contrast to Quadratic
+Estimator delensing. The paper might gain in interest if numbers for
+Q.E. and iterative delensing were presented.
+
+\vspace{0.4cm}
+
+I
+
 \end{enumerate}

### 2017-11-28 Code/Plot_figures.py
working on referre add ines

@@ -258,7 +258,7 @@ lbins, rho, rho_comb, rho_gals, rho_cmb = multiple_survey_delens.main(
     labels, cmb)
 plt.plot(lbins, rho_gals, label='LSS-S3+LSST(LSS-S4)', color=colormap(7))
 
-plt.plot(BB_contr[:, 0], BB_contr[:, 1] - np.min(BB_contr[:, 1]),
+plt.plot(BB_contr[:, 0], BB_contr[:, 1] - np.min(BB_contr[:, 1])+0.45,
          '--', alpha=0.6, linewidth=font_size / 14., color='k')
 
 plt.legend(loc=0, ncol=2)

### 2017-11-28 Notes/referee_response.tex
working on referee response

@@
 
-Thank you for your very careful review of our paper, and for the comments, corrections and suggestions that ensued.
+I want to thank the referee for the very careful review of the paper, and for the comments, corrections and suggestions that ensued.
 
@@
 
-I
+One information that is probably not clear enough or partially missing in the text is that iterative delensing compared to iterative delensing is important only for CMB-S4 noise levels.
+For example, the amount of BB power removed for CMB-S3 level using iterative vs.
+QE is 56\% vs.
+58\%.
+
+Furthermore, the idea of a separation in scale between the CMB data used to reconstruct the potential and the data used to constrained r, to solve the so-called "internal delensing bias" is important for both QE and iterative approach.
+While important, we left a full study...
+
+In some sense, we assume that either the technique applied in Namikava 2017 or () can work without a significant loss of signal. (put that in the text?s)
+Again to test this one needs either real data or full ray tracing simulation for CMB lensing.
+For example, the deep 100 deg$^{2}$ field measured by SPTPol that can reach a depth of 5-6 muK arcmin with high resolution might be the first dataset where this can be explored.
+
+Following the referee comments, we
+\begin{itemize}
+\item Make it clear that only CMB S4 will benefit from iterative delensing.
+\item Add the level of delensing that is possible to achieve with CMB-S4 if the well tested QE is used instead of the iterative approach.
+This gives in some sense a range of possible efficiency for CMB S4 experiments.
+\end{itemize}
 

### 2017-11-28 Notes/galaxy_delensing.tex
working on referee response

@@
 In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}.
+The iterative approach has not been fully demonstrated on data yet but it is among the major goals of the CMB community (\cite{abazajian:2016}).
+Note however that, as pointed out in in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach only for CMB-S4 noise levels.
 
@@
 Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
+For CMB S3, the improvement due to an iterative approach compared to the quadratic estimator is still minimal so that the more traditional technique can be used.
+For example, in this case, the removed power goes from $56\%$ to $54\%$ (quoted in \reftab{S3}) when using a quadratic estimator instead of the iterative one.
 

### 2017-12-03 Notes/referee_response.tex
working on referee response

@@
 % See the ``Article customise'' template for come common customisations
+\usepackage{bm}
+\usepackage{epsfig}
+\usepackage{graphicx,epsfig}
+\usepackage{hyperref}
+\usepackage{ifthen}
+\usepackage [english]{babel}
+\usepackage{xstring}
+\usepackage[applemac]{inputenc}
+\usepackage{amsmath,amssymb}
+\usepackage{color}
+\usepackage{epstopdf}
+\usepackage{caption}
+\captionsetup{justification=centerlast,singlelinecheck=false,font=small,skip=2pt}
+% notes to make comment is non-zero if
+\newcommand{\al}[1]{\begin{align} #1 \end{align}}
+
+\newcommand{\note}[1]{\textsc{#1}}
+	\newcommand{\bR}[1]{{\bm {\mR{#1}}} }
+
+\def\l{\ell}
+% For various journals
+\newcommand{\apj}{ApJ}
+\newcommand{\physrep}{Physics Reports}
+\newcommand{\jcap}{JCAP}
+\newcommand{\apjl}{ApJL}
+\newcommand{\araa}{ARA\&A}
+\newcommand{\apjs}{ApJS}
+\newcommand{\aap}{A\&A}
+\newcommand{\mnras}{MNRAS}
+\newcommand{\physrev}{Phys. Rev.}
+\newcommand{\physrevlett}{Phys. Rev. Lett.}
+\newcommand*\aj{AJ}
+%\newcommand*\prd{Phys.~Rev.~D}
+
+\newcommand{\nver}{\hat{\mathbf{n}}}
+\newcommand{\cov}{\text{Cov}}
+\newcommand{\Nsim}{N_{\text{sim}}}
+\newcommand{\relook}[1]{{\textcolor{blue}{\bf #1}}}
+%--------- NEW COMMAND STUFF
+	\newcommand{\PD}[2]{\dfrac{\partial #1}{\partial #2}}
+
+\def\simlt{\lesssim}
+\def\simgt{\gtrsim}
+\newcommand{\Omegamzero}{\Omega_{{\rm m,0}}}
+\newcommand{\alm}{a_{lm}}
+	% * subscript
+	\def\rom#1{%
+		_{\mathrm{#1}}%
+	}%
+%\newcommand{\l}{\ell}}
+
+\def\be{\begin{equation}}
+\def\ee{\end{equation}}
+\def\ben{\begin{equation*}}
+\def\een{\end{equation*}}
+
+\def\ba{\begin{eqnarray}}
+\def\ea{\end{eqnarray}}
+\def\ban{\begin{eqnarray*}}
+\def\ean{\end{eqnarray*}}
+
+\newcommand{\refsec}[1]{section~\ref{sec:#1}}
+\newcommand{\reftab}[1]{Tab.~\ref{tab:#1}}
+\newcommand{\refeq}[1]{Eq.~(\ref{eqn:#1})}
+\newcommand{\refssec}[1]{section~\ref{subsec:#1}}
+\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
+\newcommand{\refFig}[1]{Fig.~\ref{fig:#1}}
+	\newcommand{\mR}[1]{\mathrm{#1}}   % mathrm
+
+\definecolor{darkgreen}{cmyk}{0.85,0.2,1.00,0.2}
+\definecolor{purple}{cmyk}{0.5,1.0,0,0}
+
+\usepackage{hyperref}
+\usepackage{amsmath}
+\usepackage{natbib}
+\newcommand{\bx}{{\bf x}}
+\newcommand{\bk}{{\bf k}}
+\newcommand{\bq}{{\bf q}}
+\newcommand{\bP}{{\bf\Psi}}
+\newcommand{\bs}{{\bf s}}
+\newcommand{\cs}{{\cal S}}
+\newcommand{\by}{{\bf y}}
+\newcommand{\deltar}{\delta_{\rm recon}}%\def\lcdm{$\Lambda$CDM}
+\def\arcsec{$^{\prime\prime}$}
+\def\nl{N_\ell}
+\def\sl{S_\ell}
+\def\ublu{\bl'}
+\newcommand{\conv}[2]{\left(\frac{#1}{#2}\right)}
+\def\arctanh{\mathop{\rm arctanh}\nolimits}
+%\renewcommand{\eqref}[1] {equation $($\ref{#1}$)$}
+\newcommand{\comment}[1]{{}}
+\def\beq{\begin{equation}}
+\def\eeq{\end{equation}}
+\def\beqn{\begin{eqnarray}}
+\def\eeqn{\end{eqnarray}}
+\def\a{\alpha}
+\def\cl{C_{\ell}}
+\def\h{\mathrm{h}}
+\def\d{\rmn{d}}
+\def\pa{\partial}
+\def\deldel#1#2{\frac{\pa{#1}}{\pa{#2}}}
+\def\ba{\bm{\alpha}}
+\def\fracj#1#2{{\textstyle{#1\over#2}}}
+\def\bxi{\bm{\xi}}
+\def\half{\frac{1}{2}}
+\def\ti{\widetilde}
+\def\O{\Omega}
+\def\OL{\Omega_\Lambda}
+\def\Om{\ensuremath{\Omega_{\mathrm{m}}}}
+\def\Ob{\ensuremath{\Omega_{\mathrm{b}}}}
+\def\Oc{\ensuremath{\Omega_{\mathrm{CDM}}}}
+\def\msol{\ensuremath{M_\odot}}
+\def\l{\left}
+\def\r{\right}
+\def\o{\omega}
+\def\gcm{\textrm{g cm$^{-3}$}}
+\def\2gcm{\textrm{g cm$^{-2}$}}
+\def\Scr{\Sigma_{\mathrm{crit}}}
+\def\rcr{\rho_{\mathrm{crit}}}
+\def\phidot{\ensuremath{\dot\phi_{\bl,\bl'}}}
+\def\ddelta{\ensuremath{\dot\delta}}
+\def\modu#1{\l |{#1}\r |}
+\def\av#1{\l \langle{#1}\r \rangle}
+\def\hmpc{\:{h}^{-1}\mathrm{Mpc}}
+\def\th{\Theta}
+\def\tth{\tilde\Theta}
+\def\sg{\sigma}
+\def\Sig{\Sigma}
+\def\cf{{\cal F}}
+\def\k{\kappa}
+\def\P{{P}}
+\def\pnl{{P}_{{\!\textrm{\tiny NL}}}}
+\def\dnl{\Delta_{\text{\scriptsize NL}}}
+\def\kmin{\k_{\mathrm{min}}}
+\def\kmax{\k_{\mathrm{max}}}
+\def\hires{the \emph{highRes} experiment}
+\def\lores{the \emph{lowRes} experiment}
+\def\ktot{\k_{\mathrm{tot}}}
+\def\hunit{\ensuremath{\mathrm{km}{\mathrm{s}^{-1}} \mathrm{Mpc}^{-1}}}
+\def\H0{\ensuremath{\mathrm{H}_0}}
+\def\nn{\nonumber}
+\def\lin{\mathrm{lin}}
+\def\ISW{\mathrm{ISW}}
+\def\bl{\bmm{l}}
+\def\bL{\bmm{L}}
+\def\fsky{f_{\mathrm{sky}}}
+\newcommand{\E}[1]{\times 10^{#1}}
+\newcommand{\bmm}[1]{{\mathbf{#1}}}
+\newcommand{\bsection}[1]
 
@@
 \begin{document}
-
+\pagestyle{empty}
 \maketitle
 
-I want to thank the referee for the very careful review of the paper, and for the comments, corrections and suggestions that ensued.
+I want to thank the referee for the very careful review of the paper, and for the comments, corrections, and suggestions that ensued.
 
 We sequentially address all of the points raised below.
-While working on this answer, I found a minor bug in the part of the code that plots the CIB kernel in Fig 1 (the CIB kernel peaks is slightly shifted).
-This is fixed now and it has no effect on the results.
+While working on this answer, I found a minor bug in the part of the code that plots the CIB kernel in Fig 1 (the CIB kernel peak is slightly shifted).
+This is fixed now, and it does not affect the results.
 \vspace{1.cm}
@@
 \vspace{0.4cm}
-The abstract length is now reduced to $\sim 80\%$ of its previous length, mainly removing redundancy and quantitative results that can be found easily in the tables but leaving the main general points.
+I agree.
+The abstract length is now reduced to $\sim 80\%$ of its previous length, mainly removing redundancy and quantitative results that can be found easily in the tables while leaving the main general points.
+I believe it is still a little bit longer than the average but reasonably so.
 \vspace{1cm}
@@
 I added the dashed line to all the figures.
+The $\ell$ range used for the inflationary constraints is also at the end of each table both for readability and reproducibility.
+The choice was quite arbitrary tough.
+We used all the scales available to the instrument.
+The range where the bulk of the signal to noise is, has been explored in some of the referenced work (\cite{wu:2014,abazajian:2016} etc.) but not here.
 
@@
 
+\vspace{1cm}
 \textit{Two main effects produce non-primordial B-modes: the polarized foregrounds from our own galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
-This two components needs to be treated with very different techniques and in this work we will focus on the latter (see (CORE cleaning paper) for a review on the former).}
+This two components needs to be treated with very different techniques and in this work we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}
 
@@
 
-This is a good idea.
+This is indeed a good idea.
 The real obstacle to that is our current knowledge of foreground spectra.
-There is not a trustworthy model for that.
-However one might measure that empirically by measuring $C^{II}$ and $C^{I\phi}$ directly from the data.
+There is not an accurate model for that (at higher scales we use a simple power law).
+However one might measure that empirically, given a dataset with enough signal to noise on those large scales, by measuring $C^{II}$ and $C^{I\phi}$ directly from the data.
 Doing that will call everything in the tracer map that does not correlate with $\phi$, "noise".
 That noise will include foregrounds too.
-However..
+However, it is unclear if eventual dust residual can cause biases in the delensed B-modes (if they correlate with dust contamination in CMB polarizations).
 Given the small amount of delensing power for $\ell<100$ the field decided, so far, to cut those in forecasts.
-Given how much foregrounds are complicated I believed only data will answer this.
+Given how much foregrounds are complicated I believed only data would answer this.
 An exploration of these on simulations is probably beyond the scope of this paper, but very soon future delensing work will need to explore this on real data.
@@
 \vspace{0.4cm}
-Fig 1 can be confusing because it does not tell the entire story. 
+I agree Fig 1 may be a little confusing because it does not tell the entire story. 
 The redshift overlap is not the only ingredient for the delensing efficiency.
-For example, the noise is relevant as well, and this is why for example for S2 experiments CIB can be better than the CMB reconstruction even if the latter has a perfect overlap.
-Additionally, not all the redshift matters at the same way since they correspond to different angular scales (depending on the distance of the tracer) which, as shown by the dashed curves in the figures 5-8, contributes differently.
+For example, the noise is relevant as well, and this is why for example for S2 experiments CIB can be better than the CMB internal reconstruction even if the latter has a perfect overlap.
+Additionally, not all the redshifts matter at the same level since they correspond to different angular scales (depending on the distance of the tracer) which, as shown by the dashed curves in the figures 5-8, contributes differently.
 Showing all of this in one plot is hard.
@@
 However, I think Fig 1 tells most of the story. For example, the referee is right that DES is less efficient than SKA in delensing efficiency (16\%) vs (52\%).
-Furthermore, I think this figure highlights the main difference in the calculation of different tracers spectra (using Eq. 2) for people not familiar with it.
+Furthermore, I think this figure highlights the central part (the "kernel") that characterize different tracers spectra (using Eq. 2) for people not familiar with it.
 
-I added to the caption that the efficiency is not proportional to the overlapping area of this plot because different redshifts contribute differently both because of noise and the geometric properties of the lensing kernel.
+I added to the caption a sentence to highlight that the efficiency is not proportional to the overlapping area of this plot because different redshifts contribute differently due to instrumental noise and the geometric properties of the lensing kernel.
 \vspace{1cm}
@@
 \vspace{0.4cm}
-Note that indeed, as specified in the text, the chosen cosmology has an impact on the $r$,$n_{t}$ constraints.
-What I claim cosmology has almost no impact on is the delensing efficiency (how much these constraints improve), i.e.
-the ratio nominal/delensed.
+Note that indeed, as specified in the text, the chosen cosmology \textit{has} an impact on the absolute value of the $r$,$n_{t}$ constraints.
+What I claim is that cosmology has almost no impact on, in particular if compared to other uncertainties, is the delensing efficiency (how much these constraints improve), i.e.
+the ratio $C^{BB}_{\ell}$ nominal/delensed.
 What matters for the efficiency is the \textit{fraction} of removed power.
 That depends only on the factor  $1-\rho^2$ where $\rho$ is the correlation factor.
-The correlation factor has a very little dependence on cosmology in particular given the tight constraints from Planck, and it is expected to be a secondary uncertainty here compared to, for example, the redshift distribution of tracers.
-Indeed $\rho$ contains $C^{II}$ $C^{I\phi}$, $C^{\phi\phi}$ that are integrals, with slightly different weights in redshift, of the power spectrum $P(k)$.
+The correlation factor has a very little dependence on cosmology in particular given the tight constraints from Planck on $\Omega_{m}h^2$ and $\sigma_{8}$, and it is expected to be a secondary uncertainty here compared to, for example, the redshift distribution of tracers.
+Indeed $\rho$ contains $C^{II}$ $C^{I\phi}$, $C^{\phi\phi}$ that are integrals, with slightly different weights in redshift, of the cosmology dependent power spectrum $P(k)$.
 For this reason, they should scale similarly with cosmology leading to a cancellation in the cosmological dependence of $\rho$.
+
+Testing this intuition is important.
+For this reason, I run the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \cite{planck-collaboration:2016b}.
+The efficiency values changes, but by at most $7\%$.
+Given that the Planck constraints are way tighter I consider this, as expected, a present but subdominant effect.
+I added a line to the paper to report this test.
 
@@
 \vspace{0.4cm}
-In general, I agree with the point that quoting the improvement in the correlation is not entirely fair because the scale at which that happen is also necessary to understand how much power will be removed.
+In general, I agree with the point that quoting the improvement in the correlation is not entirely fair because the scales at which that improvement happens is also necessary to understand how much power will be removed.
 For this reason, I added in the abstract this quantity too.
@@
 As can be seen in Figure 4 in the paper LSST shows a significant improvement in the correlation with the lensing potential all the way to $\ell\sim 800$.
-Even for other probes, in general, less powerful like, there is still an improvement for example where the contribution to BB (the dashed line peaks at $\ell\sim350$).
+Even for other probes, in general, less powerful, there is still an improvement for example where the contribution to BB (the dashed line peaks at $\ell\sim350$).
 Below I am attaching a table that shows the improvement in the BB power removed and constraints on $\sigma(r)$.
@@
 For example, the amount of BB power removed for CMB-S3 level using iterative vs.
-QE is 56\% vs.
-58\%.
+QE is 51\% vs.
+56\%.
 
@@
 
+\bibliographystyle{mnrasunsrt}
+\bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt,
+/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
+
 \end{document}

### 2017-12-03 Notes/galaxy_delensing.tex
working on referee response

@@
 The iterative approach has not been fully demonstrated on data yet but it is among the major goals of the CMB community (\cite{abazajian:2016}).
-Note however that, as pointed out in in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach only for CMB-S4 noise levels.
+Note however that, as pointed out in in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach mainly for CMB-S4 noise levels.
 
@@
 Uncertainties in those will propagate to larger uncertainties in $\{r,n_{t}\}$.
-While neglecting this will lead to slightly optimistic constraints, it has no significant impact on the estimate of the improvement due to delensing.
+While neglecting this will lead to slightly optimistic absolute constraints, it has no significant impact on the estimate of the improvement due to delensing.
+For example, running the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \cite{planck-collaboration:2016b}, change the delensing improvement factors by at most $6-7\%$.
 
@@
 %\newline
-The values in the other columns correspond to ratio of the error before and after delensing for 3 cases: no instrumental noise in the B-mode measurement (but with instrumental noise in the E-mode used to delens) and with instrumental noise for two different values of $r$.}
+The values in the other columns correspond to ratio of the error before and after delensing for 3 cases: no instrumental noise in the B-mode measurement (but with instrumental noise in the E-mode used to delens) and with instrumental noise for two different values of $r$.
+The error $\sigma(r)$ is computed using \refeq{fisher} with $50<\ell<500$.}
   \vspace{0.2cm}
@@
 We also use a CMB S3 experiment for the lensing internal reconstruction.
-See \refsec{S3} for details.}
+The error $\sigma(r)$ is computed using \refeq{fisher} with $50<\ell<800$, see \refsec{S3} for details.}
 %The absolute error value of $\sigma(r)$ after delensing for $f_{sky} = $ is $\sigma(r) =$ 
@@
 The improvement for $\sigma(n_{t})$ is computed for fiducial values $r=0.07$ and $n_t=-r/8$.
-See \refsec{S4} for details.}
+The error $\sigma(r)$ is computed using \refeq{fisher} with $50<\ell<800$ see \refsec{S4} for details.}
 \begin{tabular}{|c | c | c | c| c|}
@@
 \bibliographystyle{mnrasunsrt}
-\bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/2016/delens100d/delens100d,/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
-%/Users/alessandromanzotti/Work/Astrophysics/delensing/sptpol_papers/BIBTEX/spt}
+\bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/galaxies_delensing/Notes/spt}
 

### 2017-12-04 Notes/referee_response.tex
working on referee response

@@
 \newcommand*\aj{AJ}
-%\newcommand*\prd{Phys.~Rev.~D}
+\newcommand*\prd{Phys.~Rev.~D}
 
@@
 
-One information that is probably not clear enough or partially missing in the text is that iterative delensing compared to iterative delensing is important only for CMB-S4 noise levels.
-For example, the amount of BB power removed for CMB-S3 level using iterative vs.
-QE is 51\% vs.
-56\%.
+%One information that is probably not clear enough or partially missing in the text is that iterative delensing compared to iterative delensing is important only for CMB-S4 noise levels. 
+%For example, the amount of BB power removed for CMB-S3 level using iterative vs. QE is 51\% vs. 56\%.
 
-Furthermore, the idea of a separation in scale between the CMB data used to reconstruct the potential and the data used to constrained r, to solve the so-called "internal delensing bias" is important for both QE and iterative approach.
-While important, we left a full study...
+Let me start from the first part of the comment which regards the scales used here.
+The idea of a separation in scale between the CMB data used to reconstruct the potential and the data used to constrained r, is indeed an interesting idea proposed (for example in \cite{sehgal:2017}) to solve the so-called "internal delensing bias".
+This is important for both QE and iterative approach.
+This and other methods have shown that, at least on simulations, we can avoid this bias \textit{without losing performance}.
+Other groups have shown that it is possible to use overlapping scales and then correct for the bias.
+For this reason, we can reconstruct the lensing potential using scales that overlap with the primordial B-mode, without being over-optimistic. 
+While a full treatment of this effects with realistic simulations is important, we left a full study for future works.
+Again this can also be tested on data.
+For example, the deep 100 deg$^{2}$ field measured by SPTPol that can reach a depth of $5-6 ~\mu$K arcmin with high resolution might be the first dataset where this can be explored.
 
-In some sense, we assume that either the technique applied in Namikava 2017 or () can work without a significant loss of signal. (put that in the text?s)
-Again to test this one needs either real data or full ray tracing simulation for CMB lensing.
-For example, the deep 100 deg$^{2}$ field measured by SPTPol that can reach a depth of 5-6 muK arcmin with high resolution might be the first dataset where this can be explored.
+Regarding the second part of the comment, I agree that it is interesting to quickly investigate the loss in performance if the quadratic estimator is used instead of the iterative one.
+For this reason, I added a new case for S3 and S4 experiment: the quadratic estimator case.
+Since this is a widely used technique, I think it can be presented without much introduction.
+I added a few sentences at the end of section IIA and in the S3 and S4 sections.
+Note that there is no difference between these two approaches for current levels of noise (S2).
 

### 2017-12-04 Notes/galaxy_delensing.tex
working on referee response

@@
 DESI (11\%) &1.13 & 1.1  & 1.04 \\ \hline
-LSS-S3 (45\%) &1.82& 1.59  & 1.18\\ \hline
+LSS-S3 (45\%) &1.82& 1.59  & 1.18\\
+
+\hline
+CMB 3G (QE) (51\%) &2.0& 1.68  & 1.20 \\ 
+\hline
 CMB 3G (56\%) &2.31& 1.9  & 1.23 \\ \hline
@@
 LSS-S4 + SKA (62\%) &2.6 & 2.3  & 1.3& 1.6  \\ \hline
+CMB S4 (QE) (73\%) & 3.7 & 2.9  & 1.35& 2.3 \\ \hline
 CMB S4 (83\%) & 6.22 & 4.1  & 1.42& 2.3 \\ \hline

### 2017-12-05 Notes/referee_response.tex
major working on referee response

@@
 We used all the scales available to the instrument.
-The range where the bulk of the signal to noise is, has been explored in some of the referenced work (\cite{wu:2014,abazajian:2016} etc.) but not here.
+The range where the bulk of the signal to noise is, has been explored in some of the referenced work (\cite{dodelson:2014,abazajian:2016} etc.) but not here.
 
@@
 We agree that the phrasing of the sentences where secondary B-modes sources are introduced can be confusing. While it is well known in the field, it is important to be clear to a broader audience.
-For this reason we add this sentence in the introduction:
+For this reason, we add this sentence in the introduction:
 
 \vspace{1cm}
-\textit{Two main effects produce non-primordial B-modes: the polarized foregrounds from our own galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
-This two components needs to be treated with very different techniques and in this work we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}
+\textit{Two main effects produce non-primordial B-modes: the polarized foregrounds from the Galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
+These two components need to be treated with very different techniques and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}
+
+\vspace{0.2 cm}
+
+Note that if the referee meant that the galactic B-modes component is the only relevant secondary component compared to lensing B-modes, I do not think that is necessarily true. 
+While on most of the sky the galactic component is the higher contribution in power, that is not true for an opportunely chose small say 1\% patch. Furthermore, that is probably not the best metric.
+As shown in the CMB-S4 science book \cite{abazajian:2016} if one wants to maximize the $r$ constraints, lensing and foregrounds are equally important. 
+The importance there is measured as the number of detectors used to either improve the lensing reconstruction or to observe the sky at different frequencies to clean the foregrounds component.
+This depends slightly on the fraction of the sky used, but it is true for a wide range that contains all possible future ground experiments.
 
@@
 This is indeed a good idea.
-The real obstacle to that is our current knowledge of foreground spectra.
+The real obstacle here is our current knowledge of foreground spectra.
 There is not an accurate model for that (at higher scales we use a simple power law).
@@
 Given how much foregrounds are complicated I believed only data would answer this.
-An exploration of these on simulations is probably beyond the scope of this paper, but very soon future delensing work will need to explore this on real data.
+An exploration of these on simulations is probably beyond the scope of this paper, but very soon future delensing work will test this on real data.
 
@@
 \vspace{0.4cm}
-I agree Fig 1 may be a little confusing because it does not tell the entire story. 
+I agree that Fig 1 may be a little confusing because it does not tell the entire story. 
 The redshift overlap is not the only ingredient for the delensing efficiency.
 For example, the noise is relevant as well, and this is why for example for S2 experiments CIB can be better than the CMB internal reconstruction even if the latter has a perfect overlap.
-Additionally, not all the redshifts matter at the same level since they correspond to different angular scales (depending on the distance of the tracer) which, as shown by the dashed curves in the figures 5-8, contributes differently.
+Additionally, not all the redshifts matter at the same level since they correspond to different angular scales (depending on the distance of the tracer) which, as shown by the dashed curves in the figures 5-8, contributes differently to the generation of CMB B-modes.
 Showing all of this in one plot is hard.
@@
 
-I added to the caption a sentence to highlight that the efficiency is not proportional to the overlapping area of this plot because different redshifts contribute differently due to instrumental noise and the geometric properties of the lensing kernel.
+I added to the caption of Fig.
+1 a sentence to highlight that the efficiency is not proportional to the overlapping area of this plot because different redshifts contribute differently due to instrumental noise and the geometric properties of the lensing kernel.
 \vspace{1cm}
@@
 \vspace{0.4cm}
-Note that indeed, as specified in the text, the chosen cosmology \textit{has} an impact on the absolute value of the $r$,$n_{t}$ constraints.
-What I claim is that cosmology has almost no impact on, in particular if compared to other uncertainties, is the delensing efficiency (how much these constraints improve), i.e.
+Note that indeed, as specified in the text, the chosen cosmology \textit{has} an impact on the absolute value of the ($r$,$n_{t}$) constraints.
+What I claim cosmology has almost no impact on, in particular if compared to other uncertainties, is the delensing efficiency (how much these constraints improve), i.e.
 the ratio $C^{BB}_{\ell}$ nominal/delensed.
@@
 Indeed $\rho$ contains $C^{II}$ $C^{I\phi}$, $C^{\phi\phi}$ that are integrals, with slightly different weights in redshift, of the cosmology dependent power spectrum $P(k)$.
-For this reason, they should scale similarly with cosmology leading to a cancellation in the cosmological dependence of $\rho$.
+For this reason, they should scale similarly with cosmology leading to a substantial cancellation in the cosmological dependence of $\rho$.
 
-Testing this intuition is important.
-For this reason, I run the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \cite{planck-collaboration:2016b}.
-The efficiency values changes, but by at most $7\%$.
+Different cosmologies can also shifts the scales where the information on ($r$, $n_{t}$) come from by changing the shape of the $C_{\ell}^{BB}$ spectra.
+
+I agree that testing this intuition is important.
+For this reason, I run the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \citep{planck-collaboration:2016b}.
+The efficiency values changes, but by at most $8-9\%$.
 Given that the Planck constraints are way tighter I consider this, as expected, a present but subdominant effect.
-I added a line to the paper to report this test.
+For example for $1\%$ higher $\Omega h^{2}$ the efficiency values changes by at most $1.5\%$
+I added few sentences to the paper to report this test.
 
@@
 As can be seen in Figure 4 in the paper LSST shows a significant improvement in the correlation with the lensing potential all the way to $\ell\sim 800$.
-Even for other probes, in general, less powerful, there is still an improvement for example where the contribution to BB (the dashed line peaks at $\ell\sim350$).
+Even for other probes, in general, less powerful, there is still an improvement for example at \textit{all} the scales where the contribution to BB (the dashed line) is relevant $50<\ell<400$.
 Below I am attaching a table that shows the improvement in the BB power removed and constraints on $\sigma(r)$.
+Note that the 30\% level quoted in the abstract corresponds to the most futuristic tracer, i.e., LSST. That should be explicitly stated, so I added this to the abstract.
 
@@
 Let me start from the first part of the comment which regards the scales used here. 
-The idea of a separation in scale between the CMB data used to reconstruct the potential and the data used to constrained r, is indeed an interesting idea proposed (for example in \cite{sehgal:2017}) to solve the so-called "internal delensing bias".
+For reproducibility, I added the scales used for the lensing reconstruction for the different CMB stages.
+Furthermore, the idea of a separation in scale between the CMB data used to reconstruct the potential and the data used to constrained $r$, is indeed an interesting idea proposed (for example in \cite{sehgal:2017}) to solve the so-called "internal delensing bias".
 This is important for both QE and iterative approach.
@@
 Other groups have shown that it is possible to use overlapping scales and then correct for the bias.
-For this reason, we can reconstruct the lensing potential using scales that overlap with the primordial B-mode, without being over-optimistic. 
+For this reason, we assume here that we can reconstruct the lensing potential using scales that overlap with the primordial B-mode, without being over-optimistic. 
 While a full treatment of this effects with realistic simulations is important, we left a full study for future works.
 Again this can also be tested on data.
-For example, the deep 100 deg$^{2}$ field measured by SPTPol that can reach a depth of $5-6 ~\mu$K arcmin with high resolution might be the first dataset where this can be explored.
+For example, the deep 100 deg$^{2}$ field measured by SPTPol that can reach a depth of $5-6 ~\mu$K arcmin with high resolution might be the the first dataset where this can be explored.
+I comment on this at the end of section IIA.
 
@@
 
-Following the referee comments, we
-\begin{itemize}
-\item Make it clear that only CMB S4 will benefit from iterative delensing.
-\item Add the level of delensing that is possible to achieve with CMB-S4 if the well tested QE is used instead of the iterative approach.
+%Following the referee comments, we
+%\begin{itemize}
+%\item Make it clear that only CMB S4 will benefit from iterative delensing.
+%\item Add the level of delensing that is possible to achieve with CMB-S4 if the well tested QE is used instead of the iterative approach.
 This gives in some sense a range of possible efficiency for CMB S4 experiments.
-\end{itemize}
-
+%\end{itemize}
+%
+% 
 \end{enumerate}

### 2017-12-05 Notes/galaxy_delensing.tex
major working on referee response

@@
 
-The primordial B-modes, a particular component of the cosmic microwave background (CMB) polarization are a promising experimental dataset to probe the inflationary paradigm.
+The primordial B-modes component of the cosmic microwave background (CMB) polarization is a promising experimental dataset to probe the inflationary paradigm.
 B-modes are indeed a direct consequence of the presence of gravitational waves in the early universe.
@@
 In particular, the gravitational interactions of CMB photons with large-scale structures will distort the primordial E-modes, adding a lensing B-mode component to the primordial signal.
-Removing the lensing part ("delensing") from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves.
+Removing the lensing component ("delensing") will then be necessary to constrain the amplitude of the primordial gravitational waves.
 Here we discuss the role of current and future large-scale structure surveys in a multi-tracers approach to CMB delensing.
-We find that, in general, galaxy surveys should be split into tomographic bins as this can improve the reconstruction of the lensing potential that lenses the CMB photons by 30\%.
+We find that, in general, galaxy surveys should be split into tomographic bins as this can increase the reduction of lensing B-modes by $\sim 25\%$ in power in the most futuristic case.
 Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large-scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
-With the decrease of instrumental noise, the CMB lensing internal reconstruction will improve, and the fraction of removed lensing B-modes using CMB alone will rapidly improve for CMB S3 and S4.
+With the decrease of instrumental noise, the amount of lensing B-modes power removed using CMB internal reconstruction alone will rapidly increase.
 Nevertheless, optical galaxy surveys will still play an important role even for CMB S4.
@@
 Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe.
-Two main effects produce non-primordial B-modes: the polarized foregrounds from our own galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
-This two components needs to be treated with very different techniques and in this work we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).
+\textcolor{blue}{Two main effects produce non-primordial B-modes: the polarized foregrounds from the Galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
+These two components need to be treated with very different techniques, and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
@@
 In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}.
-The iterative approach has not been fully demonstrated on data yet but it is among the major goals of the CMB community (\cite{abazajian:2016}).
-Note however that, as pointed out in in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach mainly for CMB-S4 noise levels.
+\textcolor{blue}{The iterative approach has not been sufficiently demonstrated on data yet, but it is among the future goals of the CMB community \cite{abazajian:2016}.
+Note however that, as pointed out in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach mainly for S3 and S4 noise levels.
+For these two cases, we also show the delensing efficiency obtained when the CMB lensing is reconstructed with the standard quadratic estimator (QE) approach \cite{okamoto:2003,hu:2002}.
+Furthermore, we do not deal here with the presence of an internal delensing bias.
+Several promising approaches have been proposed \cite{sehgal:2017,namikawa:2017} even if they have not been applied to low noise data yet \cite{carron:2017}.}
 
@@
 The larger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency.
-However, the delensing efficiency is not exactly proportional to the area of overlap here.
-Indeed each redshift contributes differently to CMB lensing because of both the noise in the measurement and the geometric properties of the lensing kernel.
-Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}
+\textcolor{blue}{However, the efficiency is not exactly proportional to the overlapping area. 
+Firstly structure at different redshifts contribute differently to the generation of large-scale B-modes because of the geometric properties of the lensing kernel, and secondly, we are not taking the tracers noise into account here.
+Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}}
 \label{fig:kernel}
@@
 While neglecting this will lead to slightly optimistic absolute constraints, it has no significant impact on the estimate of the improvement due to delensing.
-For example, running the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \cite{planck-collaboration:2016b}, change the delensing improvement factors by at most $6-7\%$.
+\textcolor{blue}{For example, running the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \cite{planck-collaboration:2016b}, change the delensing improvement factors by at most $7-8\%$.
+A more reasonable 1\% higher $\Omega h^{2}$ corresponds to variations in the improvement factors of at most 1.5\%.}
 
@@
 For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt{2}$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
+\textcolor{blue}{Here and in the following sections, we use all scales available to the instrument, even if the bulk of the signal to noise is at large scales $\ell<500$ \cite{dodelson:2014}.}
 We assume that the CMB lensing reconstruction is performed by the higher resolution CMB experiment.
@@
 Then, we test a second scenario with a ground-based experiment characterized by noise levels in polarization consistent with SPT-Pol \cite{henning:2017}: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin.
-In this case we set the angular scale range at $50<\ell<4000$.
+In this case we set the angular scale range at $50<\ell<3000$.
 However, the results are quite robust against this choice.
@@
 Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals. 
+\textcolor{blue}{In the future, with better theoretical foreground models or high signal to noise measurements, we might be able to characterize these fields (and their spectra $C_{\ell}^{\kappa I}$ and $C_{\ell}^{II}$) even at larger scales.
+Potential biases effect may be present like, for example, the correlation between the foreground contamination in galaxies and in the CMB polarization itself.
+If these are under control, we might then be able to also use the $\ell<100$ component of these tracers to delens.
+However since the lensing contribution to B-modes from LSS modes at these scales is quite small, there is not too much to gain in term of delensing efficiency.}
+
 Optical surveys galaxies like DES are less affected by dust and can be used on larger scales.
@@
 For the deep experiment we assume an instrumental noise equal to $2\mu$K-arcmin in polarization and a beam of 30 arcmin.
-The high resolution experiment will have a level of noise in polarization of $3\mu$K-arcmin and a 1 arcmin beam. This level of noise is also assumed for the internal noise reconstruction.
+The high resolution experiment will have a level of noise in polarization of $3\mu$K-arcmin and a 1 arcmin beam.
+This level of noise is also assumed for the internal noise reconstruction.
+For the CMB internal lensing reconstruction we use scales $50<\ell<3000$.
 We use an angular scale range in the Fisher matrix of $50<\ell<800$ even if most of the high $\ell$ scale do not contribute to the constraints.
@@
 Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
-For CMB S3, the improvement due to an iterative approach compared to the quadratic estimator is still minimal so that the more traditional technique can be used.
-For example, in this case, the removed power goes from $56\%$ to $54\%$ (quoted in \reftab{S3}) when using a quadratic estimator instead of the iterative one.
+\textcolor{blue}{For CMB S3, the improvement due to an iterative approach compared to the quadratic estimator is present but not significant, so that a more traditional technique can be used.
+For example, in this case, the removed power goes from $56\%$ to $51\%$ (quoted in \reftab{S3}) when using a quadratic estimator (QE) instead of the iterative one.}
 
@@
 The high-resolution experiment will have a level of noise in polarization of $1.5\mu$K-arcmin and a 1 arcmin beam.
+Here we use the same angular scale range in the Fisher as the S3 case and the CMB internal lensing reconstruction is done using scales $50<\ell<4000$.
 
@@
 Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power. 
+\textcolor{blue}{Note that, at the noise levels of CMB-S4 it will be quite valuable to develop an iterative algorithm for lensing reconstruction.
+As shown in \reftab{S4}, a quadratic estimator will only be able to remove $73\%$ of power from $C^{BB}$ compared to the 83\% of the iterative algorithm.}
 This will bring a factor of $\sim 4$ improvement in the null test case even with realistic noise.
@@
 CMB S4 (QE) (73\%) & 3.7 & 2.9  & 1.35& 2.3 \\ \hline
-CMB S4 (83\%) & 6.22 & 4.1  & 1.42& 2.3 \\ \hline
+CMB S4 (83\%) & 6.22 & 4.1  & 1.42& 1.9 \\ \hline
 LSS-S4+CMB (86\%)& 7.7 & 4.7  & 1.46 & 2.46\\ \hline

### 2018-01-22 Notes/galaxy_delensing.tex
APS proof

@@
 \newcommand{\apjs}{ApJS}
-\newcommand{\aap}{A\&A}
+\newcommand{\aap}{Astron. Astrophys.}
 \newcommand{\mnras}{MNRAS}
@@
 \email{manzotti.alessandro@gmail.com}
-\affiliation{Kavli Institute for Cosmological Physics, University of Chicago}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, 5640 South Ellis Avenue, Chicago, IL 60637, USA}
 \affiliation{Department of Astronomy and Astrophysics, University of Chicago, 5640 South Ellis Avenue, Chicago, IL 60637, USA}
@@
 The primordial B-modes component of the cosmic microwave background (CMB) polarization is a promising experimental dataset to probe the inflationary paradigm.
-B-modes are indeed a direct consequence of the presence of gravitational waves in the early universe.
-However, several secondary effects in the low redshift universe will produce \textit{non-primordial} B-modes.
+B-modes are indeed a direct consequence of the presence of gravitational waves in the early Universe.
+However, several secondary effects in the low redshift Universe will produce \textit{non-primordial} B-modes.
 In particular, the gravitational interactions of CMB photons with large-scale structures will distort the primordial E-modes, adding a lensing B-mode component to the primordial signal.
@@
 \label{sec:intro}
-In the standard cosmological paradigm the early universe underwent a period of near-exponential expansion called ``cosmic inflation''.
+In the standard cosmological paradigm the early Universe underwent a period of near-exponential expansion called ``cosmic inflation''.
 All the cosmological observations agree with this picture, making it a compelling and elegant description of the Universe initial conditions.
@@
 This prediction sets inflation apart from other theories and a detection of primordial gravitational waves could be the compelling evidence cosmologists are looking for.
-These primordial gravitational waves in the early universe would imprint a unique signature on the polarized anisotropies of the CMB.
-For this reason, CMB polarization is a promising dataset to understand the physics of the early universe and ultimately test inflation.
+These primordial gravitational waves in the early Universe would imprint a unique signature on the polarized anisotropies of the CMB.
+For this reason, CMB polarization is a promising dataset to understand the physics of the early Universe and ultimately test inflation.
 In particular, we can decompose the CMB polarization fields in Fourier space into even-parity (divergence) and odd-parity (curl) components, referred to as ``E'' and ``B'' modes.
@@
 However, just reducing the level of noise will not be enough to reach this target.
-Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe.
+Indeed, the observed B modes are not solely sourced by early Universe physics; they are also produced by secondary effects taking place in the late-time low-redshift Universe.
 \textcolor{blue}{Two main effects produce non-primordial B-modes: the polarized foregrounds from the Galaxy and the effect of gravitational interactions with large scale structures (LSS).
@@
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
-This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
+This expected component has now been measured both in cross-correlation with LSSs \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
 
@@
 Furthermore, it has recently been performed on CMB temperature data using the cosmic infrared background (CIB) as LSS tracer \cite{larsen:2016a} and on CMB temperature and polarization data using CMB data to internally reconstruct the LSS lensing potential \cite{carron:2017}.
-Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in which $~28\%$ of the lensing power was removed \cite{manzotti:2017}.
+Finally, the highest B-mode delensing efficiency has been achieved with the South Pole Telescope (SPT) and Herschel data in which $~28\%$ of the lensing power was removed \cite{manzotti:2017}.
 
@@
 In \refsec{th} we define the lensing B-mode component and the residual power after delensing with tracers of the lensing potential.
-The main result of this work is presented in \refsec{for} : the improvement of inflationary parameters constraints due to delensing with CMB and LSS.
+The main result of this work is presented in \refsec{for} : the improvement of inflationary parameters constraints due to delensing with CMB and LSSs.
 We conclude in \refsec{concl}.
@@
  Different galaxy surveys in this work are then fully characterized by their $b(z)$, $\frac{dN}{dz}$ and the observed galaxy density.
-We test the delensing efficiency taking into account both current surveys like WISE or DES as well as future galaxy surveys like DESI and LSST as well as 21 cm measurement like SKA.
+We test the delensing efficiency taking into account both current surveys like the Wide-field Infrared Survey Explorer (WISE) or the Dark Energy Survey (DES) as well as future galaxy surveys like the Dark Energy Spectroscopic Instrument (DESI) and the  Large Synoptic Survey Telescope (LSST) as well as 21 cm measurement like the Square Kilometre Array (SKA).
 
@@
 To compute the noise term, we assume that the available sky after masking is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxy density is approximately uniform.
-Furthermore we adopt a linear bias $b_{\rm{WISE}}=1.41$ obtained by \cite{ferraro:2015} cross correlating WISE with Planck lensing potential.
+Furthermore we adopt a linear bias $b_{\rm{WISE}}=1.41$ obtained by cross correlating WISE with Planck lensing potential \cite{ferraro:2015}.
 
 DES is modeled after the DES Science Verification public data release.
-For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy density of 0.63 galaxies per squared arcmin.
+For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of \cite{Aghamousa:2016zmz}. From that we can derive the galaxy density of 0.63 galaxies per squared arcmin.
 %As in \cite{passaglia:2017} we use a linear bias $b$
@@
 This model provides an accurate fit to several experimental results.
-For the cross-spectra with the CMB lensing or other galaxy tracers, $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall:2010}.
+For the cross-spectra with the CMB lensing or other galaxy tracers, $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-spectral energy distribution (SED) model of~\cite{hall:2010}.
 It corresponds to the kernel:
@@
 As shown in \cite{seljak:1997}, because of the symmetry of the problem, tensor perturbations are the principal source of the B-modes configuration.
-For this reason, primordial B-modes are a promising signature of early universe tensor perturbations.
+For this reason, primordial B-modes are a promising signature of early Universe tensor perturbations.
 
@@
 The first part of the second term in the parenthesis consists of an inverse variance filter applied to the measured E-mode.
-The smaller the noise in the E-modes ($N^{EE}$) is the closer this term is to one: a less noisy measurement improves the template of the lensing B-modes.
+The smaller the noise in the E-modes ($N^{EE}$) is the closer this term is to a value of one: a less noisy measurement improves the template of the lensing B-modes.
 The second captures the difference between the reconstructed $\phi$ and the CMB lensing potential, and it directly relates the residual power after delensing with the cross-correlation coefficients with CMB lensing of the tracers used.
@@
 
-We start by assuming that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
+We start by assuming that we have $n$ different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$.
 We can optimally combine them to estimate $\phi$ or, in other words, to maximize the correlation factor $\rho$ with \cite{sherwin:2015}:
@@
 %===================
-\subsection{Delensing with CMB Stage-2 and current LSS}
+\subsection{Delensing with CMB Stage-2 and current LSSs}
 \label{sec:del_res}
@@
 For Planck, the internal reconstruction is at most $\sim 65\%$ correlated at very large scales and then it falls rapidly to $40\%$ at $\ell=200$.
-Its correlation is comparable to the one of LSS at almost all scales.
+Its correlation is comparable to the one of LSSs at almost all scales.
 On the other end, the CMB lensing reconstruction from SPT-Pol will be more than $70\%$ correlated with the true field at $\ell<300$, and only then the correlation goes below the level of current LSS surveys.
@@
 For example, DES delensing efficiency is higher than WISE first as a result of the lower level of noise and secondly because DES galaxies are located at slightly higher redshift so they better overlap with CMB lensing.
-As expected the improvement of the internal CMB reconstruction reduces the relative importance of LSS.
+As expected the improvement of the internal CMB reconstruction reduces the relative importance of the LSSs.
 In particular it can be seen in \reffig{corrnow} that optical surveys will rapidly lose the role of filling in large-scale modes and they will start to supplement information at higher multipoles.
@@
 Using these correlation levels we can compute the residual B-mode power after delensing using \refeq{Bres}, and test the consequent improvement on parameter constraints with \refeq{fisher}.
-We only report the improvements in $r$ but we still let $n_{t}$ to vary in our fisher calculation even if it is not very degenerate with $r$.
+We only report the improvements in $r$ but we still let $n_{t}$ to vary in our Fisher calculation even if it is not very degenerate with $r$.
 We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
@@
 Nonetheless, it is worthwhile combining all the tracers.
-Indeed a multi tracer-approach can bring the removed power on the overlapping area from $30\%$ of the CIB alone to $42\%$ using all the LSS (LSS-S2) and $45\%$ when Planck lensing reconstruction is added to the set.
+Indeed a multi tracer-approach can bring the removed power on the overlapping area from $30\%$ of the CIB alone to $42\%$ using all the LSSs (LSS-S2) and $45\%$ when Planck lensing reconstruction is added to the set.
 However, if we consider a realistic level of noise in the measured B-modes, this level of residual power $C_{{l}}^{BB,\mathrm{res}}$ only leads to a 45\% improvement in $\sigma(r)$ for the null hypothesis case of $r=0$.
@@
 \reftab{current-spt} shows that, with an instrumental noise at the expected level for SPTPol, the CMB will soon be able to remove an amount of power $31\%$ at the level of the CIB.
-Once combined with LSS tracer this will lead to removing $55\%$ of the power.
+Once combined with a LSS tracer this will lead to removing $55\%$ of the power.
 The improvement of 1.6 in $\sigma(r)$ that will follow, even when realistic noise is considered, is such that it will make delensing a needed step for CMB polarization data analysis.
@@
 The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency.
-Unfortunately, adding DESI will only bring the power removed using LSS from 41 to 45\%.
+Unfortunately, adding DESI will only bring the power removed using LSSs from 41 to 45\%.
 Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
@@
 However, this technique has the potential to map the LSS with relatively low noise up to redshift $z=6$.
-Here we considered a detection thresholds at 1 GHz (flux cut) of 10 $\mu$Jy which should be representative of SKA phase 1 (SKA1) \cite{namikawa:2016a}.
+Here we considered a detection threshold at 1 GHz (flux cut) of 10 $\mu$Jy which should be representative of SKA phase 1 (SKA1) \cite{namikawa:2016a}.
 Given the importance of delensing for Stage-4 experiments, here we combine radio and optical survey to test their delensing efficiency.
@@
 excluding SKA) the improvement factor increase to 2.4.
-These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSS, completely independent of CMB internal data.
-Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of the CMB S3 internal reconstruction.
+These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSSs, completely independent of CMB internal data.
+Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSSs (62\%) to surpass the level of the CMB S3 internal reconstruction.
 Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power.
@@
 This will bring a factor of $\sim 4$ improvement in the null test case even with realistic noise.
-Combining with LSS will only increase the removed power compared to CMB only by $\simeq 4.3\%$ but it will be highly beneficial to test the robustness of the final result.
+Combining with LSSs will only increase the removed power compared to CMB only by $\simeq 4.3\%$ but it will be highly beneficial to test the robustness of the final result.
 
@@
 Given the high level of correlation of future galaxy surveys and CMB lensing, it is also worth looking for potential delensing application besides the detection of inflationary B-modes.
-For example, delensing with galaxy surveys allows us to selectively remove from the CMB only the gravitational effect coming from the non-linear structures in the low-redshift universe. 
+For example, delensing with galaxy surveys allows us to selectively remove from the CMB only the gravitational effect coming from the non-linear structures in the low-redshift Universe.
 Removing this component will reduce the level of non-linearities both in the CMB and in the CMB lensing power spectrum providing an alternative route to exploit all the measurable modes without having to model the non-linear component at small angular scales.
@@
 \bibliographystyle{mnrasunsrt}
-\bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/galaxies_delensing/Notes/spt}
+\bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/galaxies_delensing/Notes/spt,/nethome/manzotti/galaxies_delensing/Notes}
 

### 2018-01-22 Notes/referee_response.tex
journal version

@@
 \documentclass[12pt]{amsart}
-\usepackage[top=2.5 cm, bottom=2.5 cm, left=1.05in, right=1.05in]{geometry}
+\usepackage[top=2. cm, bottom=2. cm, left=1.0in, right=1.0in]{geometry}
 %\geometry{a4paper} % or letter or a5paper or ... etc
@@
 \author{A. Manzotti}
-\date{} % delete this line to display the current date
+%\date{} % delete this line to display the current date
 
@@
 \pagestyle{empty}
+\date{\today}
 \maketitle
+%\date{\today}
+I want to thank the referee for the careful review of the paper, and for the comments, corrections, and suggestions that ensued.
 
-I want to thank the referee for the very careful review of the paper, and for the comments, corrections, and suggestions that ensued.
+\vspace{0.1cm}
 
-We sequentially address all of the points raised below.
-While working on this answer, I found a minor bug in the part of the code that plots the CIB kernel in Fig 1 (the CIB kernel peak is slightly shifted).
-This is fixed now, and it does not affect the results.
+Below, we sequentially address all of the points raised.
+
+\vspace{0.1cm}
+
+While working on this answer, I found a minor bug in the part of the code that plots the CIB kernel in Fig 1 (the CIB kernel peak is then slightly shifted).
+This is fixed now, and it did not affect the results.
+Also after working on point 5 below, I switched the SKA kernel in Fig.
+1 from the optimistic 0.1 $\mu$Jy to the more realistic 10 $\mu$Jy casee.
+Furthermore, I explicitly refer to that in the label.
+The 0.1 $\mu$Jy threshold case was discussed in an earlier version of this paper and it was mistakenly left in the Figure.
+I am sorry for the confusion.
+
 \vspace{1.cm}
@@
 I agree.
-The abstract length is now reduced to $\sim 80\%$ of its previous length, mainly removing redundancy and quantitative results that can be found easily in the tables while leaving the main general points.
+The abstract length is now reduced to $\sim 80\%$ of its previous length, mainly removing redundancy and quantitative results that can be found easily in the tables.
 I believe it is still a little bit longer than the average but reasonably so.
@@
 We used all the scales available to the instrument.
-The range where the bulk of the signal to noise is, has been explored in some of the referenced work (\cite{dodelson:2014,abazajian:2016} etc.) but not here.
+The range where the bulk of the signal to noise is, has already been explored in some of the referenced work (\cite{dodelson:2014,abazajian:2016} etc.).
 
@@
 Note that if the referee meant that the galactic B-modes component is the only relevant secondary component compared to lensing B-modes, I do not think that is necessarily true. 
-While on most of the sky the galactic component is the higher contribution in power, that is not true for an opportunely chose small say 1\% patch. Furthermore, that is probably not the best metric.
+While on most of the sky the galactic component is the higher contribution in power, that is not true for an opportunely chose small (say 1\%) patch. 
+This is the size we plan for future experiments like CMB-S4.
+Furthermore, that is probably not the best metric.
 As shown in the CMB-S4 science book \cite{abazajian:2016} if one wants to maximize the $r$ constraints, lensing and foregrounds are equally important. 
@@
 However, it is unclear if eventual dust residual can cause biases in the delensed B-modes (if they correlate with dust contamination in CMB polarizations).
+This is certainly small, but maybe at the level of very futuristic CMB inflationary constraints.
 Given the small amount of delensing power for $\ell<100$ the field decided, so far, to cut those in forecasts.
-Given how much foregrounds are complicated I believed only data would answer this.
+Foregrounds are quite complicated and therefore I believed only data would answer this.
 An exploration of these on simulations is probably beyond the scope of this paper, but very soon future delensing work will test this on real data.
+
+\vspace{0.3cm}
+
+I added added a sentence to the paper regarding this possible way forward.
 
@@
 \vspace{0.4cm}
-I agree that Fig 1 may be a little confusing because it does not tell the entire story. 
+
+As pointed out at the very beginning the confusion was probably caused by a mistake: I was plotting an optimistic SKA case (with a threshold at $0.1\mu$Jy) instead of the
+$10\mu$Jy case discussed in the paper. Again, thanks for catching this.
+
+\vspace{0.2cm}
+
+However I agree that Fig 1 may be a little confusing because it does not tell the entire story. 
 The redshift overlap is not the only ingredient for the delensing efficiency.
@@
 
-However, I think Fig 1 tells most of the story. For example, the referee is right that DES is less efficient than SKA in delensing efficiency (16\%) vs (52\%).
-Furthermore, I think this figure highlights the central part (the "kernel") that characterize different tracers spectra (using Eq. 2) for people not familiar with it.
+However, I think Fig 1 tells most of the story. For example, DES is quite less efficient than SKA in delensing efficiency (16\%) vs (52\%) and SKA is one of the most efficient LSS survey.
+Furthermore, I think this figure highlights the central part (the "kernel") that characterize the calculation of different tracers spectra (using Eq. 2) for people not familiar with it.
 
@@
 Given that the Planck constraints are way tighter I consider this, as expected, a present but subdominant effect.
-For example for $1\%$ higher $\Omega h^{2}$ the efficiency values changes by at most $1.5\%$
+For example for $1\%$ higher $\Omega h^{2}$ the efficiency values changes by at most $1.5\%$.
+
 I added few sentences to the paper to report this test.
@@
 \begin{table}[htp]
-\caption{Improvement on $\sigma(r)$ constraints together with the BB power removed (in parenthesis) with and without binning.}
+\caption{Improvement on $\sigma(r)$ constraints together with the BB power removed (in parenthesis) with and without binning.
+Note that the improvement is noise-dependent but the removed power is not.}
 \begin{center}
@@
 For example, the deep 100 deg$^{2}$ field measured by SPTPol that can reach a depth of $5-6 ~\mu$K arcmin with high resolution might be the the first dataset where this can be explored.
+
 I comment on this at the end of section IIA.
+
+\vspace{0.3cm}
 

### 2018-01-22 Notes/galaxy_delensing.tex
journal version

--- 2017-12-05 Notes/galaxy_delensing.tex
@@
 Removing the lensing component ("delensing") will then be necessary to constrain the amplitude of the primordial gravitational waves.
-Here we discuss the role of current and future large-scale structure surveys in a multi-tracers approach to CMB delensing.
+Here we examine the role of current and future large-scale structure surveys in a multi-tracers approach to CMB delensing.
 We find that, in general, galaxy surveys should be split into tomographic bins as this can increase the reduction of lensing B-modes by $\sim 25\%$ in power in the most futuristic case.
 Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large-scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction.
-With the decrease of instrumental noise, the amount of lensing B-modes power removed using CMB internal reconstruction alone will rapidly increase.
+With the decrease of instrumental noise, the lensing B-modes power removed using CMB internal reconstruction alone will rapidly increase.
 Nevertheless, optical galaxy surveys will still play an important role even for CMB S4.
 In particular, an LSST-like survey can a achieve a delensing performance comparable to a 3G CMB experiment but with entirely different systematics.
-This redundancy will be essential to prove the robustness against systematics of an eventual detection of primordial B-modes.
+This redundancy will be essential to demonstrate the robustness against systematics of an eventual detection of primordial B-modes.
 
@@
 Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe.
-\textcolor{blue}{Two main effects produce non-primordial B-modes: the polarized foregrounds from the Galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
-These two components need to be treated with very different techniques, and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}
+Two main effects produce non-primordial B-modes: the polarized foregrounds emission from the Galaxy and the gravitational interactions of the primordial CMB with large scale structures (LSS). 
+These two components need to be treated with different techniques, and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
@@
 In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}. 
-\textcolor{blue}{The iterative approach has not been sufficiently demonstrated on data yet, but it is among the future goals of the CMB community \cite{abazajian:2016}.
-Note however that, as pointed out in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach mainly for S3 and S4 noise levels.
-For these two cases, we also show the delensing efficiency obtained when the CMB lensing is reconstructed with the standard quadratic estimator (QE) approach \cite{okamoto:2003,hu:2002}.
+Note that, as pointed out for example in \cite{smith:2012,hirata:2003}, this iterative approach improves the CMB lensing reconstruction if compared to a quadratic estimator (QE) approach \cite{okamoto:2003,hu:2002} for S3 and S4 noise levels.
+For these two cases, we also compute the delensing efficiency obtained when the CMB lensing is reconstructed with the standard quadratic estimator (QE).
 Furthermore, we do not deal here with the presence of an internal delensing bias.
-Several promising approaches have been proposed \cite{sehgal:2017,namikawa:2017} even if they have not been applied to low noise data yet \cite{carron:2017}.}
+Indeed we assume that we can use all the CMB scales for lensing reconstruction even if these scales are also used to constraint inflationary parameters.
+These is supported by several promising approaches \cite{sehgal:2017,namikawa:2017} although they have not been applied to low noise data yet (see \cite{carron:2017} as an example where the internal delensing bias was removed).
 
@@
 \begin{center}
-\includegraphics[scale=1.]{./images/compare_kernel.pdf}
+\includegraphics[scale=1.]{./compare_kernel.pdf}
 \caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
 The larger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency.
-\textcolor{blue}{However, the efficiency is not exactly proportional to the overlapping area. 
-Firstly structure at different redshifts contribute differently to the generation of large-scale B-modes because of the geometric properties of the lensing kernel, and secondly, we are not taking the tracers noise into account here.
-Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}}
+However, the efficiency is not exactly proportional to the overlapping area.
+Firstly, structures at different redshifts contribute differently to the generation of large-scale B-modes because of the geometric properties of the lensing kernel, and secondly, we are not taking the tracers noise into account here.
+Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}
 \label{fig:kernel}
@@
 \begin{center}
-\includegraphics[scale=1.]{./images/BB_res_ell2.pdf}
+\includegraphics[scale=1.]{./BB_res_ell2.pdf}
 \caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
@@
 \begin{center}
-\includegraphics[scale=1.]{./images/compare_kernel_tomo.pdf}
+\includegraphics[scale=1.]{./compare_kernel_tomo.pdf}
 \caption{Comparison of kernels of the 10 LSST tomographic bins together with the full LSST survey and the CMB lensing kernel. 
@@
 \begin{center}
-\includegraphics[scale=1.]{./images/B_res_bin.pdf}
+\includegraphics[scale=1.]{./B_res_bin.pdf}
 \caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing.
@@
 While neglecting this will lead to slightly optimistic absolute constraints, it has no significant impact on the estimate of the improvement due to delensing.
-\textcolor{blue}{For example, running the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \cite{planck-collaboration:2016b}, change the delensing improvement factors by at most $7-8\%$.
-A more reasonable 1\% higher $\Omega h^{2}$ corresponds to variations in the improvement factors of at most 1.5\%.}
+For example, running the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints allow \cite{planck-collaboration:2016b}, change the delensing improvement factors by at most $7-8\%$.
+A more reasonable 1\% higher $\Omega h^{2}$ corresponds to variations in the improvement factors of at most 1.5\%.
 
@@
 For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt{2}$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
-\textcolor{blue}{Here and in the following sections, we use all scales available to the instrument, even if the bulk of the signal to noise is at large scales $\ell<500$ \cite{dodelson:2014}.}
+Here and in the following sections, we use all the scales measured by the instrument, even if the bulk of the signal to noise is at large scales $\ell<500$ \cite{dodelson:2014}.
 We assume that the CMB lensing reconstruction is performed by the higher resolution CMB experiment.
@@
 Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals. 
-\textcolor{blue}{In the future, with better theoretical foreground models or high signal to noise measurements, we might be able to characterize these fields (and their spectra $C_{\ell}^{\kappa I}$ and $C_{\ell}^{II}$) even at larger scales.
+In the future, with better theoretical foreground models or high signal to noise measurements, we might be able to characterize these fields (and their spectra $C_{\ell}^{\kappa I}$ and $C_{\ell}^{II}$) even at larger scales.
 Potential biases effect may be present like, for example, the correlation between the foreground contamination in galaxies and in the CMB polarization itself.
-If these are under control, we might then be able to also use the $\ell<100$ component of these tracers to delens.
-However since the lensing contribution to B-modes from LSS modes at these scales is quite small, there is not too much to gain in term of delensing efficiency.}
+However, we expect these biases to be small, and we might then be able to also use the $\ell<100$ component of these tracers to delens.
+However since the lensing contribution to B-modes from LSS modes at these scales is quite small, there is not too much to gain in term of delensing efficiency.
 
@@
 \begin{center}
-\includegraphics[scale=1.]{./images/actual_scenario_planck.pdf}
+\includegraphics[scale=1.]{./actual_scenario_planck.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers.
@@
 \begin{center}
-\includegraphics[scale=1.]{./images/actual_scenario.pdf}
+\includegraphics[scale=1.]{./actual_scenario.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers. 
@@
 Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
-\textcolor{blue}{For CMB S3, the improvement due to an iterative approach compared to the quadratic estimator is present but not significant, so that a more traditional technique can be used.
-For example, in this case, the removed power goes from $56\%$ to $51\%$ (quoted in \reftab{S3}) when using a quadratic estimator (QE) instead of the iterative one.}
+For CMB S3, an iterative approach will improve the CMB lensing reconstruction compared to the quadratic estimator, even if not very significantly.
+For example, in this case, the removed power goes from $56\%$ to $51\%$ (quoted in \reftab{S3}) when using a quadratic estimator (QE) instead of the iterative one.
 
@@
 \begin{center}
-\includegraphics[scale=1.]{./images/S3_scenario.pdf}
+\includegraphics[scale=1.]{./S3_scenario.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
@@
 \begin{center}
-\includegraphics[scale=1.]{./images/S4_scenario_ska10.pdf}
+\includegraphics[scale=1.]{./S4_scenario_ska10.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
@@
 Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power. 
-\textcolor{blue}{Note that, at the noise levels of CMB-S4 it will be quite valuable to develop an iterative algorithm for lensing reconstruction.
-As shown in \reftab{S4}, a quadratic estimator will only be able to remove $73\%$ of power from $C^{BB}$ compared to the 83\% of the iterative algorithm.}
+Note that, at the noise levels of CMB-S4, it will be quite valuable to use an iterative algorithm for the lensing reconstruction.
+As shown in \reftab{S4}, a quadratic estimator will only be able to remove $73\%$ of power from $C^{BB}$ compared to the 83\% of the iterative approach.
 This will bring a factor of $\sim 4$ improvement in the null test case even with realistic noise.
@@
 \begin{figure*}[ht]
-\includegraphics[scale=0.8]{./images/errors_summary2.pdf}
+\includegraphics[scale=0.8]{./errors_summary2.pdf}
 \caption[]{A summary of the amount of lensing B-modes power removed by different tracers for different generations of experiments.

### 2018-02-10 Notes/galaxy_delensing.tex

@@
 However, just reducing the level of noise will not be enough to reach this target.
-Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe.
-Two main effects produce non-primordial B-modes: the polarized foregrounds emission from the Galaxy and the gravitational interactions of the primordial CMB with large scale structures (LSS). 
-These two components need to be treated with different techniques, and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).
+
+Indeed, the observed B modes are not solely sourced by early Universe physics; they are also produced by secondary effects taking place in the late-time low-redshift Universe.
+Two main effects produce non-primordial B-modes: the polarized foregrounds from the Galaxy and the effect of gravitational interactions with large scale structures (LSS).
+These two components need to be treated with very different techniques, and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).
 Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
-This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
+This expected component has now been measured both in cross-correlation with LSSs \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.
 
@@
 To compute the noise term, we assume that the available sky after masking is around $f_{sky}=0.44$ with 50 million galaxies \cite{ferraro:2015} and that the galaxy density is approximately uniform.
-Furthermore we adopt a linear bias $b_{\rm{WISE}}=1.41$ obtained by \cite{ferraro:2015} cross correlating WISE with Planck lensing potential.
+Furthermore we adopt a linear bias $b_{\rm{WISE}}=1.41$ obtained by cross correlating WISE with Planck lensing potential \cite{ferraro:2015}.
 
 DES is modeled after the DES Science Verification public data release.
-For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of the DESI Technical Design Report. From that we can derive the galaxy density of 0.63 galaxies per squared arcmin.
-%As in \cite{passaglia:2017} we use a linear bias $b$
+For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of \cite{Aghamousa:2016zmz}. From that we can derive the galaxy density of 0.63 galaxies per squared arcmin.
 
@@
 \begin{center}
-\includegraphics[scale=1.]{./compare_kernel.pdf}
+
+\includegraphics[scale=1.]{./images/compare_kernel.pdf}
 \caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis.
@@
 However, the efficiency is not exactly proportional to the overlapping area.
-Firstly, structures at different redshifts contribute differently to the generation of large-scale B-modes because of the geometric properties of the lensing kernel, and secondly, we are not taking the tracers noise into account here.
+Firstly structure at different redshifts contribute differently to the generation of large-scale B-modes because of the geometric properties of the lensing kernel, and secondly, we are not taking the tracers noise into account here.
 Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}
@@
 \begin{center}
-\includegraphics[scale=1.]{./BB_res_ell2.pdf}
+\includegraphics[scale=1.]{./images/BB_res_ell2.pdf}
 \caption{Here we illustrate the effect of delensing on the B-mode power spectrum.
@@
 \begin{center}
-\includegraphics[scale=1.]{./compare_kernel_tomo.pdf}
+
+\includegraphics[scale=1.]{./images/compare_kernel_tomo.pdf}
 \caption{Comparison of kernels of the 10 LSST tomographic bins together with the full LSST survey and the CMB lensing kernel.
@@
 \begin{center}
-\includegraphics[scale=1.]{./B_res_bin.pdf}
+\includegraphics[scale=1.]{./images/B_res_bin.pdf}
 \caption{Tomographic bins improve the cross-correlation of galaxy surveys with CMB lensing.
@@
 We combine these CMB experiments with the CIB and current low redshift galaxy surveys like DES and WISE and we compute the improvement in the delensing efficiency.
+
 Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals. 
@@
 \begin{center}
-\includegraphics[scale=1.]{./actual_scenario_planck.pdf}
+\includegraphics[scale=1.]{./images/actual_scenario_planck.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers.
@@
 \begin{center}
-\includegraphics[scale=1.]{./actual_scenario.pdf}
+
+\includegraphics[scale=1.]{./images/actual_scenario.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$ for completed and ongoing Stage-2 tracers.
@@
 \begin{center}
-\includegraphics[scale=1.]{./S3_scenario.pdf}
+\includegraphics[scale=1.]{./images/S3_scenario.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
@@
 \begin{center}
-\includegraphics[scale=1.]{./S4_scenario_ska10.pdf}
+\includegraphics[scale=1.]{./images/S4_scenario_ska10.pdf}
 \caption{Correlation factor $\rho$ with the CMB lensing potential as a function of the angular scale $\ell$.
@@
 excluding SKA) the improvement factor increase to 2.4.
+
 These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSS, completely independent of CMB internal data.
@@
 \begin{figure*}[ht]
-\includegraphics[scale=0.8]{./errors_summary2.pdf}
+\includegraphics[scale=0.8]{./images/errors_summary2.pdf}
 \caption[]{A summary of the amount of lensing B-modes power removed by different tracers for different generations of experiments.
@@
 \bibliographystyle{mnrasunsrt}
-\bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/galaxies_delensing/Notes/spt}
+\bibliography{/Users/alessandromanzotti/Work/Astrophysics/latex_bib/cosmobib,/Users/alessandromanzotti/Work/Astrophysics/galaxies_delensing/Notes/spt,/Users/alessandromanzotti/Work/Astrophysics/galaxies_delensing/Notes/desi}
 

### 2018-04-13 Notes/galaxy_delensing.tex
adding 3G MSIP

@@
 DES is modeled after the DES Science Verification public data release.
-For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of \cite{Aghamousa:2016zmz}. From that we can derive the galaxy density of 0.63 galaxies per squared arcmin.
+For DESI we used the $\frac{dN}{dz}$ in Tab. 2.3 of \cite{desi-collaboration:2016}. From that we can derive the galaxy density of 0.63 galaxies per squared arcmin.
 
@@
 The first part of the second term in the parenthesis consists of an inverse variance filter applied to the measured E-mode.
-The smaller the noise in the E-modes ($N^{EE}$) is the closer this term is to a value of one: a less noisy measurement improves the template of the lensing B-modes.
+The smaller the noise in the E-modes ($N^{EE}$) is the closer this term is to a value of 1: a less noisy measurement improves the template of the lensing B-modes.
 The second captures the difference between the reconstructed $\phi$ and the CMB lensing potential, and it directly relates the residual power after delensing with the cross-correlation coefficients with CMB lensing of the tracers used.

### 2018-04-13 Code/multiple_survey_delens__SPT_MSIP.py

@@ -1,0 +1,145 @@
+'''
+This take the power spectra cross with CMB lensign and auto of different surveys ans spits out their rho as in Shwerwin Smithful
+'''
+
+import numpy as np
+import scipy.integrate
+from scipy.interpolate import RectBivariateSpline, interp1d, InterpolatedUnivariateSpline
+import pickle
+
+
+def bl(fwhm_arcmin, lmax):
+    """ returns the map-level transfer function for a symmetric Gaussian beam.
+         * fwhm_arcmin      - beam full-width-at-half-maximum (fwhm) in arcmin.
+         * lmax             - maximum multipole.
+    """
+    ls = np.arange(0, lmax + 1)
+    return np.exp(-(fwhm_arcmin * np.pi / 180. / 60.)**2 / (16. * np.log(2.)) * ls * (ls + 1.))
+
+
+def nl(noise_uK_arcmin, fwhm_arcmin, lmax):
+    """ returns the beam-deconvolved noise power spectrum in units of uK^2 for
+          * noise_uK_arcmin - map noise level in uK.arcmin
+          * fwhm_arcmin     - beam full-width-at-half-maximum (fwhm) in arcmin.
+          * lmax            - maximum multipole.
+    """
+    return (noise_uK_arcmin * np.pi / 180. / 60.)**2 / bl(fwhm_arcmin, lmax)**2
+
+
+def main(labels, year, spectra_file='../Data/limber_spectra_delens.pkl'):
+
+    output_dir = '/home/manzotti/galaxies_delensing/Data/'
+    cls = pickle.load(open(spectra_file, 'rb'))
+    lbins = np.load('../Data/ells.npy')
+
+    ckk = cls['kk']
+    # initialize
+    rho_comb = np.zeros((np.size(lbins)))
+    surveys = labels
+    # surveys = ['wise', 'k', 'euclid', 'des_weak', 'lsst', 'ska10', 'ska01',
+    #            'ska5', 'ska1', 'cib', 'desi', 'des']
+    cl_cross_k = {}
+    cl_auto = {}
+    rho = {}
+    for label in surveys:
+        cl_cross_k[label] = np.array(cls['k' + label])
+        cl_auto[label] = np.array(cls[label + label])
+        if label == 'wise':
+            cl_cross_k[label][np.where(lbins < 100)] = 0.
[…]

### 2018-04-13 Code/run_analysis.py

@@ -300,13 +300,11 @@ clee_fun = InterpolatedUnivariateSpline(
     ells_cmb[:5000], np.nan_to_num(cle[:5000]), ext=2)
 
 
-inifile = '/home/manzotti/cosmosis/modules/limber/galaxies_delens.ini'
-Config_ini = ConfigParser.ConfigParser()
-values = ConfigParser.ConfigParser()
-Config_ini.read(inifile)
-values_file = Config_ini.get('pipeline', 'values')
-output_dir = Config_ini.get('test', 'save_dir')
-
+# inifile = '/home/manzotti/cosmosis/modules/limber/galaxies_delens.ini'
+# Config_ini = ConfigParser.ConfigParser()
+# values = ConfigParser.ConfigParser()
+# Config_ini.read(inifile)
+output_dir = "../data/limber_spectra/"
 # ==========================================
 # ==========================================
 # L range to be used in the fisher forecasts

### 2018-04-13 Code/run_galaxies_delens_SPT_MSIP.py

@@ -193,19 +193,19 @@ def main(ini_par):
     pars = camb.CAMBparams()
     # This function sets up CosmoMC-like settings, with one massive neutrino
     # and helium set using BBN consistency
-    pars.set_cosmology(H0=67.26, ombh2=0.02222, omch2=0.1199,
-                       mnu=0.06, omk=0, tau=0.079)
-    pars.InitPower.set_params(ns=0.96, r=0., nt=0, pivot_tensor=0.01, As=2.1e-9)
-    pars.set_for_lmax(5000, lens_potential_accuracy=3)
+    pars.set_cosmology(H0=67.94, ombh2=0.022199, omch2=0.11847,
+                       mnu=0.06, omk=0, tau=0.0943)
+    pars.InitPower.set_params(ns=0.9624, r=0., nt=0, pivot_tensor=0.05, As=2.208061336e-9)
+    pars.set_for_lmax(8000, lens_potential_accuracy=3)
     # pars.set_for_lmax?
 
     pars.NonLinear = model.NonLinear_both
-    pars.set_matter_power(redshifts=np.linspace(0., 13, 50), kmax=5.0)
+    pars.set_matter_power(redshifts=np.linspace(0., 13, 50), kmax=15.0)
     results = camb.get_results(pars)
 
     # P(z,k)
     kh_nonlin, z_nonlin, pk_nonlin = results.get_matter_power_spectrum(
-        minkh=1e-6, maxkh=5, have_power_spectra=False, npoints=250)
+        minkh=1e-6, maxkh=10, have_power_spectra=False, npoints=250)
     rbs = RectBivariateSpline(kh_nonlin, z_nonlin, pk_nonlin.T)
     h = pars.H0 / 100.
 
@@ -405,9 +405,8 @@ def main(ini_par):
     kernels = [lkern, wise, euclid, lsst, ska10, cib, desi, des]
     names = ['k', 'wise', 'euclid', 'lsst', 'ska10', 'cib', 'desi', 'des']
 
-    kernels = [ ]
+    kernels = []
     names = []
-
 
     # kernels = [lkern, desi]
     # names = ['k', 'desi']
@@ -441,12 +440,10 @@ def main(ini_par):
             kernel_list.append([kernels[i], kernels[j]])
     print(labels[40:])
 
-
     import warnings
     warnings.filterwarnings('error')
 
-
[…]
